### 2014-04-03 notes_signaling_model_01_20_14.tex
setting up signaling network repository

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}
\newtheorem{ass}{Assumption}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf \LARGE{Development of a Signaling Network}}
\vspace{10pt}
\\ Eleanor Brush
\\ updated December 12, 2012
\end{center}

\tableofcontents

\vspace{0pt}
\normalsize

\section{Introduction}
In some species of macaques, there is a stereotyped signal that one animal emits to another once it has learned, through a series of fights, that it is likely to lose subsequent fights.  The signal communicates the signaler's agreement to the subordinate role in the relationship.  The animals then use the signals they observe being transmitted between other pairs of animals in the group to estimate the relative power of the other group members.  In different species of macaque, the network of signals transmitted between every pair and power distribution are more or less skewed.  I have previously worked on ways to measure the amount of consensus present in an observed signaling network.  I am now studying the development of the signaling network.  I am developing and starting to analyze a model of stochastic fighting and signaling dynamics.  Using this model, I hope to be able to understand under what circumstances the animals should or should not signal and to explain how the traits of the individuals and the distribution of these traits in the group affect the resulting signaling network.  From encounters, fights, and signals that occur probabilistically, I can simulate and derive stochastic differential equations describing how each animal's estimate of its relative ability changes over time (Equation \ref{sdes}).   From these equations I can find the expected time until one animal signals and the probability that one or the other animal will be the first to signal.  I can then find the thresholds that optimize these pairwise dynamics and how that optimization affects the resulting network of signals among the whole population.  Many scientists have used this kind of drift-diffusion model to describe learning dynamics in neurons.  People have made analogies between social systems and networks of neurons and I would like to use this model to explore the analogy, identifying similarities, studying the consequences of differences in the dynamics, and hopefully probing the limits to which the analogy can be pushed.  The possibility of the animals or neurons communicating with each other represents a feedback between the variables not present in most models of neuronal learning and I am interested in understanding, generally, what the effect of this feedback is on the ability of individuals and groups of individuals to learn about each other and their environment.

This model is interesting for (at least) two reasons.  First, we hope to be able to explain variation we see at the group or species level in primates, namely characteristic signaling behavior and signaling networks, with differences in individual level properties, distributions of those properties, and the varying benefits or costs of fighting and storing information, i.e. to use individual level variation to explain emergent properties of species.  Second, we hope to use this model as a tool to explore the analogy between neural and social systems.

\section{Stochastic Model}
\subsection{Derivation of Stochastic Differential Equations}
We want to write down equations for how the two animal's estimates of their dominance change over time.  We assume that in the absence of any information, an animal's belief about its own dominance leaks back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.) If nothing else happened, then over a period of time of length $\tau$ an estimate would decrease as $X(t+\tau)=(1-\ell\tau)X(t)$; consequently $X(t)=\exp(-\ell t)$. Two things can happen to change an animal's estimate: it can receive a signal, boosting its estimate, or it can engage in a fight and learn about its relative dominance from the outcome of the fight.  To determine the estimate at time $t+\tau$, we can count how many times each of these events occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X(t+\tau)&=(1-\ell\tau)X(t)+b_{\text{s}}\times\# \text{ of signals received in }[t,t+\tau)
\\&+b_{\text{f}}\times\#\text{ of fights won in }[t,t+\tau)-b_{\text{f}}\times\#\text{ of fights lost in }[t,t+\tau)
\end{align*}

The animals encounter each other at a constant rate over time.  When the animals encounter each other, each animal has a probability of transmitting a signal.  (We do not preclude the possibility of both animals signaling.)  If no signals are transmitted, then the encounter will escalate to a fight with some probability.  If the animals fight, the probability that the ``stronger" animal will win is determined by how much stronger or more ``dominant" he is.  We write $N_{\text{s},i}$ for the random variable giving the number of signals transmitted by animal $i$ and $N_{\text{f},i}$ for the random variable giving the number of fights won by animal $i$.  This very general description of our model is sufficient to perform simulations of these stochastic processes.  To write down equations for the probability densities of the estimates, however, we make a few assumptions.

If the rates at which the pair of animals encounters each other, fight with each other, and signal to each other do not change from $t$ to $t+\tau$ then the numbers of these events follow Poisson distributions, as shown in the Appendix.  
\begin{ass}
There is a timescale $\tau$ over which the probabilities of the animals encountering each other, engaging in fights with each other, and transmitting signals do not change.
\end{ass}

Putting this all together gives that 
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b_{\text{s}}N_{\text{s},2}+b_{\text{f}}N_{\text{f},1}-b_{\text{f}}N_{\text{f},2}
\\ \text{ and } X_2(t+\tau)&=(1-\ell\tau)X_2(t)+b_{\text{s}}N_{\text{s},1}-b_{\text{f}}N_{\text{f},1}+b_{\text{f}}N_{\text{f},2}.
\end{align*}
Somewhat counterintuitively, even though the transmission of a signal and a fight are mutually exclusive types of encounters, the numbers of these events are independent random variables, each following a Poisson distribution, as shown in the Appendix.  $N_{\text{s},1}$, $N_{\text{s},2}$, $N_{\text{f},1}$ and $N_{\text{f},1}$ are therefore independent Poisson random variables with respective means $\tau ep_{\text{s},1}(t)$, $\tau ep_{\text{s},2}(t)$, $\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d$, and $\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)$.  If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with Normal random variables with mean and variance equal to the mean of the Poisson random variables.

\begin{ass}
In the period of time of length $\tau$ enough events happen that we can approximate the Poisson random variables with normally distributed random variables.
\end{ass}
Let $Z_{\text{s},1}$, $Z_{\text{s},2}$, $Z_{\text{f},1}$, and $Z_{\text{f},2}$ be independent standard Normal random variables, i.e. with mean $0$ and standard deviation $1$.  Then our equations for $X_i(t+\tau)$ become
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b_{\text{s}}\bigg(\tau ep_{\text{s},2}(t)+\sqrt{\tau ep_{\text{s},2}(t)}Z_{\text{s},2}\bigg)
\\&+b_{\text{f}}\bigg(\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d+\sqrt{\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d}Z_{\text{f},1}\bigg)
\\&-b_{\text{f}}\bigg(\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)+\sqrt{\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)}Z_{\text{f},2}\bigg),
\\ \text{ and } X_2(t+\tau)&=(1-\ell\tau)X_2(t)+b_{\text{s}}\bigg(\tau ep_{\text{s},1}(t)+\sqrt{\tau ep_{\text{s},1}(t)}Z_{\text{s},2}\bigg)
\\&-b_{\text{f}}\bigg(\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d+\sqrt{\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d}Z_{\text{f},1}\bigg)
\\&+b_{\text{f}}\bigg(\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)+\sqrt{\tau e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)}Z_{\text{f},2}\bigg).
\end{align*}
Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+b_{\text{s}}ep_{\text{s},2}(t)+b_{\text{f}}(2d-1)e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_\text{f}\bigg)dt+b_{\text{s}}\sqrt{ep_{\text{s},2}(t)}dW_{\text{s},2}t
\\&+b_f\sqrt{e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d}dW_{\text{f},1}t-b_{\text{f}}\sqrt{e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)}dW_{\text{f},2}t
\\dX_2&=\bigg(-\ell X_2(t)+b_{\text{s}}ep_{\text{s},1}(t)+b_{\text{f}}(1-2d)e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_\text{f}\bigg)dt+b_{\text{s}}\sqrt{ep_{\text{s},1}(t)}dW_{\text{s},1}t
\\&-b_f\sqrt{e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}d}dW_{\text{f},1}t+b_{\text{f}}\sqrt{e(1-p_{\text{s},1}(t))(1-p_{\text{s},2}(t))p_{\text{f}}(1-d)}dW_{\text{f},2}t,
\end{array}
\end{equation}
where $dW_{\text{s},i}$ and $dW_{\text{f},i}$ and independent Brownian motions representing, respectively, the signaling of animal $i$ and animal $i$ winning fights.

\begin{table}
\caption{\label{variables}}
\begin{tabular}{llll}
Variable & Meaning  \\
\hline $e$ & encounter rate
\\ $p_{\text{s},i}(t)$ & probability of animal $i$ transmitting a signal as a function of the animal's estimate at $t$
\\ $p_{\text{f}}$ & probability of fighting
\\ $d$ & probability of stronger animal winning (``dominance'')
\\ $b_{\text{s}}$ & boost to one's estimate from receiving a signal
\\ $b_{\text{f}}$ & boost to one's estimate from winning a fight
\\ $\ell$ & leak or loss rate
\end{tabular}
\end{table}

\subsection{PDEs for Hitting Probabilities and Times}

\section{Pairwise Dynamics}

\subsection{1D v 2D}
\begin{figure}
\includegraphics[width=\textwidth]{1d_vs_2d.pdf}
\caption{ For higher leak rates (right panel) or lower difference in fighting abilities (higher curves), two-dimensional decision making takes longer than one-dimensional decisions (blue vs red).}
\end{figure}

\subsection{Optimization of Threshold in Pairs}

\subsection{Optimization of Threshold in a Group}

\section{Network Dynamics}

\section{Discussion}

\section{Appendix}
\begin{fact}
If the probability of an event occurring in $[t,t+dt)$ is $rdt$ then the number of events occurring in a period of time of length $t$ is a Poisson random variable with mean $rt$, i.e. the probability of $n$ events occurring is 
$$P(n,t)=\frac{e^{-rt}(rt)^n}{n!}$$
\end{fact}
\begin{pf}
\end{pf}

\begin{fact}
$N(\lambda,t)$ is a Poisson random variable giving the number of events that happen in a time $t$ if events occur at a rate $\lambda$.  If, when one of those events occurs, it can be categorized as one of two mutually exclusive events which occur with probabilities $p_1,p_2$ with $p_1+p_2\leq 1$ and $N_i(p_i,\lambda,t)$ is the random variable giving the number of events categorized as $i$ in time $t$, then $N_1(p_i,\lambda,t),N_2(p_2,\lambda,t)$ are independent Poisson random variables with rates $p_1\lambda$ and $p_2\lambda$.
\end{fact}
\begin{pf}
To show that $N_1,N_2$ are independent it suffices to show that, for all $n,m\in\N$, $P(N_1=n,N_2=m)=P(N_1=n)P(N_2=m)$.
\end{pf}

\nocite{*}
\bibliographystyle{plain}
\bibliography{signaling_model}

\end{document}

### 2014-04-03 paper_skeleton.tex
setting up signaling network repository

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{natbib}

\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}
\newtheorem{ass}{Assumption}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf \LARGE{Stochastic Learning Dynamics with Two Applications}}
\end{center}

\tableofcontents

\section{Abstract}
Learning usually occurs in noisy environments.  In order to make a decision, one must accumulate the  evidence for different alternatives and choose one only once the uncertainty about the choice is sufficiently reduced.  There are two equivalent mathematical formulations of the optimal decision making algorithm: a sequential probability ratio test in which the likelihoods of two hypotheses are evaluated as more data points are collected and a drift-diffusion stochastic differential equation in which the belief about two hypotheses does a random walk until a threshold is reached.  In the random walk formulation, the threshold dictating when a decision has been made affects both the accuracy of the decision and the expected time until a decision is reached.  If the utility of a decision depends on accuracy and decision time, the threshold can be optimized accordingly.  There are two biological systems in which the variables used to make a decision can be described by a random walk: the firing rates of neural populations in the brain of an individual making a decision about a visual stimulus and the beliefs of monkeys about their relative dominance based on the fights they have engaged in.  The stochastic differential equations used to model these quite different systems are nearly identical, which suggests that there are universal principles underlying learning through noisy dynamical systems in biology.  However, there are (at least) three significant differences between the two models.  First, in the social system, there is a strategic component to the system since the optimal threshold of one animal depends on the threshold of the other animal, whereas in the neural system both neural populations are part of the cognitive system of the same individual trying to optimize its decision making.  Relatedly, the output of the decision has additional strategic consequences in the social system.  Second, whereas in the neural system, the differential equations are often reduced to a one-dimensional model, the social system is inherently two-dimensional.  Finally, in the social system, a monkey is not only trying to make a decision about his dominance with respect to one other monkey, but he is additionally trying to make decisions about many other partners at once, whereas in the neural case there is only one decision to be made at a time.  As a consequence of these differences... In comparing the output of the model of the social system to data, we found that ...

\section{Introduction}

\section{Observations }

\subsection{Neural }
Stochastic differential equations have been used to model, explain, and predict learning and decision making in both monkeys and humans. \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}

\subsection{Social }

\subsection{Extent and Limits of Analogy }

\begin{table}\centering
\caption{\label{analogy}{\bf  Analogy between neural and social systems.} }
\ra{1.3}
\begin{tabular}{@{}rllll@{}}
&Neural &  \phantom{abc}& Social \\
\cmidrule{2-2} \cmidrule{4-4} 
variables &  firing rates of neural populations && opinions about relative dominance
\\ input & dots moving left or right && fights won or lost
\\ 
\end{tabular}
\end{table}

\section{Derivation of Model}
With two assumptions about the timescales on which probabilistic events occur, a simple mechanistic description of the integration of information about random events can be turned into a stochastic differential equation \cite{Gillespie:2000fk}.  A drift diffusion model is equivalent to a sequential probability ratio test, a learning algorithm that is optimal with respect to its accuracy and the time required to make a decision \cite{Moehlis:2004fk,Bogacz:2006uq}.

\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+b_\text{f}(2a-1)\bigg)dt+b_f\sqrt{a}dW_{\text{f},1}t-b_{\text{f}}\sqrt{(1-a)}dW_{\text{f},2}t
\\dX_2&=\bigg(-\ell X_1(t)+b_\text{f}(1-2a)\bigg)dt-b_f\sqrt{a}dW_{\text{f},1}t+b_{\text{f}}\sqrt{(1-a)}dW_{\text{f},2}t,
\end{array}
\end{equation}

\begin{table}\centering
\caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} }
\ra{1.3}
\begin{tabular}{@{}rllll@{}}
& \multicolumn{2}{c}{Neural} & \phantom{abc}& Social \\
\cmidrule{2-3} \cmidrule{5-5} 
model  & DDM or OU & race && race with leak
\\dimensionality & $1$ & $2$ && $2$
\\decision & difference hits a threshold & one var. hits a threshold && one var. hits a threshold
\\ criterion &  reward from being ``right" &&& reward from being ``right"
\\ & reward rate &&& reward rate
\\ & Bayes risk &&& Bayes risk
\\ & &&& \fcolorbox{red}{white}{reward from  receiving signal}
\\depends on & input strength &&& input strength
\\ & noise &&& noise
\\ & leak rate &&& leak rate
\\ & &&& \fcolorbox{red}{white}{other animal's threshold}
\\threshold  algorithm & error correction
\\  & hill climbing
\end{tabular}
\end{table}

\section{Results}

\subsection{Two-Dimensional Decision Making }
In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to a one-dimensional equation describing the difference in the activity levels to make the system easier to analyze \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.  However, in the social system we are considering, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.  A ``decision," i.e. the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below a given threshold.  The social system is therefore inherently two-dimensional, whereas the neural system may be adequately described by one dimension.

We can compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other. We find that...

\subsection{Strategic Decision Making }
In the neural system, an individual wants to make both a decision that is both quick and accurate.  This is also true in the social system.  However, it may also be the case that a monkey wants to avoid emitting a subordination signal and wants to receive a signal, regardless of whether or not this is the ``correct" outcome, i.e. whether or not he is actually more likely to win a fight.  The criterion used to optimize the threshold, therefore, may be quite different in the two systems.  Additionally, the speed and outcome of the decision process depend on both animals' thresholds, so there is a game theoretic aspect to an animal's optimization in that his best behavior depends on the behavior of the other animal. Optimizing the threshold is thus strategic in the social system but not in the neural system.

When we optimize the threshold according to different criteria...

When we allow the optimization to depend on the partner's threshold...

\subsection{Social Decision Making }
A monkey can try to decide whether it is more or less likely to win a fight than another monkey using information from the fights the pair has previously engaged in.  Further, he can try to make such a decision about every other member of its social group.  Since he is trying to use the same algorithm to learn about every other animal, he is constrained to have only one decision threshold for each of these decision processes.  The optimal decision threshold in this social context depends on the distribution of fighting abilities in the group and the decision thresholds of the other animals.

\subsection{Comparison to Social Data }
Empirically, those pairs of monkeys that take the longest to establish a signaling dynamics are those with the most similar fighting abilities...

\section{Discussion}

\nocite{*}
\bibliographystyle{plain}
\bibliography{signaling_model}

\end{document}

### 2014-04-03 dimensionality_comparison.m

@@ -1,0 +1,206 @@
+deltax1=0.01;
+deltax2=0.025 ;
+b=1;
+
+lvals=0:.5:2;
+Nl=length(lvals);
+
+domvals=.5:.1:1;
+Nd=length(domvals);
+
+threshvals1=.1:.1:8;
+Nt1=length(threshvals1);
+
+threshvals2=.5:.5:2;
+Nt2=length(threshvals2);
+
+%%
+
+oneoutput=zeros(Nl,Nd,Nt1,2);
+twooutput=zeros(Nl,Nd,Nt2,2);
+
+for k=1:Nl
+    l=lvals(k);
+    for i=1:Nd
+        d=domvals(i);
+        for j=1:Nt1
+            t=threshvals1(j);
+            [x,y,~]=solve_pde_1d(-t,d,t,deltax1,l,b);
+            oneoutput(k,i,j,:)=[x y];
+        end
+        for j=1:Nt2
+            t=threshvals2(j);
+            [x,y,~]=solve_pde_2d(-t,-t,d,5,5,deltax2,deltax2,l,b);
+            twooutput(k,i,j,:)=[x y];
+        end
+    end
+end
+%%
+filename=strcat('two_v_one_output','.mat');
+save(filename,'twooutput','oneoutput','domvals','lvals','threshvals1','threshvals2');
+%%
+axislw=1.25;
+lw=3;
+labfontsz=15 ;
+textfontsz=15;
+alpha=.7;
+ticklength=[.02,.0250];
+markersz=5;
+markercol='white';
+rho=4.2/5.6;
+w=8.5;
+h=w*rho;
+
+for k=Nl:-1:1
+    figure
+    set(gcf,'Color','w')
+    v=get(gcf,'Position');
+    ratio=v(4)/v(3);
+    w=6.83*2;
+    h=w*ratio;
+    hold on
+    for i=2:Nd
+    plot(reshape(twooutput(k,i,:,1),1,[]),reshape(twooutput(k,i,:,2),1,[]),'blue','LineWidth',lw)
+    plot(reshape(oneoutput(k,i,1:40,1),1,[]),reshape(oneoutput(k,i,1:40,2),1,[]),'red','LineWidth',lw)
+    end
+    xlabel('Probability of correct output','FontSize',textfontsz)
+    ylabel('Expected time to output','FontSize',textfontsz)
[…]

### 2014-04-04 dimensionality_comparison.m
coded what i think will be the first figure

@@ -1,3 +1,8 @@
+green=importdata('green.txt');
+blue=importdata('blue.txt');
+red=importdata('red.txt');
+black=importdata('black.txt');
+%%
 deltax1=0.01;
 deltax2=0.025 ;
 b=1;
@@ -49,26 +54,57 @@ markersz=5;
 markercol='white';
 rho=4.2/5.6;
 w=8.5;
-h=w*rho;
+h=w*ratio;
+figure
+set(gcf,'Color','w')
+v=get(gcf,'Position');
+ratio=v(4)/v(3);
+w=6.83*2;
+h=w/2;
+set(gcf,'Units','inches');
+set(gcf,'Position',[v(1) v(2) w h]);
 
-for k=Nl:-1:1
-    figure
-    set(gcf,'Color','w')
-    v=get(gcf,'Position');
-    ratio=v(4)/v(3);
-    w=6.83*2;
-    h=w*ratio;
+textlabels={'a' 'b'};
+increment=floor(size(blue,1)/Nd);
+
+p=[];
+
+for k=1:2
+    subplot_tight(1,4,(1:2)+2*(k-1),[.1 .07])
+    set(gca,'FontSize',labfontsz)
     hold on
-    for i=2:Nd
-    plot(reshape(twooutput(k,i,:,1),1,[]),reshape(twooutput(k,i,:,2),1,[]),'blue','LineWidth',lw)
-    plot(reshape(oneoutput(k,i,1:40,1),1,[]),reshape(oneoutput(k,i,1:40,2),1,[]),'red','LineWidth',lw)
+    switch k
+        case 2
+            for i=2:Nd
+            p(i)=plot(reshape(twooutput(k,i,:,2),1,[]),reshape(twooutput(k,i,:,1),1,[]),'Color',blue(i,:),'LineWidth',lw);
+            p(i+Nd)=plot(reshape(oneoutput(k,i,1:40,2),1,[]),reshape(oneoutput(k,i,1:40,1),1,[]),'Color',red(i,:),'LineWidth',lw)
+            end
+        otherwise 
+            for i=2:Nd
+            plot(reshape(twooutput(k,i,:,2),1,[]),reshape(twooutput(k,i,:,1),1,[]),'Color',blue(i,:),'LineWidth',lw);
+            plot(reshape(oneoutput(k,i,1:40,2),1,[]),reshape(oneoutput(k,i,1:40,1),1,[]),'Color',red(i,:),'LineWidth',lw)
+            end
     end
[…]

### 2014-04-09 paper_skeleton.tex
coded what i think will be the first figure
fleshing out the introduction to the two systems

@@
 
-\section{Observations }
-
 \subsection{Neural }
-Stochastic differential equations have been used to model, explain, and predict learning and decision making in both monkeys and humans. \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}
+Stochastic differential equations have been used to model, explain, and predict learning and decision making in both monkeys and humans \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
+Typically, an experimental subject is presented with a visual stimulus that consists of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.
+The subject is then asked to decide in which direction the dots were moving.
+If there are two neural populations, each of which fire in response to a particular direction of motion, the firing rates of the two populations will tend to increase or decrease, depending on the strength and direction of the stimulus.
+Eventually, either the experimenter forces a decision to be made, in which case whichever population has a higher firing rate determines which decision is made, or the firing rate of one population becomes high enough that the subject decides for that direction.
+To summarize, there are two decision variables--- the firing rates of two neural populations responding to left and right motion--- that perform a biased random walk until enough evidence has been accumulated to make a decision one way or the other.
 
 \subsection{Social }
+In groups of pigtailed macaques, different animals have different fighting abilities, and when a pair of animals engages in a fight, one animal may be more likely on average to ``win" the fight.
+Over time, as one animal learns that it is likely to lose fights, it will decide to emit a formal subordination signal, communicating its agreeing to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk}.
+The outcome of a fight is affected by each animal's allies, the presence of those allies, motivation levels,  external environmental variables, and other factors.
+Additionally, the number of fights that occur in any period of time depends on similarly stochastic factors.
+Because of this stochasticity, the number of fights won and lost in any period of time is a random variable.
+Each animal's opinion about its relative dominance, therefore, will perform a biased random walk until one animal's opinion of itself is low enough to agree to being subordinate.
+
+Similar models have previously been applied to animal conflicts, although in that case the animals had perfect memory of their past interactions \citep{Froment:2010fk}
 
@@
 
-\begin{table}\centering
-\caption{\label{analogy}{\bf  Analogy between neural and social systems.} }
-\ra{1.3}
-\begin{tabular}{@{}rllll@{}}
-&Neural &  \phantom{abc}& Social \\
-\cmidrule{2-2} \cmidrule{4-4} 
-variables &  firing rates of neural populations && opinions about relative dominance
-\\ input & dots moving left or right && fights won or lost
-\\ 
-\end{tabular}
-\end{table}
+In Table \ref{analogy}, we list the inputs, decision variables, and outputs of the two systems.
 
-\section{Derivation of Model}
-With two assumptions about the timescales on which probabilistic events occur, a simple mechanistic description of the integration of information about random events can be turned into a stochastic differential equation \cite{Gillespie:2000fk}.
-A drift diffusion model is equivalent to a sequential probability ratio test, a learning algorithm that is optimal with respect to its accuracy and the time required to make a decision \cite{Moehlis:2004fk,Bogacz:2006uq}.
+\section{Derivation of Model \label{derivation}}
+Stochastic differential equations have been phenomenologically applied to learning systems, with some success \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
+However, with two assumptions about the timescales on which probabilistic events occur, a simple mechanistic description of the accumulation of information about random events can be turned into a stochastic differential equation \cite{Gillespie:2000fk}.
 
@@
 \begin{array}{ll}
-dX_1&=\bigg(-\ell X_1(t)+b_\text{f}(2a-1)\bigg)dt+b_f\sqrt{a}dW_{\text{f},1}t-b_{\text{f}}\sqrt{(1-a)}dW_{\text{f},2}t
-\\dX_2&=\bigg(-\ell X_1(t)+b_\text{f}(1-2a)\bigg)dt-b_f\sqrt{a}dW_{\text{f},1}t+b_{\text{f}}\sqrt{(1-a)}dW_{\text{f},2}t,
+dX_1&=\bigg(-\ell X_1(t)+b(2d-1)\bigg)dt+b\sqrt{d}dW_{1}t-b\sqrt{(1-d)}dW_{2}t
+\\dX_2&=\bigg(-\ell X_1(t)+b(1-2d)\bigg)dt-b\sqrt{d}dW_{1}t+b\sqrt{(1-d)}dW_{2}t,
 \end{array}
@@
 
-\begin{table}\centering
-\caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} }
-\ra{1.3}
-\begin{tabular}{@{}rllll@{}}
-& \multicolumn{2}{c}{Neural} & \phantom{abc}& Social \\
-\cmidrule{2-3} \cmidrule{5-5} 
-model  & DDM or OU & race && race with leak
-\\dimensionality & $1$ & $2$ && $2$
-\\decision & difference hits a threshold & one var. hits a threshold && one var. hits a threshold
-\\ criterion &  reward from being ``right" &&& reward from being ``right"
-\\ & reward rate &&& reward rate
-\\ & Bayes risk &&& Bayes risk
-\\ & &&& \fcolorbox{red}{white}{reward from  receiving signal}
-\\depends on & input strength &&& input strength
-\\ & noise &&& noise
-\\ & leak rate &&& leak rate
-\\ & &&& \fcolorbox{red}{white}{other animal's threshold}
-\\threshold  algorithm & error correction
-\\  & hill climbing
-\end{tabular}
-\end{table}
+In Table \ref{models}, we compare
+
+A drift diffusion model is equivalent to a sequential probability ratio test, a learning algorithm that is optimal with respect to its accuracy and the time required to make a decision \cite{Moehlis:2004fk,Bogacz:2006uq}.
+
+The outcome of the decision process is not deterministic.
+Because of the noise in this system, the utility of a decision depends on the probability of reaching the correct decision and the expected time it takes to reach a decision.
+These quantities satisfy partial differential equations with respect to the starting values of the decision variables.
+In Appendix \ref{pdes_deriv}, we derive these partial differential equations.
+While an analytical solution is known in the one-dimensional case, the two-dimensional case is analytically intractable.
+We therefore numerically solve the PDEs in the two-dimensional case.
 
@@
 A ``decision," i.e.
-the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below a given threshold.
+the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below a certain threshold.
 The social system is therefore inherently two-dimensional, whereas the neural system may be adequately described by one dimension.
@@
 We can compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
-We find that...
+We find that the two-dimensional decision is less accurate than the one-dimensional decision.
+That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with a lower probability (Figure \ref{1dv2d}).
+The difference in probability is very low when memory is perfect (i.e.
+$\ell=0$).
+However, at higher leak rates the difference becomes more pronounced.
+Further, when the strength of the input ($d$) is lower, the difference becomes greater.
+Thus, as the decision becomes harder because either memory decreases or the true state of the world becomes harder to detect, the advantage of the one-dimensional system increases.
 
 \subsection{Strategic Decision Making }
+Differences in the threshold between animals could just be explained by individual variation.
+However, there may be a strategic aspect to this variation in the social system.
 In the neural system, an individual wants to make both a decision that is both quick and accurate.
@@
 The criterion used to optimize the threshold, therefore, may be quite different in the two systems.
-Additionally, the speed and outcome of the decision process depend on both animals' thresholds, so there is a game theoretic aspect to an animal's optimization in that his best behavior depends on the behavior of the other animal.
-Optimizing the threshold is thus strategic in the social system but not in the neural system.
-
-When we optimize the threshold according to different criteria...
-
-When we allow the optimization to depend on the partner's threshold...
+Additionally, the speed and outcome of the decision process depend on both animals' thresholds.
+Since an animal's best behavior depends on the behavior of the other animal, optimizing the threshold is strategic in the social system but not in the neural system.
 
@@
 
+\section{Appendix}
+
+\subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{1d_v_2d_symmetric.pdf}
+\caption{\label{1dv2d} A two-dimensional decision process, in which the decision is made when either of the variables reaches a threshold, is less accurate than a one-dimensional decision process, in which the decision is made when the difference between the variables reaches a threshold.
+In each panel, the expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.
+Red curves represent a one-dimensional decision and red curves represent a two-dimensional decision.
+Stronger values of the input are represented in higher darker curves.
+In the left panel, there is no leak.
+In the right panel, leak rate is set to $\ell=.5$.
+Other parameters are set as follows:}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{2d_flexible_thresholds.pdf}
+\caption{\label{2dthresh} When the two thresholds in the two-dimensional case are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
+The expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.}
+\end{figure}
+
+\begin{table}
+\centering
+\caption{\label{analogy}{\bf  Analogy between neural and social systems.} }
+\ra{1.3}
+\begin{tabular}{@{}rllll@{}}
+&Neural &   Social \\
+\cmidrule{2-2} \cmidrule{3-3} 
+input & dots moving left or right & fights won or lost
+\\strength of input & coherence of dots & probability of stronger animal winning
+\\decision variables &  firing rates of neural populations & opinions about relative dominance
+\\ dist. of states of the world & left / right equally likely & distribution of fighting abilities
+\\ decision & subject indicates left or right & one animal emits a subordination signal
+\\ correct decision & subject chooses correct direction & weaker animal signals
+\end{tabular}
+\end{table}
+
+\begin{table}\centering
+\caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} }
+\ra{1.3}
+\begin{tabular}{@{}rllll@{}}
+& \multicolumn{2}{c}{Neural} & \phantom{abc}& Social \\
+\cmidrule{2-3} \cmidrule{5-5} 
+%model  & DDM or OU & race && race with leak
+%\\
+dimensionality & $1$ & $2$ && $2$
+\\decision & difference hits a threshold & one var. hits a threshold && \fcolorbox{red}{white}{one var. hits a threshold}
+\\ optimality criterion &  reward from being ``right" &&& reward from being ``right"
+\\ & reward rate &&& reward rate
+\\ & Bayes risk &&& Bayes risk
+\\ & &&& \fcolorbox{red}{white}{reward from  receiving signal}
+\\optimization depends on & input strength &&& input strength
+\\ & noise &&& noise
+\\ & leak rate &&& leak rate
+\\ & &&& \fcolorbox{red}{white}{other animal's threshold}
+\\threshold  algorithm & error correction
+\\  & hill climbing
+\end{tabular}
+\end{table}
+
 \nocite{*}

### 2014-04-10 paper_skeleton.tex
working on optimization in group setting

@@
 However, with two assumptions about the timescales on which probabilistic events occur, a simple mechanistic description of the accumulation of information about random events can be turned into a stochastic differential equation \cite{Gillespie:2000fk}.
+These assumptions are particularly reasonable in the social system.
 
@@
 We can compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
+In the neural system, there are two thresholds: if the difference in activity levels reaches the left threshold the subject decides the dots are moving left and analogously for the right threshold.
+Similarly, in the social system there are two thresholds: each animal chooses a threshold below which he will emit the subordination signal.
+To make the systems directly comparable, we assume in both cases that the thresholds are symmetric.
+
 We find that the two-dimensional decision is less accurate than the one-dimensional decision.
@@
 Thus, as the decision becomes harder because either memory decreases or the true state of the world becomes harder to detect, the advantage of the one-dimensional system increases.
+
+Without any prior knowledge of the probabilities of the state of the visual stimulus, it is reasonable to assume that the thresholds for deciding for each direction are symmetric.
+On the other hand, in the social system, the thresholds are chosen by two competing animals and there is less reason to assume that their thresholds are the same.
+The accuracy in a given amount of time depends on the ratio of the two thresholds and is improved when the weaker animal's threshold is lower than the stronger animal's threshold (Figure \ref{2dthresh}).
 
@@
 \caption{\label{2dthresh} When the two thresholds in the two-dimensional case are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
-The expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.}
+The expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.
+The curve depends on the ratio of the weaker animal's threshold to the stronger animal's threshold ($\rho$).
+Lower ratios are represented in higher darker curves.}
 \end{figure}
@@
 \begin{table}\centering
-\caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} }
+\caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} Differences are highlighted in red.}
 \ra{1.3}
 \begin{tabular}{@{}rllll@{}}
-& \multicolumn{2}{c}{Neural} & \phantom{abc}& Social \\
-\cmidrule{2-3} \cmidrule{5-5} 
+& \multicolumn{2}{c}{Neural} &  Social \\
+\cmidrule{2-3} \cmidrule{4-4} 
 %model  & DDM or OU & race && race with leak
 %\\
-dimensionality & $1$ & $2$ && $2$
-\\decision & difference hits a threshold & one var. hits a threshold && \fcolorbox{red}{white}{one var. hits a threshold}
-\\ optimality criterion &  reward from being ``right" &&& reward from being ``right"
-\\ & reward rate &&& reward rate
-\\ & Bayes risk &&& Bayes risk
-\\ & &&& \fcolorbox{red}{white}{reward from  receiving signal}
-\\optimization depends on & input strength &&& input strength
-\\ & noise &&& noise
-\\ & leak rate &&& leak rate
-\\ & &&& \fcolorbox{red}{white}{other animal's threshold}
+dimensionality & $1$  && $2$
+\\decision & difference hits a threshold  && \fcolorbox{red}{white}{one var. hits a threshold}
+\\ optimality criterion &  reward from being ``right" && reward from being ``right"
+\\ & reward rate && reward rate
+\\ & Bayes risk && Bayes risk
+\\ & && \fcolorbox{red}{white}{reward from  receiving signal}
+\\optimization depends on & input strength && input strength
+\\ & noise && noise
+\\ & leak rate && leak rate
+\\ & && \fcolorbox{red}{white}{other animal's threshold}
 \\threshold  algorithm & error correction

### 2014-04-10 effect_of_tradeoff.m

@@ -1,6 +1,5 @@
 function effect_of_tradeoff(c1,c2,c3)
 
-Nc=evalin('base','Nc');
 Nl=evalin('base','Nl');
 Nd=evalin('base','Nd');
 Nt=evalin('base','Nt');
@@ -13,7 +12,7 @@ red=evalin('base','red');
 markersz=evalin('base','markersz');
 lw=evalin('base','lw');
 
-perf=zeros(Nc,2,Nl,Nd,Nt,Nt); %criteria, individual, leak, dominance, thresholds
+perf=zeros(2,2,Nl,Nd,Nt,Nt); %criteria, individual, leak, dominance, thresholds
 
 perf(1,1,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2));
 perf(1,2,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2));
@@ -25,13 +24,13 @@ perf(1,2,:,1,:,:)=c2*(twomat(:,1,:,:,2));
 perf(2,1,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(1-twomat(:,1,:,:,1));
 perf(2,2,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(twomat(:,1,:,:,1));
 
-nasheq_thresh=cell(Nc,Nl,Nd);
-nasheq_ind=cell(Nc,Nl,Nd);
+nasheq_thresh=cell(2,Nl,Nd);
+nasheq_ind=cell(2,Nl,Nd);
 
-T1_best=zeros(Nc,Nl,Nd,Nt); %criteria, leak, dominance, threshold
-T2_best=zeros(Nc,Nl,Nd,Nt);
+T1_best=zeros(2,Nl,Nd,Nt); %criteria, leak, dominance, threshold
+T2_best=zeros(2,Nl,Nd,Nt);
 
-for i=1:Nc
+for i=1:2
     for j=1:Nl
         for k=1:Nd
             for u=1:Nt

### 2014-04-10 optimization_for_group_distribution.m

@@ -1,169 +1,125 @@
-Lxvals=-1:.05:-.1;
-Lyvals=-1:.05:-.1;
-deps=.1;
-dvals=0:deps:1;
-NLx=length(Lxvals);
-NLy=length(Lyvals);
-Nd=length(dvals);
-
-l=1;
-b=1;
 
 %%
-f=zeros(NLx,NLy,Nd,3);
-
-for i=1:NLx
-    for j=1:NLy
-        for k=1:Nd
-            Lx=Lxvals(i);
-            Ly=Lyvals(j);
-            d=dvals(k);
-            [x,y,~]=solve_pde(Lx,Ly,d,5,5,.05,.05,l,b);
-            f(i,j,k,:)=[x,1-x,y];
-        end
-    end
-end
-
-
-%%
-Lx_best=zeros(NLy,Nd,2);
-
-for j=1:NLy
-    for k=1:Nd
-        d=dvals(k);
-        perfvals1=-f(:,j,k,3)+f(:,j,k,1).*(1-f(:,j,k,3)); %receiving a signal 
-        if d<.5
-            perfvals2=-f(:,j,k,3)+f(:,j,k,2).*(1-f(:,j,k,3)); %getting it 'right'
-        else perfvals2=perfvals1;
-        end
-        [~,w1]=max(perfvals1);
-        [~,w2]=max(perfvals2);
-        Lx_best(j,k,:)=Lxvals([w1 w2]);
-    end
-end
-
-%%
-NEx={};
-NEy={};
-
-for i=0:(Nd-1)
-    pp1=interp1(Lyvals,Lx_best(:,1+i,1),'linear','pp');
-    pp2=interp1(Lyvals,Lx_best(:,Nd-i,1),'linear','pp');
-    Tystar=Lyvals(abs(ppval(pp2,ppval(pp1,Lyvals))-Lyvals)<0.001);
-    Txstar=ppval(pp1,Tystar);
-    NEx{i+1}=Txstar;
-    NEy{i+1}=Tystar;
-end
-
-
-%%
-its=100;
-N=50;
+its=10;
+N=30;
 
 fighting_abilities=20*rand(1,N);
+% fighting_abilities=normrnd(0,10,1,N);
+% fighting_abilities=lognrnd(0,1,1,N);
 fighting_abilities=sort(fighting_abilities,'descend');
 
-dmat=zeros(N);
-v=1:Nd;
[…]

### 2014-04-14 paper_skeleton.tex
reorganizing code

@@
 
+%note about applicability to quorum sensing?!?
+
 \begin{equation}
@@
 
+Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.
+The sign and size of the two variables indicate how much evidence there is for each possible decision outcome.
+When a lot of evidence for the first possibility has been observed, $X_1$ will be large and positive, and similarly for the second possibility.
+If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility.
+If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.
+Specifically, there are two thresholds, $-T_2$ and $T_1$ such that if $Y<-T_2$ the decision is for $X_2$ and if $Y>T_1$ the decision is for $X_1$.
+However, it may be the case that the difference in the variables cannot be observed.
+In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.
+Again, there are two thresholds, now $-T_2$ and $-T_1$ and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.
+
+In either case, the thresholds affect the expected time until a decision is reached (decision time, DT), the probability that an incorrect decision is made (error rate, ER), and, in the social case, the probability for each animal that he will emit a subordination signal (signaling probability, $\text{SP}_i$).
+These quantities, DT, ER, and $\text{SP}_i$, satisfy partial differential equations with respect to the starting values of the decision variables.
+In Appendix \ref{pdes_deriv}, we derive these partial differential equations.
+While an analytical solution is known in the one-dimensional case, the two-dimensional case is analytically intractable.
+We therefore numerically solve the PDEs in the two-dimensional case.
+
 In Table \ref{models}, we compare
@@
 A drift diffusion model is equivalent to a sequential probability ratio test, a learning algorithm that is optimal with respect to its accuracy and the time required to make a decision \cite{Moehlis:2004fk,Bogacz:2006uq}.
-
-The outcome of the decision process is not deterministic.
-Because of the noise in this system, the utility of a decision depends on the probability of reaching the correct decision and the expected time it takes to reach a decision.
-These quantities satisfy partial differential equations with respect to the starting values of the decision variables.
-In Appendix \ref{pdes_deriv}, we derive these partial differential equations.
-While an analytical solution is known in the one-dimensional case, the two-dimensional case is analytically intractable.
-We therefore numerically solve the PDEs in the two-dimensional case.
 
@@
 
+When waiting for a decision to be made is more costly, so that minimizing DT is prioritized, the resulting distribution of the number of signals received is more skewed with a heavier right tail (Figure \ref{skewness}).
+
 \subsection{Comparison to Social Data }
@@
 Lower ratios are represented in higher darker curves.}
+\end{figure}
+
+\begin{figure}
+
+\caption{\label{givingup} }
+\end{figure}
+
+\begin{figure}
+
+\caption{\label{skewness} }
 \end{figure}

### 2014-04-14 dimensionality_comparison.m

@@ -13,47 +13,6 @@ markersz=5;
 markercol='white';
 
 %%
-deltax1=0.01;
-deltax2=0.025 ;
-b=1;
-
-lvals=0:.5:2;
-Nl=length(lvals);
-
-domvals=.5:.1:1;
-Nd=length(domvals);
-
-threshvals1=.1:.1:8;
-Nt1=length(threshvals1);
-
-threshvals2=.5:.5:2;
-Nt2=length(threshvals2);
-
-%%
-
-oneoutput=zeros(Nl,Nd,Nt1,2);
-twooutput=zeros(Nl,Nd,Nt2,2);
-
-for k=1:Nl
-    l=lvals(k);
-    for i=1:Nd
-        d=domvals(i);
-        for j=1:Nt1
-            t=threshvals1(j);
-            [x,y,~]=solve_pde_1d(-t,d,t,deltax1,l,b);
-            oneoutput(k,i,j,:)=[x y];
-        end
-        for j=1:Nt2
-            t=threshvals2(j);
-            [x,y,~]=solve_pde_2d(-t,-t,d,5,5,deltax2,deltax2,l,b);
-            twooutput(k,i,j,:)=[x y];
-        end
-    end
-end
-%%
-filename=strcat('two_v_one_output','.mat');
-save(filename,'twooutput','oneoutput','domvals','lvals','threshvals1','threshvals2');
-%%
 
 figure
 set(gcf,'Color','w')
@@ -75,13 +34,13 @@ for k=1:2
     switch k
         case 2
             for i=2:Nd
-            p(i)=plot(reshape(twooutput(k,i,:,2),1,[]),reshape(twooutput(k,i,:,1),1,[]),'Color',blue(i,:),'LineWidth',lw);
-            p(i+Nd)=plot(reshape(oneoutput(k,i,1:40,2),1,[]),reshape(oneoutput(k,i,1:40,1),1,[]),'Color',red(i,:),'LineWidth',lw);
+            p(i)=plot(reshape(diag(reshape(twooutput(k,i,:,:,2),Nt,[])),1,[]),reshape(diag(reshape(twooutput(k,i,:,:,1),Nt,[])),1,[]),'Color',blue(i,:),'LineWidth',lw);
+            p(i+Nd)=plot(reshape(onemat(k,i,1:40,2),1,[]),reshape(onemat(k,i,1:40,1),1,[]),'Color',red(i,:),'LineWidth',lw);
             end
[…]

### 2014-04-14 nash_eq.m

@@ -26,6 +26,9 @@ indices=1:Nt;
 b=1;
 deltax2=0.025;
 
+threshvals1=.1:.1:8;
+Nt1=length(threshvals1);
+
 %%
 twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time
 
@@ -44,6 +47,23 @@ for k=1:Nl
     end
 end
 
+%%
+onemat=zeros(Nl,Nd,Nt1,2);
+
+for k=1:Nl
+    l=lvals(k);
+    for i=1:Nd
+        d=domvals(i);
+        for j=1:Nt1
+            t=threshvals1(j);
+            [x,y,~]=solve_pde_1d(-t,d,t,deltax1,l,b);
+            onemat(k,i,j,:)=[x y];
+        end
+    end
+end
+%%
+filename=strcat('twomat_long_run','.mat');
+save(filename,'twomat','lvals','domvals','threshvals','threshvals');
 %%
 
 c1=0;

### 2014-04-14 write_and_save_data.m

@@ -1,0 +1,55 @@
+%%
+lvals=0:.1:1;
+Nl=length(lvals);
+
+domvals=.5:.05:1;
+Nd=length(domvals);
+
+threshvals=.5:.1:2;
+Nt=length(threshvals);
+
+indices=1:Nt;
+
+b=1;
+deltax2=0.025;
+
+deltax1=0.01;
+threshvals1=.1:.1:8;
+Nt1=length(threshvals1);
+
+%%
+twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time
+
+for k=1:Nl
+    l=lvals(k);
+    for u=1:Nd
+        d=domvals(u);
+        for i=1:Nt
+            T1=threshvals(i);
+            for j=1:Nt
+                T2=threshvals(j);
+                [x,y,~]=solve_pde_2d(-T1,-T2,d,5,5,deltax2,deltax2,l,b);
+                twomat(k,u,i,j,:)=[x y];
+            end
+        end
+    end
+end
+
+%%
+onemat=zeros(Nl,Nd,Nt1,2);
+
+for k=1:Nl
+    l=lvals(k);
+    for i=1:Nd
+        d=domvals(i);
+        for j=1:Nt1
+            t=threshvals1(j);
+            [x,y,~]=solve_pde_1d(-t,d,t,deltax1,l,b);
+            onemat(k,i,j,:)=[x y];
+        end
+    end
+end
+
+%%
+filename=strcat('twomat_long_run','.mat');
+save(filename,'twomat','lvals','domvals','threshvals','threshvals');

### 2014-04-17 paper_skeleton.tex
focusing on accuracy of 2d process now that i think theyre the same

@@
 
-We find that the two-dimensional decision is less accurate than the one-dimensional decision.
-That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with a lower probability (Figure \ref{1dv2d}).
-The difference in probability is very low when memory is perfect (i.e.
-$\ell=0$).
-However, at higher leak rates the difference becomes more pronounced.
-Further, when the strength of the input ($d$) is lower, the difference becomes greater.
-Thus, as the decision becomes harder because either memory decreases or the true state of the world becomes harder to detect, the advantage of the one-dimensional system increases.
+We find that the two-dimensional decision is as accurate as the one-dimensional decision.
+That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability.
+Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($d$) is higher (Figure \ref{2dacc}).
 
@@
 On the other hand, in the social system, the thresholds are chosen by two competing animals and there is less reason to assume that their thresholds are the same.
-The accuracy in a given amount of time depends on the ratio of the two thresholds and is improved when the weaker animal's threshold is lower than the stronger animal's threshold (Figure \ref{2dthresh}).
+The accuracy in a given amount of time depends on the ratio of the two thresholds and is improved when the weaker animal's threshold is lower than the stronger animal's threshold (Figure \ref{2dflex}).
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{1d_v_2d_symmetric.pdf}
-\caption{\label{1dv2d} A two-dimensional decision process, in which the decision is made when either of the variables reaches a threshold, is less accurate than a one-dimensional decision process, in which the decision is made when the difference between the variables reaches a threshold.
-In each panel, the expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.
-Red curves represent a one-dimensional decision and red curves represent a two-dimensional decision.
-Stronger values of the input are represented in higher darker curves.
-In the left panel, there is no leak.
-In the right panel, leak rate is set to $\ell=.5$.
-Other parameters are set as follows:}
+\includegraphics[width=\textwidth]{2d_accuracy.pdf}
+\caption{\label{2dacc} }
 \end{figure}
@@
 \includegraphics[width=\textwidth]{2d_flexible_thresholds.pdf}
-\caption{\label{2dthresh} When the two thresholds in the two-dimensional case are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
+\caption{\label{2dflex} When the two thresholds in the two-dimensional case are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
 The expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.

### 2014-04-17 nash_eq.m

@@ -1,7 +1,4 @@
-green=importdata('green.txt');
-blue=importdata('blue.txt');
-red=importdata('red.txt');
-black=importdata('black.txt');
+
 %%
 axislw=1.25;
 lw=3;
@@ -11,59 +8,8 @@ alpha=.7;
 ticklength=[.02,.0250];
 markersz=5;
 markercol='white';
-%%
-lvals=0:.1:1;
-Nl=length(lvals);
 
-domvals=.5:.05:1;
-Nd=length(domvals);
 
-threshvals=.5:.1:2;
-Nt=length(threshvals);
-
-indices=1:Nt;
-
-b=1;
-deltax2=0.025;
-
-threshvals1=.1:.1:8;
-Nt1=length(threshvals1);
-
-%%
-twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time
-
-for k=1:Nl
-    l=lvals(k);
-    for u=1:Nd
-        d=domvals(u);
-        for i=1:Nt
-            T1=threshvals(i);
-            for j=1:Nt
-                T2=threshvals(j);
-                [x,y,~]=solve_pde_2d(-T1,-T2,d,5,5,deltax2,deltax2,l,b);
-                twomat(k,u,i,j,:)=[x y];
-            end
-        end
-    end
-end
-
-%%
-onemat=zeros(Nl,Nd,Nt1,2);
-
-for k=1:Nl
-    l=lvals(k);
-    for i=1:Nd
-        d=domvals(i);
-        for j=1:Nt1
-            t=threshvals1(j);
-            [x,y,~]=solve_pde_1d(-t,d,t,deltax1,l,b);
-            onemat(k,i,j,:)=[x y];
-        end
-    end
-end
-%%
-filename=strcat('twomat_long_run','.mat');
-save(filename,'twomat','lvals','domvals','threshvals','threshvals');
 %%
 
 c1=0;
@@ -72,26 +18,36 @@ c3=1;
 c2vals=0:.1:1;
 Nc2=length(c2vals);
 
-nasheq_ind=cell(Nc2,Nl,Nd);
-giveup=zeros(Nc2,Nl,Nd);
+nasheq_ind2=cell(Nc2,Nl,Nd);
+giveup2=zeros(Nc2,Nl,Nd);
+
+objective_best2=cell(Nc2,Nl,Nd);
 
 for i=1:Nc2
     c2=c2vals(i);
[…]

### 2014-04-17 output_of_2d_process.m

@@ -30,7 +30,7 @@ set(gcf,'Position',[v(1) v(2) w h]);
 
 textlabels={'a' 'b'};
 
-p=zeros(1,2*Nd);
+p=zeros(1,Nd);
 
 currblue=mycolors(Nd,'blue');
 currred=mycolors(Nd,'red');
@@ -42,39 +42,33 @@ for k=1:2
     hold on
     switch k
         case 2
-            for i=2:Nd
+            for i=3:2:Nd
             p(i)=plot(reshape(diag(reshape(twomat(k,i,:,:,2),Nt,[])),1,[]),reshape(diag(reshape(twomat(k,i,:,:,1),Nt,[])),1,[]),'Color',currblue(i,:),'LineWidth',lw);
-            p(i+Nd)=plot(reshape(diag(reshape(onemat(k,i,:,:,2),Nt1,[])),1,[]),reshape(diag(reshape(onemat(k,i,:,:,1),Nt1,[])),1,[]),'Color',currred(i,:),'LineWidth',lw);
+%             p(i+Nd)=plot(reshape(diag(reshape(onemat(k,i,:,:,2),Nt1,[])),1,[]),reshape(diag(reshape(onemat(k,i,:,:,1),Nt1,[])),1,[]),'Color',currred(i,:),'LineWidth',lw);
             end
         otherwise 
-            for i=2:Nd
+            for i=2:2:Nd
             plot(reshape(diag(reshape(twomat(k,i,:,:,2),Nt,[])),1,[]),reshape(diag(reshape(twomat(k,i,:,:,1),Nt,[])),1,[]),'Color',currblue(i,:),'LineWidth',lw);
-            plot(reshape(diag(reshape(onemat(k,i,:,:,2),Nt1,[])),1,[]),reshape(diag(reshape(onemat(k,i,:,:,1),Nt1,[])),1,[]),'Color',currred(i,:),'LineWidth',lw);
+%             plot(reshape(diag(reshape(onemat(k,i,:,:,2),Nt1,[])),1,[]),reshape(diag(reshape(onemat(k,i,:,:,1),Nt1,[])),1,[]),'Color',currred(i,:),'LineWidth',lw);
             end
     end
     set(gca,'xlim',[0 5],'ylim',[.5 1]);
     ylabel('Probability of correct output','FontSize',textfontsz)
     xlabel('Expected time to output','FontSize',textfontsz)
[…]

### 2014-06-05 paper_skeleton.tex
trying to change language from animals to more general individuals

@@
 However, there are (at least) three significant differences between the two models.
-First, in the social system, there is a strategic component to the system since the optimal threshold of one animal depends on the threshold of the other animal, whereas in the neural system both neural populations are part of the cognitive system of the same individual trying to optimize its decision making.
+First, whereas in the neural system, the differential equations are often reduced to a one-dimensional model, the social system is inherently two-dimensional.
+Second, in the social system, there is a strategic component to the system since the optimal threshold of one animal depends on the threshold of the other animal, whereas in the neural system both neural populations are part of the cognitive system of the same individual trying to optimize its decision making.
 Relatedly, the output of the decision has additional strategic consequences in the social system.
-Second, whereas in the neural system, the differential equations are often reduced to a one-dimensional model, the social system is inherently two-dimensional.
 Finally, in the social system, a monkey is not only trying to make a decision about his dominance with respect to one other monkey, but he is additionally trying to make decisions about many other partners at once, whereas in the neural case there is only one decision to be made at a time.
-As a consequence of these differences...
+First, we found that the dimensionality of the decision process does not affect its accuracy.
+The costs of the waiting time until a decision affect the Nash equilibrium thresholds for a pair of animals, the optimal thresholds for a group of animals, and the skewness and information content of the resulting distribution of power.
 In comparing the output of the model of the social system to data, we found that ...
@@
 
-In Table \ref{analogy}, we list the inputs, decision variables, and outputs of the two systems.
+In Table \ref{variables}, we list the inputs, decision variables, and outputs of the two systems.
 
-\section{Derivation of Model \label{derivation}}
-Stochastic differential equations have been phenomenologically applied to learning systems, with some success \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
-However, with two assumptions about the timescales on which probabilistic events occur, a simple mechanistic description of the accumulation of information about random events can be turned into a stochastic differential equation \cite{Gillespie:2000fk}.
-These assumptions are particularly reasonable in the social system.
+\section{The model \label{derivation}}
+\subsection{Stochastic differential equations }
+Learning systems have been successfully modeled with stochastic differential equations \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}, although the application of this type of model has been quite phenomenological.
+With two assumptions about the timescales on which probabilistic events occur, a simple mechanistic description of the accumulation of information about random events can be turned into a stochastic differential equation \cite{Gillespie:2000fk}.
 
@@
 
+In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.
+(A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period of time of length $\tau$ a decision variable would decrease as $X(t+\tau)=(1-\ell\tau)X(t)$.
+There are two types of inputs: in the neural system, a dot can move either left or right and in the social system a fight can be won by one animal or another.
+For clarity in the following we will refer to these left or right dots, but the biological interpretation of the two types of inputs depends on the system.
+Each decision variable responds positively to one type of input and negatively to the other, with the signs of their responds the opposite of each other.
+Specifically, in the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.
+In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite.
+
+To determine the estimate at time $t+\tau$, we can count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
+\begin{align*}
+X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
+\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
+\end{align*}
+where $b$ is the change caused by a single input.
+
+Our first assumption is that the rates at which the two types of inputs occur is constant.
+We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$.
+We rewrite our basic equations to give
+\begin{align*}
+X_1(t+\tau)&=(1-\ell\tau)X_1(t)+bN_\text{L}-bN_\text{R}
+\\ X_1(t+\tau)&=(1-\ell\tau)X_2(t)-bN_\text{L}+bN_\text{R}.
+\end{align*}
+If events happen at a rate $r$ and the event is a left dot with probability $c$ and a right dot with probability $1-c$, then the expectation of $N_L$ and $N_R$ in a period of length $\tau$ are, respectively, $\tau r c$ and $\tau r(1-c)$.
+In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.
+A $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.
+
+If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with Normal random variables with mean and variance equal to the mean of the Poisson random variables.
+Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation.
+Let $Z_\text{L}$ and $Z_\text{R}$, be independent standard Normal random variables, i.e.
+with mean $0$ and standard deviation $1$.
+Then our equations for $X_i(t+\tau)$ become
+\begin{align*}
+X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)-b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg)
+\\X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)+b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg).
+\end{align*}
+Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
 \begin{equation}
 \begin{array}{ll}
-dX_1&=\bigg(-\ell X_1(t)+b(2d-1)\bigg)dt+b\sqrt{d}dW_{1}t-b\sqrt{(1-d)}dW_{2}t
-\\dX_2&=\bigg(-\ell X_1(t)+b(1-2d)\bigg)dt-b\sqrt{d}dW_{1}t+b\sqrt{(1-d)}dW_{2}t,
+dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{L}t-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
+\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{L}t+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
 \end{array}
 \end{equation}
+where $dW_{\text{L}}$ and $dW_{\text{R}}$ are independent Brownian motions representing, respectively, the left and right inputs.
+The assumptions about the timescales on which inputs occur are particularly reasonable in the social system, although the successful application of this type of model to neural populations implies they are not unreasonable in that system.
 
+%\begin{equation}
+%\begin{array}{ll}
+%dX_1&=\bigg(-\ell X_1(t)+b(2d-1)\bigg)dt+b\sqrt{d}dW_{1}t-b\sqrt{(1-d)}dW_{2}t
+%\\dX_2&=\bigg(-\ell X_1(t)+b(1-2d)\bigg)dt-b\sqrt{d}dW_{1}t+b\sqrt{(1-d)}dW_{2}t,
+%\end{array}
+%\end{equation}
+\subsection{Reaching a decision }
 Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.
@@
 If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.
-Specifically, there are two thresholds, $-T_2$ and $T_1$ such that if $Y<-T_2$ the decision is for $X_2$ and if $Y>T_1$ the decision is for $X_1$.
+Specifically, there are two thresholds, $T_2$ and $T_1$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.
 However, it may be the case that the difference in the variables cannot be observed.
 In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.
-Again, there are two thresholds, now $-T_2$ and $-T_1$ and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.
+Again, there are two thresholds, $T_2$ and $T_1$, and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.
 
-In either case, the thresholds affect the expected time until a decision is reached (decision time, DT), the probability that an incorrect decision is made (error rate, ER), and, in the social case, the probability for each animal that he will emit a subordination signal (signaling probability, $\text{SP}_i$).
-These quantities, DT, ER, and $\text{SP}_i$, satisfy partial differential equations with respect to the starting values of the decision variables.
-In Appendix \ref{pdes_deriv}, we derive these partial differential equations.
-While an analytical solution is known in the one-dimensional case, the two-dimensional case is analytically intractable.
-We therefore numerically solve the PDEs in the two-dimensional case.
-
-In Table \ref{models}, we compare
-
-A drift diffusion model is equivalent to a sequential probability ratio test, a learning algorithm that is optimal with respect to its accuracy and the time required to make a decision \cite{Moehlis:2004fk,Bogacz:2006uq}.
-
-\section{Results}
-
-\subsection{Two-Dimensional Decision Making }
-In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to a one-dimensional equation describing the difference in the activity levels to make the system easier to analyze \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.
+In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels to make the system easier to analyze \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.
 However, in the social system we are considering, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.
@@
 
-We can compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
-In the neural system, there are two thresholds: if the difference in activity levels reaches the left threshold the subject decides the dots are moving left and analogously for the right threshold.
-Similarly, in the social system there are two thresholds: each animal chooses a threshold below which he will emit the subordination signal.
-To make the systems directly comparable, we assume in both cases that the thresholds are symmetric.
+\subsection{Utility of the decision }
+The thresholds affect the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER).
+The thresholds also affect the probability that either variable will ``give up'' first.
+In the social case, this is the probability that each animal will emit a subordination signal (signaling probability, $\text{SP}_i$).
+In the neural case, ???
+These quantities, DT, ER, and $\text{SP}_i$, satisfy partial differential equations with respect to the starting values of the decision variables.
+In Appendix \ref{pdes_deriv}, we derive these partial differential equations.
+While an analytical solution is known in the one-dimensional case, the two-dimensional case is analytically intractable.
+We therefore numerically solve the PDEs in the two-dimensional case.
+In Table \ref{models}, we compare how the decision is made and how the decision process is optimized in the two systems.
+
+The quality of the decision process depends on each of these three quantities.
+The lower the decision time and error rate, the better the decision.
+However, both cannot be decreased simultaneously so there is a tradeoff that depends on how the two quantities are prioritized.
+Additionally, in the social case, each animal may benefit from lower its signaling probability, which may or may not conflict with lowering the error rate, depending on whether the animal is weaker or stronger than its opponent.
+To quantify the utility of the decision process, we introduce three weights, $w_1$, $w_2$, and $w_3$, which determine how the three quantities are prioritized.
+For animal $i$, we define
+\begin{equation*}
+U_i=w_1\text{DT}+w_2\text{ER}+w_3\text{SP}_i.
+\end{equation*}
+The decision is better if $U_i$ is lower.
+In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.
+Similarly, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.
+
+%A drift diffusion model is equivalent to a sequential probability ratio test, a learning algorithm that is optimal with respect to its accuracy and the time required to make a decision \cite{Moehlis:2004fk,Bogacz:2006uq}.
+
+\section{Results}
+
+\subsection{Decisions are as accurate without an arbiter. }
+We compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
+To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e.
+$T_1=T_2$.
 
@@
 That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability.
-Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($d$) is higher (Figure \ref{2dacc}).
+Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher (Figure \ref{2dacc}).
 
-Without any prior knowledge of the probabilities of the state of the visual stimulus, it is reasonable to assume that the thresholds for deciding for each direction are symmetric.
-On the other hand, in the social system, the thresholds are chosen by two competing animals and there is less reason to assume that their thresholds are the same.
+Without any prior knowledge of the probabilities of the state of the external stimulus, it is reasonable to assume that the thresholds for deciding for each direction are symmetric.
+On the other hand, the thresholds are chosen by two competing individuals-- neural populations or animals-- and it is possible that the two individuals have different thresholds.
 The accuracy in a given amount of time depends on the ratio of the two thresholds and is improved when the weaker animal's threshold is lower than the stronger animal's threshold (Figure \ref{2dflex}).
@@
 \subsection{Strategic Decision Making }
-Differences in the threshold between animals could just be explained by individual variation.
+Differences in the thresholds could be explained by individual variation.
 However, there may be a strategic aspect to this variation in the social system.
@@
 
-\subsection{Social Decision Making }
+For any pair of individuals, each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.
+These thresholds depend on the relative dominance of the two animals and how costly waiting is.
+If waiting is not costly and the error rate does not matter ($w_1=0$, $w_2=0$, $w_3>0$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so they will set high thresholds in order to wait until the other individual gives up.
+However, as waiting costs increase, the pair can reach a decision more quickly if the weaker individual lowers its threshold.
+If waiting costs are high enough, the stronger individual will also lower its threshold to avoid a long decision time (Figure \ref{nasheq_thresholds}).
+By increasing the costs of waiting, the weaker individual has an incentive to give up more quickly, which can actually improve the accuracy of the decision process (Figure \ref{nasheq_thresholds}).
+For low to moderate waiting costs, pairs of individuals making a difficult decision (low $c$, e.g.
+animals with similar fighting abilities) will take longer to reach a decision than pairs making a difficult decision (high $c$, e.g.
+animals with disparate fighting abilities).
+However, when costs are high, the pairs that take the longest to reach a decision are those making moderately difficult decisions (Figure \ref{nasheq_thresholds}).
+This is because pairs making difficult decisions will both set their thresholds quite low and in pairs making difficult decisions the weaker will set its threshold much lower, both of which decrease the time it takes until reaching a decision.
+
+\subsection{Costs affect the shape and the information content of the power distribution.  }
 A monkey can try to decide whether it is more or less likely to win a fight than another monkey using information from the fights the pair has previously engaged in.
@@
 
-When waiting for a decision to be made is more costly, so that minimizing DT is prioritized, the resulting distribution of the number of signals received is more skewed with a heavier right tail (Figure \ref{skewness}).
+When waiting for a decision becomes more costly ($c_2$ increases), the resulting distribution of the number of signals received first becomes more skewed, with a heavier right tail, and then loses it skewness again (Figure \ref{power}).
+Changing $c_1$ and $c_3$ have less of an effect, although when $c_1$ is higher the skewness slightly increases.
 
-\subsection{Comparison to Social Data }
+\subsection{Comparison to social data }
 Empirically, those pairs of monkeys that take the longest to establish a signaling dynamics are those with the most similar fighting abilities...
@@
 \section{Discussion}
-
-\section{Appendix}
-
-\subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}
 
@@
 \includegraphics[width=\textwidth]{2d_flexible_thresholds.pdf}
-\caption{\label{2dflex} When the two thresholds in the two-dimensional case are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
+\caption{\label{2dflex} When the two thresholds in the two-dimensional system are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
 The expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.
@@
 \begin{figure}
-
-\caption{\label{givingup} }
+\includegraphics[width=\textwidth]{nasheq_thresholds.pdf}
+\caption{\label{nasheq_thresholds} If a pair of individuals optimize their thresholds in response to each other, as the costs of waiting increase, first the weaker individual will start to lower his threshold to avoid waiting for a long time and then the stronger individual will start to lower his threshold as well.
+As the costs of waiting increase, first the accuracy with which a pair can make a difficult decision increases and then it decreases.
+At low waiting costs, difficult decisions take the most time.
+At intermediate costs, intermediate decisions take the most time.
+At high costs, easy decisions take the most time.
+The first row of panels shows the Nash equilibrium thresholds as a function of the difficulty of the decision ($c$ close to $.5$ is a more difficult decision).
+The second and third rows show, respectively, the probability of the pair reaching the correct decision and the expected time until a decision is reached, if the pair of individuals uses the Nash equilibrium thresholds.
+Each column corresponds to a particular waiting cost and the cost increases from left to right ($w_1=0, w_3=1-w_2$).
+}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{groupeq_thresholds.pdf}
+\caption{\label{groupeq_thresholds} If the animals optimize their thresholds with respect to the thresholds and fighting abilities of the whole group of animals, as the costs of fighting increase, first the weaker animals in the group lower their thresholds to avoid fighting for a long time with the stronger animals and then the stronger animals lower their thresholds as well.
+The ability of to learn from an animal's signaling behavior decreases as the costs of fighting increase.
+When fighting is not costly, the pairs of animals that take longest to decide on a relationship are those with similar fighting abilities.
+As costs increase pairs of animals with similarly low fighting abilities resolve their relationships quickly, whereas pairs of animals with similarly high fighting abilities take a long time to resolve their relationships.
+The first row of panels shows the optimal threshold as a function of the animal's true rank with respect to fighting ability.
+The second row shows the mutual information between the probability of emitting a signal and the probability of being the weaker animal, as a function of the animal's true rank.
+The third row shows the expected time until each pair of animals resolves their relationship.
+Each column corresponds to a particular waiting cost and the cost increases from left to right ($w_1=0, w_3=1-w_2$).
+}
+\end{figure}
 
-\caption{\label{skewness} }
+\begin{figure}
+\begin{subfigure}{3.415in}
+\caption{}
+\includegraphics[width=\textwidth]{skewness_heatmap.pdf}
+\end{subfigure}
+\begin{subfigure}{3.415in}
+\caption{}
+\includegraphics[width=\textwidth]{mutinfo_heatmap.pdf}
+\end{subfigure}
+\caption{\label{power} As the costs of waiting increase, the distribution of the number of relationships in which an animal is dominant becomes skewed to the right and then loses its skewness as costs increase further.
+As the costs of waiting increase, the information content of the power distribution always decreases and if dominance is more important than accuracy the information content is higher.
+The simplex shows the possible combinations of priorities, $w_1$, $w_2$, and $w_3$ such that the three weights sum to $1$.
+In the lower left corner, $w_1=1$ (only accuracy matters), in the upper corner $w_2=1$ (only waiting time matters), and in the lower right $w_3=1$ (only the probability of receiving the signal matters).
+The heatmaps show (a) the skewness of the power distribution and (b) the mutual information of the power distribution when the animals in the group optimize their thresholds according to those weights.
+%Increasing or decreasing $c_2$ has the strongest effect.  Increasing $c_1$ slightly increases the skewness. 
+}
 \end{figure}
@@
 \centering
-\caption{\label{analogy}{\bf  Analogy between neural and social systems.} }
+\caption{\label{variables}{\bf  Variables and interpretations in neural and social systems.} }
 \ra{1.3}
 \begin{tabular}{@{}rllll@{}}
-&Neural &   Social \\
-\cmidrule{2-2} \cmidrule{3-3} 
-input & dots moving left or right & fights won or lost
-\\strength of input & coherence of dots & probability of stronger animal winning
-\\decision variables &  firing rates of neural populations & opinions about relative dominance
-\\ dist. of states of the world & left / right equally likely & distribution of fighting abilities
-\\ decision & subject indicates left or right & one animal emits a subordination signal
-\\ correct decision & subject chooses correct direction & weaker animal signals
+Variable & Interpretation & Neural &   Social \\
+\cmidrule{1-4} 
+& input & dots moving left or right & fights won or lost
+\\$c$ & strength of input & coherence of dots & probability of stronger animal winning
+\\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
+\\ & dist. of states of the world & left / right equally likely & distribution of fighting abilities
+\\ & decision & subject indicates left or right & one animal emits a subordination signal
+\\ & correct decision & subject chooses correct direction & weaker animal signals
 \end{tabular}
@@
 
+\pagebreak
+\section{Appendix}
+
+\subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}
+
+\pagebreak
 \nocite{*}

### 2014-07-16 paper_skeleton.tex
im about to rewrite the paper_skeleton quite a bit

@@
 In order to make a decision, one must accumulate the  evidence for different alternatives and choose one only once the uncertainty about the choice is sufficiently reduced.
-There are two equivalent mathematical formulations of the optimal decision making algorithm: a sequential probability ratio test in which the likelihoods of two hypotheses are evaluated as more data points are collected and a drift-diffusion stochastic differential equation in which the belief about two hypotheses does a random walk until a threshold is reached.
-In the random walk formulation, the threshold dictating when a decision has been made affects both the accuracy of the decision and the expected time until a decision is reached.
-If the utility of a decision depends on accuracy and decision time, the threshold can be optimized accordingly.
 There are two biological systems in which the variables used to make a decision can be described by a random walk: the firing rates of neural populations in the brain of an individual making a decision about a visual stimulus and the beliefs of monkeys about their relative dominance based on the fights they have engaged in.
-The stochastic differential equations used to model these quite different systems are nearly identical, which suggests that there are universal principles underlying learning through noisy dynamical systems in biology.
-However, there are (at least) three significant differences between the two models.
-First, whereas in the neural system, the differential equations are often reduced to a one-dimensional model, the social system is inherently two-dimensional.
-Second, in the social system, there is a strategic component to the system since the optimal threshold of one animal depends on the threshold of the other animal, whereas in the neural system both neural populations are part of the cognitive system of the same individual trying to optimize its decision making.
-Relatedly, the output of the decision has additional strategic consequences in the social system.
-Finally, in the social system, a monkey is not only trying to make a decision about his dominance with respect to one other monkey, but he is additionally trying to make decisions about many other partners at once, whereas in the neural case there is only one decision to be made at a time.
-First, we found that the dimensionality of the decision process does not affect its accuracy.
-The costs of the waiting time until a decision affect the Nash equilibrium thresholds for a pair of animals, the optimal thresholds for a group of animals, and the skewness and information content of the resulting distribution of power.
+The same stochastic differential equations can be used to model these quite different systems, which suggests that there are universal principles underlying learning through noisy dynamical systems.
+In the neural system, the differential equations are often reduced to a one-dimensional model, but the social system is inherently two-dimensional.
+The thresholds dictating when a decision has been made affect both the accuracy of the decision and the expected time until a decision is reached.
+We find that the dimensionality of the decision process does not affect its accuracy.
+We find that if a pair of individuals optimize their thresholds to avoid waiting, a pair making an easy decision will take more time than a pair making a difficult decision.
+In a group of individuals optimizing their thresholds to avoid waiting, those pairs making the difficult decisions about high value inputs will have to wait the longest.
+Finally, the skewness of the power distribution in the social system is maximized at intermediate waiting costs and the information content of the power distribution is higher when the individual maximize their probability of becoming dominant rather than the accuracy of their decisions.
 In comparing the output of the model of the social system to data, we found that ...
@@
 
-Similar models have previously been applied to animal conflicts, although in that case the animals had perfect memory of their past interactions \citep{Froment:2010fk}
+Similar models have previously been applied to animal conflicts, although in that case the model assumed that the animals had perfect memory of their past interactions \citep{Froment:2010fk}
 
@@
 In Table \ref{variables}, we list the inputs, decision variables, and outputs of the two systems.
+
+%In the neural system, an individual wants to make both a decision that is both quick and accurate.
+This is also true in the social system.
+However, it may also be the case that a monkey wants to avoid emitting a subordination signal and wants to receive a signal, regardless of whether or not this is the ``correct" outcome, i.e.
+whether or not he is actually more likely to win a fight.
+The criterion used to optimize the threshold, therefore, may be quite different in the two systems.
+Additionally, the speed and outcome of the decision process depend on both individuals' thresholds.
+Since an individuals' best behavior depends on the behavior of the other animal, optimizing the threshold has a strategic aspect.
 
@@
 \begin{equation*}
-U_i=w_1\text{DT}+w_2\text{ER}+w_3\text{SP}_i.
+U_{i}=w_1\text{DT}+w_2\text{ER}+w_3\text{SP}_i.
 \end{equation*}
@@
 
+\subsection{Optimal thresholds }
+An individual-- either a neural population or a monkey-- can be engaged in either a single decision or in many decisions at once.
+In the social system, for example, an animal wants to decide whether it is dominant or subordinate to every other member of its social group.
+For any pair of individuals, each has a utility, $U_1$ and $U_2$, and consequently each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.
+Similarly, in a group of individuals, each animal $i$ has a utility $U_{ij}$ from its decision process with animal $j$.
+These lead to an optimized set of thresholds such that no individual has an incentive to choose another threshold.
+
+Since an individual is trying to use the same algorithm to learn about every other animal, he is constrained to have only one decision threshold for each of these decision processes.
+The optimal decision threshold in this group context depends on the distribution of difficulties ($c$) of the decision an individual will have to make and the thresholds of the other individuals.
+In the social system, the distribution of difficulties ($c$) comes from the distribution of fighting abilities of the animals.
+
+ We use the following procedure to find optimal thresholds as a function of position in the group.
+We repeatedly draw a set of fighting abilities from a uniform distribution; we find the optimal thresholds for each set, and then find the average optimal threshold as a function of an animal's order in the set of fighting abilities.
+To find the optimal thresholds for a given set of fighting abilities, we turn the difference in fighting abilities between each pair into the difficulty of the decision they have to make ($c$).
+For an initial set of thresholds, we find $U_{ij}$ for each animal $i$ in each of its decisions and then average $\langle U_{ij}\rangle_j$ across all the animals with which it is engaged in a decision process.
+We then allow each animal to choose a threshold that optimizes $\langle U_{ij}\rangle_j$ with respect to the rest of the group's thresholds and we iterate these choices until no animal chooses a new threshold.
+
+\subsection{Consensus }
+Consensus about the value of each individual in a group emerges out of the decision process.
+In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Flack:2004oq,Flack:2006uq,Brush:2013fk}.
+Given the decisions that each pair of individuals reaches, there are measures of the consensus in the group that successfully predict an animal's social power and also extend to other social systems \cite{Brush:2013fk}.
+For each random set of fighting abilities, we let the animals optimize their thresholds, which give the probability that each pair will reach either decision, and we then generate a set of decisions given by these probabilities.
+These decisions then give a distribution of power among the group.
+For many random sets of fighting abilities, we can find the mutual information between the power distribution and the true fighting abilities, which we call the information content of the power distribution.
+It tells us how informative observing the outcomes of the decisions is about the underlying fighting abilities.
+We also find the average skewness of the power distributions.
+
 \section{Results}
@@
 $T_1=T_2$.
-
 We find that the two-dimensional decision is as accurate as the one-dimensional decision.
@@
 
-\subsection{Strategic Decision Making }
-Differences in the thresholds could be explained by individual variation.
-However, there may be a strategic aspect to this variation in the social system.
-In the neural system, an individual wants to make both a decision that is both quick and accurate.
-This is also true in the social system.
-However, it may also be the case that a monkey wants to avoid emitting a subordination signal and wants to receive a signal, regardless of whether or not this is the ``correct" outcome, i.e.
-whether or not he is actually more likely to win a fight.
-The criterion used to optimize the threshold, therefore, may be quite different in the two systems.
-Additionally, the speed and outcome of the decision process depend on both animals' thresholds.
-Since an animal's best behavior depends on the behavior of the other animal, optimizing the threshold is strategic in the social system but not in the neural system.
+\subsection{Increasing waiting costs make easy decisions take longer. }
 
-For any pair of individuals, each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.
-These thresholds depend on the relative dominance of the two animals and how costly waiting is.
-If waiting is not costly and the error rate does not matter ($w_1=0$, $w_2=0$, $w_3>0$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so they will set high thresholds in order to wait until the other individual gives up.
+If waiting is not costly and the error rate does not matter ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so the Nash equilibrium thresholds are high and both individuals wait until the other gives up.
 However, as waiting costs increase, the pair can reach a decision more quickly if the weaker individual lowers its threshold.
 If waiting costs are high enough, the stronger individual will also lower its threshold to avoid a long decision time (Figure \ref{nasheq_thresholds}).
-By increasing the costs of waiting, the weaker individual has an incentive to give up more quickly, which can actually improve the accuracy of the decision process (Figure \ref{nasheq_thresholds}).
-For low to moderate waiting costs, pairs of individuals making a difficult decision (low $c$, e.g.
+Pairs making difficult decisions make more accurate decisions when the costs of waiting are higher since the weaker animal has an incentive to give up more quickly (Figure \ref{nasheq_thresholds}).
+For low waiting costs, pairs of individuals making a difficult decision (low $c$, e.g.
 animals with similar fighting abilities) will take longer to reach a decision than pairs making a difficult decision (high $c$, e.g.
 animals with disparate fighting abilities).
-However, when costs are high, the pairs that take the longest to reach a decision are those making moderately difficult decisions (Figure \ref{nasheq_thresholds}).
-This is because pairs making difficult decisions will both set their thresholds quite low and in pairs making difficult decisions the weaker will set its threshold much lower, both of which decrease the time it takes until reaching a decision.
+At intermediate costs, pairs of individuals making intermediate decisions will take the longest to reach a decision.
+When costs are high, the pairs that take the longest to reach a decision are those making the easiest decisions (Figure \ref{nasheq_thresholds}).
+
+As with the Nash equilibrium thresholds, if waiting is not costly and the error rate does not matter ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so they will all set high thresholds.
+As waiting costs increase, first weaker animals will start to lower their thresholds and then strong animals will as well (Figure \ref{groupeq_thresholds}).
+For low waiting costs, pairs of animals with similar fighting abilities, and thus making difficult decisions, take the longest to reach a decision.
+For high waiting costs, pairs of animals with similar and high fighting abilities take much longer, and those animals with similar and low fighting abilities will reach a decision more quickly.
 
 \subsection{Costs affect the shape and the information content of the power distribution.  }
-A monkey can try to decide whether it is more or less likely to win a fight than another monkey using information from the fights the pair has previously engaged in.
-Further, he can try to make such a decision about every other member of its social group.
-Since he is trying to use the same algorithm to learn about every other animal, he is constrained to have only one decision threshold for each of these decision processes.
-The optimal decision threshold in this social context depends on the distribution of fighting abilities in the group and the decision thresholds of the other animals.
 
-When waiting for a decision becomes more costly ($c_2$ increases), the resulting distribution of the number of signals received first becomes more skewed, with a heavier right tail, and then loses it skewness again (Figure \ref{power}).
-Changing $c_1$ and $c_3$ have less of an effect, although when $c_1$ is higher the skewness slightly increases.
+The average skewness of the power distribution is maximized at intermediate waiting costs ($w_2$) (Figure \ref{power}).
+Changing $w_1$ and $w_3$ have less of an effect, although when $w_1$ is higher the skewness slightly increases.
+The information content decreases as waiting costs increase.
+Counterintuitively, if the individuals maximize the probability of being dominant in each of its pairs rather than the accuracy of those decisions, the information content of the power distribution increases (Figure \ref{power}).
+In other words, the information in all of the group's decisions is higher when the accuracy of pairwise decisions decreases.
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{2d_accuracy.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{2d_flexible_thresholds.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{nasheq_thresholds.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=\textwidth]{groupeq_thresholds.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \begin{subfigure}{3.415in}
 \caption{}
-\includegraphics[width=\textwidth]{skewness_heatmap.pdf}
+\includegraphics[width=\textwidth]{skewness_heatmap_poster.pdf}
 \end{subfigure}
@@
 \caption{}
-\includegraphics[width=\textwidth]{mutinfo_heatmap.pdf}
+\includegraphics[width=\textwidth]{mutinfo_heatmap_poster.pdf}
 \end{subfigure}
-\caption{\label{power} As the costs of waiting increase, the distribution of the number of relationships in which an animal is dominant becomes skewed to the right and then loses its skewness as costs increase further.
-As the costs of waiting increase, the information content of the power distribution always decreases and if dominance is more important than accuracy the information content is higher.
+\caption{\label{power} Average skewness of the power distribution is maximized at intermediate waiting costs.
+The information content of the power distribution always increases as waiting costs decrease and when the benefits of being dominant increase.
 The simplex shows the possible combinations of priorities, $w_1$, $w_2$, and $w_3$ such that the three weights sum to $1$.
@@
 
-\begin{table}
+\begin{table}[ht]
 \centering
@@
 
-\begin{table}\centering
+\begin{table}[ht]
+\centering
 \caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} Differences are highlighted in red.}

### 2014-07-21 paper_skeleton.tex
there is now a complete results section

@@
 \section{Introduction}
+
+Collective computation
+
+decision making previously and why social systems are different
+
+specifically, how do monkey groups compute their power structure?
+
+the model, which extends neural models, reduces to game theory, and can be derived from first principles with only a few assumptions!  aren't we great?
 
@@
 For each random set of fighting abilities, we let the animals optimize their thresholds, which give the probability that each pair will reach either decision, and we then generate a set of decisions given by these probabilities.
-These decisions then give a distribution of power among the group.
-For many random sets of fighting abilities, we can find the mutual information between the power distribution and the true fighting abilities, which we call the information content of the power distribution.
-It tells us how informative observing the outcomes of the decisions is about the underlying fighting abilities.
+Given this signaling network, we use various formalisms to measure the consensus about each node in the network, i.e.
+a power score for each animal in the group.
+For many random sets of fighting abilities, we can find the mutual information between the power score and the true fighting abilities.
+This tells us how informative observing the outcomes of the decisions is about the underlying fighting abilities.
 We also find the average skewness of the power distributions.
+
+In the following we consider four types of consensus formalisms.
+The first and simplest is simply the number of signalers submitting to each animal.
+The second is the entropy of the distribution of the number of signals sent to each animal, which measures how uniformly all other animals in the group behave with respect to a focal animal.
+The third is the eigenvector centrality of the signaling network, which measures how central each node is in the global structure of the network.
+Finally, there are a number of ways to Incorporate the number of signals received into a measure of power, including just counting that number or multiplying that number by the entropy.
+However, these measures all perform very similarly with respect to how informative they are so we group them together in the following analysis.
+For more details see \cite{Brush:2013fk}.
 
@@
 That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability.
-Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher (Figure \ref{2dacc}).
+Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.
 
-Without any prior knowledge of the probabilities of the state of the external stimulus, it is reasonable to assume that the thresholds for deciding for each direction are symmetric.
-On the other hand, the thresholds are chosen by two competing individuals-- neural populations or animals-- and it is possible that the two individuals have different thresholds.
-The accuracy in a given amount of time depends on the ratio of the two thresholds and is improved when the weaker animal's threshold is lower than the stronger animal's threshold (Figure \ref{2dflex}).
+\subsection{Prioritizing signaling probability does not always decrease pairwise decision accuracy. }
+We first look at the Nash equilibrium thresholds when the animals only care about one property of the decision (error rate, signaling property, or decision time).
+In these extremes, the Nash thresholds do not depend on how difficult the decision is.
+The pair can make the most accurate decision if the weaker animal chooses a low threshold and the strong animal chooses a high one, so if only error rate matters ($w_1=1$, $w_2=0$, $w_3=0$), those will be the Nash  thresholds.
+If only signaling probability matters ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so both individuals wait until the other gives up and the Nash  thresholds are both high.
+If only decision time matters ($w_1=0$, $w_2=1$, $w_3=1$), both animals have an incentive to set their thresholds low and reach a decision quickly, whatever that decision may be (Figure \ref{nasheq_thresholds}).
+Notice that in changing from prioritizing signaling probability ($w_3=1$) to prioritizing decision time ($w_2=1$), first the weaker animal lowers its threshold and \emph{then} the stronger animal lowers his.
 
-\subsection{Increasing waiting costs make easy decisions take longer. }
+For intermediate optimization weights, the Nash  thresholds depend on the difficulty of the decision being made.
+The Nash thresholds for two animals making a difficult decision ($c$ close to $.5$) are more similar to each other than those for two animals making an easy decision.
+For a pair making a difficult decision, their similar Nash thresholds decrease as waiting costs become more important (Figure \ref{nasheq_thresholds}).
+Once waiting costs are sufficiently high, the tradeoff between error rate and signaling probability does not affect the Nash thresholds.
 
-If waiting is not costly and the error rate does not matter ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so the Nash equilibrium thresholds are high and both individuals wait until the other gives up.
-However, as waiting costs increase, the pair can reach a decision more quickly if the weaker individual lowers its threshold.
-If waiting costs are high enough, the stronger individual will also lower its threshold to avoid a long decision time (Figure \ref{nasheq_thresholds}).
-Pairs making difficult decisions make more accurate decisions when the costs of waiting are higher since the weaker animal has an incentive to give up more quickly (Figure \ref{nasheq_thresholds}).
-For low waiting costs, pairs of individuals making a difficult decision (low $c$, e.g.
-animals with similar fighting abilities) will take longer to reach a decision than pairs making a difficult decision (high $c$, e.g.
-animals with disparate fighting abilities).
-At intermediate costs, pairs of individuals making intermediate decisions will take the longest to reach a decision.
-When costs are high, the pairs that take the longest to reach a decision are those making the easiest decisions (Figure \ref{nasheq_thresholds}).
+The thresholds affect the accuracy and decision time with which a pair can reach a decision.
+For a difficult decision ($c$ close to $.5$), the accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and lowest when only signaling probability matters ($w_3=1$).
+Accuracy decreases as the weight given to decision time increases (Figure \ref{nasheq_thresholds}).
+Since, for sufficiently high waiting costs, the Nash thresholds do not depend on the tradeoff between error rate and signaling probability, neither does the accuracy for a pair using Nash thresholds.
+In other words, once waiting costs are high, a pair of animals each minimizing error rate and a pair minimizing signaling probability will reach a decision with the same accuracy.
 
-As with the Nash equilibrium thresholds, if waiting is not costly and the error rate does not matter ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so they will all set high thresholds.
-As waiting costs increase, first weaker animals will start to lower their thresholds and then strong animals will as well (Figure \ref{groupeq_thresholds}).
-For low waiting costs, pairs of animals with similar fighting abilities, and thus making difficult decisions, take the longest to reach a decision.
-For high waiting costs, pairs of animals with similar and high fighting abilities take much longer, and those animals with similar and low fighting abilities will reach a decision more quickly.
+\subsection{Intermediately difficult decisions sometimes take the longest. }
 
-\subsection{Costs affect the shape and the information content of the power distribution.  }
+For most optimization weights, it takes more time to make a difficult decision than to make an easy one.
+However, at intermediate costs, pairs of individuals making intermediate decisions will take the longest to reach a decision.
+An easy decision can always be reached quickly.
+Higher waiting costs incentivize pairs making difficult decisions to lower their thresholds more than pairs making intermediately difficult decisions, with the result that they can actually reach a decision more quickly.
+It is the pairs making intermediately difficult decisions using intermediate thresholds that will take the longest in this case (Figure \ref{nasheq_thresholds}).
 
-The average skewness of the power distribution is maximized at intermediate waiting costs ($w_2$) (Figure \ref{power}).
-Changing $w_1$ and $w_3$ have less of an effect, although when $w_1$ is higher the skewness slightly increases.
-The information content decreases as waiting costs increase.
-Counterintuitively, if the individuals maximize the probability of being dominant in each of its pairs rather than the accuracy of those decisions, the information content of the power distribution increases (Figure \ref{power}).
-In other words, the information in all of the group's decisions is higher when the accuracy of pairwise decisions decreases.
+\subsection{Decisions in a group context take longest between animals with similarly high abilities. }
 
-\subsection{Comparison to social data }
+After analyzing Nash equilibrium thresholds between pairs of animals, we found the optimal thresholds for a group of animals ($N=20$) such that no animal had an incentive to choose a different threshold.
+As with pairwise Nash thresholds, when the animals only care about signaling probability ($w_3=1$), all animals in the group choose high thresholds, whereas if they only care about decision time ($w_2=1$), all animals choose low thresholds.
+When the animals only care about accuracy ($w_1=1$), all animals choose high thresholds, except the weakest who choose a low threshold.
+At intermediate weights, thresholds decrease as position in the group decreases (Figure \ref{groupeq_thresholds}).
+
+When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with less similar abilities also take a fairly long time.
+When waiting costs are high, most pairs can reach a decision quite quickly, except those who have similar and high abilities (Figure \ref{groupeq_thresholds}).
+The average accuracy decreases as waiting costs increase (Figure \ref{groupeq_thresholds}).
+
+As soon as any waiting costs are introduced, as with the pairwise Nash thresholds, the weaker animals have incentives to lower their thresholds to make quicker decisions at the expense of accuracy.
+This incentive is stronger when error rate is more important than signaling probability.
+The result is that all the weakest animals set their thresholds as low as possible, making the accuracy of decisions between those animals very poor and decreasing the overall accuracy of the group.
+In other words, average accuracy \emph{increases} as signaling probability is prioritized over error rate, except when there are no waiting costs (Figure \ref{groupeq_thresholds}).
+The mutual information of any measure of consensus in the signaling network increases (more or less) with accuracy so it too increases as signaling probability becomes more important.
+
+\subsection{The skewness of the power distribution is maximized at intermediate waiting costs.  }
+For each measure of consensus in the signaling network, we find the average skewness of the distribution of power scores of a group using optimal thresholds, as they depend on optimization weight.
+We find that for all measures skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and signaling probability (Figure \ref{skewness}).
+(Results for number of signalers are shown in Figure \ref{skewness} and others are similar).
+When waiting costs are low and accuracy is high, the signals accurately represent the animals' true fighting abilities so the distribution of number of signalers accurately represents the distribution of abilities, which is not very right-skewed.
+When waiting costs are high and accuracy is low, all animals receive signals from a similar number of others, giving a distribution that is quite uniform and not right-skewed.
+At intermediate waiting costs, the signals sent to animals with low to middle abilities are very noisy so that the animals in the bottom and middle of the group receive signals from similar numbers of animals.
+However, the group accurately signals to the top couple of animals, giving them high scores, resulting in a right-skewed distribution.
+
+\subsection{Which measure of consensus is most informative depends on how accurate pairwise decisions are. }
+We consider four types of measures of consensus in the signaling network to evaluate which is most informative about the underlying fighting abilities.
+When pairwise decisions are quite accurate, the number of signals received is the most informative about the underlying fighting ability (Figure \ref{bestmetric}).
+This makes sense since when each signal accurately reflects the different in ability between receiver and sender, the number of signals is informative about an animal's position in the group.
+When the pairwise decisions are less accurate, global structure of the signaling network provides information about ability that the individual relationships do not, so eigenvector centrality is the measure that is the most informative about ability.
+
+The number of signalers consistently overestimates animals' ability.
+When accuracy is fairly high, this bias makes it uninformative about ability.
+However, when accuracy is quite low, this consistent bias actually make the number of signalers more informative than measures that both underestimate and overestimate ability.
+When accuracy is lower still, none of the measures are very informative, but entropy correctly gives the lowest animals low scores, giving it an edge over the other measures.
+
+\section{Discussion}
+what's interesting and exciting and new?
+
+we can derive sde's from first principles
+
+prioritizing signaling probability sometimes makes decision more accurate
+
+prioritizing waiting costs change which decisions take the longest
+
+optimization weights change the overall social structure
+
+extension of neural model by removing third party comparison and allow individuals to minimize signaling probability
+
+extension of game theory by introducing heterogeneous abilities
+
+extension of David's paper
+
+explanation for variation across species of macaque
+
+evolutionary implications
+
+%comparison to social data
 Empirically, those pairs of monkeys that take the longest to establish a signaling dynamics are those with the most similar fighting abilities...
 
-\section{Discussion}
+skewness
+they're using eigenvector centrality
 
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{2d_accuracy.pdf}
-\caption{\label{2dacc} }
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/pairwise_accuracy.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/nasheq_thresholds.pdf}
+\caption{\label{nasheq_thresholds}  The accuracy of a pair using Nash equilibrium thresholds decreases as the weight given to decision time increases.
+(a) In the upper panel, the color indicates the accuracy of a decision made by a pair making a difficult decision ($c=.55$) using Nash equilibrium thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
+In the lower left corner of the simplex, only error rate matters ($w_1=1$).
+In the upper corner, only decision time matters ($w_2=1$).
+In the lower right corner, only signaling probability matters ($w_3=1$).
+Accuracy decreases as $w_2$ increases but is not greatly affected by a tradeoff between $w_1$ and $w_3$.
+In (b)-(g) we show how the Nash equilibrium thresholds (top row) and decision time for a pair using those thresholds (bottom row) depends on the difficulty of the decision ($c$).
+The optimization weights for each panel are indicated in the simplex with the corresponding letter.
+At intermediate waiting costs (f), intermediately difficult decision take the longest.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/groupaccuracy_heatmap.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/groupeq_thresholds.pdf}
+\caption{\label{groupeq_thresholds}   The accuracy of a group using optimal thresholds decreases as the weight given to decision time increases.
+In the upper panels, the color indicates the average accuracy of the decisions between individuals in the bottom quartile (a) and the average accuracy of the whole group (b) using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
+In the lower left corner of the simplex, only error rate matters ($w_1=1$).
+In the upper corner, only decision time matters ($w_2=1$).
+In the lower right corner, only signaling probability matters ($w_3=1$).
+Both bottom-quartile and total accuracy decrease as $w_2$ increases, and the increase in bottom-quartile accuracy as $w_3$ increases drives a slight increase in total accuracy.
+In (c)-(h) we show how the group-context optimal thresholds (top row) and decision time for each pair (bottom row) depends on the true order of ability where $1$ is the strongest animal and $20$ is the weakest.
+The optimization weights for each panel are indicated in the simplex with the corresponding letter.
+For low waiting costs (c and d), it takes longest for pairs with similar ability to reach a decision.
+For slightly higher waiting costs (f and g), it takes longer for pairs with similar and high ability to reach a decision.}
 \end{figure}
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{2d_flexible_thresholds.pdf}
-\caption{\label{2dflex} When the two thresholds in the two-dimensional system are allowed to vary freely, the decision process can be arbitrarily accurate in a given amount of time, depending on how much lower the weaker animal's threshold is than the stronger animal's threshold.
-The expected time to reach a decision is on the x-axis and the probability of reaching the correct decision in that amount of time is on the y-axis.
-The curve depends on the ratio of the weaker animal's threshold to the stronger animal's threshold ($\rho$).
-Lower ratios are represented in higher darker curves.}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/skewness_heatmap.pdf}
+\caption{\label{skewness} The skewness of the distribution of the number of signalers is maximized at intermediate waiting costs.
+In the lower left corner of the simplex, only error rate matters ($w_1=1$).
+In the upper corner, only decision time matters ($w_2=1$).
+In the lower right corner, only signaling probability matters ($w_3=1$).}
 \end{figure}
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{nasheq_thresholds.pdf}
-\caption{\label{nasheq_thresholds} If a pair of individuals optimize their thresholds in response to each other, as the costs of waiting increase, first the weaker individual will start to lower his threshold to avoid waiting for a long time and then the stronger individual will start to lower his threshold as well.
-As the costs of waiting increase, first the accuracy with which a pair can make a difficult decision increases and then it decreases.
-At low waiting costs, difficult decisions take the most time.
-At intermediate costs, intermediate decisions take the most time.
-At high costs, easy decisions take the most time.
-The first row of panels shows the Nash equilibrium thresholds as a function of the difficulty of the decision ($c$ close to $.5$ is a more difficult decision).
-The second and third rows show, respectively, the probability of the pair reaching the correct decision and the expected time until a decision is reached, if the pair of individuals uses the Nash equilibrium thresholds.
-Each column corresponds to a particular waiting cost and the cost increases from left to right ($w_1=0, w_3=1-w_2$).
-}
-\end{figure}
-
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{groupeq_thresholds.pdf}
-\caption{\label{groupeq_thresholds} If the animals optimize their thresholds with respect to the thresholds and fighting abilities of the whole group of animals, as the costs of fighting increase, first the weaker animals in the group lower their thresholds to avoid fighting for a long time with the stronger animals and then the stronger animals lower their thresholds as well.
-The ability of to learn from an animal's signaling behavior decreases as the costs of fighting increase.
-When fighting is not costly, the pairs of animals that take longest to decide on a relationship are those with similar fighting abilities.
-As costs increase pairs of animals with similarly low fighting abilities resolve their relationships quickly, whereas pairs of animals with similarly high fighting abilities take a long time to resolve their relationships.
-The first row of panels shows the optimal threshold as a function of the animal's true rank with respect to fighting ability.
-The second row shows the mutual information between the probability of emitting a signal and the probability of being the weaker animal, as a function of the animal's true rank.
-The third row shows the expected time until each pair of animals resolves their relationship.
-Each column corresponds to a particular waiting cost and the cost increases from left to right ($w_1=0, w_3=1-w_2$).
-}
-\end{figure}
-
-\begin{figure}[ht]
-\begin{subfigure}{3.415in}
-\caption{}
-\includegraphics[width=\textwidth]{skewness_heatmap_poster.pdf}
-\end{subfigure}
-\begin{subfigure}{3.415in}
-\caption{}
-\includegraphics[width=\textwidth]{mutinfo_heatmap_poster.pdf}
-\end{subfigure}
-\caption{\label{power} Average skewness of the power distribution is maximized at intermediate waiting costs.
-The information content of the power distribution always increases as waiting costs decrease and when the benefits of being dominant increase.
-The simplex shows the possible combinations of priorities, $w_1$, $w_2$, and $w_3$ such that the three weights sum to $1$.
-In the lower left corner, $w_1=1$ (only accuracy matters), in the upper corner $w_2=1$ (only waiting time matters), and in the lower right $w_3=1$ (only the probability of receiving the signal matters).
-The heatmaps show (a) the skewness of the power distribution and (b) the mutual information of the power distribution when the animals in the group optimize their thresholds according to those weights.
-%Increasing or decreasing $c_2$ has the strongest effect.  Increasing $c_1$ slightly increases the skewness. 
-}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/mostinformative_heatmap.pdf}
+\caption{\label{bestmetric}  Which measure of consensus is most informative about fighting ability depends on how accurate decision are on average.
+When decision are very accurate or very inaccurate, measures of the number of signals received are most informative.
+When accuracy is slightly lower, eigenvector centrality of the signaling network is most informative.
+When accuracy is slightly lower again, the number of signalers is most informative.
+When accuracy is lower still, the entropy of signals received is most informative.
+In the lower left corner of the simplex, only error rate matters ($w_1=1$).
+In the upper corner, only decision time matters ($w_2=1$).
+In the lower right corner, only signaling probability matters ($w_3=1$).}
 \end{figure}
@@
 \\ optimality criterion &  reward from being ``right" && reward from being ``right"
-\\ & reward rate && reward rate
-\\ & Bayes risk && Bayes risk
+\\ & decision time && decision time
 \\ & && \fcolorbox{red}{white}{reward from  receiving signal}
@@
 \\ & && \fcolorbox{red}{white}{other animal's threshold}
-\\threshold  algorithm & error correction
-\\  & hill climbing
 \end{tabular}

### 2014-10-22 ms_draft_07_22.tex
ive added code for tracking mutual information over time and changed figures a bit

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{natbib}

\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}
\newtheorem{ass}{Assumption}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf \LARGE{Conflict and Strategy in Collective Computation (???)}}
\end{center}

\section{Abstract}
In neural and social systems, an individual's position in an interaction network can be informative about its function within the group. The development of the network and scores given by each node's position is a collective computation whose output is available to group members.  For example, in social groups, power is defined as the degree of consensus in the network about each node's value. We use a leaky integrator model for a noisy decision process that gives the direction and magnitude of each edge in the network. We find that differences in the model when it is applied to neural and social systems do not affect the accuracy of decisions. We find that a group of individuals prioritizing personal preference over accuracy make more accurate decisions. We find that when pairwise decisions are accurate, local properties of the network are the most informative, but when accuracy decreases global properties are more informative. Finally, we find that the shape of the distribution of power is determined by waiting costs of those decisions. Our model can be applied to a number of systems and our results suggest that conflict and strategy are important in the collective computations performed by all of them.

\section{Keywords }
collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power, social structure

\section{Introduction}

When individuals come together in groups, their computational abilities can be combined in nonlinear ways, giving rise to collective computations that are better than would be expected from the individuals' abilities (ref XXX).  This collective computation cannot be decomposed into parts that are caused or performed by individuals. In schools and flocks, the group can sense its environment and navigate more accurately than any of its individual members can (ref XXX).  Similarly, a group of quorum sensing bacteria can change their behavior, for example color or phosphorescence, based on the properties of the environment that no individual cell would be able to detect (ref XXX). The brain performs computations and makes decisions of which no individual neuron would be capable (ref XXX). Social groups of animals create social structures, which provide information about their members to which no individual would otherwise have access \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}. As external observers to these systems, the integration of individual behaviors into a performance that is much greater than the sum of its parts can seem an amazing feat. To the members of the systems, for example birds in a flock or monkeys in a social group, the group level output can be a new source of information and can be used to make decisions about how to behave.

Some of these collective computations rely on an interaction network between the individual members of the group. There are many measurements of position in a network that assign a value to each node, for example a node's in-degree or centrality within the network.  Any of these can be thought of as a collective computation, a group level output that depends on the structure of the network and whose dependence on any individual behavior is difficult to describe.  If each individual in the group has a functional role or value, as well as a position in the interaction network, then a useful collective computation would give individuals scores that are informative about these underlying values.

An animal's power is defined as the degree of consensus in its group about its fighting ability and the distribution of these power scores is referred to as the power structure \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq,Preuschoft:2004ly}. In macaques, power is a useful variable since it reflects the animals' true abilities and the animals themselves do use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.  In pigtailed macaques, the animals have a stereotyped signal, a formal subordination signal, that one animal will emit to another to communicate that it agrees to be subordinate \cite{Flack:2004oq,Flack:2006uq,Flack:2007kx,Preuschoft:2004ly}. Since a subordination signal indicates one animal believes another to have a higher fighting ability, a network of subordination signals can be used to determine the degree of consensus in the whole group about any individual's ability.  In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node \cite{Brush:2013fk,Flack:2006uq}.  We found a few that produced scores that were predictive of individuals' functional behaviors, which suggests they provide information about the animals' true fighting abilities.  This kind of consensus may be important in neural systems, as well.  A number of neural populations may respond to or encode different environmental stimuli or variables.  The consensus in a network of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.

In each of these systems, the individuals-- monkeys or neural populations-- are trying to learn and make decisions in a noisy environment.  In such a situation, an individual acquires evidence that may or may not reflect the true state of the environment and only makes a decision once it has acquired enough evidence to reduce its uncertainty sufficiently.  If the environment can be in one of the two possible states, the optimal way to decide between these alternatives is an algorithm called a sequential probability ratio test (SPRT) (ref XXX).  An individual using this learning method gathers a sequence of pieces of evidence that he uses to update his belief about the state of the world until this belief is strong enough to make a decision.  This updating process can equivalently be written as a set of stochastic differential equations (ref XXX), which are more flexible in allowing for memory loss.  This type of model, sometimes referred to as a leaky integrator model, is applicable to any system in which an individual gathers evidence and chooses the alternative for which he has gathered more support.

Noisy learning models can be used to model animal conflicts, and the SPRT has been previously used to describe animal conflicts over resources \citep{Froment:2010fk}. In a group of macaques, animals have different fighting abilities, and when a pair of animals engages in a fight, one animal is more likely on average to win.  Over time, as one animal learns that it is likely to lose fights, it will decide to emit the formal subordination signal, communicating its agreeing to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk}.  The outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other factors.  Because of this stochasticity, the number of fights won and lost in any period of time is a random variable.  Each animal's opinion about its relative dominance, therefore, will perform a biased random walk until one animal's opinion of itself is low enough to agree to being subordinate, and this process can be describe by a leaky integrator model.

Leaky integrator models have been used to describe decision making in both monkeys and humans \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.  Often an experimental subject is presented with a visual stimulus that consists of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.  The subject is then asked to decide the direction in which the dots are moving.  If there are two neural populations, each of which fires in response to a particular direction of motion, one firing rate will tend to increase and the other will tend to decrease, depending on the direction of the stimulus.  The random dots make the stimulus noisy and sometimes the firing rates will change in the wrong direction. Eventually, either the experimenter forces a decision to be made, in which case whichever population has a higher firing rate determines which decision is made, or the firing rate of one population becomes high enough that the subject decides for that direction.  In the model, the two decision variables--- the firing rates of two neural populations responding to left and right motion--- perform a biased random walk until enough evidence has been accumulated to make a decision one way or the other.  This model has successfully explained and predicted performance in a number of experiments \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.

The use of stochastic differential equations to model noisy learning has been rather phenomenological.  Here we turn a simple mechanistic description of the accumulation of information into a set of stochastic differential equations, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}. We then use this model to answer the following questions. Are there differences between decision making in the neural and social systems? What method of measuring position in an interaction network is the most informative about individuals' true values? How should individuals make decisions such that the resulting collective computation is most accurate? How is the shape of the power structure affected by how individuals make decisions? How well does our model describe empirical social systems?

\section{Model} 
\subsection{Stochastic differential equations }    
%% make debt to Gillespie clearer

There are two possible types of environmental evidence: in the neural system, a dot can move either left or right and in the social system, a fight can be won by one animal or another.  For clarity in the following we will refer to these left or right dots, but the biological interpretation of the two types of inputs depends on the system. There are two decision variables, $X_1$ and $X_2$. In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$.

Each decision variable responds positively to one type of input and negatively to the other.  In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.  In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite. Neither decision variable has perfect memory.  Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.     
To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
\end{align*}

Our first assumption is that the rates at which the two types of inputs occur is constant over time.  We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$.  We rewrite our basic equations to give 
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+bN_\text{L}-bN_\text{R}
\\ X_1(t+\tau)&=(1-\ell\tau)X_2(t)-bN_\text{L}+bN_\text{R}.
\end{align*}
If events happen at a rate $r$ and the event is a left dot with probability $c$ and a right dot with probability $1-c$, then the expectation of $N_L$ and $N_R$ in a period of length $\tau$ are, respectively, $\tau r c$ and $\tau r(1-c)$. In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.  A $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.

If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with Normal random variables with mean and variance equal to the mean of the Poisson random variables.  Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation. Let $Z_\text{L}$ and $Z_\text{R}$, be independent standard Normal random variables, i.e. with mean $0$ and standard deviation $1$.  Then our equations for $X_i(t+\tau)$ become
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)-b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg)
\\X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)+b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg).
\end{align*}
Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{L}t-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{L}t+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\end{array}
\end{equation}
where $dW_{\text{L}}$ and $dW_{\text{R}}$ are independent Brownian motions representing, respectively, the left and right inputs.  The assumptions about the timescales on which inputs occur are particularly reasonable in the social system, although the successful application of this type of model to neural populations implies they are not unreasonable in that system.  In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the neural and social systems.

\subsection{Reaching a decision }
Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.  The sign and size of the two variables indicate how much evidence there is for each possible decision outcome.  When a lot of evidence for the first possibility has been observed, $X_1$ will be large and positive, and similarly for the second possibility.  If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility. If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.  Specifically, there are two thresholds, $T_2$ and $T_1$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_2$ and $T_1$, and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.

In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels to make the system easier to analyze \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.  However, in the social system we are considering, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.  A ``decision," i.e. the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below a certain threshold.  The social system is therefore inherently two-dimensional, whereas the neural system may be adequately described by one dimension.  In Table \ref{differences}, we compare how the decision is made and how the decision process is optimized in the two systems.

\subsection{Utility of the decision }
The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.  A good decision is one that reaches the correct output quickly, i.e. the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.   However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.  It may also be the case that each individual prefers one output over the other, regardless of which is ``correct''.  In the social case, an animal would prefer that its partner agreed to be subordinate, whether or not it is truly stronger.  In the neural case, different neural populations may receive different rewards if the decision reaches different outputs, regardless of which of the alternatives is ``true."  In these cases, a variable wants to maximize the probability that his preference is reached (probability of preference being chosen, $\text{PP}_i$).

To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{PP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  Similarly, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair. In both cases, $w_3$ represents how stubborn the individuals are about their preferences. If $w_3\gneq 0$ and the individuals prefer different outcomes, neither can simultaneously be satisfied since $\text{PP}_1=1-\text{PP}_2$.

\subsection{Optimal thresholds }
An individual-- either a neural population or a monkey-- can be engaged in either a single decision or in many decisions at once.  In the social system, for example, an animal wants to decide whether it is dominant or subordinate to every other member of its social group.  For any pair of individuals, each has a utility, $U_1$ and $U_2$, and consequently each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.  Similarly, in a group of individuals, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$.  His total utility is given by the average of these, $\langle U_{ij}\rangle _j$. These lead to an optimized set of thresholds such that no individual has an incentive to choose another threshold.

Since an individual is trying to use the same algorithm to learn about every one else, he is constrained to have only one decision threshold for each of these decision processes.  In order to consistently assign the difficulties with which all pairs in a group make decisions, we assign each individual a value $a_i$ and the difficulty with which the pair $i,j$ makes a decision increases with the difference in abilities, i.e. $c_{ij}$ is an increasing function of $|a_i-a_j|$. In the social system, these values can be interpreted as the fighting abilities of the animals.The optimal decision threshold in this group context depends on the distribution of difficulties ($c_{ij}$) of the decision an individual will have to make and the thresholds of the other individuals.

 We use the following procedure to find optimal thresholds as a function of position in the group.  We repeatedly draw a set of fighting abilities from a uniform distribution; we find the optimal thresholds for each set, and then find the average optimal threshold as a function of an animal's order in the set of fighting abilities. To find the optimal thresholds for a given set of fighting abilities, we turn the difference in fighting abilities between each pair into the difficulty of the decision they have to make ($c$).  For an initial set of thresholds, we find $U_{ij}$ for each animal $i$ in each of its decisions and then average $\langle U_{ij}\rangle_j$ across all the animals with which it is engaged in a decision process.  We then allow each animal to choose a threshold that optimizes $\langle U_{ij}\rangle_j$ with respect to the rest of the group's thresholds and we iterate these choices until no animal chooses a new threshold.

\subsection{Collective computation }
We are interested in how the consensus about the value of each individual emerges out of the decision process.  (In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq}.)  Given an interaction network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}. These formalisms assign a consensus score to each node in the network (interpreted as a power score in the social system).  In the following we consider four  consensus formalisms.  The first and simplest is simply the number of nodes sending decisions to each animal (number of signalers).  The second is the number of decisions received by each node, which is a finer measure than the first.  The third is the entropy of the distribution of the number of decisions sent to each individual, which measures gives a coarse measurement of how uniformly all other animals in the group behave with respect to a focal individual.  The fourth is the eigenvector centrality of the signaling network, which measures how central each node is in the global structure of the network.  For more details see \cite{Brush:2013fk}.

To use these measures, we need to construct an interaction network between individuals in the group.  For each random set of fighting abilities, we let the animals optimize their thresholds, which give the probability that each pair will reach either decision and the expected time it will take.  We use these probabilities to generate a set of decisions between all pairs.  We use the convention that a decision is sent from $i$ to $j$ if in their decision $j$'s preference prevailed.  (In the social system, $i$ sends a subordination signal to $j$ to communicate his agreement to be subordinate.) To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be sent, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.  The decision network is then formed such that the edge from $i$ to $j$ is the number of signals sent from $i$ to $j$.  For many random sets of fighting abilities and for each consensus formalism, we can find the mutual information between the consensus scores and the true values and  the average skewness of the set of consensus scores.

\section{Results}
\subsection{Decision making is as accurate without a third party arbiter. }
We compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.  To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e. $T_1=T_2$.  We find that the two-dimensional decision is as accurate as the one-dimensional decision.  That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability.  Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.

\subsection{The average accuracy in a group of stubborn individuals is higher than in a group of accurate individuals. }
The best measure of consensus depends on the accuracy of the decisions, so we address accuracy before moving on to discuss consensus. The Nash equilibrium thresholds for a pair of individuals depend on whether the individuals are minimizing error rate, maximizing probability of their preference, or minimizing decision time.  The pair can make the most accurate decision if the weaker individual chooses a low threshold and the strong individual chooses a high one, so if only error rate matters ($w_1=1$, $w_2=0$, $w_3=0$), those will be the Nash  thresholds.   If only preference matters ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so both individuals wait until the other gives up and the Nash  thresholds are both high.  If only decision time matters ($w_1=0$, $w_2=1$, $w_3=1$), both individuals have an incentive to set their thresholds low and reach a decision quickly, whatever that decision may be (Figure \ref{nasheq}).

The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and lowest when only decision time matters ($w_2=1$).  When waiting costs are low, accuracy decreases as the weight given to preference increases (Figure \ref{nasheq}).  It makes sense that as the individuals prioritize preference over error rate, they make increasingly inaccurate decisions.  For sufficiently high waiting costs, however, the Nash thresholds do not depend on the tradeoff between error rate and preference, so neither does the accuracy for a pair using Nash thresholds.  In other words, once waiting costs are high, a pair of individuals minimizing error rate and a pair each of whom maximizes the probability of reaching his preference will reach a decision with the same accuracy.

Prioritizing preference at the expense of error rate has a different effect as soon as the individuals are in a group with more than two members.  In a group, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.  If the individuals value error rate, as soon as any waiting costs are introduced individuals at the bottom of the group lower their thresholds because in so doing, they make much faster decisions without losing much accuracy.  On the other hand, if the individuals value preference and waiting costs are introduced, individuals at the bottom of the group will keep their thresholds higher because they can increase their chances of receiving a decision, even though it costs them time (Figure \ref{groupeq}).

When individuals in the bottom of the group have slightly higher thresholds and wait slightly longer to make a decision, the accuracy of decisions between individuals both in the bottom of the group increases.  Even though the accuracy between two individuals in the top of the group is not strongly affected by this tradeoff and the accuracy between an individual in the bottom and an individual in the top decreases slightly, the net effect is that the average accuracy of all pairs in the group increases (Figure \ref{groupeq}).  For every measure of consensus, the mutual information between the consensus scores and the true values increases with average pairwise accuracy.  Thus, the information content of the consensus scores is also slightly higher when the individuals minimizes preference rather than error rate.

\subsection{Global network structure provides the most information when accuracy is high but imperfect. }
When pairwise decisions are on average quite accurate, the number of decisions received is the most informative about the underlying values (Figure \ref{bestmetric}).  This makes sense since when each edge in the decision network accurately reflects the difference in value between individuals, the number of edges received is informative about an individual's position in the group.  When the pairwise decisions are less accurate, global structure of the decision network provides information about values that the individual relationships do not, so eigenvector centrality is the measure that is the most informative about values.

The number of signalers consistently overestimates individuals' value.  When accuracy is fairly high, this bias makes it uninformative about value.  However, when accuracy is quite low, this consistent bias actually make the number of signalers more informative than measures whose errors are less predictable. When accuracy is lower still, none of the measures are very informative, but entropy correctly gives the lowest individuals low scores, giving it an edge over the other measures.

\subsection{The skewness of the power distribution is maximized when decisions are moderately accurate. }
For each measure of consensus in the decision network, we find the average skewness of the distribution of consensus scores of a group using optimal thresholds, as they depend on optimization weight.  We find that for all measures skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and preference (Figure \ref{skewness}). (Results for number of signalers are shown in Figure \ref{skewness} and others are similar).  When waiting costs are low and accuracy is high, the decisions accurately represent the individuals' true  abilities so the distribution of number of signalers accurately represents the distribution of abilities, which is not very right-skewed.  When waiting costs are high and accuracy is low, all individuals receive decisions from a similar number of others, giving a distribution that is quite uniform and not right-skewed.  At intermediate waiting costs, the decisions sent to individuals with low to middle abilities are very noisy so that the individuals in the bottom and middle of the group receive signals from similar numbers of individuals.  However, all individuals make accurately decisions with the top few individuals, giving them high scores, resulting in a right-skewed distribution.

\subsection{Decisions in a group take longest between individuals with similarly high abilities. }
Difficult decisions tend to (but do not always) take longer than easier decisions.  In a social system, the difficult of a decision increases as a pair becomes more similar.  In a group of decision makers, the absolute value of two individuals, as well as the difference in their values, affects how quickly they can make a decision.  When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with different abilities also take a fairly long time.  When waiting costs are high, all pairs reach decisions quickly.  At intermediate costs, most pairs can reach a decision quite quickly, except those who have similar and high abilities (Figure \ref{groupeq}).  Pairs with similar and high abilities always take as long or longer to make a decision than any other pairs do.

\section{Discussion}

We derive a pair of stochastic differential equations that describe how two individuals accumulate evidence from a noisy environment.  The model can be equally well applied to the firing rates of neural populations observing a visual stimulus and the beliefs of two monkeys about their relative dominance based on their fighting history.   We first compare the simplified version of the model that is frequently used to describe the neural system and the full version we need to describe the social system.  We then consider a population of individuals using this model to make decisions to investigate the best way of making a collective computation, how to improve the accuracy of this computation and how the shape of the power structure in the social system depends on the accuracy of decisions.  Finally, we find a model output that is comparable to empirical observations.

Either the absolute value of the decision variables or the difference in their values can determine the outcome of the decision.  We find that, for a given decision time, the accuracy of the decision does not depend on which of these mechanisms is used.  In order for the difference in variables to be used, a third party arbiter needs to be present to evaluate the value of each decision variable and find their difference.  In the brain, it might be the case that there is a third neural population whose activity depends on the two populations gathering external evidence, but in a social group, there is not a third party informing how a pair of animals should establish their subordination-dominance relationship.  The fact, then, that the model does not depend on the existence of this third party arbiter means that it can be equally well applied whether or not it is present.

The decisions between pairs of individuals give rise to a network of decisions between all members of a group. A group of individuals will have some degree of consensus about the status of each individual. This consensus is the result of a collective computation that depends on individual behaviors.  There are many different ways of measuring this consensus and each represents a different way of performing the collective computation \cite{Brush:2013fk,Flack:2006uq}. When the accuracy of pairwise decisions is relatively high but imperfect, then a global metric like eigenvector centrality is the most informative about the true states of the nodes. When pairwise accuracy is quite low, then a coarse measure like the number of signalers or the entropy of decisions received outperforms a finer measure like the number of decisions received.

As network scientists, if we are presented with an interaction network, we would like to know how to measure the consensus present in the network about each node.  If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes, this result gives a rule of thumb for how to pick a measure of consensus. As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities. In previous work, we found that both eigenvector centrality and measures of the number of decisions received produced scores that were informative about animals' functional behavior in a group of pigtailed macaques.  This may be informative about the optimization weights the animals in that group are using.  If so, it would suggest that decision time is less important than a combination of error rate and the probability of being dominant. If animals in different species do prioritize fighting costs and the benefits from being dominant (i.e. decision time and probability of preference)  differently, we might expect to find that those animals use different properties of their social networks to estimate the power structure of the group.

Individuals in biological systems are likely to be selfish and stubborn, both in order to maximize their own welfare and because prioritizing accuracy requires knowing which the ``right'' answer is.  The accuracy of the collective computation depends on how the individuals make the tradeoff between personal preference and accuracy, regardless of which measure is used to make the computation.  We expected to find that individuals valuing their preference at the expense of the correct output would tend to make decisions that were less accurate.  In a pair, this holds true.  However, when there is a group of individuals all making this tradeoff, valuing preferences over accuracy improves the accuracy with which individuals in the bottom of the group resolve their differences, increasing the average accuracy of the whole group. This overall accuracy translates into a more accurate collective computation.

Usually only two neural populations are included in models of neural decision making (although see XXX).  Further, it is usually assumed that neural populations in the brain value accuracy, since the whole organism receives a benefit from making a correct decision in a short time.  But it is possible that different neural populations receive different rewards from making different decisions, regardless of which decision is the ``correct'' one.  Only by considering a group of decision makers do we find the surprising result that the accuracy of a collective computation can be improved by introducing conflicting interests.  This might suggest that in situations where a group of neural populations are recruited to make a decision, we could expect to find that they have different preferences, rather than all valuing accuracy. Our result also has interesting implications for designing networks whose properties we would like to use to make computations.

The skewness of the power structure that results from this collective computation in social groups of macaques has important functional consequences \cite{Flack:2004oq,Flack:2006fk}. We find that for most measures of power the skewness of the distribution of power scores is maximized at intermediate waiting costs.  In our data, we observe a right-skewed distribution of power \cite{Brush:2013fk,Flack:2006uq}.  To the extent that our model recapitulates the empirical system, this would suggest that waiting costs are neither the only priority nor are they ignored completely.

Different species of macaque have stereotypical power structures with different amounts of skewness \cite{Flack:2004oq,Preuschoft:2004ly,Waal:1985fk}.  One explanation might be that the distribution of fighting abilities is different from species to species, for instance if there is more or less variation in body size.  However, our model suggests that even if the underlying distribution of fighting abilities are identical, the priorities given to preference, error rate, and waiting costs can affect the distribution of power.  The benefit of having a partner agree to be subordinate, the benefit of reaching the ``correct" agreement, and the costs of fighting may vary from species to species, causing variation in the shape of the power structure.

In a more complicated model, the ``aggressiveness'' of animals moving in a spatially explicit area affected the shape of the dominance hierarchy \cite{Hemelrijk:2011fk}.  While the mechanisms of that model are quite different, ``aggressiveness'' could be argued to be similar to disregarding fighting costs and prioritizing preference.  Neither the authors of that work nor we found that a difference in the underlying distribution of fighting abilities is required to change the distribution of dominance or power.  In other words, in both models, the collective computation that results in a power structure depends more on how the individuals interact and collect information than on the distribution of fighting abilities.

It is difficult to assess whether decisions in empirical systems are being made accurately, but it is easier to measure how long they take.  In a group, each pair of individuals experiences the difference in their abilities, but their absolute positions within the group also matter.  In data from a captive population of macaques, we observe that pairs of animals with high abilities reach agreements about their relative dominance but that pairs of animals with low abilities sometimes do not reach such agreements (JC Flack, unpublished data).  In our model, decisions between pairs who have similar and high abilities always  take as long or longer than decisions between any other pairs.  This discrepancy indicates that our simple model is missing a level of social complexity that could explain this empirical pattern.

There are (at least) two modifications to the model that we expect would produce results closer to the empirical observations.  First, in the model, the rates at which animals fight each other do not depend on the abilities of the animals.  However, it is likely that animals with lower abilities are less mobile and encounter their peers at a lower rate than animals with higher abilities who are more free to range throughout the group.  In the model, this heterogeneity in encounter rate would make the decisions between animals with low encounter rates take longer.  Second, in the model, all animals value preference, error rate, and waiting costs equally.  However, it is likely that the benefits from receiving a subordination signal from a partner and the costs of fighting depend on both animals' abilities.  If animals with low ability  do not value an agreement with other animals with similar abilities, there would be less incentive for them to reach an agreement.  These types of heterogeneity would make our model more biologically realistic when applied to the social system, but without them it is general enough to be applied to a number of systems in which collective computations are being performed.

We consider two systems, a neural and a social, and our model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two individuals are equal.  The similarities between these three frameworks-- neural decision making, social decision making, and game theory-- suggest that there are common principles of decision making through conflict that are applicable to a number of systems. Conflict and strategy are not usually ascribed to the neural model, and the war of attrition is not usually thought of as a process through which the pair of competitors reaches a decision about an external truth.  Our work is a first step toward unifying these different frameworks and better understanding the importance of conflict and strategy in collective computation.
%%other applications?

\begin{figure}[ht]
\includegraphics[]{pairwise_accuracy.pdf}
\includegraphics[]{nasheq_thresholds.pdf}
\caption{\label{nasheq} The accuracy of a pair using Nash equilibrium thresholds decreases as the weight given to decision time increases.  (a) In the upper panel, the color indicates the accuracy of a decision made by a pair making a difficult decision ($c=.55$) using Nash equilibrium thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.  In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$). Accuracy decreases as $w_2$ increases but is not greatly affected by a tradeoff between $w_1$ and $w_3$. In (b)-(g) we show how the Nash equilibrium thresholds depend on the difficulty of the decision ($c$). The optimization weights for each panel are indicated in the simplex with the corresponding letter.  }
\end{figure}

\begin{figure}[ht]
\includegraphics[width=\textwidth]{groupaccuracy_heatmap.pdf}
\includegraphics[width=\textwidth]{groupeq_thresholds.pdf}
\caption{\label{groupeq}   The accuracy of a group using optimal thresholds decreases as the weight given to decision time increases.  In the upper panels, the color indicates the average accuracy of the decisions between individuals in the bottom quartile (a) and the average accuracy of the whole group (b) using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.  In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$). Both bottom-quartile and total accuracy decrease as $w_2$ increases, and the increase in bottom-quartile accuracy as $w_3$ increases drives a slight increase in total accuracy.  In (c)-(h) we show how the group-context optimal thresholds (top row) and decision time for each pair (bottom row) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest. The optimization weights for each panel are indicated in the simplex with the corresponding letter.  For low waiting costs (c and d), it takes longest for pairs with similar values to reach a decision.  For slightly higher waiting costs (f and g), it takes longer for pairs with similar and high values to reach a decision.}

\end{figure}

\section{Figures}
\begin{figure}[ht]
\includegraphics[width=\textwidth]{mostinformative_heatmap.pdf}
\caption{\label{bestmetric} The best method to use for measuring consensus in the decision network depends on how accurate decisions are on average.  When decision are very accurate or very inaccurate, measures of the number of decisions received are most informative.  When accuracy is slightly lower, eigenvector centrality of the decision network is most informative.  When accuracy is slightly lower again, the number of signalers is most informative.  When accuracy is lower still, the entropy of decisions received is most informative.  In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$).}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=\textwidth]{skewness_heatmap.pdf}
\caption{\label{skewness} The average skewness of the distribution of the number of signalers is maximized at intermediate waiting costs. In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$).}
\end{figure}

\begin{table}[ht]
\centering
\caption{\label{variables}{\bf  Variables and interpretations in neural and social systems.} }
\ra{1.3}
\begin{tabular}{@{}lllll@{}}
Variable & Interpretation & Neural &   Social \\
\cmidrule{1-4} 
& input & dots moving left or right & fights won or lost
\\$c$ & strength of input & coherence of dots & prob. of stronger animal winning
\\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
\\ $w_1$ & error rate weight & reward from being ``right" & reward from being ``right"
\\ $w_2$ & decision time weight & penalty for taking a long time & costs of fighting
\\ $w_3$ & prob. of preference weight & ? & benefit from receiving signal
\\ & decision & subject indicates left or right & one animal emits subordination signal
\\ & correct decision & subject chooses correct direction & weaker animal signals
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{\label{differences}{\bf  Comparison of models applied to neural and social systems.} Differences are highlighted in red.}
\ra{1.3}
\begin{tabular}{@{}lllll@{}}
& \multicolumn{2}{c}{Neural} &  Social \\
\cmidrule{2-3} \cmidrule{4-4} 
dimensionality & $1$  && $2$
\\decision & difference hits a threshold  && \fcolorbox{red}{white}{one var. hits a threshold}
\\ optimality criterion &  reward from being ``right" && reward from being ``right"
\\ & decision time && decision time
\\ & && \fcolorbox{red}{white}{reward from  receiving signal}
\\optimization depends on & input strength && input strength
\\ & noise && noise
\\ & leak rate && leak rate
\\ & && \fcolorbox{red}{white}{other animal's threshold}
\end{tabular}
\end{table}

\pagebreak
\section{Appendix}

\subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}

\pagebreak
\nocite{*}
\bibliographystyle{plain}
\bibliography{signaling_model}

\end{document}

### 2014-10-22 ms_draft_10_10_14.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 ms_draft_07_22.tex
@@
 \begin{center}
-{\bf \LARGE{Conflict and Strategy in Collective Computation (???)}}
+{\bf \LARGE{Conflict and Strategy in Collective Computation}}
 \end{center}
@@
 \section{Abstract}
-In neural and social systems, an individual's position in an interaction network can be informative about its function within the group.
-The development of the network and scores given by each node's position is a collective computation whose output is available to group members.
-For example, in social groups, power is defined as the degree of consensus in the network about each node's value.
-We use a leaky integrator model for a noisy decision process that gives the direction and magnitude of each edge in the network.
+Groups of individuals making decisions in noisy environments perform collective computations in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys.
+In each of these systems, pairs of individuals make decisions about their relative values.
+We use a leaky integrator model, a set of stochastic differential equations describing how decision variables change over time, to model the decision process.
+The decisions made between each pair form a directed network over the whole group.
+The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values.
 We find that differences in the model when it is applied to neural and social systems do not affect the accuracy of decisions.
-We find that a group of individuals prioritizing personal preference over accuracy make more accurate decisions.
-We find that when pairwise decisions are accurate, local properties of the network are the most informative, but when accuracy decreases global properties are more informative.
-Finally, we find that the shape of the distribution of power is determined by waiting costs of those decisions.
-Our model can be applied to a number of systems and our results suggest that conflict and strategy are important in the collective computations performed by all of them.
+We find that conflicts of interest in the decision process affect individuals with low values the most and that conflicts of interest can incentivize these individuals to make more accurate decisions.
+We find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative.
+Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.
 
@@
 \section{Introduction} 
+Many biological systems perform collective computation.
+When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether the dots are moving left or right, even if the difference in the sizes of the groups moving either direction is not very large.
+A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules only when the density is sufficiently high.
+Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion.
+In social groups of primates, the distribution of the degree of consensus about each monkeyÕs ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave.
+In each of these examples, the individual members of the group make decisions based on noisy environmental information and a function of or algorithm acting on those individual decisions determines the groupÕs output.
+This two-step process allows the group to produce a successful output or behavior, despite computational limitations of its members and the conflicts of interest between them.
+While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them.
 
-When individuals come together in groups, their computational abilities can be combined in nonlinear ways, giving rise to collective computations that are better than would be expected from the individuals' abilities (ref XXX).
-This collective computation cannot be decomposed into parts that are caused or performed by individuals.
-In schools and flocks, the group can sense its environment and navigate more accurately than any of its individual members can (ref XXX).
-Similarly, a group of quorum sensing bacteria can change their behavior, for example color or phosphorescence, based on the properties of the environment that no individual cell would be able to detect (ref XXX).
-The brain performs computations and makes decisions of which no individual neuron would be capable (ref XXX).
-Social groups of animals create social structures, which provide information about their members to which no individual would otherwise have access \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.
-As external observers to these systems, the integration of individual behaviors into a performance that is much greater than the sum of its parts can seem an amazing feat.
-To the members of the systems, for example birds in a flock or monkeys in a social group, the group level output can be a new source of information and can be used to make decisions about how to behave.
+ The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment in the sense that for a given amount of time the SPRT is the most accurate algorithm and for a given error rate the SPRT reaches a decision most quickly.
+The algorithm works by keeping track of the amount of evidence supporting one alternative and deciding in favor of that alternative once that amount is sufficiently high or deciding against it once that amount is sufficiently low.
+While this algorithm assumes the decider has perfect memory, in reality this is rarely the case and a simple way to model memory loss is to assume that the amount of evidence decays back toward 0 at teach point in time.
+This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time.
+The random walk described by the SDEs is biased by the true state of the environment.
+This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which he has gathered more support.
+The leaky integrator model, while a powerful for modeling noisy learning dynamics, has been applied rather phenomenologically, and we seek here to justify its use with a mechanistic derivation.
 
-Some of these collective computations rely on an interaction network between the individual members of the group.
-There are many measurements of position in a network that assign a value to each node, for example a node's in-degree or centrality within the network.
-Any of these can be thought of as a collective computation, a group level output that depends on the structure of the network and whose dependence on any individual behavior is difficult to describe.
-If each individual in the group has a functional role or value, as well as a position in the interaction network, then a useful collective computation would give individuals scores that are informative about these underlying values.
+In this paper, we will study a group of individuals, each pair of which is trying to make a decision about their relative values.
+Each individual will have a decision variable indicating how much evidence it has accumulated and a threshold.
+If the decision variable reaches the threshold, the individual will signal that a decision has been made.
+Both individuals involved in the decision value making a correct and fast decision.
+Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed.
+Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher value.
+This introduces a conflict of interest.
+The thresholds are the strategies they use to negotiate accuracy, speed, and their personal preferences.
 
-An animal's power is defined as the degree of consensus in its group about its fighting ability and the distribution of these power scores is referred to as the power structure \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq,Preuschoft:2004ly}.
+The leaky integrator model has been applied to neural decision-making, and it successfully explains and predict performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
+In this application, each of two neural populations responds to evidence for either left or right motion.
+The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high.
+Usually, the visual stimulus driving this decision process is a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.
+The bias in the SDEs is in favor of the direction in which the coherent dots are moving.
+However, this process could be used by neural populations deciding between any two possibilities, for example between up and down motion or between two colors, and a group of neural populations could be used to distinguish between or rank a whole set of possibilities.
+
+While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been.
+In this application, each of two animals has a belief about its dominance with respect to the other which increases or decreases as the animal wins or loses fights against its opponent.
+The SDEs describe how these beliefs change over time.
+The source of evidence is the fight history between the two animals.
+Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise.
+The bias in the SDEs is in favor of the animal that is more likely to win.
+Once an animals' opinion goes below a threshold, it decides to agree to be subordinate to the other.
+A correct decision is one in which the weaker individual makes this agreement.
+In pigtailed macaques ({\it Macaca nemestrina}) and other species of macaques, there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.
+This signal marks the end of the decision process.
+
+The leaky integrator model can thus be used to describe a decision being made between each pair of individuals in a group.
+These decisions form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects.
+In many systems, the degree of consensus in a group about each individuals' state is informative about the individuals'  true states and affects the individuals' functional behavior \cite{Brush:2013fk}.
+In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functional behaviors \cite{Brush:2013fk,Flack:2006uq}.
+The output of one of these consensus measures being applied to a decision network would be a set of consensus scores, a collective computation which would be informative about the individuals' true states.
+
+In the social case, this consensus is referred to as social power.
 In macaques, power is a useful variable since it reflects the animals' true abilities and the animals themselves do use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.
-In pigtailed macaques, the animals have a stereotyped signal, a formal subordination signal, that one animal will emit to another to communicate that it agrees to be subordinate \cite{Flack:2004oq,Flack:2006uq,Flack:2007kx,Preuschoft:2004ly}.
-Since a subordination signal indicates one animal believes another to have a higher fighting ability, a network of subordination signals can be used to determine the degree of consensus in the whole group about any individual's ability.
-In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node \cite{Brush:2013fk,Flack:2006uq}.
-We found a few that produced scores that were predictive of individuals' functional behaviors, which suggests they provide information about the animals' true fighting abilities.
+The distribution of power scores in an empirical group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions in the social group than the rest of its members \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.
+Further, the shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms.
 This kind of consensus may be important in neural systems, as well.
 A number of neural populations may respond to or encode different environmental stimuli or variables.
-The consensus in a network of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.
+The consensus in a group of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.
 
-In each of these systems, the individuals-- monkeys or neural populations-- are trying to learn and make decisions in a noisy environment.
-In such a situation, an individual acquires evidence that may or may not reflect the true state of the environment and only makes a decision once it has acquired enough evidence to reduce its uncertainty sufficiently.
-If the environment can be in one of the two possible states, the optimal way to decide between these alternatives is an algorithm called a sequential probability ratio test (SPRT) (ref XXX).
-An individual using this learning method gathers a sequence of pieces of evidence that he uses to update his belief about the state of the world until this belief is strong enough to make a decision.
-This updating process can equivalently be written as a set of stochastic differential equations (ref XXX), which are more flexible in allowing for memory loss.
-This type of model, sometimes referred to as a leaky integrator model, is applicable to any system in which an individual gathers evidence and chooses the alternative for which he has gathered more support.
-
-Noisy learning models can be used to model animal conflicts, and the SPRT has been previously used to describe animal conflicts over resources \citep{Froment:2010fk}.
-In a group of macaques, animals have different fighting abilities, and when a pair of animals engages in a fight, one animal is more likely on average to win.
-Over time, as one animal learns that it is likely to lose fights, it will decide to emit the formal subordination signal, communicating its agreeing to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk}.
-The outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other factors.
-Because of this stochasticity, the number of fights won and lost in any period of time is a random variable.
-Each animal's opinion about its relative dominance, therefore, will perform a biased random walk until one animal's opinion of itself is low enough to agree to being subordinate, and this process can be describe by a leaky integrator model.
-
-Leaky integrator models have been used to describe decision making in both monkeys and humans \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
-Often an experimental subject is presented with a visual stimulus that consists of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.
-The subject is then asked to decide the direction in which the dots are moving.
-If there are two neural populations, each of which fires in response to a particular direction of motion, one firing rate will tend to increase and the other will tend to decrease, depending on the direction of the stimulus.
-The random dots make the stimulus noisy and sometimes the firing rates will change in the wrong direction.
-Eventually, either the experimenter forces a decision to be made, in which case whichever population has a higher firing rate determines which decision is made, or the firing rate of one population becomes high enough that the subject decides for that direction.
-In the model, the two decision variables--- the firing rates of two neural populations responding to left and right motion--- perform a biased random walk until enough evidence has been accumulated to make a decision one way or the other.
-This model has successfully explained and predicted performance in a number of experiments \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
-
-The use of stochastic differential equations to model noisy learning has been rather phenomenological.
-Here we turn a simple mechanistic description of the accumulation of information into a set of stochastic differential equations, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}.
+There are thus three factors affecting the collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.
+In this paper, first we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}.
 We then use this model to answer the following questions.
-Are there differences between decision making in the neural and social systems?
-What method of measuring position in an interaction network is the most informative about individuals' true values?
-How should individuals make decisions such that the resulting collective computation is most accurate?
-How is the shape of the power structure affected by how individuals make decisions?
+Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process?
+What threshold strategies should individuals with conflicting interests use to make decisions in a group context?
+How do the individualsÕ threshold strategies affect the collective computation?
+What function of the decision network produces the most informative collective computation?
 How well does our model describe empirical social systems?
@@
 
+As show in the Appendix, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model.
+There are analytical solutions to these equations when the system of SDEs is reduced to one dimension.
+However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.
+
 \subsection{Optimal thresholds }
@@
 (In the social system, $i$ sends a subordination signal to $j$ to communicate his agreement to be subordinate.) To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be sent, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.
-The decision network is then formed such that the edge from $i$ to $j$ is the number of signals sent from $i$ to $j$.
-For many random sets of fighting abilities and for each consensus formalism, we can find the mutual information between the consensus scores and the true values and  the average skewness of the set of consensus scores.
+At each point in time $t$, the weight of the edge from $i$ to $j$ is $0$ if the expected decision time is greater than $t$ (i.e.
+a decision has not been made) and is given by the number of signals sent from $i$ to $j$ otherwise.
+Using many random sets of fighting abilities, for each consensus formalism and each point in time $t$, we can find the mutual information between the consensus scores from the network at $t$ and the true values.
+We can also find the average skewness of the set of consensus scores.
 
@@
 \subsection{Decision making is as accurate without a third party arbiter. }
-We compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
+First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
 To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e.
@@
 That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability.
-Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.
+Because they perform equally well and because the two-dimensional system is more appropriate for the social system (and perhaps the neural system as well), we use the two-dimensional system in the rest of our analyses.
+While the results come from an analysis of the two-dimensional system, they should also apply to the reduced system.
+%Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.
 
-\subsection{The average accuracy in a group of stubborn individuals is higher than in a group of accurate individuals. }
-The best measure of consensus depends on the accuracy of the decisions, so we address accuracy before moving on to discuss consensus.
+\subsection{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
 The Nash equilibrium thresholds for a pair of individuals depend on whether the individuals are minimizing error rate, maximizing probability of their preference, or minimizing decision time.
-The pair can make the most accurate decision if the weaker individual chooses a low threshold and the strong individual chooses a high one, so if only error rate matters ($w_1=1$, $w_2=0$, $w_3=0$), those will be the Nash  thresholds.
-If only preference matters ($w_1=0$, $w_2=0$, $w_3=1$), there is no incentive for the individuals to stop accumulating evidence (fighting, in the social case) so both individuals wait until the other gives up and the Nash  thresholds are both high.
-If only decision time matters ($w_1=0$, $w_2=1$, $w_3=1$), both individuals have an incentive to set their thresholds low and reach a decision quickly, whatever that decision may be (Figure \ref{nasheq}).
+If accuracy of the decision is the only thing that matters ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible since if the decision is correct the weaker animal gives up first.
+When personal preference matters ($w_3>0)$ and there is a conflict of interest between the individuals, both will set their threshold high since there is no incentive for the individuals to stop accumulating evidence.
+As the importance of decision time increases (i.e.
+waiting costs increase), both individuals will lower their thresholds in order to reach the decision more quickly, whatever that decision may be (Figure \ref{nasheq}).
+Notice that decision time and personal preference lead both animals to have the same thresholds whereas the individuals have different thresholds when accuracy is important.
+The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and decreases as either decision time or personal preference become important (Figure \ref{nasheq}).
 
-The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and lowest when only decision time matters ($w_2=1$).
-When waiting costs are low, accuracy decreases as the weight given to preference increases (Figure \ref{nasheq}).
-It makes sense that as the individuals prioritize preference over error rate, they make increasingly inaccurate decisions.
-For sufficiently high waiting costs, however, the Nash thresholds do not depend on the tradeoff between error rate and preference, so neither does the accuracy for a pair using Nash thresholds.
-In other words, once waiting costs are high, a pair of individuals minimizing error rate and a pair each of whom maximizes the probability of reaching his preference will reach a decision with the same accuracy.
+In a group, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.
+As in a pair, prioritizing accuracy tends to lead animal's at the top of the group to set high thresholds and those at the bottom to set low thresholds.
+Prioritizing personal preferences, on the other hand, tends to lead all individuals to set high thresholds (Figure \ref{groupeq}).
+The difference in an individual's threshold between when accuracy is important and when personal preferences are important is largest for the individuals at the bottom of the group.
+Individuals towards the top will set high thresholds in either case.
+Individuals toward the bottom, however, will set lower thresholds when they prioritize accuracy then when they prioritize personal preferences.
+This is because choosing a higher threshold would increase waiting time in decisions with respect to all individuals and decrease the accuracy of its decisions with respect to higher animals, and it would only slightly increase the accuracy of its decisions with respect to other low individuals.
+In other words, individuals at the bottom of the group sacrifice some accuracy in their decisions with respect to each other in order to decrease their overall decision times.
 
-Prioritizing preference at the expense of error rate has a different effect as soon as the individuals are in a group with more than two members.
-In a group, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.
-If the individuals value error rate, as soon as any waiting costs are introduced individuals at the bottom of the group lower their thresholds because in so doing, they make much faster decisions without losing much accuracy.
-On the other hand, if the individuals value preference and waiting costs are introduced, individuals at the bottom of the group will keep their thresholds higher because they can increase their chances of receiving a decision, even though it costs them time (Figure \ref{groupeq}).
+This leads to the slightly counterintuitive result that making personal preferences less important (i.e.
+having a larger conflict of interest) and making accuracy more important actually decreases the average accuracy of the decisions made between individuals at the bottom of the group.
+This in turn leads to a very slight increase in the average accuracy of all pairs in the group (Figure \ref{groupeq}).
+For every measure of consensus, the mutual information between the consensus scores and the true values increases with average pairwise accuracy.
+Thus, the information content of the consensus scores is also slightly higher when the individuals prioritize personal preference rather than error rate.
 
-When individuals in the bottom of the group have slightly higher thresholds and wait slightly longer to make a decision, the accuracy of decisions between individuals both in the bottom of the group increases.
-Even though the accuracy between two individuals in the top of the group is not strongly affected by this tradeoff and the accuracy between an individual in the bottom and an individual in the top decreases slightly, the net effect is that the average accuracy of all pairs in the group increases (Figure \ref{groupeq}).
-For every measure of consensus, the mutual information between the consensus scores and the true values increases with average pairwise accuracy.
-Thus, the information content of the consensus scores is also slightly higher when the individuals minimizes preference rather than error rate.
-
-\subsection{Global network structure provides the most information when accuracy is high but imperfect. }
+\subsection{Global network structure provides the most information when accuracy is high but imperfect and when the network is developing. }
+First, we consider the mutual information between the consensus formalisms applied to a fully formed decision network and the true values.
+It is interesting to note that for each of the four types of formalisms we consider, there are regions of the simplex given by the optimization weights that make that the most informative formalism.
 When pairwise decisions are on average quite accurate, the number of decisions received is the most informative about the underlying values (Figure \ref{bestmetric}).
@@
 
+The differences in mutual information given by each formalism are very small, but the clear pattern that the best formalism depends on waiting costs needs explanation.
 The number of signalers consistently overestimates individuals' value.
@@
 When accuracy is lower still, none of the measures are very informative, but entropy correctly gives the lowest individuals low scores, giving it an edge over the other measures.
+
+When the number of signals or number of signalers out-perform the other formalisms, the differences in performance are very small.
+When eigenvector centrality is the best, on the other hand, it clearly outperforms the others (Figure \ref{bestmetric}).
+Eigenvector centrality is consistently the most informative formalism on a network that is not fully formed (Figure \ref{bestmetric}).
 
@@
 If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes, this result gives a rule of thumb for how to pick a measure of consensus.
+However, since eigenvector centrality always performs optimally or close to optimally on a fully formed network and it performs optimally on a network that is still forming, eigenvector centrality would would generally be a good choice.
+
 As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities.
@@
 
+Individual-based models and computer simulations have previous been used to explain the social structures of macaques \cite{Hemelrijk:2011fk,Hausfater:1981fk}.
 In a more complicated model, the ``aggressiveness'' of animals moving in a spatially explicit area affected the shape of the dominance hierarchy \cite{Hemelrijk:2011fk}.
@@
 \begin{figure}[ht]
-\includegraphics[]{pairwise_accuracy.pdf}
+\includegraphics[]{pairwiseaccuracy.pdf}
 \includegraphics[]{nasheq_thresholds.pdf}
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{groupaccuracy_heatmap.pdf}
+\includegraphics[width=\textwidth]{groupaccuracy.pdf}
 \includegraphics[width=\textwidth]{groupeq_thresholds.pdf}
 \caption{\label{groupeq}   The accuracy of a group using optimal thresholds decreases as the weight given to decision time increases.
-In the upper panels, the color indicates the average accuracy of the decisions between individuals in the bottom quartile (a) and the average accuracy of the whole group (b) using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
+In the upper panels, the color indicates (a) the average accuracy of the whole group and (b) the average accuracy of the decisions between individuals in the bottom quartile using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
@@
 Both bottom-quartile and total accuracy decrease as $w_2$ increases, and the increase in bottom-quartile accuracy as $w_3$ increases drives a slight increase in total accuracy.
-In (c)-(h) we show how the group-context optimal thresholds (top row) and decision time for each pair (bottom row) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest.
+In (c)-(f) we show how the group-context optimal thresholds (left panels) and decision time for each pair (right panels) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest.
 The optimization weights for each panel are indicated in the simplex with the corresponding letter.
-For low waiting costs (c and d), it takes longest for pairs with similar values to reach a decision.
-For slightly higher waiting costs (f and g), it takes longer for pairs with similar and high values to reach a decision.}
+For low waiting costs (c and e), it takes longest for pairs with similar values to reach a decision.
+For slightly higher waiting costs (d and f), it takes longer for pairs with similar and high values to reach a decision.}
 
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{mostinformative_heatmap.pdf}
+%\begin{minipage}{.66\linewidth}
+%\includegraphics[width=\linewidth]{mostinformative.pdf}
+%\end{minipage}
+%\begin{minipage}{.44\linewidth}
+%\includegraphics[width=.66\linewidth]{infogrowth.pdf}
+%\end{minipage}
+\includegraphics[width=\textwidth]{info.pdf}
 \caption{\label{bestmetric} The best method to use for measuring consensus in the decision network depends on how accurate decisions are on average.
-When decision are very accurate or very inaccurate, measures of the number of decisions received are most informative.
-When accuracy is slightly lower, eigenvector centrality of the decision network is most informative.
-When accuracy is slightly lower again, the number of signalers is most informative.
-When accuracy is lower still, the entropy of decisions received is most informative.
+(a) In the left panel, the color indicates the most informative consensus formalisms.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
 In the upper corner, only decision time matters ($w_2=1$).
-In the lower right corner, only preference matters ($w_3=1$).}
+In the lower right corner, only preference matters ($w_3=1$).
+In (b)-(d) we show how the mutual information of each consensus formalisms increases over time.
+The optimization weights for each panels are indicated in the simplex with the corresponding letter.}
 \end{figure}
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{skewness_heatmap.pdf}
+\includegraphics[width=\textwidth]{skewness.pdf}
+\includegraphics[width=\textwidth]{skewness_histograms.pdf}
 \caption{\label{skewness} The average skewness of the distribution of the number of signalers is maximized at intermediate waiting costs.
+In the upper panel, the color indicates the average skewness of the distribution of a group using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
 In the upper corner, only decision time matters ($w_2=1$).
-In the lower right corner, only preference matters ($w_3=1$).}
+In the lower right corner, only preference matters ($w_3=1$).
+In (b)-(d) we show a histogram for a representative distribution of the number of signalers for a group using the optimization weights as indicated in the simplex with the corresponding letter.}
 \end{figure}
+
+\begin{table}[ht]
+\centering
+\caption{\label{variables}{\bf  Collective computations.} }
+\ra{1.3}
+\begin{tabular}{@{}lllll@{}}
+System & Environmental information & Individual &   Individual decision & Collective computation \\
+\cmidrule{1-5} 
+brain & dots moving left or right & neural population & to stop firing & brain-level decision for left or right
+\\  social group & fights won or lost & monkey & to emit a subordination signal & power structure
+\\  quorum sensing & signaling molecule & bacteria & to produce signaling molecule & toxicity or phosphorescence
+\\ flock movement & environment & bird & where to fly and how quickly & flock movement and navigation
+\end{tabular}
+\end{table}
 
@@
 \centering
-\caption{\label{differences}{\bf  Comparison of models applied to neural and social systems.} Differences are highlighted in red.}
+\caption{\label{differences}{\bf  Comparison of model in its original application to neural systems, in an extended application to neural systems, and its application to social systems.} Differences are highlighted in red.}
 \ra{1.3}
 \begin{tabular}{@{}lllll@{}}
-& \multicolumn{2}{c}{Neural} &  Social \\
+&Original neural & & Social \\
 \cmidrule{2-3} \cmidrule{4-4} 

### 2014-10-22 old_notes/eleanor_signaling_model_03_10.tex
ive added code for tracking mutual information over time and changed figures a bit

\documentclass{article}
%\documentclass{book}
%\documentstyle[amssymb,amsmath]{article}
%\documentclass[aps,tightenlines,amscd,amsmath,amssymb,verbatim,12pt]{revtex4}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
\usepackage{subfig}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\captionsetup{justification=RaggedRight, singlelinecheck=false}
%  \newcommand{backslash the  command}{what it should do}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf Preliminary ideas and comparison to Karen's work}
\\ Eleanor Brush
\\ March 10, 2012
\end{center}

\vspace{0pt}
\normalsize
\section{Draft of a model}

The model that I'm thinking about right now is the following.  The basic idea is that time is discrete, measured in when fighting occurs.  At each point in time (=each time a fight occurs), each individual gathers evidence about its dominance with respect to its opponent.  At each point in time, the individual's estimate ``leaks" some amount, meant to represent the individuals forgetting the estimate at the last point in time and making a conservative lower estimate.  When the evidence that an individual has about its being subordinate reaches/ passes some threshold $T$, it emits a signal.

\underline{\bf Variables:}
\begin{enumerate}
\item{$x_{ij}(t)=$ the estimate $i$ has of its dominance with respect to $j$}
\item{$e_{ij}(t)=$ the evidence $i$ receives at time $t$ about its dominance wrt to $j$

In the simplest case, this would be the result of a fight between $i$ and $j$, drawn from a normal distribution, $N(a_i-a_j,\sigma^2)$ where $a_i$,$a_j$ are their respective fighting abilities.

In a more complicated model, individual $i$ could also gather evidence about $j$ from $j$'s interactions with other individuals.}
\item{$s_{ij}(t)=$ indicates whether $i$ receives a signal from $j$ at time $t$}
\end{enumerate}

\underline{\bf Difference equations for the variables:}
\begin{align*}
x_{ij}(t+1)&=(1+\lambda)x_{ij}(t)+e_{ij}(t)+\alpha s_{ij}(t)
\\ s_{ij}(t)&=\mathbb{I}_{x_{ji}(t)\leq T< x_{ji}(t-1)}
\end{align*}

\underline{\bf Parameters:}
\begin{enumerate}
\item{$a_i$: $i$'s ``true" fighting ability}
\item{$\sigma^2$: variance or noise in the fight outcomes / evidence gathered by each individual}
\item{$\lambda$: ``leak" or forgetting parameter}
\item{$\alpha$: indicates how informative receiving a signal is compared to winning a fight}
\item{$T$: threshold dictating when an individual signals}
\end{enumerate}

\section{What we can ask / answer by looking at such a model}
There are very similar models out there for different situations.

The leaky accumulator model / drift diffusion model in neuroscience / psychology describes the dynamics of neuron's firing rates over time as a function of the visual stimulus they're exposed to:
\begin{equation} \frac{dx}{dt}=(\lambda x+A(t))dt+cdW_t \label{neuro} \end{equation}
where $A(t)$ is the visual stimulus at time $t$ and $W_t$ is a Weiner process / Brownian motion / white noise.

Similar models have been used to model animals' decision making processes about backing down or continuing to fight in a particular contest given how the fight's been going so far:
\begin{equation} x_{ij}(t+1)=x_{ij}(t)+e_{ij}(t) \label{adrian} \end{equation}
where $e_{ij}\sim N(a_i-a_j,\sigma^2)$, if $a_i,a_j$ are the contestants relative ``resource holding potential" (or ``RHP").

Equations (\ref{neuro}) with $\lambda=0$ is the continuous time version of (\ref{adrian}), which is equivalent to performing a sequential probability ratio test to find the maximum likelihood estimate of $a_i-a_j$.  This can be shown to be an optimal decision-making procedure in the sense that if you set your error threshold at some level, this process will come to a decision in the shortest time and if you set the length of the decision-making process, this process will be the most accurate.  What's new / interesting about the model that I wrote above is that the outcome of the decision process-- the emission of a signal-- feeds back to influence future decisions.

\begin{table}[h] 
	\caption{\label{matrices} Similar models: decision making in neuroscience / psychology, e.g. Newsome et al; animals deciding to back down in a particular contest, e.g. Adrian de Froment's thesis; this model.}
		\begin{tabular}{lllllll} 
			& \vline & Neuro  & \vline & Backing down  & \vline & Here \\
			\hline 
			Variable & \vline & firing rate & \vline & estimate of ``RHP" & \vline & estimate of fighting abilities
			\\Time & \vline & continuous & \vline & discrete & \vline & discrete
			\\Input / Evidence & \vline & visual stimulus & \vline & rams butting heads & \vline & fight outcomes
			\\ Distribution to Learn & \vline & dots' movements & \vline & RHP & \vline & fighting ability
			\\ $\lambda$ & \vline & leak in activity & \vline &  NA & \vline &  forgetting 
			\\ Decision & \vline & left or right & \vline & back down or keep fighting & \vline & signal or not
			\end{tabular}
\end{table}

\underline{\bf Empirical First Steps:}
\begin{enumerate}
\item{Does this model explain the data at all?}
\item{Can we fit parameters to data?  How?  Perhaps by looking at the ratio of the rates of fighting and signaling at steady state.
\begin{enumerate}
\item{$a_i-a_j$: difference in ``true" fighting ability}
\item{$\sigma^2$: noise of evidence gathering process}
\item{$\lambda$: rate at which estimates are forgotten}
\item{$\alpha$: weight with which signals affect estimates}
\item{$T$: threshold of estimates at which signals are emitted}
\end{enumerate}
}
\end{enumerate}

\underline{\bf Theoretical Questions:}
\begin{enumerate}
\item{When there is ``leak" or forgetting ($\lambda\neq 0$), by having signaling feedback to influence future decision ($\alpha\neq 0$), can individual's learn better / are estimates of fighting abilities more stable / is uncertainty about fighting abilities reduced? }
\item{If there is forgetting ($\lambda\neq 0$), is there a feedback weight ($\alpha$) that optimizes the rate at which individuals learn whether they're subordinate or dominant / minimizes the error with which individuals signal / minimizes the number of fights a pair has to engage in / maximizes a reward coming from signaling correctly?  Is this what we observe in the parameters when we fit them to the data?}
\item{Hypothesis: The stead state ratio of signaling rate to fighting rate is a decreasing function of $\lambda$, (increasing in $-\lambda$) and a decreasing function of $\alpha$.}
\end{enumerate}

\section{Comparison to Karen's work}
This is closely related to the modeling work in the Karen paper, but I think the question and the methods are different.  The big questions here are:

\begin{enumerate}
\item{Can we describe how signaling behavior between a pair of individual develops as a function of the fights that occur between them?  Does a learning model where individuals integrate evidence from successive fights and signal when the integrated evidence passes some threshold provide an adequate description?  Can we fit some of the parameters of the model to data?}
\item{How do forgetting and signaling feedback interact to affect how individuals learn whether they're dominant or subordinate?}
\end{enumerate}

In the Karen paper, as here, fighting abilities follow some probability distribution that dictates the outcomes of fights.  With perfect learning and no errors in signaling, this distribution of fight outcomes is translated directly into a distribution of signals emitted.  With errors in signaling or with imperfect learning, a history of fight outcomes becomes a signal when it reaches some threshold.  The question in Karen's work, as I understand it, is to reproduce the observed number of signalers / number of signals distributions by changing the underlying distribution of fighting abilities, whether there are errors in signaling or not, whether there is imperfect learning, and the threshold beyond which an individual emits a signal.

There's a difference in the questions in that Karen's work is looking to reproduce the observed heavy-tailed distribution of power, whereas the above model would be looking at the development of the signaling network.  Instead of making probabilistic statements about fight outcomes and signaling behavior that lead to a distribution of signals received over the group, I would be using a dynamical model to track the relationship between a particular pair over time.  Finally, the above model allows for the signaling behavior to feedback and influence the decision-making process.

\end{document}

### 2014-10-22 old_notes/model_04_19.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/eleanor_signaling_model_03_10.tex
@@
 \\ Eleanor Brush
-\\ March 10, 2012
+\\ April 19, 2012
 \end{center}
@@
 
+\section{Model: SDE framework}
+There are three processes that can cause an animal's estimate of its dominance with respect to another to change:
+\begin{enumerate}
+\item an error can be made so that the animal's estimate randomly changes from one point in time to the next without an external stimulus
+
+\item external observations provide evidence about each animal's dominance, e.g. by fighting each other they gather evidence about their relative strength
+
+\item receiving a signal provides evidence that the receiver is dominant to the signaler
+
+\end{enumerate}
+{\bf Assumption:} For now, we will assume that the two animals have access to the same external observations, which would be the case, for instance, if they only gather evidence from fights that they both are engaged in.
+We can relax this assumption in the future.
+
+Let $X_t^{(i)}$ denote animal $i$'s estimate of its dominance at time $t$.  We can therefore right down equations describing how these estimates change over time:
+\begin{align*}
+X_{t+\tau}^{(1)}=X_t^{(1)}+\sum_{i=0}^{N_{e_1}(\tau)}E_i^{(1)}+\sum_{i=0}^{N_f(\tau)}F_i+b_{s_2}S_2(X_t^{(2)},\tau)
+\\ X_{t+\tau}^{(2)}=X_t^{(2)}+\sum_{i=0}^{N_{e_2}(\tau)}E_i^{(2)}-\sum_{i=0}^{N_f(\tau)}F_i+b_{s_1}S_1(X_t^{(1)},\tau)
+\end{align*} 
+where
+\begin{itemize}
+\item $E_i^{(j)}$ describes the magnitude of an error in estimate, which are identically distributed over time; $N_{e_j}(\tau)$ describes the number of error events in an interval of length $\tau$; and therefore $\sum_{i=0}^{N_{e_j}(\tau)}E_i^{(j)}$ gives the total magnitude of errors in an interval of length $\tau$
+
+\item $F_i$ describes the magnitude of external evidence, which are identically distributed over time; $N_f(\tau)$ describes the number of external observations in an interval of length $\tau$; and therefore $\sum_{i=0}^{N_f(\tau)}F_i$ gives the total magnitude of external evidence in an interval of length $\tau$ (and the evidence gathered changes the animal's estimate in opposite directions, i.e.
+if a fight is won, one animal's estimate is increased while the other's is decreased by the same amount)
+
+\item $S_j(X_t^{(j)},\tau)$ gives the number of signals emitted by individual $j$ in an interval of length $\tau$ and $b_{s_j}$ describes the boost that each of those signals would give to animal $j+1'$s estimate of its dominance
+
+\end{itemize}
+The fact that the two animals are privy to the same external observations is captured by the identity of the second sums in the above equations.
+{\bf Assumption:} The above equations assume that $\tau$ is small enough so that $X_t$ does not change significantly enough in the interval $[t,t+\tau)$ that signaling rates would change.
+(Error and fighting rates do not depend on the estimates, another assumption.)
+
+It may seem strange that evidence from errors and fights are summed over several events in the interval $[t,t+\tau)$ while the evidence from signals is just a number of events times a scaling constant.
+We could write the evidence from signals as a sum over time, but this would be equivalent to our formulation above since we're assuming the strength of a boost from receiving a signal is constant over time.
+
+Now we can give the distributions from which these events are drawn:
+\begin{itemize}
+\item $N_{e_j}(\tau)\sim \scr{P}(r_{e_j},\tau)$
+\item $E_i^{(j)}\sim \scr{N}(\mu_{e_j},\sigma^2_{e_j}) \text{ , i.i.d.}$
+\item $N_f(\tau)\sim\scr{P}(r_f,\tau)$
+\item $F_i\sim\scr{N}(\mu_f,\sigma^2_f)$
+\item $S_j(X_t^{(j)},\tau)\sim\scr{P}(f(X_t^{(j)}),\tau)$
+\end{itemize}
+where $\scr{P}(\lambda,\tau)$ denotes a Poisson process with rate $\lambda$ and $\scr{N}(\mu,\sigma^2)$ denotes a Normal distribution with mean $\mu$ and $\sigma^2$.
+$f(x)$ gives the rate (probability) of emitting a signal as a function of the estimate $x$, which is decreasing in $x$.
+
+We now make two approximations:
+\begin{enumerate}
+\item If $Y_i$, i.i.d.  are drawn from some distribution and $N(\tau)\sim\scr{P}(\lambda,\tau)$, then $Z(\tau)=\sum_{i=0}^{N(\tau)}Y_i$ is a compound Poisson process with 
+\begin{align*}
+\E[Z(\tau)]&=\E[N(\tau)]\E[Y]=\lambda\tau\E[Y]
+\\ \text{ and } Var(Z(\tau))&=\E[N(\tau)]\E[Y^2]=\lambda\tau\E[Y^2]
+\end{align*}
+For computational convenience, we will approximate the distribution of $Z(\tau)$ by $\scr{N}(\lambda\tau\E[Y],\lambda\tau\E[Y^2])$. (HOW VALID IS THIS?)
+
+\item If $\tau$ is big enough that enough events happen, then we can approximate the Poisson process $\scr{P}(\lambda,\tau)$ with $\scr{N}(\lambda\tau,\lambda\tau)$.
+
+\end{enumerate}
+
+This allows us to rewrite our equations for $X_t$ as:
+
+\begin{align*}
+X_{t+\tau}^{(1)}&=X_t^{(1)}+Y_{e_1}(\tau)+Y_f(\tau)+b_{s_2}Y_{s_2}(X_t^{(2)}\tau)
+\\ X_{t+\tau}^{(2)}&=X_t^{(2)}+Y_{e_2}(\tau)-Y_f(\tau)+b_{s_1}Y_{s_1}(X_t^{(1)}\tau)
+\end{align*}
+where 
+\begin{itemize}
+\item $Y_{e_j}(\tau)\sim\scr{N}(r_{e_j}\tau\mu_{e_j},r_{e_j}\tau(\sigma_{e_j}^2+\mu_{e_j}^2))$
+
+\item $Y_f(\tau)\sim N(r_f\tau\mu_f,r_f\tau(\sigma^2_f+\mu_f^2)$
+
+\item $Y_{s_j}(\tau)\sim\scr{N}(f(X_t^{(j+1)})\tau,f(X_t^{(j+1)})\tau)$
+
+\end{itemize}
+Finally, if we define the following constants:
+\begin{align*}
+m_{e_j}&=r_{e_j}\mu_{e_j},
+\\ n_{e_j}^2&=r_{e_j}(\sigma_{e_j}^2+\mu_{e_j}^2),
+\\ m_f&=r_f\mu_f,
+\\ n_f^2&=r_f(\sigma_f^2+\mu_f^2),
+\end{align*}
+then we get the following equations:
+\begin{align*}
+X_{t+\tau}^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]\tau+n_f\sqrt{\tau}Z^{(1)}+n_{e_1}\sqrt{\tau}Z^{(2)}+b_{s_2}f(X_t^{(2)})\sqrt{\tau}Z^{(3)}
+\\ dX_t^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}f(X_t^{(1)})\sqrt{\tau}Z^{(5)}
+\end{align*}
+where $Z^{(1)},\dots,Z^{(5)}\sim \scr{N}(0,1),$ i.i.d., which are equivalent to the stochastic differential equations:
+\begin{align*}
+dX_t^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}dW_t^{(3)}
+\\ dX_t^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
+\end{align*}
+where $W_t^{(1)},\dots,dW_t^{(5)}$ are independent Brownian motions.
+
+We could make a few simplifying assumptions to reduce the number of parameters.  {\bf Assumptions:}
+\begin{itemize}
+\item $r_{e_1}=r_{e_2}$
+\item $\mu_{e_1}=\mu_{e_2}$, and even simpler would be to have both equal $0$ (expected error is $0$)
+\item $\sigma^2_{e_1}=\sigma^2_{e_2}$
+\item $b_{s_1}=b_{s_2}$ 
+\item and combining the above would give $m_{e_1}=m_{e_2}$ and $n_{e_1}=n_{e_2}$
+\end{itemize}
+
+More generally, consider a system
+$$ dX_t=a(X_t)dt+\sum_{i=1}^5\sigma_i(X_t)dW_t^{(i)}$$
+with $W_t^{(1)},\dots,W_t^{(5)}$ independent Brownian motions.
+If $a,\sigma_1,\dots,\sigma_5:\R^2\to\R^2,$ are ``nice" enough functions, and the random variable $X_t$ has a ``nice" enough density function (i.e.
+$\E[g(X_t)]=\int g(x)p(t,x)dx$ for every bounded measurable function $g$), then according to the Fokker-Planck / Kolmogorov forward equation
+\begin{align*}
+\frac{\partial p}{\partial t}&=-\sum_{i=1}^2\frac{\partial}{\partial x_i}(a^{(i)}(x)p(t,x))+\frac{1}{2}\sum_{i,j=1}^2\sum_{k=1}^5\frac{\partial^2}{\partial x_i\partial x_j}(\sigma_k^{(i)}\sigma_k^{(j)}p(t,x)
+\\ \Rightarrow \frac{\partial p}{\partial t}&=-(m_{e_1}+m_f+b_{s_2}f(x_2))\frac{\partial p}{\partial x_1}-(m_{e_2}-m_f+b_{s_1}f(x_1))\frac{\partial p}{\partial x_2}
+\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}+n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
+\end{align*}
+using the functions $a$ and $\sigma_i$ as for our particular model.  Note that this requires the function $f:\R\to\R$ be $C^2$.
 \end{document}

### 2014-10-22 old_notes/model_04_22.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_04_19.tex
@@
 \begin{center}
-{\bf Preliminary ideas and comparison to Karen's work}
+{\bf Development of Signaling Network}
 \\ Eleanor Brush
-\\ April 19, 2012
+\\ April 22, 2012
 \end{center}
@@
 \normalsize
-\section{Draft of a model}
-
-The model that I'm thinking about right now is the following.
-The basic idea is that time is discrete, measured in when fighting occurs.
-At each point in time (=each time a fight occurs), each individual gathers evidence about its dominance with respect to its opponent.
-At each point in time, the individual's estimate ``leaks" some amount, meant to represent the individuals forgetting the estimate at the last point in time and making a conservative lower estimate.
-When the evidence that an individual has about its being subordinate reaches/ passes some threshold $T$, it emits a signal.
-
-\underline{\bf Variables:}
-\begin{enumerate}
-\item{$x_{ij}(t)=$ the estimate $i$ has of its dominance with respect to $j$}
-\item{$e_{ij}(t)=$ the evidence $i$ receives at time $t$ about its dominance wrt to $j$
-
-In the simplest case, this would be the result of a fight between $i$ and $j$, drawn from a normal distribution, $N(a_i-a_j,\sigma^2)$ where $a_i$,$a_j$ are their respective fighting abilities.
-
-In a more complicated model, individual $i$ could also gather evidence about $j$ from $j$'s interactions with other individuals.}
-\item{$s_{ij}(t)=$ indicates whether $i$ receives a signal from $j$ at time $t$}
-\end{enumerate}
-
-\underline{\bf Difference equations for the variables:}
-\begin{align*}
-x_{ij}(t+1)&=(1+\lambda)x_{ij}(t)+e_{ij}(t)+\alpha s_{ij}(t)
-\\ s_{ij}(t)&=\mathbb{I}_{x_{ji}(t)\leq T< x_{ji}(t-1)}
-\end{align*}
-
-\underline{\bf Parameters:}
-\begin{enumerate}
-\item{$a_i$: $i$'s ``true" fighting ability}
-\item{$\sigma^2$: variance or noise in the fight outcomes / evidence gathered by each individual}
-\item{$\lambda$: ``leak" or forgetting parameter}
-\item{$\alpha$: indicates how informative receiving a signal is compared to winning a fight}
-\item{$T$: threshold dictating when an individual signals}
-\end{enumerate}
-
-\section{What we can ask / answer by looking at such a model}
-There are very similar models out there for different situations.
-
-The leaky accumulator model / drift diffusion model in neuroscience / psychology describes the dynamics of neuron's firing rates over time as a function of the visual stimulus they're exposed to:
-\begin{equation} \frac{dx}{dt}=(\lambda x+A(t))dt+cdW_t \label{neuro} \end{equation}
-where $A(t)$ is the visual stimulus at time $t$ and $W_t$ is a Weiner process / Brownian motion / white noise.
-
-Similar models have been used to model animals' decision making processes about backing down or continuing to fight in a particular contest given how the fight's been going so far:
-\begin{equation} x_{ij}(t+1)=x_{ij}(t)+e_{ij}(t) \label{adrian} \end{equation}
-where $e_{ij}\sim N(a_i-a_j,\sigma^2)$, if $a_i,a_j$ are the contestants relative ``resource holding potential" (or ``RHP").
-
-Equations (\ref{neuro}) with $\lambda=0$ is the continuous time version of (\ref{adrian}), which is equivalent to performing a sequential probability ratio test to find the maximum likelihood estimate of $a_i-a_j$.
-This can be shown to be an optimal decision-making procedure in the sense that if you set your error threshold at some level, this process will come to a decision in the shortest time and if you set the length of the decision-making process, this process will be the most accurate.
-What's new / interesting about the model that I wrote above is that the outcome of the decision process-- the emission of a signal-- feeds back to influence future decisions.
-
-\begin{table}[h] 
-	\caption{\label{matrices} Similar models: decision making in neuroscience / psychology, e.g.
-Newsome et al; animals deciding to back down in a particular contest, e.g.
-Adrian de Froment's thesis; this model.}
-		\begin{tabular}{lllllll} 
-			& \vline & Neuro  & \vline & Backing down  & \vline & Here \\
-			\hline 
-			Variable & \vline & firing rate & \vline & estimate of ``RHP" & \vline & estimate of fighting abilities
-			\\Time & \vline & continuous & \vline & discrete & \vline & discrete
-			\\Input / Evidence & \vline & visual stimulus & \vline & rams butting heads & \vline & fight outcomes
-			\\ Distribution to Learn & \vline & dots' movements & \vline & RHP & \vline & fighting ability
-			\\ $\lambda$ & \vline & leak in activity & \vline &  NA & \vline &  forgetting 
-			\\ Decision & \vline & left or right & \vline & back down or keep fighting & \vline & signal or not
-			\end{tabular}
-\end{table}
-
-\underline{\bf Empirical First Steps:}
-\begin{enumerate}
-\item{Does this model explain the data at all?}
-\item{Can we fit parameters to data?  How?  Perhaps by looking at the ratio of the rates of fighting and signaling at steady state.
-\begin{enumerate}
-\item{$a_i-a_j$: difference in ``true" fighting ability}
-\item{$\sigma^2$: noise of evidence gathering process}
-\item{$\lambda$: rate at which estimates are forgotten}
-\item{$\alpha$: weight with which signals affect estimates}
-\item{$T$: threshold of estimates at which signals are emitted}
-\end{enumerate}
-}
-\end{enumerate}
-
-\underline{\bf Theoretical Questions:}
-\begin{enumerate}
-\item{When there is ``leak" or forgetting ($\lambda\neq 0$), by having signaling feedback to influence future decision ($\alpha\neq 0$), can individual's learn better / are estimates of fighting abilities more stable / is uncertainty about fighting abilities reduced?
-}
-\item{If there is forgetting ($\lambda\neq 0$), is there a feedback weight ($\alpha$) that optimizes the rate at which individuals learn whether they're subordinate or dominant / minimizes the error with which individuals signal / minimizes the number of fights a pair has to engage in / maximizes a reward coming from signaling correctly?
-Is this what we observe in the parameters when we fit them to the data?}
-\item{Hypothesis: The stead state ratio of signaling rate to fighting rate is a decreasing function of $\lambda$, (increasing in $-\lambda$) and a decreasing function of $\alpha$.}
-\end{enumerate}
-
-\section{Comparison to Karen's work}
-This is closely related to the modeling work in the Karen paper, but I think the question and the methods are different.  The big questions here are:
-
-\begin{enumerate}
-\item{Can we describe how signaling behavior between a pair of individual develops as a function of the fights that occur between them?
-Does a learning model where individuals integrate evidence from successive fights and signal when the integrated evidence passes some threshold provide an adequate description?
-Can we fit some of the parameters of the model to data?}
-\item{How do forgetting and signaling feedback interact to affect how individuals learn whether they're dominant or subordinate?}
-\end{enumerate}
-
-In the Karen paper, as here, fighting abilities follow some probability distribution that dictates the outcomes of fights.
-With perfect learning and no errors in signaling, this distribution of fight outcomes is translated directly into a distribution of signals emitted.
-With errors in signaling or with imperfect learning, a history of fight outcomes becomes a signal when it reaches some threshold.
-The question in Karen's work, as I understand it, is to reproduce the observed number of signalers / number of signals distributions by changing the underlying distribution of fighting abilities, whether there are errors in signaling or not, whether there is imperfect learning, and the threshold beyond which an individual emits a signal.
-
-There's a difference in the questions in that Karen's work is looking to reproduce the observed heavy-tailed distribution of power, whereas the above model would be looking at the development of the signaling network.
-Instead of making probabilistic statements about fight outcomes and signaling behavior that lead to a distribution of signals received over the group, I would be using a dynamical model to track the relationship between a particular pair over time.
-Finally, the above model allows for the signaling behavior to feedback and influence the decision-making process.
-
 \section{Model: SDE framework}
@@
 \begin{align*}
-dX_t^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}dW_t^{(3)}
-\\ dX_t^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
+dX_t^{(1)}&=\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}dW_t^{(3)}
+\\ dX_t^{(2)}&=\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
 \end{align*}

### 2014-10-22 old_notes/model_05_14.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_04_22.tex
@@
 \\ Eleanor Brush
-\\ April 22, 2012
+\\ May 20, 2012
 \end{center}
@@
 \normalsize
-\section{Model: SDE framework}
+\section{Model as a Stochastic Differential Equation}
+
+\subsection{Derivation of the SDE from microscopic events }
 There are three processes that can cause an animal's estimate of its dominance with respect to another to change:
@@
 \begin{align*}
-X_{t+\tau}^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]\tau+n_f\sqrt{\tau}Z^{(1)}+n_{e_1}\sqrt{\tau}Z^{(2)}+b_{s_2}f(X_t^{(2)})\sqrt{\tau}Z^{(3)}
-\\ dX_t^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}f(X_t^{(1)})\sqrt{\tau}Z^{(5)}
+X_{t+\tau}^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]\tau+n_f\sqrt{\tau}Z^{(1)}+n_{e_1}\sqrt{\tau}Z^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}\sqrt{\tau}Z^{(3)}
+\\ dX_t^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}\sqrt{\tau}Z^{(5)}
 \end{align*}
@@
 
+\subsection{Evolution of probability distribution over time }
 More generally, consider a system
@@
 using the functions $a$ and $\sigma_i$ as for our particular model.  Note that this requires the function $f:\R\to\R$ be $C^2$.
+
+\subsection{Expected Hitting Time }
+(See ``The Theory of Stochastic Processes," Cox and Miller.)  Let $p(x_0,x,y_0,y;t)$ be the probability of ending up at $x,y$ at time $t$ given that the process started at $x_0,y_0$.
+Then $p$ satisfies the backward equation:
+\begin{align}
+-\frac{\partial}{\partial t}p(x_0,x,y_0,y;t)&=(m_{e_1}+m_f+b_{s_2}f(y_0))\frac{\partial}{\partial x_0}p(x_0,x,y_0,y;t)+(m_{e_1}-m_f+b_{s_2}f(x_0))\frac{\partial}{\partial y_0}p(x_0,x,y_0,y;t) \notag
+\\&+\frac{1}{2}(n_f^2+n_{e_1}^2+b_{s_2}^2f(y_0))\frac{\partial^2}{\partial x_0^2}p(x_0,x,y_0,y;t)+n_f^2\frac{\partial^2}{\partial x_0\partial y_0}p(x_0,x,y_0,y;t) \notag
+\\&+\frac{1}{2}(n_f^2+n_{e_2}^2+b_{s_1}^2f(x_0))\frac{\partial^2}{\partial y_0^2}p(x_0,x,y_0,y;t) \label{back}
+\end{align}
+
+We would like to know the expected time until either estimate reaches the signaling threshold $-T$.  Let 
+\begin{equation}
+\tau=\inf\{t:X_t^{(1)}\leq -T\} \label{stopping}
+\end{equation}
+Let $p(x_0,x;t)$ be the probability density that $X_t^{(1)}=x$ given that $X_0^{(1)}=x_0$ and that the $X_s^{(1)}\gneq -T$ for all $s\in[0,t]$.
+Then, assuming $x,x_0>-T$, this modified $p$ also satisfies the backward equation.
+
+Let $G(t|x_0,y_0)=P[\tau\leq t]$ so that 
+\begin{align*}
+1-G(t|x_0,y_0)&=\int_{-\infty}^\infty\int_{-b}^ap(x_0,x,y_0,y;t)dxdy
+\\ \Rightarrow \frac{\partial}{\partial t}G(t|x_0,y_0)&=
+\end{align*}
 \end{document}

### 2014-10-22 old_notes/model_07_04.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_04_22.tex
@@
 \documentclass{article}
-%\documentclass{book}
-%\documentstyle[amssymb,amsmath]{article}
-%\documentclass[aps,tightenlines,amscd,amsmath,amssymb,verbatim,12pt]{revtex4}
-\usepackage{latexsym}
-\usepackage{amssymb,amsmath}
 \usepackage{custom2}
 \usepackage{graphicx} % for figures
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage{epstopdf} % so can use EPS or PDF figures
@@
 \captionsetup{justification=RaggedRight, singlelinecheck=false}
-%  \newcommand{backslash the  command}{what it should do}
 
@@
 \begin{center}
-{\bf Development of Signaling Network}
+{\bf \LARGE{Development of a Signaling Network}}
+\vspace{10pt}
 \\ Eleanor Brush
-\\ April 22, 2012
+\\ July 9, 2012
 \end{center}
+
+%\title{Development of Signaling Network}
+%\author{Eleanor Brush}
+%\maketitle
 
@@
 
-Let $X_t^{(i)}$ denote animal $i$'s estimate of its dominance at time $t$.  We can therefore right down equations describing how these estimates change over time:
+Let $X_t^{(i)}$ denote animal $i$'s estimate of its dominance at time $t$.  We can write down equations describing how these estimates change over time:
 \begin{align*}
@@
 \begin{itemize}
-\item $E_i^{(j)}$ describes the magnitude of an error in estimate, which are identically distributed over time; $N_{e_j}(\tau)$ describes the number of error events in an interval of length $\tau$; and therefore $\sum_{i=0}^{N_{e_j}(\tau)}E_i^{(j)}$ gives the total magnitude of errors in an interval of length $\tau$
+\item $E_i^{(j)}$ describes the magnitude of an error in estimate for animal $j$, which are identically distributed over time; $N_{e_j}(\tau)$ describes the number of error events in an interval of length $\tau$ so that $\sum_{i=0}^{N_{e_j}(\tau)}E_i^{(j)}$ gives the total magnitude of errors in an interval of length $\tau$
 
-\item $F_i$ describes the magnitude of external evidence, which are identically distributed over time; $N_f(\tau)$ describes the number of external observations in an interval of length $\tau$; and therefore $\sum_{i=0}^{N_f(\tau)}F_i$ gives the total magnitude of external evidence in an interval of length $\tau$ (and the evidence gathered changes the animal's estimate in opposite directions, i.e.
+\item $F_i$ describes the magnitude of a piece of external evidence, which are identically distributed over time; $N_f(\tau)$ describes the number of external observations in an interval of length $\tau$ so that $\sum_{i=0}^{N_f(\tau)}F_i$ gives the total magnitude of external evidence in an interval of length $\tau$ (and the evidence gathered changes the animal's estimate in opposite directions, i.e.
 if a fight is won, one animal's estimate is increased while the other's is decreased by the same amount)
@@
 {\bf Assumption:} The above equations assume that $\tau$ is small enough so that $X_t$ does not change significantly enough in the interval $[t,t+\tau)$ that signaling rates would change.
-(Error and fighting rates do not depend on the estimates, another assumption.)
+(Error and fighting rates do not depend on the estimates-- another assumption.)  We could write the evidence from signals as a sum over several events in the interval $[t,t+\tau)$ as with errors and observations, but this is equivalent to the formulation above since we assume the size of a signal and the strength of a boost from receiving a signal is constant over time.
 
-It may seem strange that evidence from errors and fights are summed over several events in the interval $[t,t+\tau)$ while the evidence from signals is just a number of events times a scaling constant.
-We could write the evidence from signals as a sum over time, but this would be equivalent to our formulation above since we're assuming the strength of a boost from receiving a signal is constant over time.
-
-Now we can give the distributions from which these events are drawn:
+We can now describe the distributions from which these events are drawn:
 \begin{itemize}
@@
 \item $N_f(\tau)\sim\scr{P}(r_f,\tau)$
-\item $F_i\sim\scr{N}(\mu_f,\sigma^2_f)$
+\item $F_i\sim\scr{N}(\mu_f,\sigma^2_f) \text{ , i.i.d.}$
 \item $S_j(X_t^{(j)},\tau)\sim\scr{P}(f(X_t^{(j)}),\tau)$
@@
 \begin{enumerate}
-\item If $Y_i$, i.i.d.  are drawn from some distribution and $N(\tau)\sim\scr{P}(\lambda,\tau)$, then $Z(\tau)=\sum_{i=0}^{N(\tau)}Y_i$ is a compound Poisson process with 
+\item If $Y_i$, i.i.d., are drawn from some distribution and $N(\tau)\sim\scr{P}(\lambda,\tau)$, then $Z(\tau)=\sum_{i=0}^{N(\tau)}Y_i$ is a compound Poisson process with 
 \begin{align*}
@@
 \begin{align*}
-X_{t+\tau}^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]\tau+n_f\sqrt{\tau}Z^{(1)}+n_{e_1}\sqrt{\tau}Z^{(2)}+b_{s_2}f(X_t^{(2)})\sqrt{\tau}Z^{(3)}
-\\ dX_t^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}f(X_t^{(1)})\sqrt{\tau}Z^{(5)}
+X_{t+\tau}^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]\tau+n_f\sqrt{\tau}Z^{(1)}+n_{e_1}\sqrt{\tau}Z^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}\sqrt{\tau}Z^{(3)}
+\\ X_{t+\tau}^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}\sqrt{\tau}Z^{(5)}
 \end{align*}
@@
 using the functions $a$ and $\sigma_i$ as for our particular model.  Note that this requires the function $f:\R\to\R$ be $C^2$.
+
+\subsection{Statistics of the Process WITHOUT Signaling Feedback }
+%(See ``The Theory of Stochastic Processes," Cox and Miller.)  Let $p(x_0,x,y_0,y;t)$ be the probability of ending up at $x,y$ at time $t$ given that the process started at $x_0,y_0$.
+Then $p$ satisfies the backward equation:
+%\begin{align}
+%-\frac{\partial}{\partial t}p(x_0,x,y_0,y;t)&=(m_{e_1}+m_f+b_{s_2}f(y_0))\frac{\partial}{\partial x_0}p(x_0,x,y_0,y;t)+(m_{e_1}-m_f+b_{s_2}f(x_0))\frac{\partial}{\partial y_0}p(x_0,x,y_0,y;t) \notag
+%\\&+\frac{1}{2}(n_f^2+n_{e_1}^2+b_{s_2}^2f(y_0))\frac{\partial^2}{\partial x_0^2}p(x_0,x,y_0,y;t)+n_f^2\frac{\partial^2}{\partial x_0\partial y_0}p(x_0,x,y_0,y;t) \notag
+%\\&+\frac{1}{2}(n_f^2+n_{e_2}^2+b_{s_1}^2f(x_0))\frac{\partial^2}{\partial y_0^2}p(x_0,x,y_0,y;t) \label{back}
+%\end{align}
+%
+%We would like to know the expected time until either estimate reaches the signaling threshold $-T$.  
+For simplicity, we start with the case where there is no feedback from signaling to the estimates so that the model becomes
+\begin{align*}
+dX_t^{(1)}&=\left[m_{e_1}+m_f\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}
+\\ dX_t^{(2)}&=\left[m_{e_2}-m_f\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}
+\end{align*}
+
+This is a simple drift-diffusion model.
+We can consider the dynamics of each estimate separately.
+We will show results for $X_t^{(1)}$, which follow similarly for $X_t^{(2)}$.
+As derived in \cite{Bogacz:2006uq}, the expected time until $X_t^{(1)}$ reaches either $-T$ or $T$ is
+\begin{align*}
+\frac{T}{m_e+m_f}\tanh\left(\frac{T(m_e+m_f)}{n_f^2+n_e^2}\right)
+\end{align*}
+and the probability that $X_t^{(1)}$ reaches $-T$ before it reaches $T$ is
+\begin{align*}
+\frac{1}{1+\exp\left(\frac{2T(m_e+m_f)}{n_f^2+n_e^2}\right)}
+\end{align*}
+(assuming $X_t^{(1)}=0$).
+
+\subsection{Ornstein-Uhlenbeck Variant }
+The dynamics are more interesting if the estimates don't just keep increasing or decreasing more or less deterministically.
+To add negative feedback and prevent this from happening, we can make the size of an error proportional to the current estimate:
+\begin{align*}
+dX_t^{(1)}&=\left[m_{e_1}X_t^{(1)}+m_f+b_{s_2}f(X_t^{(2)})\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}dW_t^{(3)}
+\\ dX_t^{(2)}&=\left[m_{e_2}X_t^{(2)}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
+\end{align*}
+The Fokker-Planck equation now becomes
+\begin{align*}
+ \frac{\partial p}{\partial t}&=-(m_{e_1}x_1+m_f+b_{s_2}f(x_2))\frac{\partial p}{\partial x_1}-m_{e_1}p-(m_{e_2}x_2-m_f+b_{s_1}f(x_1))\frac{\partial p}{\partial x_2}-m_{e_2}p
+\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}+n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
+\end{align*}
+If we ignore signaling feedback for the moment and look for a stationary distribution, we want to solve the equation
+\begin{align*}
+0&=(m_{e_1}x_1+m_f)\frac{\partial p}{\partial x_1}-m_{e_1}p-(m_{e_2}x_2-m_f)\frac{\partial p}{\partial x_2}-m_{e_2}p
+\\&+\frac{1}{2}(n_{e_1}^2+n_f^2)\frac{\partial^2 p}{\partial x_1^2}+n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2)\frac{\partial^2 p}{\partial x_2^2}
+\\&=(m_ex+m_f)\frac{\partial p}{\partial x}-(m_ey-m_f)\frac{\partial p}{\partial y}-2m_ep+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial x^2}+n_f^2\frac{\partial p}{\partial x\partial y}+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial y^2}
+\\& \text{ if we assume $m_{e_1}=m_{e_2}$}
+\end{align*}
+
+\section{Simulating the Learning / Signaling Process }
+A Gillespie algorithm allows us to simulate the process as described by the model above (ORNSTEIN-UHLENBECK VARIANT).
+We are interested in the effect of signaling feedback, i.e.
+the $b_s$ parameter, on the dynamics of the model.
+We run the algorithm $1000$ times so that we obtain, at each point in time, a distribution of each animal's estimates.
+We then compare these distributions when there is and isn't signaling.
+In both regimes, each animal's estimates reach a relatively stationary distribution fairly quickly so that we can pick a typical time point at which the distributions are not changing any more.
+(A Kolmogorov-Smirnov test for the differences between distributions at different points in time does not allow us to reject the null hypothesis that the distributions are the same, at a significance level of $p=0.05$.)  In Figure (\ref{diff_dists}), we plot the distribution of each animal's estimates at $t=200$ when there is and isn't signaling feedback ($b_{s_1}=b_{s_2}=0$ and $b_{s_1}=b_{s_2}\gneq 0$).
+We find that the distributions with and without signaling feedback are quite different.
+They are, in fact, statistically distinguishable (K-S test statistic $D=0.68,$ $p<2.2\times 10^{-16}$ for Animal 2 and $D=0.21$, $p<0.05$ for Animal 2).
+In both cases, the most noticeable affect of increasing the signaling feedback is increasing the variance of the distributions (from $1.55$ to $81.44$ for Animal 1 and from $1.23$ to $7.02$ for Animal 2).
+While the mean for Animal 2 is similar in the two cases ($-10.03$ without feedback and $-9.36$ with feedback), the mean for Animal 1 noticeably increases from $10.04$ to $17.25$.
+
+By comparing distributions of the estimates at a given point in time, we compare the behavior of the stochastic process over many realizations.
+We can also compare the behavior of individual trajectories over time.
+We hypothesize that by increasing the strength of signaling feedback, we should increase the stability of signaling behavior and it should be less likely that the dominant animal's estimate sink to such an extent that it becomes subordinate.
+To test this hypothesis, we count, for each run of the simulation, the number of switches between dominance regimes.
+A dominance regime is defined by one animal's estimate being below the signaling threshold $-T$ and the other animal's estimate being above the positive threshold $T$.
+If a dominance regime is established with Animal $i$'s estimate above the threshold (dominant) and Animal $i+1$'s estimate below the threshold (subordinate) and subsequently a dominance regime is established with Animal $i$ subordinate and Animal $i+1$ dominant, we count this as one switch.
+Fixing the other parameters of the model, for each difference in fighting ability $m_f$ and level of signaling feedback $b_{s_1}=b_{s_2}$, we can find the average number of switches that occur in a run of the model.
+This is shown in Figure (\ref{diff_switches}).
+As expected, as the difference in fighting ability increases fewer switches occur since Animal 1 is dominant the vast majority of the time.
+However, what is less expected is that increasing signaling feedback actually decreases stability.
+
+\begin{figure}
+%\begin{subfigure}{.5 \textwidth}
+\begin{center}
+\includegraphics[width=.85\textwidth]{feedback_estimate_distributions_v1.pdf}
+\end{center}
+%\caption{}
+%\end{subfigure}
+\caption{\label{diff_dists} The distribution of estimates for each animal at time $200$, for no signaling feedback and for signaling feedback $b_{s_1}=b_{s_2}=3$.
+Other parameters are set as follows: threshold $T=5$, $m_{e_1}=m_{e_2}=-.5$, $n_{e_1}=n_{e_2}=1$, $m_f=5$, $n_f=.5$.
+Vertical red lines indicate predicted mean for process without feedback, $-m_f/m_{e_1}$ for Animal 1 and $m_f/m_{e_2}$ for Animal 2.
+}
+\end{figure}
+
+\begin{figure}
+\begin{subfigure}{.5 \textwidth}
+\includegraphics[width=1\textwidth]{feedback_stability_of_dominance_v1.pdf}
+\caption{}
+\end{subfigure}
+%\begin{subfigure}{.5 \textwidth}
+%\includegraphics[width=1\textwidth]{feedback_stability_of_dominance_v2.pdf}
+%\caption{}
+%\end{subfigure}
+\caption{\label{diff_switches} The number of switches between dominance regimes as a function of the difference in fighting ability, $m_f$, between the animals and the degree to which receiving a signal affects an animal's estimate of its dominance, $b_{s_1}=b_{s_2}$.
+Lighter colors correspond to a higher number of switches.
+}
+\end{figure}
+
+\nocite{*}
+\bibliographystyle{plain}
+\bibliography{signaling_model}
+
 \end{document}

### 2014-10-22 old_notes/model_07_10.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_07_04.tex
@@
 \\ Eleanor Brush
-\\ July 9, 2012
+\\ July 11, 2012
 \end{center}
@@
 X_{t+\tau}^{(1)}&=X_t^{(1)}+\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]\tau+n_f\sqrt{\tau}Z^{(1)}+n_{e_1}\sqrt{\tau}Z^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}\sqrt{\tau}Z^{(3)}
-\\ X_{t+\tau}^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}\sqrt{\tau}Z^{(5)}
+\\ X_{t+\tau}^{(2)}&=X_t^{(2)}+\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]\tau-n_f\sqrt{\tau}Z^{(1)}+n_{e_2}\sqrt{\tau}Z^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}\sqrt{\tau}Z^{(5)}
 \end{align*}
@@
 dX_t^{(1)}&=\left[m_{e_1}+m_f+b_{s_2}f(X_t^{(2)})\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}dW_t^{(3)}
-\\ dX_t^{(2)}&=\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
+\\ dX_t^{(2)}&=\left[m_{e_2}-m_f+b_{s_1}f(X_t^{(1)})\right]dt-n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
 \end{align*}
@@
 \\ \Rightarrow \frac{\partial p}{\partial t}&=-(m_{e_1}+m_f+b_{s_2}f(x_2))\frac{\partial p}{\partial x_1}-(m_{e_2}-m_f+b_{s_1}f(x_1))\frac{\partial p}{\partial x_2}
-\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}+n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
+\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}-n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
 \end{align*}
@@
 dX_t^{(1)}&=\left[m_{e_1}+m_f\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}
-\\ dX_t^{(2)}&=\left[m_{e_2}-m_f\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}
+\\ dX_t^{(2)}&=\left[m_{e_2}-m_f\right]dt-n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}
 \end{align*}
@@
 
+The stationary distribution of the estimate will be 
+\begin{equation}
+X^{(1)}\sim  \label{ddm_stat}
+\end{equation}
+
 \subsection{Ornstein-Uhlenbeck Variant }
+As shown in (\ref{ddm_stat}), both the mean and variance of the distribution of estimates increase with time for a pure drift-diffusion model.
 The dynamics are more interesting if the estimates don't just keep increasing or decreasing more or less deterministically.
@@
 dX_t^{(1)}&=\left[m_{e_1}X_t^{(1)}+m_f+b_{s_2}f(X_t^{(2)})\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}+b_{s_2}\sqrt{f(X_t^{(2)})}dW_t^{(3)}
-\\ dX_t^{(2)}&=\left[m_{e_2}X_t^{(2)}-m_f+b_{s_1}f(X_t^{(1)})\right]dt+n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
+\\ dX_t^{(2)}&=\left[m_{e_2}X_t^{(2)}-m_f+b_{s_1}f(X_t^{(1)})\right]dt-n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}+b_{s_1}\sqrt{f(X_t^{(1)})}dW_t^{(5)}
 \end{align*}
@@
  \frac{\partial p}{\partial t}&=-(m_{e_1}x_1+m_f+b_{s_2}f(x_2))\frac{\partial p}{\partial x_1}-m_{e_1}p-(m_{e_2}x_2-m_f+b_{s_1}f(x_1))\frac{\partial p}{\partial x_2}-m_{e_2}p
-\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}+n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
+\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}-n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
 \end{align*}
@@
 0&=(m_{e_1}x_1+m_f)\frac{\partial p}{\partial x_1}-m_{e_1}p-(m_{e_2}x_2-m_f)\frac{\partial p}{\partial x_2}-m_{e_2}p
-\\&+\frac{1}{2}(n_{e_1}^2+n_f^2)\frac{\partial^2 p}{\partial x_1^2}+n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2)\frac{\partial^2 p}{\partial x_2^2}
-\\&=(m_ex+m_f)\frac{\partial p}{\partial x}-(m_ey-m_f)\frac{\partial p}{\partial y}-2m_ep+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial x^2}+n_f^2\frac{\partial p}{\partial x\partial y}+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial y^2}
+\\&+\frac{1}{2}(n_{e_1}^2+n_f^2)\frac{\partial^2 p}{\partial x_1^2}-n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2)\frac{\partial^2 p}{\partial x_2^2}
+\\&=(m_ex+m_f)\frac{\partial p}{\partial x}-(m_ey-m_f)\frac{\partial p}{\partial y}-2m_ep+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial x^2}-n_f^2\frac{\partial p}{\partial x\partial y}+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial y^2}
 \\& \text{ if we assume $m_{e_1}=m_{e_2}$}
 \end{align*}
+We can find a stationary distribution for each estimate individually:
+\begin{align}
+X^{(1)}& \sim N\left(-\frac{m_f}{m_e},\frac{n_f^2+n_e^2}{-2m_e}\right)\ \label{ou_stat1}
+\\ \text{ and } X^{(2)} &\sim N\left(\frac{m_f}{m_e},\frac{n_f^2+n_e^2}{-2m_e}\right) \label{ou_stat2}
+\end{align}
+and the covariance between the two estimates at steady state is
+\begin{align*}
+Cov(X^{(1)},X^{(2)})=\frac{-n_f^2}{-2m_e}.
+\end{align*}
 
-\section{Simulating the Learning / Signaling Process }
-A Gillespie algorithm allows us to simulate the process as described by the model above (ORNSTEIN-UHLENBECK VARIANT).
+\section{Analyzing the Model without Stochasticity }
+As a first pass, we can ignore the Brownian motion / the noise and treat the model as a dynamical system:
+
+\begin{align*}
+\frac{dx}{dt}&=m_ex+m_f+bf(y)
+\\ \frac{dy}{dt}&=m_ey-m_f+bf(x)
+\end{align*}
+If $b=0$, there is an equilibrium at $(x,y)=\left(-\frac{m_f}{m_e},\frac{m_f}{m_e}\right).$  If $b\neq 0$, we can find equilibrium by solving $\frac{dx}{dt}=\frac{dy}{dt}=0$.
+We assume that $f(x)=1$ for $x<-T$ and $f(x)=0$ for $x>-T$.
+\begin{enumerate}
+\item $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{-m_f}{m_e}>-T$ and $\frac{m_f}{m_e}>-T$
+\item $\left(\frac{-m_f}{m_e},\frac{m_f-b}{m_e}\right)$ if $\frac{-m_f}{m_e}<-T$ and $\frac{m_f-b}{m_e}>-T$
+\item $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{-m_f-b}{m_e}>-T$ and $\frac{m_f}{m_e}<-T$
+\item $\left(\frac{-m_f-b}{m_e},\frac{m_f-b}{m_e}\right)$ if $\frac{-m_f-b}{m_e}<-T$ and $\frac{m_f-b}{m_e}<-T$
+\end{enumerate}
+Errors only provide negative feedback and prevent escalation when $m_e\leq 0$.
+In this case, equilibria $2.$ and $4.$ are impossible since $-m_f,-m_f-b<0$ $\Rightarrow$ $\frac{-m_f}{m_e},\frac{-m_f-b}{m_e}>0$ while $-T<0$.
+Therefore, no equilibria where Animal 1 is signaling is possible.
+So we are left with
+\begin{enumerate}
+\item $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{m_f}{m_e}>-T$
+\item $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{m_f}{m_e}<-T$
+\end{enumerate}
+Note that if $b=0$, the equilibria are identical.
+Equilibrium $1.$ is only possible if $\frac{m_f}{m_e}>-T$ while $2.$ is only possible if $\frac{m_f}{m_e}<-T$ so that the relative strength of the effects from fights and errors determines which equilibria is possible.
+If $m_f$ is small, an equilibrium with neither animal signaling is possible, whereas is $m_f$ is sufficiently large at equilibrium Animal 2 will signal Animal 1 and Animal 1's estimate is boosted by $\frac{b}{-m_e}$.
+In either case, Animal 2's equilibrium estimate is the same.
+
+The Jacobian of this system at an equilibrium is
+$$
+J=\left[
+\begin{array}{ccccc}
+m_e & bf'(y^*) \\
+bf'(x^*) & m_e
+\end{array}\right] $$
+with eigenvalues
+$$\lambda=m_e\pm  \frac{b}{2}\sqrt{f'(y^*)f'(x^*)}.$$
+
+Since we have assumed that $f$ is non-increasing everywhere, $f'(x)<0$ everywhere and both eigenvalues are real.  If $m_e<0$, the eigenvalue with greater absolute value will be 
+$$\lambda=m_e- \frac{b}{2} \sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(x^*\right)}$$
+since for both possible equilibria $y^*=\frac{m_f}{m_e}$.  In addition, $\lambda<0$ so that both equilibria are stable.
+
+Let $\lambda_s=m_e- \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f}{m_e}\right)}$ and $\lambda_{ns}=m_e-  \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f-b}{m_e}\right)}$ so that $\lambda_s$ is the eigenvalue associated with the equilibrium where Animal 2 signals and $\lambda_{ns}$ is the eigenvalue associated with the equilibrium where Animal 2 is not signaling.
+If we further assume that $f$ is a decreasing sigmoid, then $f'$ is decreasing for $x<-T$ and increasing for $x>-T$ so that $0<\left|f'\left(\frac{-m_f-b}{m_e}\right)\right|<\left|f'\left(\frac{-m_f}{m_e}\right)\right|$.
+Therefore, $0>\lambda_s>\lambda_{ns}$.
+In general, this suggests that (SMALL!) perturbations around the non-signaling equilibrium, $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$, will shrink faster than (SMALL!) perturbations around the signaling equilibrium, $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$.
+
+\section{Simulating the Stochastic Process }
+A simple updating algorithm allows us to simulate the process as described by the model above (ORNSTEIN-UHLENBECK VARIANT):
+
+\noindent Fix $\Delta t>0$.  Set $t=0$ and initialize $X_0^{(1)}=X_0^{(2)}=0$.
+
+\begin{enumerate}
+\item Draw $Z_1,\dots,Z_5$ i.i.d. from $N(0,1)$.
+\item \begin{align*}
+X_{t+\Delta t}^{(1)}&=X_t^{(1)}+\left(m_{e_1}X_t^{(1)}+m_f+b_{s_2}f(X_t^{(2)}\right)\Delta t+n_f\sqrt{\Delta t}Z_1+n_{e_1}\sqrt{\Delta t}Z_2+b_{s_2}\sqrt{f(X_t^{(2)}}\sqrt{\Delta t}Z_3\\
+X_{t+\Delta t}^{(2)}&=X_t^{(2)}+\left(m_{e_2}X_t^{(2)}+m_f+b_{s_1}f(X_t^{(1)}\right)\Delta t+n_f\sqrt{\Delta t}Z_1+n_{e_2}\sqrt{\Delta t}Z_4+b_{s_1}\sqrt{f(X_t^{(1)}}\sqrt{\Delta t}Z_5
+\end{align*}
+\item $t=t+\Delta t$
+
+\end{enumerate}
+
+To confirm that our simulation is performing reasonably, we let the model run without signaling feedback until the estimates reach stationary distributions and compare these to the analytical prediction, as given by (\ref{ou_stat1}) and (\ref{ou_stat2}).
+This comparison is shown in Figure \ref{ou_stat_dist}.
+As can be seen in the figure, the observed probability distribution of estimates is statistically indistinguishable from the prediction (for the first, K-S test $D=0.0243$, $p=0.5969$ and for the second, $D=0.0211$, $p=0.766$).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=.65\textwidth]{ou_stat_distribution.pdf} \end{center}
+\caption{\label{ou_stat_dist} A comparison between the stationary distribution from a simulation of the Ornstein-Uhlenbeck model and the analytically predicted stationary distribution.}
+\end{figure}
+
 We are interested in the effect of signaling feedback, i.e.
@@
 \begin{figure}
-\begin{subfigure}{.5 \textwidth}
-\includegraphics[width=1\textwidth]{feedback_stability_of_dominance_v1.pdf}
-\caption{}
+\begin{subfigure}{1 \textwidth}
+\begin{center}
+\includegraphics[width=.75\textwidth]{feedback_stability_of_dominance_v1.pdf}
+\end{center}
+%\caption{}
 \end{subfigure}
@@
 
+\section{To Do }
+
+\begin{itemize}
+\item More mathematical analysis.
+\item Embed the signaling dynamics in a network.  
+\begin{itemize}
+\item What's the relationship between the distribution of fighting abilities and the distribution of signaling behaviors?
+Between fighting abilities and the distribution of power?
+And how does that depend on the metric used to calculate power?
+\item What's the timescale separation between fighting abilities / fight outcomes / signaling behaviors and the power distribution?
+Tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
+\end{itemize}
+\end{itemize}
+
 \nocite{*}

### 2014-10-22 old_notes/model_07_13.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_07_10.tex
@@
 
-The stationary distribution of the estimate will be 
+The stationary distribution of the estimate will be  (I THINK)
 \begin{equation}
-X^{(1)}\sim  \label{ddm_stat}
+X^{(1)}\sim N(\exp((m_{e_1}+m_f)t),\frac{n_f^2+n_e^2}{2}t) \label{ddm_stat}
 \end{equation}
@@
 Let $\lambda_s=m_e- \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f}{m_e}\right)}$ and $\lambda_{ns}=m_e-  \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f-b}{m_e}\right)}$ so that $\lambda_s$ is the eigenvalue associated with the equilibrium where Animal 2 signals and $\lambda_{ns}$ is the eigenvalue associated with the equilibrium where Animal 2 is not signaling.
+There are two effects of signaling feedbacks on these eigenvalues.
+First, when $b\gneq 0$, both $\lambda_s$ and $\lambda_{ns}$ decrease so that the decay to the equilibria is faster when the estimates are influenced by signaling behaviors.
+
 If we further assume that $f$ is a decreasing sigmoid, then $f'$ is decreasing for $x<-T$ and increasing for $x>-T$ so that $0<\left|f'\left(\frac{-m_f-b}{m_e}\right)\right|<\left|f'\left(\frac{-m_f}{m_e}\right)\right|$.
@@
 In general, this suggests that (SMALL!) perturbations around the non-signaling equilibrium, $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$, will shrink faster than (SMALL!) perturbations around the signaling equilibrium, $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$.
+This effect is greater when the transition from non-signaling to signaling behavior as a function of $x$ is shallower: when the transition is steep, $f'(x)$ is essentially zero for $x>-T$ so that there is only a negligible difference between $f'\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ and $f'\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$
 
@@
 In both regimes, each animal's estimates reach a relatively stationary distribution fairly quickly so that we can pick a typical time point at which the distributions are not changing any more.
-(A Kolmogorov-Smirnov test for the differences between distributions at different points in time does not allow us to reject the null hypothesis that the distributions are the same, at a significance level of $p=0.05$.)  In Figure (\ref{diff_dists}), we plot the distribution of each animal's estimates at $t=200$ when there is and isn't signaling feedback ($b_{s_1}=b_{s_2}=0$ and $b_{s_1}=b_{s_2}\gneq 0$).
-We find that the distributions with and without signaling feedback are quite different.
-They are, in fact, statistically distinguishable (K-S test statistic $D=0.68,$ $p<2.2\times 10^{-16}$ for Animal 2 and $D=0.21$, $p<0.05$ for Animal 2).
-In both cases, the most noticeable affect of increasing the signaling feedback is increasing the variance of the distributions (from $1.55$ to $81.44$ for Animal 1 and from $1.23$ to $7.02$ for Animal 2).
-While the mean for Animal 2 is similar in the two cases ($-10.03$ without feedback and $-9.36$ with feedback), the mean for Animal 1 noticeably increases from $10.04$ to $17.25$.
+(A Kolmogorov-Smirnov test for the differences between distributions at different points in time does not allow us to reject the null hypothesis that the distributions are the same, at a significance level of $p=0.05$.)  If we allow the simulation to reach a stationary state, we find that the stationary distributions with and without signaling feedback ($b_{s_1}=b_{s_2}=0$ and $b_{s_1}=b_{s_2}\gneq 0$) are quite different for Animal 1(K-S test statistic, $D=0.995,p<2.2\times10^{-16}$), whereas they are statistically indistinguishable for Animal 2 ($D=0.048, p=0.1995$) (Figure (\ref{diff_dists})).
 
@@
 Fixing the other parameters of the model, for each difference in fighting ability $m_f$ and level of signaling feedback $b_{s_1}=b_{s_2}$, we can find the average number of switches that occur in a run of the model.
-This is shown in Figure (\ref{diff_switches}).
-As expected, as the difference in fighting ability increases fewer switches occur since Animal 1 is dominant the vast majority of the time.
-However, what is less expected is that increasing signaling feedback actually decreases stability.
+As the difference in fighting ability increases fewer switches occur since Animal 1 is dominant the vast majority of the time.
+Additionally, the average number of switches between dominance regimes decreases as $b_{s}$ increases, and the effect is greater for when the difference in fighting ability is smaller (Figure \ref{diff_switches}).
 
@@
 \begin{center}
-\includegraphics[width=.85\textwidth]{feedback_estimate_distributions_v1.pdf}
+\includegraphics[width=.85\textwidth]{feedback_estimate_distributions_v2.pdf}
 \end{center}
@@
 Other parameters are set as follows: threshold $T=5$, $m_{e_1}=m_{e_2}=-.5$, $n_{e_1}=n_{e_2}=1$, $m_f=5$, $n_f=.5$.
-Vertical red lines indicate predicted mean for process without feedback, $-m_f/m_{e_1}$ for Animal 1 and $m_f/m_{e_2}$ for Animal 2.
 }
@@
 \begin{subfigure}{1 \textwidth}
-\begin{center}
-\includegraphics[width=.75\textwidth]{feedback_stability_of_dominance_v1.pdf}
-\end{center}
-%\caption{}
+\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v2.pdf}
+\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v5.pdf}
+\caption{\label{low_error}}
+\end{subfigure}
+\begin{subfigure}{1 \textwidth}
+\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v4.pdf}
+\caption{\label{high_error}}
 \end{subfigure}
@@
 %\end{subfigure}
-\caption{\label{diff_switches} The number of switches between dominance regimes as a function of the difference in fighting ability, $m_f$, between the animals and the degree to which receiving a signal affects an animal's estimate of its dominance, $b_{s_1}=b_{s_2}$.
-Lighter colors correspond to a higher number of switches.
-}
+\caption{\label{diff_switches} The average number of switches between dominance regimes as a function of the difference in fighting ability, $m_f$, between the animals and the degree to which receiving a signal affects an animal's estimate of its dominance, $b_{s_1}=b_{s_2}$.
+In general, even for small differences in fighting ability there are very few switches in dominance regimes.
+However, fighting ability can be made small enough to allow switches to occur, in which case the strength of signaling feedbacks affects how frequently they happen.
+When the leak rate is relatively low, increases in signaling feedback decrease the number of switches (\ref{low_error}).
+When the error rate is high / estimates leak back to $0$ very quickly, increasing the signaling feedback actually increases the number of switches between dominance regimes / decreases the stability of a dominance regime (\ref{high_error}).}
 \end{figure}
+
+\section{Linear Learning Formulation}
+
+So far, we have considered how the processes of engaging in fights and sending and receiving signals might affect the estimates two animals make of each other's dominance.
+We are interested in understanding these processes on the population level and therefore would like to embed these dynamics in a model of a social network.
+In order to make this more analytically tractable, we modify our model into a linear learning model and, for the moment, ignore feedback from signaling into the learning process.
+Our main question is how the structure of a social network can facilitate individuals' learning about the states of other nodes in the network.
+
+Suppose each individual has a true fighting ability $a_1,\dots,a_N$ and let $f_{jk}=a_j-a_k$.
+Consider an individual in a social network observing its own interactions with other members of the group and the interactions between other members of the group and trying to learn each individual's true fighting ability.
+Let $x_j(t)$ be the individual's estimate of individual $j$'s fighting ability at time $t$.
+The focal individual will be able to learn about other individual's fighting ability by engaging in fights with individuals it interacts with and observing fights between pairs of individuals both of whom it interacts with.
+Let
+$$I_{jk}=\left\{
+\begin{array}{cccc}
+1 & , & \text{ if the focal individual interacts with both $j$ and $k$}\\
+0 & , & \text{ otherwise}
+\end{array}\right..
+$$
+Now, $x_j$ will change according to
+\begin{align*}
+\dot{x}_j(t)&=\sum_kI_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right] \text{ for } j\lneq N
+\\ \text{ and } x_N(t)&=T-\sum_{j\lneq N}x_j(t),
+\end{align*}
+where we impose a total on the estimates so that they do not diverge.  Therefore,
+\begin{align*}
+\dot{x}_j(t)&=\sum_{k\neq N}I_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right]+I_{jN}\left[f_{jN}-(x_j(t)-x_N(t))\right]
+\\&=\sum_{k\neq N}I_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right]+I_{jN}\left[f_{jN}-(x_j(t)-(T-\sum_{k\neq N}x_k(t)))\right]
+\\&=\sum_kI_{jk}f_{jk}-x_j(t)\sum_{k}I_{jk}+\sum_{k\neq N}I_{jk}x_k(t)+TI_{jN}-\sum_{k\neq N}I_{jN}x_k(t)
+\\&=\sum_kI_{jk}f_{jk}+TI_{jN}-x_j(t)\sum_kI_{jk}+\sum_{k\neq N}(I_{jk}-I_{jN})x_k(t)
+\\&=C_j+\sum_{k}L_{jk}x_k(t) \text{ where } C_j=\sum_{k\neq N}I_{jk}f_{jk}+TI_{jN} \text{ and } 
+\end{align*}
+$L_{jk}$  is an $n-1\times n-1$ matrix given by
+$$
+L_{jk}=\left\{
+\begin{array}{cccc}
+-\sum_lI_{jl}+I_{jj}-I_{jN} & , & \text{ if }j=k
+\\ I_{jk}-I_{jN} & , & \text{ if } j\neq k
+\end{array}\right..
+$$
+If $x_k(t)=a_k+\frac{T-\sum_ka_k}{N}$  for all $k$, $\dot{x}(t)=0$.  (If $T=\sum_ka_k$ then $x_k(t)=a_k$ is an equilibrium.)
+
+Stability of the equilibrium is determined by the eigenvalues of $L$.
 
@@
 \item More mathematical analysis.
+\begin{itemize}
+\item We know that if signaling is constantly present or absent, the stationary distribution of the Ornstein-Uhlenbeck process is just shifted and the variance does not change.
+However, we do not know what the likelihood of switching from non-signaling to signaling behavior (or vice versa) and whether the estimates would stay in a non-signaling / signaling regime long enough for a stationary distribution to be established.
+\end{itemize}
 \item Embed the signaling dynamics in a network.  

### 2014-10-22 old_notes/model_08_29_12.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_07_13.tex
@@
 \\ Eleanor Brush
-\\ July 11, 2012
+\\ August 29, 2012
 \end{center}
@@
 \item What's the timescale separation between fighting abilities / fight outcomes / signaling behaviors and the power distribution?  
-Tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
+	\begin{itemize}
+	\item when formulated as a Langevin equation: measure timescales through the expected escape time using martingales
+	\item when formulated as a dynamical system: measure timescales through time constants of potential wells.
+	\end{itemize}
+\item What's the / is there a tradeoff between mutual information between fighting abilities and power distribution and stability of the power distribution?
+\end{itemize}
+\item Compare to data.
+	\begin{itemize}
+	\item What are the relative rates of fighting and signals observed / predicted?
+	\item Can we explain the data with our model?
+	\item Do the monkeys weight fights won and lost equally?  Are they trying to maximize reward, minimize punishment, or are the two weighted equally?
+	\item Are the monkeys trying to minimize error tae or decision time? Can we tell if they're being forced to signal?
+	\item Is there a sense in which the monkeys are performing optimally? 
+	\item (ALSO THEORY QUESTION) How efficiently are the monkeys representing /using the information they're observing about the world?
+	\item If we can estimate parameters like the thresholds for signaling, how are they related to other variables in the system?
 	\end{itemize}

### 2014-10-22 old_notes/model_10_03_12.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_08_29_12.tex
@@
 \documentclass{article}
+\usepackage{latexsym}
+\usepackage{amssymb,amsmath}
 \usepackage{custom2}
 \usepackage{graphicx} % for figures
+\usepackage{epstopdf} % so can use EPS or PDF figures
 \usepackage{caption}
 \usepackage{subcaption}
-\usepackage{epstopdf} % so can use EPS or PDF figures
-\usepackage{subfig}
 \usepackage{url}
@@
 \usepackage{mathrsfs}
+\usepackage{booktabs}
+\usepackage[pdftex]{hyperref}
+\usepackage{lscape}
 \captionsetup{justification=RaggedRight, singlelinecheck=false}
+\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
+\newcommand{\argmax}{\text{argmax}}
+\newcommand{\Tr}{\text{Tr}}
+%\newtheorem{claim}{Claim}
+
+\addtolength{\evensidemargin}{-.5in}
+\addtolength{\oddsidemargin}{-.5in}
+\addtolength{\textwidth}{1.4in}
+\addtolength{\textheight}{1.4in}
+\addtolength{\topmargin}{-.5in}
+
+\pagestyle{empty}
 
@@
 \\ Eleanor Brush
-\\ August 29, 2012
+\\ October 03, 2012
 \end{center}
+
+\tableofcontents
 
@@
 \begin{align*}
-\frac{dx}{dt}&=m_ex+m_f+bf(y)
-\\ \frac{dy}{dt}&=m_ey-m_f+bf(x)
+\frac{dx}{dt}&=-ex+f+br(y+T)
+\\ \frac{dy}{dt}&=-ey-f+br(x+T)
 \end{align*}
-If $b=0$, there is an equilibrium at $(x,y)=\left(-\frac{m_f}{m_e},\frac{m_f}{m_e}\right).$  If $b\neq 0$, we can find equilibrium by solving $\frac{dx}{dt}=\frac{dy}{dt}=0$.
-We assume that $f(x)=1$ for $x<-T$ and $f(x)=0$ for $x>-T$.
+
+The isoclines of this system are given by
+\begin{align*}
+\dot{x}=0 \Rightarrow x&=\frac{f}{e}+\frac{b}{e}r(y+T) 
+\\\dot{y}=0 \Rightarrow y&=-\frac{f}{e}+\frac{b}{e}r(x+T)
+\end{align*}
+
+The Jacobian of this system is given by
+$$J(x,y)=\left(\begin{array}{cccc} 
+-e & br'(y+T)
+\\ br'(x+T) & -e
+\end{array}\right)$$
+Letting $\tau=Tr(J(x,y))$ and $\Delta=\text{det}(J(x,y)),$ we see that $\tau=-2e$ and $\Delta=e^2-b^2r'(y+T)r'(x+T)$.
+Since $e>0$, $\tau<0$ and any fixed point will be a saddle point, a stable sink, or a stable spiral, depending on the sign of $\Delta$.
+\begin{align*}
+\Delta<0 &\Leftrightarrow e^2-b^2r'(y+T)r'(x+T)<0 
+\\&\Leftrightarrow e<b\sqrt{r'(y+T)r'(x+T)} \text{ since } e,b, \text{ and } r'(y+T)r'(x+T)\geq0
+\\ &\Leftrightarrow \left\{\begin{array}{cccc}
+b>0 & \text{ if } e=0
+\\ b>2\epsilon e & \text{ if } e>0
+\end{array}\right.
+\end{align*}
+Therefore, if there is a fixed point with $r'(y+T)\neq 0$ and $r'(x+T)\neq 0$ (i.e.
+the isoclines cross where their slopes are not zero), the fixed point is a saddle and there are two other stable fixed points only if $b>2\epsilon e$.
+Other conditions are required, as well.
+In the following, we let $c=\frac{f}{e}$.
+Then we also need
+\begin{align}
+-T+\epsilon&<-c+\frac{b}{e}r(c+T) \label{cond1} 
+\\ -T-\epsilon&>-c+\frac{b}{e}r(c+\frac{b}{e}+T) \label{cond2}
+\end{align}
+
+These lead to the following conditions on $\frac{f}{e}=c$:
+\begin{align}
+(\ref{cond1}) \Rightarrow -T+\epsilon&<-c-\frac{b}{2\epsilon e}c-\frac{bT}{2\epsilon e}+\frac{b}{2e} \notag
+\\ \Rightarrow c&<\left(\frac{2\epsilon e-b}{2\epsilon e+b}\right)T-\frac{2\epsilon^2e}{2\epsilon e+b}+\frac{\epsilon b}{2\epsilon e+b}=:M \label{xmax}
+\\ (\ref{cond2}) \Rightarrow -T-\epsilon&>-c-\frac{b}{2\epsilon e}c-\frac{b^2}{2\epsilon e^2}-\frac{bT}{2\epsilon e}+\frac{b}{2e} \notag
+\\ \Rightarrow c&>\left(\frac{2\epsilon e-b}{2\epsilon e+b}\right)T+\frac{2\epsilon^2e-b^2/e}{2\epsilon e +b}+\frac{\epsilon b}{2\epsilon e+b}=:m \label{xmin}
+\end{align}
+Note that $m<M$ if and only if $b>2\epsilon e$, so that $m$ and $M$ are valid lower and upper bounds on $x$ only if $b$ is sufficiently large for a saddle point to exist.
+Equations (\ref{xmax}) and (\ref{xmin}) can be rewritten to provide bounds on $T$ that allow a bifurcation to occur:
+\begin{align}
+(\ref{xmax}) \Rightarrow(2\epsilon e-b)T&>  (2\epsilon e+b)c +2\epsilon^2e-\epsilon b \notag
+\\ \Rightarrow  T&<\frac{(2\epsilon e+b)c+2\epsilon^2e-\epsilon b}{2\epsilon e-b}  \notag
+\\ \Rightarrow T&< \frac{ 2\epsilon e(c+\epsilon)+b(c-\epsilon)}{2\epsilon e-b}\label{tmax}
+\\(\ref{xmin}) \Rightarrow (2\epsilon e-b)T&< (2\epsilon e+b)c-2\epsilon^2e+b^2/e-\epsilon b \notag
+\\ \Rightarrow  T&>\frac{(2\epsilon e+b)c-2\epsilon^2e+b^2/e-\epsilon b}{2\epsilon e-b}\notag
+\\ \Rightarrow T&> \frac{2\epsilon e(c-\epsilon)+b(c-\epsilon+b/e)}{2\epsilon e-b} \label{tmin}
+\end{align}
+If the equivalent conditions (\ref{cond1}),(\ref{xmax}),(\ref{tmax}) are violated, there can be no stable fixed point with animal 1 signaling at a nonzero rate and animal 2 not signaling because the difference in fighting ability is too great or the threshold below which signaling occurs is too low.
+
+Another interesting phenomenon is the possibility of each animal signaling the other at a non-zero rate.
+In order for this to occur, we need $\dot{x}(x,-\frac{f}{e})<0$ for al $x>x^*$ and in particular for $x=-T+\epsilon$ so that
+\begin{align*}
+-e(-T+\epsilon)+f+br(-\frac{f}{e}+T)&<0
+\\ \Rightarrow-T+\epsilon&>c+\frac{b}{e}r(-\frac{f}{e}+T)
+\\ \Rightarrow -T+\epsilon&>c+\frac{b}{e}\left(\frac{c}{2\epsilon}-\frac{T}{2\epsilon}+\frac{1}{2}\right)
+\\ \Rightarrow \frac{(b-2\epsilon e)}{2\epsilon e}T&>c-\epsilon+\frac{bc}{2\epsilon e}+\frac{b}{2e}
+\\ \Rightarrow (b-2\epsilon e)T&>2\epsilon e(c-\epsilon)+bc+b\epsilon
+\\ \Rightarrow T&<\frac{2\epsilon e(\epsilon -c)-b(c+\epsilon)}{2\epsilon e-b} 
+\end{align*}
+where the last inequality follows because this can only happen if $b<2\epsilon e$.
+(The condition $b<2\epsilon e$ is equivalent to the slope of the $x$ isocline being steeper than the slop of the $y$ isocline where they have non-zero slopes.
+Otherwise, if $x$ and $y$ are both signaling at non-zero rates, a bifurcation will occur.)
+
+Note that if $f>0$ so that $\frac{f}{e}>-T-\epsilon,$ there can be no steady state with animal $1$ signaling at rate $=1$.  The possible steady states are as follows:
 \begin{enumerate}
-\item $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{-m_f}{m_e}>-T$ and $\frac{m_f}{m_e}>-T$
-\item $\left(\frac{-m_f}{m_e},\frac{m_f-b}{m_e}\right)$ if $\frac{-m_f}{m_e}<-T$ and $\frac{m_f-b}{m_e}>-T$
-\item $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{-m_f-b}{m_e}>-T$ and $\frac{m_f}{m_e}<-T$
-\item $\left(\frac{-m_f-b}{m_e},\frac{m_f-b}{m_e}\right)$ if $\frac{-m_f-b}{m_e}<-T$ and $\frac{m_f-b}{m_e}<-T$
+\item Neither animal signaling:
+\begin{align*}
+\Rightarrow -T+\epsilon< \frac{f}{e} \text{ and } -T+\epsilon<-\frac{f}{e}
+\\ \Rightarrow T>\epsilon-\frac{f}{e} \text{ and } T> \epsilon+\frac{f}{e}
+\\ \Rightarrow T>\epsilon+\frac{f}{e}
+\end{align*}
+\item Animal 1 not signaling, animal 2 signaling at a rate $\in(0,1)$:
+\begin{enumerate}
+\item 
+\begin{align*}
+-T-\epsilon<-\frac{f}{e}
+\\ \Rightarrow T>\frac{f}{e}-\epsilon
+\end{align*}
+\item \begin{align*}
+\frac{f}{e}+\frac{b}{e}r\left(-\frac{f}{e}+T\right)>-T+\epsilon&>\max\left\{-\frac{f}{e}+\frac{b}{e}r\left(\frac{f}{e}+T\right),-\frac{f}{e}\right\}
+\\ \Rightarrow \epsilon-\frac{f}{e}-\frac{b}{e}r\left(-\frac{f}{e}+T\right)<T&<\epsilon-\max\left\{-\frac{f}{e}+\frac{b}{e}r\left(\frac{f}{e}+T\right),-\frac{f}{e}\right\}
+\\ \Rightarrow \epsilon-\frac{f}{e}-\frac{b}{e}r\left(-\frac{f}{e}+T\right)<T&<\min\left\{\epsilon+\frac{f}{e}-\frac{b}{e}r\left(\frac{f}{e}+T\right),\epsilon+\frac{f}{e}\right\}
+\\ \Rightarrow \left\{\begin{array}{cccc} 
+\frac{2\epsilon e(\epsilon -\frac{f}{e})+b(-\frac{f}{e}-\epsilon)}{2\epsilon e-b}<T<\epsilon+\frac{f}{e} &\text{ if } b<2\epsilon e
+\\\frac{2\epsilon e(\epsilon +\frac{f}{e})+b(\frac{f}{e}-\epsilon)}{2\epsilon e-b}<T<\epsilon+\frac{f}{e} &\text{ if } b>2\epsilon e
+\end{array}\right.
+\end{align*}
 \end{enumerate}
-Errors only provide negative feedback and prevent escalation when $m_e\leq 0$.
-In this case, equilibria $2.$ and $4.$ are impossible since $-m_f,-m_f-b<0$ $\Rightarrow$ $\frac{-m_f}{m_e},\frac{-m_f-b}{m_e}>0$ while $-T<0$.
-Therefore, no equilibria where Animal 1 is signaling is possible.
-So we are left with
+\item Animal 1 not signaling, animal 2 signaling at rate $=1$:
 \begin{enumerate}
-\item $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{m_f}{m_e}>-T$
-\item $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{m_f}{m_e}<-T$
+\item 
+\begin{align*}
+-T-\epsilon&>-\frac{f}{e}
+\\ \Rightarrow T&<\frac{f}{e}-\epsilon
+\end{align*}
+Therefore, for the following condition we assume that $\frac{f}{e}-\epsilon>0$ so that $\epsilon-\frac{f}{e}<0$.
+\item 
+\begin{align*}
+\frac{f}{e}+\frac{b}{e}>-T+\epsilon&>-\frac{f}{e}+\frac{b}{e}r\left(\frac{f}{e}+T\right)
+\\ \Rightarrow \epsilon-\frac{f}{e}-\frac{b}{e}<T&<\epsilon+\frac{f}{e}-\frac{b}{e}r\left(\frac{f}{e}+T\right)
+\\ \Rightarrow T&\left\{\begin{array}{cccc}
+<\frac{2\epsilon e(\epsilon +\frac{f}{e})+b(\frac{f}{e}-\epsilon)}{2\epsilon e-b} & \text{ if } b<2\epsilon e
+\\ >\frac{2\epsilon e(\epsilon +\frac{f}{e})+b(\frac{f}{e}-\epsilon)}{2\epsilon e-b} & \text{ if } b>2\epsilon e
+\end{array}\right.
+\\ \Rightarrow
+0<T&<\frac{f}{e}-\epsilon
+\end{align*}
 \end{enumerate}
-Note that if $b=0$, the equilibria are identical.
-Equilibrium $1.$ is only possible if $\frac{m_f}{m_e}>-T$ while $2.$ is only possible if $\frac{m_f}{m_e}<-T$ so that the relative strength of the effects from fights and errors determines which equilibria is possible.
-If $m_f$ is small, an equilibrium with neither animal signaling is possible, whereas is $m_f$ is sufficiently large at equilibrium Animal 2 will signal Animal 1 and Animal 1's estimate is boosted by $\frac{b}{-m_e}$.
-In either case, Animal 2's equilibrium estimate is the same.
+\item Both animals signaling at a non-zero rate, but only one steady state:
+\begin{enumerate}
+\item $b<2\epsilon e$ ensures that there is no unstable equilibrium, i.e. there is only one equilibrium
+\item 
+\begin{align*}
+-T+\epsilon&>-\frac{f}{e}+\frac{b}{e}r(-\frac{f}{e}+T)
+\\ \Rightarrow T&<\frac{2\epsilon e(\epsilon -\frac{f}{e})-b(\frac{f}{e}+\epsilon)}{2\epsilon e-b} 
+\end{align*}
 
-The Jacobian of this system at an equilibrium is
-$$
-J=\left[
-\begin{array}{ccccc}
-m_e & bf'(y^*) \\
-bf'(x^*) & m_e
-\end{array}\right] $$
-with eigenvalues
-$$\lambda=m_e\pm  \frac{b}{2}\sqrt{f'(y^*)f'(x^*)}.$$
+\end{enumerate}
 
-Since we have assumed that $f$ is non-increasing everywhere, $f'(x)<0$ everywhere and both eigenvalues are real.  If $m_e<0$, the eigenvalue with greater absolute value will be 
-$$\lambda=m_e- \frac{b}{2} \sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(x^*\right)}$$
-since for both possible equilibria $y^*=\frac{m_f}{m_e}$.  In addition, $\lambda<0$ so that both equilibria are stable.
+\item Two stable equilibria, one with animal 1 signaling at a non-zero rate and animal 2 not signaling and one with animal 1 not signaling and animal 2 signaling at a non-zero rate:
+\begin{enumerate}
+\item $b>2\epsilon e$
+\item 
+\begin{align*}
+-T+\epsilon&<-\frac{f}{e}+\frac{b}{e}r\left(\frac{f}{e}+T\right)
+\\ \Rightarrow T&>\epsilon+\frac{f}{e}-\frac{b}{e}r\left(\frac{f}{e}+T\right)
+\\ \Rightarrow T&<\frac{ 2\epsilon e(\epsilon+\frac{f}{e})+b(\frac{f}{e}-\epsilon)}{2\epsilon e-b}\end{align*}
+\item 
+\begin{align*}
+-T-\epsilon&>-\frac{f}{e}+\frac{b}{e}r\left(\frac{f}{e}+\frac{b}{e}+T\right)
+\\ \Rightarrow T&>\frac{2\epsilon e(\frac{f}{e}-\epsilon)+b(\frac{f}{e}+\frac{b}{e}-\epsilon)}{2\epsilon e-b}
+\\ \Rightarrow T&>0 \text{ if } b>2\epsilon e \text{ and } \frac{f}{e}<\epsilon 
+\end{align*}
+\end{enumerate}
+(Two stable equilibria, one with animal 1 signaling at a non zero rate and animal 2 not signaling and the other with animal 1 not signaling and animal 2 signaling at rate $=1$ would require $T>\epsilon-\frac{f}{e}$ and $T<\epsilon-\frac{f}{e}$, which is clearly not possible.)
+\item A single steady state with animal 1 signaling at a non-zero rate is not possible.
+\end{enumerate}
 
-Let $\lambda_s=m_e- \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f}{m_e}\right)}$ and $\lambda_{ns}=m_e-  \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f-b}{m_e}\right)}$ so that $\lambda_s$ is the eigenvalue associated with the equilibrium where Animal 2 signals and $\lambda_{ns}$ is the eigenvalue associated with the equilibrium where Animal 2 is not signaling.
-There are two effects of signaling feedbacks on these eigenvalues.
-First, when $b\gneq 0$, both $\lambda_s$ and $\lambda_{ns}$ decrease so that the decay to the equilibria is faster when the estimates are influenced by signaling behaviors.
+Let $$\psi(b)=\frac{2\epsilon e(\epsilon -\frac{f}{e})-b(\frac{f}{e}+\epsilon)}{2\epsilon e-b} \text{ and } \varphi(b)=\frac{ 2\epsilon e(\epsilon+\frac{f}{e})+b(\frac{f}{e}-\epsilon)}{2\epsilon e-b}$$
+Then, there are three general cases, depending on the magnitude of $\frac{f}{e}$ (Figure \ref{bifurcation}):
+\begin{enumerate}
+\item $\frac{f}{e}<\epsilon$:
+\begin{enumerate}
+\item neither animal signaling if $T>\epsilon+\frac{f}{e}$
+\item animal 1 not signaling and animal 2 signaling at a non-zero rate if $T<\epsilon+\frac{f}{e}$ and 
+$$T>\left\{\begin{array}{cccc}
+\varphi(b) & \text{ if } b>2\epsilon e
+\\ \psi(b) & \text{ if } b<2\epsilon e
+\end{array}\right.$$
 
-If we further assume that $f$ is a decreasing sigmoid, then $f'$ is decreasing for $x<-T$ and increasing for $x>-T$ so that $0<\left|f'\left(\frac{-m_f-b}{m_e}\right)\right|<\left|f'\left(\frac{-m_f}{m_e}\right)\right|$.
-Therefore, $0>\lambda_s>\lambda_{ns}$.
-In general, this suggests that (SMALL!) perturbations around the non-signaling equilibrium, $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$, will shrink faster than (SMALL!) perturbations around the signaling equilibrium, $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$.
-This effect is greater when the transition from non-signaling to signaling behavior as a function of $x$ is shallower: when the transition is steep, $f'(x)$ is essentially zero for $x>-T$ so that there is only a negligible difference between $f'\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ and $f'\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$
+\item both animals signal at a non-zero rate if $b<2\epsilon e$ and $T<\psi(b)$
+\item two stable equilibria if $b>2\epsilon e$ and $T<\varphi(b)$
+\end{enumerate}
+\item $\frac{f}{e}>\epsilon$:
+\begin{enumerate}
+\item neither animal signaling if $T>\epsilon+\frac{f}{e}$
+\item animal 1 not signaling and animal 2 signaling at a non-zero rate if $T<\epsilon+\frac{f}{e}$ and $T>\frac{f}{e}-\epsilon$
+\item animal 1 not signaling and animal 2 signaling at a rate $=1$ if $T<\frac{f}{e}-\epsilon$
+\end{enumerate}
+\end{enumerate}
+In summary, when the difference in fighting ability is low, there is more uncertainty about which animal is dominant and there can even be two stable equilibria.
+Once the difference in fighting ability is great enough, there can never be a stable equilibrium where the stronger animal thinks he is subordinate, but the equilibrium estimate of the weaker animal depends on the threshold below which signaling occurs.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=1\textwidth]{bifurc_diagram.pdf}
+\end{center}
+\caption{{\bf The equilibria of the signaling model as a function of feedback level, $b$, and signaling threshold, $T$.} For both plots, $e=.5$, $\epsilon=.1$.
+As $f$ increases from less than $\epsilon e$ to greater than $\epsilon e$, the possible equilibria change.
+\label{bifurcation}}
+\end{figure}
+
+%\frac{(2\epsilon e)(\epsilon-\frac{f}{e})-b(\frac{f}{e}+\epsilon)}{2\epsilon e-b}
+%\begin{table}
+%\caption{Possible equilibrium of the signaling model without stochastic dynamics.}
+%\begin{tabular}{@{}lllll@{}}
+%\toprule
+%State & Condition on $b$ & Condition on $T$
+%\\\midrule animal $1$ not signaling, animal $2$ signaling at a non-zero rate & & $T>\frac{f}{e}-\epsilon$
+%\end{tabular}
+%\end{table}
 

### 2014-10-22 old_notes/notes_signaling_model_10_03_12.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_10_03_12.tex
@@
 %\newtheorem{claim}{Claim}
-
-\addtolength{\evensidemargin}{-.5in}
-\addtolength{\oddsidemargin}{-.5in}
-\addtolength{\textwidth}{1.4in}
-\addtolength{\textheight}{1.4in}
-\addtolength{\topmargin}{-.5in}
-
-\pagestyle{empty}
 
@@
 \end{enumerate}
-(Two stable equilibria, one with animal 1 signaling at a non zero rate and animal 2 not signaling and the other with animal 1 not signaling and animal 2 signaling at rate $=1$ would require $T>\epsilon-\frac{f}{e}$ and $T<\epsilon-\frac{f}{e}$, which is clearly not possible.)
+(Two stable equilibria, one with animal 1 signaling at a non zero rate and animal 2 not signaling and the other with animal 1 not signaling and animal 2 signaling at rate $=1$ would require $T<\frac{f}{e}-\epsilon$, which is negative if there is a bifurcation, so that this is not possible.
+
+Two stable equilibria, one with animal 1 signaling at a non zero rate and animal 2 not signaling and the other with animal 1 signaling at a non-zero rate and animal 2 signaling at rate $=1$ would require $T<\epsilon-\frac{f}{e}-\frac{b}{e}\Rightarrow T<-\epsilon-\frac{f}{e}$, which is negative if there is a bifurcation, so that this is not possible.)
 \item A single steady state with animal 1 signaling at a non-zero rate is not possible.

### 2014-10-22 old_notes/notes_signaling_model_11_01_12.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/model_08_29_12.tex
@@
 \usepackage{mathrsfs}
+\usepackage{booktabs}
+\usepackage[pdftex]{hyperref}
+\usepackage{lscape}
 \captionsetup{justification=RaggedRight, singlelinecheck=false}
+\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
+\newcommand{\argmax}{\text{argmax}}
+\newcommand{\Tr}{\text{Tr}}
+%\newtheorem{claim}{Claim}
 
@@
 \\ Eleanor Brush
-\\ August 29, 2012
+\\ November 01, 2012
 \end{center}
+
+\tableofcontents
 
@@
 \begin{align*}
-\frac{dx}{dt}&=m_ex+m_f+bf(y)
-\\ \frac{dy}{dt}&=m_ey-m_f+bf(x)
+\frac{dx}{dt}&=-ex+f+br(y+T)
+\\ \frac{dy}{dt}&=-ey-f+br(x+T)
 \end{align*}
-If $b=0$, there is an equilibrium at $(x,y)=\left(-\frac{m_f}{m_e},\frac{m_f}{m_e}\right).$  If $b\neq 0$, we can find equilibrium by solving $\frac{dx}{dt}=\frac{dy}{dt}=0$.
-We assume that $f(x)=1$ for $x<-T$ and $f(x)=0$ for $x>-T$.
+
+The isoclines of this system are given by
+\begin{align*}
+\dot{x}=0 \Rightarrow x&=\frac{f}{e}+\frac{b}{e}r(y+T) 
+\\\dot{y}=0 \Rightarrow y&=-\frac{f}{e}+\frac{b}{e}r(x+T)
+\end{align*}
+
+The Jacobian of this system is given by
+$$J(x,y)=\left(\begin{array}{cccc} 
+-e & br'(y+T)
+\\ br'(x+T) & -e
+\end{array}\right)$$
+Letting $\tau=Tr(J(x,y))$ and $\Delta=\text{det}(J(x,y)),$ we see that $\tau=-2e$ and $\Delta=e^2-b^2r'(y+T)r'(x+T)$.
+Since $e>0$, $\tau<0$ and any fixed point will be a saddle point, a stable sink, or a stable spiral, depending on the sign of $\Delta$.
+\begin{align*}
+\Delta<0 &\Leftrightarrow e^2-b^2r'(y+T)r'(x+T)<0 
+\\&\Leftrightarrow e<b\sqrt{r'(y+T)r'(x+T)} \text{ since } e,b, \text{ and } r'(y+T)r'(x+T)\geq0
+\\ &\Leftrightarrow \left\{\begin{array}{cccc}
+b>0 & \text{ if } e=0
+\\ b>2\epsilon e & \text{ if } e>0
+\end{array}\right.
+\end{align*}
+Therefore, if there is a fixed point with $r'(y+T)\neq 0$ and $r'(x+T)\neq 0$ (i.e.
+the isoclines cross where their slopes are not zero), the fixed point is a saddle and there are two other stable fixed points only if $b>2\epsilon e$.
+Figure \ref{basins} shows the basins of attraction for the two stable nodes when the system is bistable.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in]{basins_of_attraction.eps}
+\end{center}
+\caption{{\bf The basins of attraction for each of the two stable nodes in the case when the system is bistable.} The fixed points are marked in red circles.
+The isoclines are marked as blue lines.
+The red line separates the basin of attraction for the fixed point where the stronger animal is signaling (above the red line) and the fixed point where the weaker animal is signaling (below the red line).
+\label{basins}}
+\end{figure}
+
+Let $$\psi(b)=\frac{2\epsilon e(\epsilon -\frac{f}{e})-b(\frac{f}{e}+\epsilon)}{2\epsilon e-b} \text{ and } \varphi(b)=\frac{ 2\epsilon e(\epsilon+\frac{f}{e})+b(\frac{f}{e}-\epsilon)}{2\epsilon e-b}$$
+Then, there are three general cases, depending on the magnitude of $\frac{f}{e}$ (Figures \ref{bifurcation}-\ref{wireframe}):
 \begin{enumerate}
-\item $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{-m_f}{m_e}>-T$ and $\frac{m_f}{m_e}>-T$
-\item $\left(\frac{-m_f}{m_e},\frac{m_f-b}{m_e}\right)$ if $\frac{-m_f}{m_e}<-T$ and $\frac{m_f-b}{m_e}>-T$
-\item $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{-m_f-b}{m_e}>-T$ and $\frac{m_f}{m_e}<-T$
-\item $\left(\frac{-m_f-b}{m_e},\frac{m_f-b}{m_e}\right)$ if $\frac{-m_f-b}{m_e}<-T$ and $\frac{m_f-b}{m_e}<-T$
+\item $\frac{f}{e}<\epsilon$:
+\begin{enumerate}
+\item neither animal signaling if $T>\epsilon+\frac{f}{e}$
+\item animal 1 not signaling and animal 2 signaling at a non-zero rate if $T<\epsilon+\frac{f}{e}$ and 
+$$T>\left\{\begin{array}{cccc}
+\varphi(b) & \text{ if } b>2\epsilon e
+\\ \psi(b) & \text{ if } b<2\epsilon e
+\end{array}\right.$$
+
+\item both animals signal at a non-zero rate if $b<2\epsilon e$ and $T<\psi(b)$
+\item two stable equilibria if $b>2\epsilon e$ and $T<\varphi(b)$
 \end{enumerate}
-Errors only provide negative feedback and prevent escalation when $m_e\leq 0$.
-In this case, equilibria $2.$ and $4.$ are impossible since $-m_f,-m_f-b<0$ $\Rightarrow$ $\frac{-m_f}{m_e},\frac{-m_f-b}{m_e}>0$ while $-T<0$.
-Therefore, no equilibria where Animal 1 is signaling is possible.
-So we are left with
+\item $\frac{f}{e}>\epsilon$:
 \begin{enumerate}
-\item $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{m_f}{m_e}>-T$
-\item $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$ if $\frac{m_f}{m_e}<-T$
+\item neither animal signaling if $T>\epsilon+\frac{f}{e}$
+\item animal 1 not signaling and animal 2 signaling at a non-zero rate if $T<\epsilon+\frac{f}{e}$ and $T>\frac{f}{e}-\epsilon$
+\item animal 1 not signaling and animal 2 signaling at a rate $=1$ if $T<\frac{f}{e}-\epsilon$
 \end{enumerate}
-Note that if $b=0$, the equilibria are identical.
-Equilibrium $1.$ is only possible if $\frac{m_f}{m_e}>-T$ while $2.$ is only possible if $\frac{m_f}{m_e}<-T$ so that the relative strength of the effects from fights and errors determines which equilibria is possible.
-If $m_f$ is small, an equilibrium with neither animal signaling is possible, whereas is $m_f$ is sufficiently large at equilibrium Animal 2 will signal Animal 1 and Animal 1's estimate is boosted by $\frac{b}{-m_e}$.
-In either case, Animal 2's equilibrium estimate is the same.
+\end{enumerate}
+In summary, when the difference in fighting ability is low, there is more uncertainty about which animal is dominant and there can even be two stable equilibria.
+Once the difference in fighting ability is great enough, there can never be a stable equilibrium where the stronger animal thinks he is subordinate, but the equilibrium estimate of the weaker animal depends on the threshold below which signaling occurs.
 
-The Jacobian of this system at an equilibrium is
-$$
-J=\left[
-\begin{array}{ccccc}
-m_e & bf'(y^*) \\
-bf'(x^*) & m_e
-\end{array}\right] $$
-with eigenvalues
-$$\lambda=m_e\pm  \frac{b}{2}\sqrt{f'(y^*)f'(x^*)}.$$
+In the above analysis, we considered fixed values of $b$ and $f$ and determined how the value of the threshold $T$ affected the behavior of the system.
+In practice, it seems more reasonable for there to be fixed values of $T$ and $b$ (i.e.
+psychologically all animals in the population are more or less the same), but the difference in fighting ability will vary from pair to pair.
+If this is the case, then we need to determine how the value of $f$ affects the behavior of the system (Figures \ref{f_heatmap} and \ref{wireframe}):
+\begin{enumerate}
+\item $T>\epsilon$ (so that no animal ever signals when he has a positive estimate of its own dominance):
+\begin{enumerate}
+\item neither animal signals if $f<e(T-\epsilon)$ 
+\item the weaker animal signals at a non-zero rate if $f>e(T-\epsilon)$ and $f<e(T+\epsilon)$
+\item the weaker animal signals at a maximal rate if $f>e(T+\epsilon)$
+\end{enumerate}
+\item $T<\epsilon$
+\begin{enumerate}
+\item both signal if $b<2\epsilon e$ and $f<\frac{(2\epsilon e-b)e(\epsilon-T)}{2\epsilon e+b}$
+\item the system is bistable with two equilibria, and in either case one animal doesn't signal and the other signals at a non-zero rate if $b>2\epsilon e$ and $f<\frac{(b-2\epsilon e)e(\epsilon-T)}{2\epsilon e+b}$
+\item the weaker animal signals at a non-zero rate if $f>\left|\frac{(b-2\epsilon e)e(\epsilon-T)}{2\epsilon e+b}\right|$ and $f<e(T+\epsilon)$
+\item the weaker animal signals at a maximal rate if $f>e(T+\epsilon)$
+\end{enumerate}
+\end{enumerate}
 
-Since we have assumed that $f$ is non-increasing everywhere, $f'(x)<0$ everywhere and both eigenvalues are real.  If $m_e<0$, the eigenvalue with greater absolute value will be 
-$$\lambda=m_e- \frac{b}{2} \sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(x^*\right)}$$
-since for both possible equilibria $y^*=\frac{m_f}{m_e}$.  In addition, $\lambda<0$ so that both equilibria are stable.
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=1\textwidth]{bifurc_diagram.pdf}
+%\end{center}
+%\caption{{\bf The equilibria of the signaling model as a function of feedback level, $b$, and signaling threshold, $T$.} For both plots, $e=.5$, $\epsilon=.1$.
+As $f$ increases from less than $\epsilon e$ to greater than $\epsilon e$, the possible equilibria change.
+\label{bifurcation}}
+%\end{figure}
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=6in]{heatmap_bifurcation_v1.eps}
+%\end{center}
+%\caption{{\bf The critical value of the parameter $f$ as a function of $b$ and $T$.}  For a given pair of values of $b$ and $T$, if $f$ is above the value indicated by the heatmap the equilibrium of the system is such that the weaker animal signals at a non-zero rate and the stronger animal doesn't signal.
+If $f$ is below the value indicated by the heatmap, the equilibria is as indicated by the text.
+\label{f_heatmap}}
+%\end{figure}
+%
+%\begin{figure}
+%%\includegraphics[]{wireframe_bifurcation_v1.eps}
+%\includegraphics[width=\textwidth]{wireframe_bifurcation_v2.eps}
+%\caption{{\bf The dependence of the equilibria of the system on all three parameters, $b$, $T$, and $f$.} $T$ is on the $x$ axis, $b$ is on the $y$ axis and $f$ is on the $z$ axis.
+If $f$ is above the upper plane, the equilibrium is such that the weaker animal signals at a maximal rate.
+If $f$ is between the two planes, the equilibrium is such that the weaker animal signals at a non-zero rate.
+If $f$ is below the lower plane, the equilibria depends on $b$ and $T$ as indicated by Figures \ref{bifurcation} and \ref{f_heatmap}.
+\label{wireframe}}
+%\end{figure}
 
-Let $\lambda_s=m_e- \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f}{m_e}\right)}$ and $\lambda_{ns}=m_e-  \frac{b}{2}\sqrt{f'\left(\frac{m_f}{m_e}\right)f'\left(\frac{-m_f-b}{m_e}\right)}$ so that $\lambda_s$ is the eigenvalue associated with the equilibrium where Animal 2 signals and $\lambda_{ns}$ is the eigenvalue associated with the equilibrium where Animal 2 is not signaling.
-There are two effects of signaling feedbacks on these eigenvalues.
-First, when $b\gneq 0$, both $\lambda_s$ and $\lambda_{ns}$ decrease so that the decay to the equilibria is faster when the estimates are influenced by signaling behaviors.
+Suppose these dynamics are going on between every pair of animals in a group where the fighting abilities are drawn from some distribution.
+How will a network of signals exchanged develop?
+Let's first look at a simple metric of such a network: the number of signalers sending signals to a given animal.
+\begin{align*}
+P(n \text{ signalers})=\int P(x)P(n \text{ signalers}|x)dx \text{ where $x$ is the fighting ability of a focal animal}
+\end{align*}
+If $T>\epsilon$ then animal $i$ signals to animal $j$ if and only if $f_j-f_i>e(T-\epsilon)$.  Therefore,
+\begin{align*}
+P(n \text{ signalers}|x)={N-1 \choose n}P(x-y>e(T-\epsilon))^nP(x-y<e(T-\epsilon))^{(N-1-n)}
+\end{align*}
+where $N$ is the total number of animals in the group and we use the binomial distribution to express the probability of $n$ signalers as a function of the probability of one of the other $N-1$ animals signaling, which we assume to be independent events.
 
-If we further assume that $f$ is a decreasing sigmoid, then $f'$ is decreasing for $x<-T$ and increasing for $x>-T$ so that $0<\left|f'\left(\frac{-m_f-b}{m_e}\right)\right|<\left|f'\left(\frac{-m_f}{m_e}\right)\right|$.
-Therefore, $0>\lambda_s>\lambda_{ns}$.
-In general, this suggests that (SMALL!) perturbations around the non-signaling equilibrium, $\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$, will shrink faster than (SMALL!) perturbations around the signaling equilibrium, $\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$.
-This effect is greater when the transition from non-signaling to signaling behavior as a function of $x$ is shallower: when the transition is steep, $f'(x)$ is essentially zero for $x>-T$ so that there is only a negligible difference between $f'\left(\frac{-m_f}{m_e},\frac{m_f}{m_e}\right)$ and $f'\left(\frac{-m_f-b}{m_e},\frac{m_f}{m_e}\right)$
+At first, we will assume that fighting ability is drawn uniformly from some range $[-L,L]$.  In order for signaling to be possible, we need $e(T-\epsilon)<2L$.
+\begin{align*}
+\\ \Rightarrow P(x-y>e(T-\epsilon))&=\frac{L+x-e(T-\epsilon)}{2L} \\ \text{ and } P(x-y<e(T-\epsilon))&=\frac{L-x+e(T-\epsilon))}{2L}
+\\ \Rightarrow P(n \text{ signalers}|x)&={N-1\choose n}\left(\frac{L+x-e(T-\epsilon)}{2L}\right)^n\left(\frac{L-x+e(T-\epsilon))}{2L}\right)^{(N-1-n)}
+\\ \Rightarrow P(n \text{ signalers})&=\int_{-L+e(T-\epsilon)}^L\frac{1}{2L}{N-1\choose n}\left(\frac{L+x-e(T-\epsilon)}{2L}\right)^n\left(\frac{L-x+e(T-\epsilon))}{2L}\right)^{(N-1-n)}dx
+\\ &=\int_0^{1-\frac{e(T-\epsilon)}{2L}}{N-1\choose n}z^n(1-z)^{(N-1-n)}dz \text{ using $z=\frac{L+x-e(T-\epsilon)}{2L}$}
+\\ &={N-1\choose n}\int_0^{1-\frac{e(T-\epsilon)}{2L}}z^n(1-z)^{(N-1-n)}
+\\ &={N-1 \choose n} B\left(1-\frac{e(T-\epsilon)}{2L};n+1,N-n\right) \text{ where $B(c;x,y)=\int_0^cz^{x-1}(1-z)^{y-1}dz$}  
+\\&={N-1\choose n}B(n+1,N-n)F\left(1-\frac{e(T-\epsilon)}{2L};n+1,N-n\right)
+\\&=\frac{1}{N}F\left(1-\frac{e(T-\epsilon)}{2L};n+1,N-n\right) \tag{$\star$} \label{STAR}
+\end{align*} 
+where $F(x;\alpha,\beta)=\int_0^x\frac{z^{\alpha-1}(1-z)^{\beta-1}}{B(\alpha,\beta)}dz$ is the cumulative distribution function of the Beta distribution and $B$ is the Beta function with shape parameters $\alpha$ and $\beta$.
+
+If $n=0$, we have to make a small correction:
+\begin{align*}
+P(0 \text{ signalers})&=\frac{1}{N}F\left(1-\frac{e(T-\epsilon)}{2L};1,N\right)+\frac{e(T-\epsilon)}{2L}
+\end{align*}
+since if $x\in [-L,-L+\frac{e(T+\epsilon)}{2L}],$ the animal will receive no signals.
+
+Suppose $T=\epsilon$ so signaling at a non-zero rate starts as soon as an animal's estimate is negative.  Then 
+\begin{align*}
+P(n \text{ signalers})&=\frac{1}{N}F(1-0;n+1,N-n)
+\\&=\frac{1}{N}F(1;n+1,N-n)
+\\&=\frac{1}{N}
+\end{align*}
+so the distribution of number of signalers is uniform over the possibilities $0,1,\dots,N-1$.
+If $T\gneq \epsilon$, having fewer signalers is more probable and having more signalers is less probably.
+The probability distribution of number of signalers for a small set of parameters is shown in Figure \ref{derived_sig_probs}.
+
+\begin{figure}
+\includegraphics[width=6in]{distributions_of_signals_received_uniform.pdf}
+\caption{The probability distribution of number of signalers, assuming a uniform distribution of fighting abilities, as derived in Eq. (\ref{STAR}).  \label{derived_sig_probs}}
+\end{figure}
 

### 2014-10-22 old_notes/notes_signaling_model_12_11_12.tex
ive added code for tracking mutual information over time and changed figures a bit

--- 2014-10-22 old_notes/notes_signaling_model_11_01_12.tex
@@
 \\ Eleanor Brush
-\\ November 01, 2012
+\\ updated December 12, 2012
 \end{center}
@@
 
-%\title{Development of Signaling Network}
-%\author{Eleanor Brush}
-%\maketitle
-
 \vspace{0pt}
 \normalsize
+
+\section{Motivation }
+
+\begin{table*}\centering
+$$\text{The Model : }dX_t=\left(f-eX_t+bf(Y_t)\right)dt+\sigma^2 dW_t^X$$
+\ra{1.3}
+\begin{tabular}{@{}lll@{}}\toprule
+& Behavioral  & Neural 
+\\\midrule 
+\emph{Objects: } \\
+Individual / node & animal & neuron(al population) with preferences
+\\ Variable ($X,Y$) & estimate of dominance  & firing rate
+\\ Drift ($f$) &  trend in fights  & trend in firing rates
+\\ Source of drift & difference in fighting ability & difference in response to stimulus
+\\ Leak / error ($e$) & errors in cognitive process & loss of firing activity
+\\ Feedback ($b$) & importance of information from other animal & strength of connection between neurons
+\\ Threshold ($T$) & estimate below which an animal signals & firing rate at which a decision is made
+\\ \midrule
+\emph{Interpretation:} \\
+Conflict dynamics & fights (and other ways of gathering evidence) & firing rates
+\\ Outcome for pair & agreement about which is dominant & agreement about which is firing faster
+\\ Outcome for population & signaling network & comparisons between pairs of neurons
+\\ Consequence for pair & increased affiliation & ?
+\\ Consequence for pop. & power distribution & decision between left and right
+\\ \midrule 
+\emph{Measure of performance:} \\
+For individual & incorrectly signaling, time to decision & 
+\\ \bottomrule
+\end{tabular}
+\caption{Summary of analogy between the model as applied to behavioral and neural systems.}
+\end{table*}
+
 \section{Model as a Stochastic Differential Equation}
@@
 
-\subsection{Evolution of probability distribution over time }
-More generally, consider a system
-$$ dX_t=a(X_t)dt+\sum_{i=1}^5\sigma_i(X_t)dW_t^{(i)}$$
-with $W_t^{(1)},\dots,W_t^{(5)}$ independent Brownian motions.
-If $a,\sigma_1,\dots,\sigma_5:\R^2\to\R^2,$ are ``nice" enough functions, and the random variable $X_t$ has a ``nice" enough density function (i.e.
-$\E[g(X_t)]=\int g(x)p(t,x)dx$ for every bounded measurable function $g$), then according to the Fokker-Planck / Kolmogorov forward equation
-\begin{align*}
-\frac{\partial p}{\partial t}&=-\sum_{i=1}^2\frac{\partial}{\partial x_i}(a^{(i)}(x)p(t,x))+\frac{1}{2}\sum_{i,j=1}^2\sum_{k=1}^5\frac{\partial^2}{\partial x_i\partial x_j}(\sigma_k^{(i)}\sigma_k^{(j)}p(t,x)
-\\ \Rightarrow \frac{\partial p}{\partial t}&=-(m_{e_1}+m_f+b_{s_2}f(x_2))\frac{\partial p}{\partial x_1}-(m_{e_2}-m_f+b_{s_1}f(x_1))\frac{\partial p}{\partial x_2}
-\\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}-n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
-\end{align*}
-using the functions $a$ and $\sigma_i$ as for our particular model.  Note that this requires the function $f:\R\to\R$ be $C^2$.
-
-\subsection{Statistics of the Process WITHOUT Signaling Feedback }
-%(See ``The Theory of Stochastic Processes," Cox and Miller.)  Let $p(x_0,x,y_0,y;t)$ be the probability of ending up at $x,y$ at time $t$ given that the process started at $x_0,y_0$.
-Then $p$ satisfies the backward equation:
-%\begin{align}
-%-\frac{\partial}{\partial t}p(x_0,x,y_0,y;t)&=(m_{e_1}+m_f+b_{s_2}f(y_0))\frac{\partial}{\partial x_0}p(x_0,x,y_0,y;t)+(m_{e_1}-m_f+b_{s_2}f(x_0))\frac{\partial}{\partial y_0}p(x_0,x,y_0,y;t) \notag
-%\\&+\frac{1}{2}(n_f^2+n_{e_1}^2+b_{s_2}^2f(y_0))\frac{\partial^2}{\partial x_0^2}p(x_0,x,y_0,y;t)+n_f^2\frac{\partial^2}{\partial x_0\partial y_0}p(x_0,x,y_0,y;t) \notag
-%\\&+\frac{1}{2}(n_f^2+n_{e_2}^2+b_{s_1}^2f(x_0))\frac{\partial^2}{\partial y_0^2}p(x_0,x,y_0,y;t) \label{back}
-%\end{align}
-%
-%We would like to know the expected time until either estimate reaches the signaling threshold $-T$.  
-For simplicity, we start with the case where there is no feedback from signaling to the estimates so that the model becomes
-\begin{align*}
-dX_t^{(1)}&=\left[m_{e_1}+m_f\right]dt+n_fdW_t^{(1)}+n_{e_1}dW_t^{(2)}
-\\ dX_t^{(2)}&=\left[m_{e_2}-m_f\right]dt-n_fdW_t^{(1)}+n_{e_2}dW_t^{(4)}
-\end{align*}
-
-This is a simple drift-diffusion model.
-We can consider the dynamics of each estimate separately.
-We will show results for $X_t^{(1)}$, which follow similarly for $X_t^{(2)}$.
-As derived in \cite{Bogacz:2006uq}, the expected time until $X_t^{(1)}$ reaches either $-T$ or $T$ is
-\begin{align*}
-\frac{T}{m_e+m_f}\tanh\left(\frac{T(m_e+m_f)}{n_f^2+n_e^2}\right)
-\end{align*}
-and the probability that $X_t^{(1)}$ reaches $-T$ before it reaches $T$ is
-\begin{align*}
-\frac{1}{1+\exp\left(\frac{2T(m_e+m_f)}{n_f^2+n_e^2}\right)}
-\end{align*}
-(assuming $X_t^{(1)}=0$).
-
-The stationary distribution of the estimate will be  (I THINK)
-\begin{equation}
-X^{(1)}\sim N(\exp((m_{e_1}+m_f)t),\frac{n_f^2+n_e^2}{2}t) \label{ddm_stat}
-\end{equation}
-
 \subsection{Ornstein-Uhlenbeck Variant }
-As shown in (\ref{ddm_stat}), both the mean and variance of the distribution of estimates increase with time for a pure drift-diffusion model.
+Both the mean and variance of the distribution of estimates increase with time for a pure drift-diffusion model.
 The dynamics are more interesting if the estimates don't just keep increasing or decreasing more or less deterministically.
@@
 \\&+\frac{1}{2}(n_{e_1}^2+n_f^2+b_{s_2}^2f(x_2))\frac{\partial^2 p}{\partial x_1^2}-n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2+b_{s_1}^2f(x_1))\frac{\partial^2 p}{\partial x_2^2}
-\end{align*}
-If we ignore signaling feedback for the moment and look for a stationary distribution, we want to solve the equation
-\begin{align*}
-0&=(m_{e_1}x_1+m_f)\frac{\partial p}{\partial x_1}-m_{e_1}p-(m_{e_2}x_2-m_f)\frac{\partial p}{\partial x_2}-m_{e_2}p
-\\&+\frac{1}{2}(n_{e_1}^2+n_f^2)\frac{\partial^2 p}{\partial x_1^2}-n_f^2\frac{\partial p}{\partial x_1\partial x_2}+\frac{1}{2}(n_{e_2}^2+n_f^2)\frac{\partial^2 p}{\partial x_2^2}
-\\&=(m_ex+m_f)\frac{\partial p}{\partial x}-(m_ey-m_f)\frac{\partial p}{\partial y}-2m_ep+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial x^2}-n_f^2\frac{\partial p}{\partial x\partial y}+\frac{1}{2}(n_e^2+n_f^2)\frac{\partial^2 p}{\partial y^2}
-\\& \text{ if we assume $m_{e_1}=m_{e_2}$}
 \end{align*}
@@
 
+\subsection{Equilibria }
 The isoclines of this system are given by
@@
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=1\textwidth]{bifurc_diagram.pdf}
-%\end{center}
-%\caption{{\bf The equilibria of the signaling model as a function of feedback level, $b$, and signaling threshold, $T$.} For both plots, $e=.5$, $\epsilon=.1$.
+\begin{figure}
+\begin{center}
+\includegraphics[width=1\textwidth]{bifurc_diagram.pdf}
+\end{center}
+\caption{{\bf The equilibria of the signaling model as a function of feedback level, $b$, and signaling threshold, $T$.} For both plots, $e=.5$, $\epsilon=.1$.
 As $f$ increases from less than $\epsilon e$ to greater than $\epsilon e$, the possible equilibria change.
 \label{bifurcation}}
-%\end{figure}
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=6in]{heatmap_bifurcation_v1.eps}
-%\end{center}
-%\caption{{\bf The critical value of the parameter $f$ as a function of $b$ and $T$.}  For a given pair of values of $b$ and $T$, if $f$ is above the value indicated by the heatmap the equilibrium of the system is such that the weaker animal signals at a non-zero rate and the stronger animal doesn't signal.
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in]{heatmap_bifurcation_v1.eps}
+\end{center}
+\caption{{\bf The critical value of the parameter $f$ as a function of $b$ and $T$.}  For a given pair of values of $b$ and $T$, if $f$ is above the value indicated by the heatmap the equilibrium of the system is such that the weaker animal signals at a non-zero rate and the stronger animal doesn't signal.
 If $f$ is below the value indicated by the heatmap, the equilibria is as indicated by the text.
 \label{f_heatmap}}
-%\end{figure}
-%
-%\begin{figure}
-%%\includegraphics[]{wireframe_bifurcation_v1.eps}
-%\includegraphics[width=\textwidth]{wireframe_bifurcation_v2.eps}
-%\caption{{\bf The dependence of the equilibria of the system on all three parameters, $b$, $T$, and $f$.} $T$ is on the $x$ axis, $b$ is on the $y$ axis and $f$ is on the $z$ axis.
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{wireframe_bifurcation_v2.eps}
+\caption{{\bf The dependence of the equilibria of the system on all three parameters, $b$, $T$, and $f$.} $T$ is on the $x$ axis, $b$ is on the $y$ axis and $f$ is on the $z$ axis.
 If $f$ is above the upper plane, the equilibrium is such that the weaker animal signals at a maximal rate.
@@
 \label{wireframe}}
-%\end{figure}
+\end{figure}
 
+\subsection{Development of a Signaling Network }
 Suppose these dynamics are going on between every pair of animals in a group where the fighting abilities are drawn from some distribution.
@@
 \begin{figure}
-\includegraphics[width=6in]{distributions_of_signals_received_uniform.pdf}
-\caption{The probability distribution of number of signalers, assuming a uniform distribution of fighting abilities, as derived in Eq. (\ref{STAR}).  \label{derived_sig_probs}}
+\includegraphics[width=6in]{distributions_of_signals_received.pdf}
+\caption{The probability distribution of number of signalers.
+The lines represent the analytically derived distribution Eq.
+(\ref{STAR}), assuming a uniform distribution of fighting abilities.
+The dots represent probabilities from simulated signaling networks where fighting abilities are drawn from a normal distribution.
+\label{derived_sig_probs}}
 \end{figure}
 
-\section{Simulating the Stochastic Process }
-A simple updating algorithm allows us to simulate the process as described by the model above (ORNSTEIN-UHLENBECK VARIANT):
+\section{Parameter Optimization for Individuals }
+The only case where the parameters really affect the dynamics is when the system exhibits bistability.
 
-\noindent Fix $\Delta t>0$.  Set $t=0$ and initialize $X_0^{(1)}=X_0^{(2)}=0$.
+\subsection{Minimizing Error Rate or Decision Time }
 
-\begin{enumerate}
-\item Draw $Z_1,\dots,Z_5$ i.i.d. from $N(0,1)$.
-\item \begin{align*}
-X_{t+\Delta t}^{(1)}&=X_t^{(1)}+\left(m_{e_1}X_t^{(1)}+m_f+b_{s_2}f(X_t^{(2)}\right)\Delta t+n_f\sqrt{\Delta t}Z_1+n_{e_1}\sqrt{\Delta t}Z_2+b_{s_2}\sqrt{f(X_t^{(2)}}\sqrt{\Delta t}Z_3\\
-X_{t+\Delta t}^{(2)}&=X_t^{(2)}+\left(m_{e_2}X_t^{(2)}+m_f+b_{s_1}f(X_t^{(1)}\right)\Delta t+n_f\sqrt{\Delta t}Z_1+n_{e_2}\sqrt{\Delta t}Z_4+b_{s_1}\sqrt{f(X_t^{(1)}}\sqrt{\Delta t}Z_5
-\end{align*}
-\item $t=t+\Delta t$
+\subsection{Optimizing Response to Changes in Fighting Ability }
 
-\end{enumerate}
+\section{Effects of Individual Optimization on the Signaling Network }
 
-To confirm that our simulation is performing reasonably, we let the model run without signaling feedback until the estimates reach stationary distributions and compare these to the analytical prediction, as given by (\ref{ou_stat1}) and (\ref{ou_stat2}).
-This comparison is shown in Figure \ref{ou_stat_dist}.
-As can be seen in the figure, the observed probability distribution of estimates is statistically indistinguishable from the prediction (for the first, K-S test $D=0.0243$, $p=0.5969$ and for the second, $D=0.0211$, $p=0.766$).
+\subsection{Mutual Information between Fight Distribution and Signaling Network }
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=.65\textwidth]{ou_stat_distribution.pdf} \end{center}
-\caption{\label{ou_stat_dist} A comparison between the stationary distribution from a simulation of the Ornstein-Uhlenbeck model and the analytically predicted stationary distribution.}
-\end{figure}
-
-We are interested in the effect of signaling feedback, i.e.
-the $b_s$ parameter, on the dynamics of the model.
-We run the algorithm $1000$ times so that we obtain, at each point in time, a distribution of each animal's estimates.
-We then compare these distributions when there is and isn't signaling.
-In both regimes, each animal's estimates reach a relatively stationary distribution fairly quickly so that we can pick a typical time point at which the distributions are not changing any more.
-(A Kolmogorov-Smirnov test for the differences between distributions at different points in time does not allow us to reject the null hypothesis that the distributions are the same, at a significance level of $p=0.05$.)  If we allow the simulation to reach a stationary state, we find that the stationary distributions with and without signaling feedback ($b_{s_1}=b_{s_2}=0$ and $b_{s_1}=b_{s_2}\gneq 0$) are quite different for Animal 1(K-S test statistic, $D=0.995,p<2.2\times10^{-16}$), whereas they are statistically indistinguishable for Animal 2 ($D=0.048, p=0.1995$) (Figure (\ref{diff_dists})).
-
-By comparing distributions of the estimates at a given point in time, we compare the behavior of the stochastic process over many realizations.
-We can also compare the behavior of individual trajectories over time.
-We hypothesize that by increasing the strength of signaling feedback, we should increase the stability of signaling behavior and it should be less likely that the dominant animal's estimate sink to such an extent that it becomes subordinate.
-To test this hypothesis, we count, for each run of the simulation, the number of switches between dominance regimes.
-A dominance regime is defined by one animal's estimate being below the signaling threshold $-T$ and the other animal's estimate being above the positive threshold $T$.
-If a dominance regime is established with Animal $i$'s estimate above the threshold (dominant) and Animal $i+1$'s estimate below the threshold (subordinate) and subsequently a dominance regime is established with Animal $i$ subordinate and Animal $i+1$ dominant, we count this as one switch.
-Fixing the other parameters of the model, for each difference in fighting ability $m_f$ and level of signaling feedback $b_{s_1}=b_{s_2}$, we can find the average number of switches that occur in a run of the model.
-As the difference in fighting ability increases fewer switches occur since Animal 1 is dominant the vast majority of the time.
-Additionally, the average number of switches between dominance regimes decreases as $b_{s}$ increases, and the effect is greater for when the difference in fighting ability is smaller (Figure \ref{diff_switches}).
-
-\begin{figure}
-%\begin{subfigure}{.5 \textwidth}
-\begin{center}
-\includegraphics[width=.85\textwidth]{feedback_estimate_distributions_v2.pdf}
-\end{center}
-%\caption{}
-%\end{subfigure}
-\caption{\label{diff_dists} The distribution of estimates for each animal at time $200$, for no signaling feedback and for signaling feedback $b_{s_1}=b_{s_2}=3$.
-Other parameters are set as follows: threshold $T=5$, $m_{e_1}=m_{e_2}=-.5$, $n_{e_1}=n_{e_2}=1$, $m_f=5$, $n_f=.5$.
-}
-\end{figure}
-
-\begin{figure}
-\begin{subfigure}{1 \textwidth}
-\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v2.pdf}
-\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v5.pdf}
-\caption{\label{low_error}}
-\end{subfigure}
-\begin{subfigure}{1 \textwidth}
-\includegraphics[width=.5\textwidth]{feedback_stability_of_dominance_v4.pdf}
-\caption{\label{high_error}}
-\end{subfigure}
-%\begin{subfigure}{.5 \textwidth}
-%\includegraphics[width=1\textwidth]{feedback_stability_of_dominance_v2.pdf}
-%\caption{}
-%\end{subfigure}
-\caption{\label{diff_switches} The average number of switches between dominance regimes as a function of the difference in fighting ability, $m_f$, between the animals and the degree to which receiving a signal affects an animal's estimate of its dominance, $b_{s_1}=b_{s_2}$.
-In general, even for small differences in fighting ability there are very few switches in dominance regimes.
-However, fighting ability can be made small enough to allow switches to occur, in which case the strength of signaling feedbacks affects how frequently they happen.
-When the leak rate is relatively low, increases in signaling feedback decrease the number of switches (\ref{low_error}).
-When the error rate is high / estimates leak back to $0$ very quickly, increasing the signaling feedback actually increases the number of switches between dominance regimes / decreases the stability of a dominance regime (\ref{high_error}).}
-\end{figure}
-
-\section{Linear Learning Formulation}
-
-So far, we have considered how the processes of engaging in fights and sending and receiving signals might affect the estimates two animals make of each other's dominance.
-We are interested in understanding these processes on the population level and therefore would like to embed these dynamics in a model of a social network.
-In order to make this more analytically tractable, we modify our model into a linear learning model and, for the moment, ignore feedback from signaling into the learning process.
-Our main question is how the structure of a social network can facilitate individuals' learning about the states of other nodes in the network.
-
-Suppose each individual has a true fighting ability $a_1,\dots,a_N$ and let $f_{jk}=a_j-a_k$.
-Consider an individual in a social network observing its own interactions with other members of the group and the interactions between other members of the group and trying to learn each individual's true fighting ability.
-Let $x_j(t)$ be the individual's estimate of individual $j$'s fighting ability at time $t$.
-The focal individual will be able to learn about other individual's fighting ability by engaging in fights with individuals it interacts with and observing fights between pairs of individuals both of whom it interacts with.
-Let
-$$I_{jk}=\left\{
-\begin{array}{cccc}
-1 & , & \text{ if the focal individual interacts with both $j$ and $k$}\\
-0 & , & \text{ otherwise}
-\end{array}\right..
-$$
-Now, $x_j$ will change according to
-\begin{align*}
-\dot{x}_j(t)&=\sum_kI_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right] \text{ for } j\lneq N
-\\ \text{ and } x_N(t)&=T-\sum_{j\lneq N}x_j(t),
-\end{align*}
-where we impose a total on the estimates so that they do not diverge.  Therefore,
-\begin{align*}
-\dot{x}_j(t)&=\sum_{k\neq N}I_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right]+I_{jN}\left[f_{jN}-(x_j(t)-x_N(t))\right]
-\\&=\sum_{k\neq N}I_{jk}\left[f_{jk}-(x_j(t)-x_k(t))\right]+I_{jN}\left[f_{jN}-(x_j(t)-(T-\sum_{k\neq N}x_k(t)))\right]
-\\&=\sum_kI_{jk}f_{jk}-x_j(t)\sum_{k}I_{jk}+\sum_{k\neq N}I_{jk}x_k(t)+TI_{jN}-\sum_{k\neq N}I_{jN}x_k(t)
-\\&=\sum_kI_{jk}f_{jk}+TI_{jN}-x_j(t)\sum_kI_{jk}+\sum_{k\neq N}(I_{jk}-I_{jN})x_k(t)
-\\&=C_j+\sum_{k}L_{jk}x_k(t) \text{ where } C_j=\sum_{k\neq N}I_{jk}f_{jk}+TI_{jN} \text{ and } 
-\end{align*}
-$L_{jk}$  is an $n-1\times n-1$ matrix given by
-$$
-L_{jk}=\left\{
-\begin{array}{cccc}
--\sum_lI_{jl}+I_{jj}-I_{jN} & , & \text{ if }j=k
-\\ I_{jk}-I_{jN} & , & \text{ if } j\neq k
-\end{array}\right..
-$$
-If $x_k(t)=a_k+\frac{T-\sum_ka_k}{N}$  for all $k$, $\dot{x}(t)=0$.  (If $T=\sum_ka_k$ then $x_k(t)=a_k$ is an equilibrium.)
-
-Stability of the equilibrium is determined by the eigenvalues of $L$.
+\subsection{Timescale Separation }
 
@@
 \begin{itemize}
-\item We know that if signaling is constantly present or absent, the stationary distribution of the Ornstein-Uhlenbeck process is just shifted and the variance does not change.
-However, we do not know what the likelihood of switching from non-signaling to signaling behavior (or vice versa) and whether the estimates would stay in a non-signaling / signaling regime long enough for a stationary distribution to be established.
+\item expected time for an animal to start signaling
+\item probability of one or the other animal to be the first to start signaling
+\item optimize error rate / decision time with respect to the threshold for each pair and then see how that affects the resulting distribution of signaling behavior
 \end{itemize}

### 2014-10-22 paper_skeleton.tex
ive added code for tracking mutual information over time and changed figures a bit

@@
 
-\tableofcontents
+%\tableofcontents
 
 \section{Abstract}
-Learning usually occurs in noisy environments.
+%Learning usually occurs in noisy environments.
 In order to make a decision, one must accumulate the  evidence for different alternatives and choose one only once the uncertainty about the choice is sufficiently reduced.
@@
 Each animal's opinion about its relative dominance, therefore, will perform a biased random walk until one animal's opinion of itself is low enough to agree to being subordinate.
-
-Similar models have previously been applied to animal conflicts, although in that case the model assumed that the animals had perfect memory of their past interactions \citep{Froment:2010fk}
-
-\subsection{Extent and Limits of Analogy }
-
-In Table \ref{variables}, we list the inputs, decision variables, and outputs of the two systems.
-
-%In the neural system, an individual wants to make both a decision that is both quick and accurate.
-This is also true in the social system.
-However, it may also be the case that a monkey wants to avoid emitting a subordination signal and wants to receive a signal, regardless of whether or not this is the ``correct" outcome, i.e.
-whether or not he is actually more likely to win a fight.
-The criterion used to optimize the threshold, therefore, may be quite different in the two systems.
-Additionally, the speed and outcome of the decision process depend on both individuals' thresholds.
-Since an individuals' best behavior depends on the behavior of the other animal, optimizing the threshold has a strategic aspect.
+A similar model has previously been applied to animal conflicts, although in that case the model assumed that the animals had perfect memory of their past interactions \citep{Froment:2010fk}.
 
@@
 
-%note about applicability to quorum sensing?!?
+%note about applicability to quorum sensing
 
@@
 The assumptions about the timescales on which inputs occur are particularly reasonable in the social system, although the successful application of this type of model to neural populations implies they are not unreasonable in that system.
+In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the neural and social systems.
 
-%\begin{equation}
-%\begin{array}{ll}
-%dX_1&=\bigg(-\ell X_1(t)+b(2d-1)\bigg)dt+b\sqrt{d}dW_{1}t-b\sqrt{(1-d)}dW_{2}t
-%\\dX_2&=\bigg(-\ell X_1(t)+b(1-2d)\bigg)dt-b\sqrt{d}dW_{1}t+b\sqrt{(1-d)}dW_{2}t,
-%\end{array}
-%\end{equation}
 \subsection{Reaching a decision }
@@
 The social system is therefore inherently two-dimensional, whereas the neural system may be adequately described by one dimension.
+In Table \ref{differences}, we compare how the decision is made and how the decision process is optimized in the two systems.
 
@@
 We therefore numerically solve the PDEs in the two-dimensional case.
-In Table \ref{models}, we compare how the decision is made and how the decision process is optimized in the two systems.
 
@@
 Consensus about the value of each individual in a group emerges out of the decision process.
-In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Flack:2004oq,Flack:2006uq,Brush:2013fk}.
+In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq}.
 Given the decisions that each pair of individuals reaches, there are measures of the consensus in the group that successfully predict an animal's social power and also extend to other social systems \cite{Brush:2013fk}.
@@
 
-\subsection{Prioritizing signaling probability does not always decrease pairwise decision accuracy. }
+\subsection{Prioritizing signaling probability does not always decrease pairwise decision accuracy. Intermediately difficult decisions sometimes take the longest. }
 We first look at the Nash equilibrium thresholds when the animals only care about one property of the decision (error rate, signaling property, or decision time).
@@
 The thresholds affect the accuracy and decision time with which a pair can reach a decision.
-For a difficult decision ($c$ close to $.5$), the accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and lowest when only signaling probability matters ($w_3=1$).
-Accuracy decreases as the weight given to decision time increases (Figure \ref{nasheq_thresholds}).
-Since, for sufficiently high waiting costs, the Nash thresholds do not depend on the tradeoff between error rate and signaling probability, neither does the accuracy for a pair using Nash thresholds.
-In other words, once waiting costs are high, a pair of animals each minimizing error rate and a pair minimizing signaling probability will reach a decision with the same accuracy.
-
-\subsection{Intermediately difficult decisions sometimes take the longest. }
+For a difficult decision ($c$ close to $.5$), the accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and lowest when only decision time matters ($w_2=1$).
+When waiting costs are low, accuracy decreases as the weight given to signaling probability increases (Figure \ref{nasheq_thresholds}).
+For sufficiently high waiting costs, the Nash thresholds do not depend on the tradeoff between error rate and signaling probability, so neither does the accuracy for a pair using Nash thresholds.
+In other words, once waiting costs are high, a pair of animals minimizing error rate and a pair each of whom minimizes his signaling probability will reach a decision with the same accuracy.
 
@@
 
-\subsection{Decisions in a group context take longest between animals with similarly high abilities. }
+\subsection{Average group accuracy increases as signaling probability becomes more important. Decisions in a group context take longest between animals with similarly high abilities. }
 
-After analyzing Nash equilibrium thresholds between pairs of animals, we found the optimal thresholds for a group of animals ($N=20$) such that no animal had an incentive to choose a different threshold.
+To study decision making in a group, we find the optimal thresholds for a group of animals ($N=20$) such that no animal had an incentive to choose a different threshold.
 As with pairwise Nash thresholds, when the animals only care about signaling probability ($w_3=1$), all animals in the group choose high thresholds, whereas if they only care about decision time ($w_2=1$), all animals choose low thresholds.
@@
 
-When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with less similar abilities also take a fairly long time.
-When waiting costs are high, most pairs can reach a decision quite quickly, except those who have similar and high abilities (Figure \ref{groupeq_thresholds}).
 The average accuracy decreases as waiting costs increase (Figure \ref{groupeq_thresholds}).
-
 As soon as any waiting costs are introduced, as with the pairwise Nash thresholds, the weaker animals have incentives to lower their thresholds to make quicker decisions at the expense of accuracy.
@@
 The mutual information of any measure of consensus in the signaling network increases (more or less) with accuracy so it too increases as signaling probability becomes more important.
+%% hit harder that different parts of the group are making the tradeoff differently and that accuracy really does increase!!
+that's a difference between pairs and groups!!!
+and hit harder in the discussion too.
+
+When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with less similar abilities also take a fairly long time.
+When waiting costs are high, all pairs reach decision quickly.
+At intermediate costs, most pairs can reach a decision quite quickly, except those who have similar and high abilities (Figure \ref{groupeq_thresholds}).
 
@@
 \section{Discussion}
-what's interesting and exciting and new?
 
-we can derive sde's from first principles
+Starting with a simple and reasonable description of how fights occur between pairs of animals in a social group and how the outcomes of those fights affect the animals' opinions of their relative dominance, we derived  a model in which those opinions change according to stochastic differential equations.
+A nearly identical model has previously been used to describe decision making in primate brains where the decision is determined by the firing rates of two neural populations.
+The model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two animals are the same.
+Conflict and strategy are not usually ascribed to the neural model and the war of attrition is not usually thought of as a process through which the pair of competitors reaches a decision about an external environment.
+The similarities between these three frameworks-- social decision making, neural decision making, and a game-- suggest that there are common principles of decision making through conflict that are applicable to a number of systems.
 
-prioritizing signaling probability sometimes makes decision more accurate
+In the kind of leaky integrator models we studied here, either the absolute value of the decision variables or the difference in their values can determine the outcome of the decision.
+We find that, for a given decision time, the accuracy of the decision does not depend on which of these mechanisms is used.
+In order for the difference in variables to be used, a third party arbiter needs to be present to evaluate the value of each decision variable and find their difference.
+In the brain, it might be the case that there is a third neural population whose activity depends on the two populations gathering input, but in a social group, there is not a third party informing how a pair of animals should establish their subordination-dominance relationship.
+The fact, then, that the model does not depend on the existence of this third party arbiter means that it can be equally well applied whether or not it is present.
 
-prioritizing waiting costs change which decisions take the longest
+Any method for making a decision reaches each output with a given probability, reaches the ``correct" output with an error rate, and makes a decision in an expected decision time.
+If each individual involved in the decision prefers a different output, the pair is in conflict and choosing appropriate strategies has a strategic aspect.
+The individuals can make a tradeoff between how valuing their personal preference and valuing the correct output.
+We expected to find that individuals valuing their personal preference at the expense of the correct output would tend to make decisions that were less accurate.
+However, if decision time is sufficiently important, than the tradeoff between the other two variables does not affect the strategies the pair chooses.
+In fact, when there is a group of animals all making this tradeoff, increasing the importance of personal preferences over accuracy improves the accuracy with which animals in the bottom of the group resolve their differences, increasing the average accuracy of the whole group.
+It is usually assumed that neural populations in the brain are in agreement about the desired outcome, since the whole organism receives a benefit from making a correct decision in a short time.
+But it is not implausible that different neural populations would prefer to make different decisions and we find that introducing this conflict into the model of decision making could actually improve overall accuracy.
 
-optimization weights change the overall social structure
+In a pair of individuals making a decision, the difference in their abilities, not their absolute values, determines how quickly they can reach a decision.
+When waiting costs are high or low, more difficult decisions take longer to make.
+However, at intermediate waiting costs we find that it takes the longest to reach decisions of intermediate difficulty.
+The waiting costs affect pairs making difficult decisions and pairs making intermediate decisions differently, such that a pair making a difficult decision does better off by ignoring the output of the decision and just reaching the decision quickly.
 
-extension of neural model by removing third party comparison and allow individuals to minimize signaling probability
+The costs of decision time, error rate, and signaling probability also affect the decision times of decisions in a group context.
+In a group context, each pair of animals experiences the difference in their abilities, but their absolute positions within the group also matter.
+In this context, at intermediate waiting costs, the longest decisions are those between pairs who have similar and high abilities.
+In data from a captive population of macaques, we observe that pairs of animals with high abilities \emph{do} reach agreements about their relative dominance but that pairs of animals with low abilities sometimes do not reach such agreements.
+This indicates that our simple model is missing a level of social complexity that could explain this empirical pattern.
 
-extension of game theory by introducing heterogeneous abilities
+There are (at least) two modifications to the model that we expect would produce results closer to the empirical observations.
+First, in the model, the rates at which animals fight each other do not depend on the abilities of the animals.
+However, it is likely that animals with lower abilities are less mobile and encounter their peers at a lower rate than animals with higher abilities who are more free to range throughout the group.
+In the model, this heterogeneity in encounter rate would make the decisions between animals with low encounter rates take longer.
+Second, in the model, all animals value signaling probability, error rate, and waiting costs equally.
+However, it is likely that the benefits from receiving a subordination signal from a partner and the costs of fighting depend on the ability on both animals' ability.
+If animals with low ability  do not value an agreement with other animals with similar abilities, there would be less incentive for them to reach an agreement.
 
-extension of David's paper
+With the goal of keeping the model simple and interpretable, we did not allow parameters like encounter rates and optimization weights to vary across the group of animals, although they are likely to in the empirical system.
+Even with the small number of parameters we did consider, we find systematic variation in group-level properties that are observable in the data.
+We find that the skewness of the distribution of power scores, given by a number of measures of power \cite{Brush:2013fk,Flack:2006uq}, is maximized at intermediate waiting costs.
+In our data, we observe a right-skewed distribution of power.
+To the extent that our model recapitulates the empirical system, this would suggest that waiting costs are neither the only priority nor are they ignored completely.
 
-explanation for variation across species of macaque
+In a more complicated model, the ``aggressiveness'' of animals moving in a spatially explicit area affected the shape of the dominance hierarchy \cite{Hemelrijk:2011fk}.
+While the mechanisms of that model are quite different, ``aggressiveness'' could be argued to be similar to disregarding waiting costs and prioritizing signaling probability.
+Neither the authors of that work nor we found that a difference in the underlying distribution of fighting abilities is required to change the distribution of dominance or power.
+In other words, in both models, the collective computation that results in a power structure emerges out of how the individuals interact and collect information rather than from the distribution of fighting abilities.
 
-evolutionary implications
+The variation in skewness of the power distribution is interesting because the power structures typical to different species of macaque vary in their average skewness \cite{Flack:2004oq,Preuschoft:2004ly,Waal:1985fk}.
+One explanation might be that the distribution of fighting abilities is different from species to species, for instance because there is more or less variation in body size.
+However, our model suggests that even if the underlying distribution of fighting abilities are identical, the priorities given to signaling probability, error rate, and waiting costs can affect the distribuiton of power.
+The benefit of having a partner agree to be subordinate, the benefit or reaching the ``correct" agreement, and the costs of fighting may vary from species to species, causing variation in the shape of the power structure.
 
-%comparison to social data
-Empirically, those pairs of monkeys that take the longest to establish a signaling dynamics are those with the most similar fighting abilities...
+The other feature of the signaling network that changes as a function of the optimization weight is which measure of power is most informative about the underlying fighting abilities.
+There are many measures of a network that might seem like reasonable ways to measure the consensus about each individual's ability.
+As network scientists, we would like to know when which measure is the most informative and we find that it depends on how accurately a decision between paris reflects the difference in their abilities.
+If we expect relatively high but imperfect pairwise accuracy then a global metric like eigenvector centrality is the best.
+As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities.
+In previous work, we used both eigenvector centrality and measures of the number of signals received to measure the consensus about node status in empirical networks and found that both were produced scores that were informative about the nodes' functional behavior.
+This suggests...
+%%%
 
-skewness
-they're using eigenvector centrality
+%extension of game theory by introducing heterogeneous abilities
+%
+%extension of David's paper
+%
+%evolutionary implications
+%
+%they're using eigenvector centrality
 
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/pairwise_accuracy.pdf}
-\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/nasheq_thresholds.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Dropbox/signaling_network/pairwise_accuracy.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Dropbox/signaling_network/nasheq_thresholds.pdf}
 \caption{\label{nasheq_thresholds}  The accuracy of a pair using Nash equilibrium thresholds decreases as the weight given to decision time increases.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/groupaccuracy_heatmap.pdf}
-\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/groupeq_thresholds.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Dropbox/signaling_network/groupaccuracy_heatmap.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Dropbox/signaling_network/groupeq_thresholds.pdf}
 \caption{\label{groupeq_thresholds}   The accuracy of a group using optimal thresholds decreases as the weight given to decision time increases.
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/skewness_heatmap.pdf}
-\caption{\label{skewness} The skewness of the distribution of the number of signalers is maximized at intermediate waiting costs.
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Dropbox/signaling_network/skewness_heatmap.pdf}
+\caption{\label{skewness} The average skewness of the distribution of the number of signalers is maximized at intermediate waiting costs.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{/Users/eleanorbrush/Desktop/mostinformative_heatmap.pdf}
+\includegraphics[width=\textwidth]{/Users/eleanorbrush/Dropbox/signaling_network/mostinformative_heatmap.pdf}
 \caption{\label{bestmetric}  Which measure of consensus is most informative about fighting ability depends on how accurate decision are on average.
@@
 \ra{1.3}
-\begin{tabular}{@{}rllll@{}}
+\begin{tabular}{@{}lllll@{}}
 Variable & Interpretation & Neural &   Social \\
@@
 & input & dots moving left or right & fights won or lost
-\\$c$ & strength of input & coherence of dots & probability of stronger animal winning
+\\$c$ & strength of input & coherence of dots & prob. of stronger animal winning
 \\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
-\\ & dist. of states of the world & left / right equally likely & distribution of fighting abilities
-\\ & decision & subject indicates left or right & one animal emits a subordination signal
+\\ $w_1$ & error rate weight & reward from being ``right" & reward from being ``right"
+\\ $w_2$ & decision time weight & penalty for taking a long time & costs of fighting
+\\ $w_3$ & signaling prob. weight & ? & benefit from receiving signal
+\\ & decision & subject indicates left or right & one animal emits subordination signal
 \\ & correct decision & subject chooses correct direction & weaker animal signals
@@
 \centering
-\caption{\label{models}{\bf  Comparison of models applied to neural and social systems.} Differences are highlighted in red.}
+\caption{\label{differences}{\bf  Comparison of models applied to neural and social systems.} Differences are highlighted in red.}
 \ra{1.3}
-\begin{tabular}{@{}rllll@{}}
+\begin{tabular}{@{}lllll@{}}
 & \multicolumn{2}{c}{Neural} &  Social \\
 \cmidrule{2-3} \cmidrule{4-4} 
-%model  & DDM or OU & race && race with leak
-%\\
 dimensionality & $1$  && $2$

### 2014-10-22 acoeffunction.m

@@ -1,0 +1,20 @@
+function A = acoeffunction(p,t,l,a)
+
+N = 3; % Number of equations
+% Triangle point indices
+it1=t(1,:);
+it2=t(2,:);
+it3=t(3,:);
+
+% Find centroids of triangles
+xpts=(p(1,it1)+p(1,it2)+p(1,it3))/3;
+ypts=(p(2,it1)+p(2,it2)+p(2,it3))/3;
+
+nt = size(t,2); % Number of columns
+A = zeros(N^2,nt); % Allocate a
+
+% Now the particular functional form of f
+A(4,:) = 1;
+A(7,:) = 1;
+A(6,:) = -l*xpts+a;
+A(9,:) = -l*ypts-a;

### 2014-10-22 average_threshold.m

@@ -1,0 +1,115 @@
+function threshvec=average_threshold(c1,c2,c3,its,N,dist)
+
+if nargin==3
+    N=30;
+    its=25;
+end
+
+Nl=evalin('base','Nl');
+Nd=evalin('base','Nd');
+Nt=evalin('base','Nt');
+twomat=evalin('base','twomat');
+domvals=evalin('base','domvals');
+threshvals=evalin('base','threshvals');
+
+Nb1=10;
+Nb2=10;
+infomat=zeros(Nb1,Nb2);
+skewvals=zeros(1,its);
+finalthresh=zeros(N,its);
+for c=1:its
+
+    opt_its=10;
+
+    fighting_abilities=zeros(1,N);
+    
+    switch dist
+        case 'unif'
+            fighting_abilities=20*rand(1,N);
+        case 'norm'
+            fighting_abilities=normrnd(0,10,1,N);
+        case 'lognorm'
+            fighting_abilities=lognrnd(0,1,1,N);
+    end
+    
+    fighting_abilities=sort(fighting_abilities,'descend');
+
+    dmat=zeros(N); %turn difference in fighting abilities into probabilitise of winning
+    extendeddomvals=[1-domvals(end:-1:2) domvals];
+    v=1:(Nd+Nd-1);
+    deps=.05;
+
+    for i=1:N
+        for j=[1:(i-1),(i+1):N]
+            diff=fighting_abilities(i)-fighting_abilities(j);
+            d=exp(diff)/(exp(diff)+1);
+            d=floor(round(d/deps))*deps;
+            dmat(i,j)=v(abs(d-extendeddomvals)<=deps/2);
+        end
+    end
+
+    %size(twomat)=[Nl,Nd,Nt,Nt,2];
+    
+    perf=zeros(2,Nl,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+    perf(1,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2))+c3*(1-twomat(:,2:Nd,:,:,1));
+    perf(2,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2))+c3*(twomat(:,2:Nd,:,:,1));
+
[…]

### 2014-10-22 calcCMatPiezoActuator.m

@@ -1,0 +1,34 @@
+function c = calcCMatPiezoActuator( p, t, c11, c12, c22, c13, c23, c33 )
+%CALCCMATPIEZOACTUATOR C-matrix for piezoelectric actuator example
+%   c = CALCCMATPIEZOACTUATOR( p, t, c11, c12, c22, c13, c23, c33 )
+%   returns the 'c' coefficient matrix for the piezoelectric actuator
+%   example given the point and element matrices along with the
+%   constitutive submatrices (cij) for the PVDF material.
+
+% Copyright 2012 The MathWorks, Inc.
+
+numElems = size(t,2);
+c=zeros(21,numElems);
+%
+% Although the material in both layers is PVDF, in the top layer
+% the polarization direction points down (minus y direction) and in the
+% bottom layer, it points up. That is, the top layer has d-coefficients
+% that are the negative of those in the bottom layer.
+%
+% The code below examines the y-location of the centroid of each
+% triangular element and assigns the correct material properties to
+% element depending on whether it is in the top or bottom layer.
+%
+ctop = [c11(:); c12(:); c22(:); -c13(:); -c23(:); -c33(:)];
+cbot = [c11(:); c12(:); c22(:);  c13(:);  c23(:); -c33(:)];
+% calculate y-coordinate of triangle centers
+yCenter=(p(2,t(1,:)) + p(2,t(2,:)) + p(2,t(3,:)))/3;
+for i=1:numElems
+  if(yCenter(i) < 0)
+    c(:,i) = cbot;
+  else
+    c(:,i) = ctop;
+  end
+end
+
+end

### 2014-10-22 checking_the_paradox_of_mutual_info.m

@@ -10,7 +10,7 @@
 % c13=0;c23=.3;c33=.7;
 time=0.05;
 c11=0;c21=time;c31=1-time;
-c12=1-time;c22=time;c33=0;
+c12=1-time;c22=time;c32=0;
 
 cvals=[c11 c21 c31; c12 c22 c32];
 % cvals=[c12,c22,c32];
@@ -36,17 +36,18 @@ dmat=reshape(dmatssaved(c,:,:),N,N);
 
         perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
 
-        perf(1,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(1-twomat(l,2:Nd,:,:,1));
-        perf(2,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(twomat(l,2:Nd,:,:,1));
+        perf(1,2:Nd,:,:)=c1*(1-twomat(leak,2:Nd,:,:,1))+c2*(twomat(leak,2:Nd,:,:,2))+c3*(1-twomat(leak,2:Nd,:,:,1));
+        perf(2,2:Nd,:,:)=c1*(1-twomat(leak,2:Nd,:,:,1))+c2*(twomat(leak,2:Nd,:,:,2))+c3*(twomat(leak,2:Nd,:,:,1));
 
-        perf(1,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(1-twomat(l,1,:,:,1));
-        perf(2,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(twomat(l,1,:,:,1));
+        perf(1,1,:,:)=c2*(twomat(leak,1,:,:,2))+c3*(1-twomat(leak,1,:,:,1));
+        perf(2,1,:,:)=c2*(twomat(leak,1,:,:,2))+c3*(twomat(leak,1,:,:,1));
 
         Tvals=zeros(N,opt_its);
 %         Tvals(:,1)=2*ones(N,1);
         % Tvals(:,1)=randi([1 Nt],N,1);
-        Tvals(:,1)=reshape(seqoptthresh(c,ic,:),[],1);
-        Tvals(3,1)=seqoptthresh(c,otheric,3);
+%         Tvals(:,1)=reshape(seqoptthresh(c,ic,:),[],1);
[…]

### 2014-10-22 compare_nasheq.m

@@ -1,0 +1,22 @@
+comp=zeros(Nc,Nd,2,2);
+
+for i=1:Nc
+    for j=1:Nd
+        n=nasheq_ind{i,j};
+        if isempty(n)
+            n=[0; 0];
+        end
+        n2=nasheq_ind2{i,j};
+        if isempty(n2)
+            n2=[0;0];
+        end
+        comp(i,j,1,:)=n;
+        comp(i,j,2,:)=n2;
+    end
+end
+%%
+j=2;
+plot(col(comp(:,j,1,2)),'-o')
+hold on
+plot(col(comp(:,j,2,2)),'r')
+hold off

### 2014-10-22 compare_to_tech_rept.m

@@ -1,0 +1,12 @@
+function [theirs mine]=compare_to_tech_rept(z,d,l,b,deltax)
+
+ztilde=z/(2*b*(2*d-1));
+atilde=(2*d-1)^2;
+lambda=-l;
+
+theirs=(erf(sqrt(atilde/lambda)*(1+lambda*ztilde))-erf(sqrt(atilde/lambda)))/(erf(sqrt(atilde/lambda)*(1+lambda*ztilde))-erf(sqrt(atilde/lambda)*(1-lambda*ztilde)));    
+
+[x,~,~]=solve_pde_1d(-z,d,z,deltax,l,b);
+mine=x;
+
+end

### 2014-10-22 confusion1.m

@@ -1,0 +1,36 @@
+perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+perf(1,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(1-twomat(l,2:Nd,:,:,1));
+perf(2,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(twomat(l,2:Nd,:,:,1));
+
+perf(1,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(1-twomat(l,1,:,:,1));
+perf(2,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(twomat(l,1,:,:,1));
+
+perfvals=zeros(N,1);
+indperfvals=zeros(N,N);
+for i=1:N
+    opp_thresh=Tvals([1:(i-1),(i+1):N]);
+    ds=dmat(i,[1:(i-1),(i+1):N]);
+    perfsum=0;
+            for q=1:(N-1)
+                if ds(q)<=Nd-1
+                    perfsum=perfsum+perf(2,Nd-ds(q)+1,opp_thresh(q),Tvals(i)); 
+                else 
+                    perfsum=perfsum+perf(1,ds(q)-Nd+1,Tvals(i),opp_thresh(q));
+                end
+            end  
+    perfvals(i)=perfsum;
+    for j=[1:(i-1), (i+1):N]
+        ds=dmat(i,j);
+        if ds<=Nd-1
+            indperfvals(i,j)=perf(2,Nd-ds+1,Tvals(j),Tvals(i)); 
+        else 
+            indperfvals(i,j)=perf(1,ds-Nd+1,Tvals(i),Tvals(j));
+        end
+    end
+end
+
+[probmat, timemat, accmat]=thresh2mat(Tvals,twomat,dmat,l);
+
+v=c1*(1-accmat)+c2*timemat+c3*(1-probmat);
+v=sum(v,2);

### 2014-10-22 confusion2.m

@@ -1,0 +1,87 @@
+its=2;
+N=20;
+dist='unif';
+extendeddomvals=[1-domvals(end:-1:2) domvals];
+v=1:(Nd+Nd-1);
+deps=.05;
+l=1;
+opt_its=50;
+
+% c11=.4;c21ç=.6;c31=0;
+% c12=.9;c22=.1;c32=0;
+% c11=0;c21=.4;c31=.6;
+% c12=.9;c22=.1;c32=0;
+% c11=0;c21=.55;c31=.45;
+% c12=.85;c22=.2;c32=.05;
+
+c11=0;c21=.3;c31=.7;
+c12=.7;c22=.3;c32=0;
+c13=0;c23=.3;c33=.7;
+
+cvals=[c11 c21 c31; c12 c22 c32];
+% cvals=[c12,c22,c32];
+% cvals=[c11 c21 c31];
+totalc=size(cvals,1);
+
+seqoptthresh=zeros(its,totalc,N);
+dmatssaved=zeros(its,N,N);
+perfvalssaved=zeros(its,totalc,N);
+
+for c=1:its
+
+    fighting_abilities=20*rand(1,N);
+    fighting_abilities=sort(fighting_abilities,'descend');
+
+    dmat=zeros(N);
+    
+    for i=1:N
+        for j=[1:(i-1),(i+1):N]
+            diff=fighting_abilities(i)-fighting_abilities(j);
+            d=exp(diff)/(exp(diff)+1);
+            d=floor(round(d/deps))*deps;
+            dmat(i,j)=v(abs(d-extendeddomvals)<=deps/2);
+        end
+    end
+
+    dmatssaved(c,:,:)=dmat;
+    
+    for ic=1:totalc
+ 
+        c1=cvals(ic,1);
+        c2=cvals(ic,2);
+        c3=cvals(ic,3);
+        
+        Tvals=zeros(N,opt_its);
+        Tvals(:,1)=2*ones(N,1);
+
+        count=1;
+        while count<=opt_its
+            for i=1:N
+                testthreshvals=Tvals(:,count);
+                perftest=zeros(Nt,1);
+                for j=1:Nt
+                    testthreshvals(i)=j;
+                    testperfvec=thresh2perf([c1 c2 c3],testthreshvals,twomat,dmat,l);
+                    perftest(j)=testperfvec(i);
+                end
+                
+                [p,n]=min(perftest);
+                Tvals(i,count+1)=n;  
+                
+            end
+            
[…]

### 2014-10-22 data2norm.m

@@ -1,0 +1,14 @@
+function [normx,normy]=data2norm(x,y)
+hAxes=gca;
+x=[x,x];
+y=[y,y];
+scribepin = scribe.scribepin('parent',hAxes,'DataAxes',hAxes,'DataPosition',[x;y;[0,0]]');
+figPixelPos = scribepin.topixels;
+hFig = ancestor(hAxes,'figure');
+figPos = getpixelposition(hFig);
+figPixelPos(:,2) = figPos(4) - figPixelPos([2,1],2);
+figNormPos = hgconvertunits(hFig,[figPixelPos(1,1:2),diff(figPixelPos)],'pixels','norm',hFig);
+annotationX = figNormPos([1,1]);
+annotationY = figNormPos([2,2]) + figNormPos(4)*[1,0];
+normx=annotationX(1);
+normy=annotationY(1);

### 2014-10-22 dividecontours.m

@@ -1,0 +1,12 @@
+function L = dividecontours(C)
+L=cell(1);
+q=size(C,2);
+location=1;
+i=1;
+while location<=q
+    go=C(2,location);
+    L{i}=C(:,location+(1:go));
+    location=location+1+go;
+    i=i+1;
+end
+    

### 2014-10-22 effect_of_costs.m

@@ -1,0 +1,118 @@
+function effect_of_costs(c1,c2,c3)
+
+Nl=evalin('base','Nl');
+Nd=evalin('base','Nd');
+Nt=evalin('base','Nt');
+twomat=evalin('base','twomat');
+threshvals=evalin('base','threshvals');
+indices=1:Nt;
+domvals=evalin('base','domvals');
+blue=evalin('base','blue');
+red=evalin('base','red');
+markersz=evalin('base','markersz');
+lw=evalin('base','lw');
+
+perf=zeros(2,2,Nl,Nd,Nt,Nt); %criteria, individual, leak, dominance, thresholds
+
+perf(1,1,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2));
+perf(1,2,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2));
+perf(2,1,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2))+c3*(1-twomat(:,2:Nd,:,:,1));
+perf(2,2,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2))+c3*(twomat(:,2:Nd,:,:,1));
+
+perf(1,1,:,1,:,:)=c2*(twomat(:,1,:,:,2));
+perf(1,2,:,1,:,:)=c2*(twomat(:,1,:,:,2));
+perf(2,1,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(1-twomat(:,1,:,:,1));
+perf(2,2,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(twomat(:,1,:,:,1));
+
+nasheq_thresh=cell(2,Nl,Nd);
+nasheq_ind=cell(2,Nl,Nd);
+
+T1_best=zeros(2,Nl,Nd,Nt); %criteria, leak, dominance, threshold
+T2_best=zeros(2,Nl,Nd,Nt);
+
+for i=1:2
+    for j=1:Nl
+        for k=1:Nd
+            for u=1:Nt
+            [~,m]=min(perf(i,1,j,k,:,u));
+            T1_best(i,j,k,u)=m;
+            [~,m]=min(perf(i,2,j,k,u,:));
[…]

### 2014-10-22 eigcentrality.m

@@ -1,0 +1,19 @@
+function v=eigcentrality(M,weight)
+N=size(M,2);
+totalsum=sum(sum(M));
+if totalsum==0
+    v=zeros(N,1);
+else
+    stochM=M/totalsum;
+    stochM(1:N+1:end)=1-sum(stochM,1);
+
+    unifmat=1/N*ones(N);
+
+    tofindeigenvectorsof=(1-weight)*stochM+weight*unifmat;
+    [vecs,vals]=eig(tofindeigenvectorsof);
+    f=find(abs(real(diag(vals))-1)<1e-5,1,'first');
+    v=vecs(:,f);
+    v=abs(v);
+end
+
+

### 2014-10-22 eigen_optimization.R

@@ -1,0 +1,75 @@
+setwd("/Users/eleanorbrush/Dropbox/signaling_network")
+source("gillespie.R")
+
+tau=1
+e=5
+bs=2
+bf=1
+rf=1
+dominance=.75
+leak=0
+ramp=.5
+thresh=1
+
+xmax=5
+possible=seq(-xmax,xmax,by=1)
+s=length(possible)
+
+states=array(,c(s^2,2))
+colnames(states)=c("1 state","2 state")
+rates=array(,c(s^2,4))
+colnames(rates)=c("1 signals","2 signals","1 wins","2 wins")
+transitions=array(0,c(s^2,s^2))
+
+v<-possible
+states[,1]=as.vector(sapply(v, function (x) rep(x,s)))
+states[,2]=rep(v,s)
+rates[,1]=apply(as.matrix(states[,1],ncol=1),1,sig_rate,ramp=ramp,thresh=thresh)
+rates[,2]=apply(as.matrix(states[,2],ncol=1),1,sig_rate,ramp=ramp,thresh=thresh)
+rates[,3]=rf*dominance*(1-rates[,1])*(1-rates[,2])
+rates[,4]=rf*(1-dominance)*(1-rates[,1])*(1-rates[,2])
+rates=rates
+
+for(i in 1:s^2){
+	state=states[i,]
+	leakedstate=(1-leak*tau)*state
+	newi=which(apply(states==leakedstate,1,sum)==2)
+	probs=cbind(dpois(0:s,rates[i,1]),dpois(0:s,rates[i,2]),dpois(0:s,rates[i,3]),dpois(0:s,rates[i,4]))
+	for(n1 in 1:s){
+		for(n2 in 1:s){
+			for(n3 in 1:s){
+				for(n4 in 1:s){
+					p=probs[n1+1,1]*probs[n2+1,2]*probs[n3+1,3]*probs[n4+1,4]
+					change=c((n2)*bs+(n3)*bf-(n4)*bf,(n1)*bs-(n3)*bf+(n4)*bf)
+					if(round(p,3)>0){
+					print(c(round(p,3),n1,n2,n3,n4,change))}
+					changedstate=state+change
+					if(changedstate[1]>xmax){changedstate[1]=xmax}
+					if(changedstate[2]>xmax){changedstate[2]=xmax}
+					if(changedstate[1]<(-xmax)){changedstate[1]=-xmax}
+					if(changedstate[2]<(-xmax)){changedstate[2]=-xmax}
[…]

### 2014-10-22 figures.m

@@ -1,0 +1,449 @@
+%% plotting parameters
+axislw=1.25;
+lw=3;
+textfontsz=12;
+labfontsz=10;
+fontname='Times New Roman';
+yellow=[1 1 0];
+red=[.9 0 0];
+green=[0 .7 0];
+blue=[0 0 .9];
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+numpowerfuns=length(powerfuns);
+%% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+%% find how nash equilibria depend on c1,c2,c3
+indices=1:Nt;
+leak=1;
+
+nasheq_ind=cell(Nc,Nd);
+giveup=zeros(Nc,Nd);
+nashaccuracy=zeros(Nc,Nd);
+
+for i=1:Nc
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    
+    perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+    perf(1,2:Nd,:,:)=c1*(1-twomat(leak,2:Nd,:,:,1))+c2*(twomat(leak,2:Nd,:,:,2))+c3*(1-twomat(leak,2:Nd,:,:,1));
+    perf(2,2:Nd,:,:)=c1*(1-twomat(leak,2:Nd,:,:,1))+c2*(twomat(leak,2:Nd,:,:,2))+c3*(twomat(leak,2:Nd,:,:,1));
+
+    perf(1,1,:,:)=c2*(twomat(leak,1,:,:,2))+c3*(1-twomat(leak,1,:,:,1));
+    perf(2,1,:,:)=c2*(twomat(leak,1,:,:,2))+c3*(twomat(leak,1,:,:,1));
+
+        for k=2:Nd
[…]

### 2014-10-22 figures_for_qcb_poster.m

@@ -1,0 +1,489 @@
+% %% plotting parameters
+axislw=1.25;
+lw=3;
+textfontsz=15;
+labfontsz=12;
+fontname='Arial';
+yellow=[1 1 0];
+red=[.9 0 0];
+green=[0 .7 0];
+blue=[0 0 .9];
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+numpowerfuns=length(powerfuns);
+%% coordinate set up
+load /Users/eleanorbrush/Dropbox/signaling_network/variables.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/twomat.mat
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+%% find how nash equilibria depend on c1,c2,c3
+indices=1:Nt;
+leak=1;
+
+nasheq_ind=cell(Nc,Nd);
+giveup=zeros(Nc,Nd);
+nashaccuracy=zeros(Nc,Nd);
+
+for i=1:Nc
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    
+    perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+    perf(1,2:Nd,:,:)=c1*(1-twomat(leak,2:Nd,:,:,1))+c2*(twomat(leak,2:Nd,:,:,2))+c3*(1-twomat(leak,2:Nd,:,:,1));
+    perf(2,2:Nd,:,:)=c1*(1-twomat(leak,2:Nd,:,:,1))+c2*(twomat(leak,2:Nd,:,:,2))+c3*(twomat(leak,2:Nd,:,:,1));
+
+    perf(1,1,:,:)=c2*(twomat(leak,1,:,:,2))+c3*(1-twomat(leak,1,:,:,1));
[…]

### 2014-10-22 figures_fortalks.m

@@ -1,0 +1,449 @@
+%% plotting parameters
+axislw=1.25;
+lw=3;
+textfontsz=12;
+labfontsz=10;
+fontname='Arial';
+yellow=[1 1 0];
+red=[.9 0 0];
+green=[0 .7 0];
+blue=[0 0 .9];
+% mycolormap=cbrewer('seq', 'YlOrRd', 100);
+% mycolormap=cbrewer('seq','BuPu',100);
+mycolormap=cbrewer('seq','Reds',64);
+mycolormap=mycolormap(1:end,:);
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+numpowerfuns=length(powerfuns);
+load /Users/eleanorbrush/Dropbox/signaling_network/variables
+%% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+indices=1:Nt;
+leak=2;
+
+%% find how nash equilibria depend on c1,c2,c3
+
+nasheq_ind=cell(Nc,Nd);
+giveup=zeros(Nc,Nd);
+nashaccuracy=zeros(Nc,Nd);
+
+for i=1:Nc
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    
+    perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+    perf(1,2:Nd,:,:)=c1*(1-twomat2(leak,2:Nd,:,:,1))+c2*(twomat2(leak,2:Nd,:,:,2))+c3*(1-twomat2(leak,2:Nd,:,:,1));
[…]

### 2014-10-22 figures_nonzeroleak.m

@@ -1,0 +1,571 @@
+%% plotting parameters
+axislw=1.25;
+lw=3;
+markersz=3;
+textfontsz=12;
+labfontsz=10;
+fontname='Times New Roman';
+yellow=[1 1 0];
+red=[.9 0 0];
+green=[0 .7 0];
+blue=[0 0 .9];
+% mycolormap=cbrewer('seq', 'YlOrRd', 100);
+% mycolormap=cbrewer('seq','BuPu',100);
+mycolormap=cbrewer('seq','Reds',64);
+mycolormap=mycolormap(1:end,:);
+% fourcols=[yellow;red;green;blue];
+fourcols=cbrewer('qual','Set1',4);
+fourcols=fourcols([3 1 2 4],:);
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+fourpowerfuns={'D','R','H','C'};
+funtypes={'number of signals','entropy','number of signalers','eigenvector centrality'};
+cornerlabels={'Error rate','Personal preference','Decision time'};
+markertypes={'-o','-d','-v','-s'};
+funstoplot=[1 2 4 6];
+colorperm=[1 2 4 3];
+numpowerfuns=length(powerfuns);
+load /Users/eleanorbrush/Dropbox/signaling_network/variables
+%% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+endvals=zeros(1,Nc);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    endvals(i+1)=layers{i+1}(end);
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+indices=1:Nt;
+leak=2;
+
+%% find how nash equilibria depend on c1,c2,c3
+
+nasheq_ind=cell(Nc,Nd);
+giveup=zeros(Nc,Nd);
[…]

### 2014-10-22 findthresh.m

@@ -1,0 +1,14 @@
+function t=findthresh(d,l,b,deltax,p)
+
+myfun = @(t,d,l,b,deltax,p) solve_pde_1d_prob(t,d,l,b,deltax)-p;  % parameterized function
+
+fun = @(x) myfun(x,d,l,b,deltax,p);
+if sign(fun(.1))~=sign(fun(10))
+    t = fzero(fun,[.1 10]);
+else 
+    t=fzero(fun,2);
+    v=[.1 t 10];
+    q=[fun(.1) fun(t) fun(10)];
+    [~,m]=min(abs(q));
+    t=v(m);
+end

### 2014-10-22 gillespie.R

@@ -1,0 +1,103 @@
+sig_rate<-function(x,ramp=1,thresh=1){
+	1/(1+exp(4*ramp*(x+thresh)))
+}
+
+sims<-function(e,maxtime,ramp,thresh1,thresh2,rf,dominance,leak,bs,bf){
+
+	its=100
+	boost=matrix(c(0,bs,bs,0,bf,-bf,-bf,bf,0,0),nrow=2)
+	X=list()
+	TIMES=list()
+	RXNS=list()
+
+	for(j in 1:its){
+		x=array(0,c(2,1))
+		x.now=x
+		times=0
+		time.now=times
+		rxns=array(0,5)
+		while(time.now<maxtime){
+			rxn.now=array(0,5)
+			a0=e
+			s1=sig_rate(x.now[1],ramp=ramp,thresh=thresh1)
+			s2=sig_rate(x.now[2],ramp=ramp,thresh=thresh2)
+			s=s1+s2-s1*s2
+			rates=c(s1,s2,(1-s)*rf*dominance,(1-s)*rf*(1-dominance))
+			r=runif(2,0,1)
+			tau=1/a0*log(1/r[1])
+			rxn.now[min(which(r[2]<=c(cumsum(rates),1)))]=1
+			deltax=boost%*%rxn.now
+			x.now=(1-leak*tau)*x.now+deltax
+			time.now=time.now+tau
+			x=cbind(x,x.now)
+			times=c(times,time.now)
+			rxns=cbind(rxns,rxn.now)
+		}
+	X[[j]]=cbind(x,x.now)
+	TIMES[[j]]=c(times,maxtime)
+	RXNS[[j]]=cbind(rxns,rep(1,5))
+	dimnames(RXNS[[j]])[[2]]=NULL
+	}	
+return(list(estimates=X,times=TIMES,rxns=RXNS))
+}
+
+stats<-function(simulation){
+	estimates=simulation$estimates
+	times=simulation$times
+	rxns=simulation$rxns
+	L=length(times) #number of simulations 
+	num.rxns=array(,L)
+	waits=array(,L)
+	firstsignaler=array(0,c(2,L))
+	switches.count=array(0,L)
+	switches.times=list()
+	transitions=list()
+	regimes=NULL
+	for(i in 1:L){
+		num.rxns[i]=length(times[[i]])-2
+		
+		signaling=rxns[[i]][1:2,]
+		signaling.happened=which(apply(signaling,2,sum)>0)
[…]

### 2014-10-22 group_props_N_dependence.m

@@ -1,0 +1,503 @@
+
+Nb1=min(10,N);
+Nb2=min(10,N);
+l=1;
+weight=.05;
+
+keeptrackofinfoD=zeros(its,Nb1*Nb2);
+keeptrackofinfoR=zeros(its,Nb1*Nb2);
+keeptrackofinfoDelta=zeros(its,Nb1*Nb2);
+keeptrackofinfoH=zeros(its,Nb1*Nb2);
+keeptrackofinfoPi=zeros(its,Nb1*Nb2);
+keeptrackofinfoC=zeros(its,Nb1*Nb2);
+keeptrackofinfoprobs=zeros(its,Nb1*Nb2);
+infomatD=zeros(Nb1,Nb2);
+infomatR=zeros(Nb1,Nb2);
+infomatDelta=zeros(Nb1,Nb2);
+infomatH=zeros(Nb1,Nb2);
+infomatPi=zeros(Nb1,Nb2);
+infomatC=zeros(Nb1,Nb2);
+infomatprobs=zeros(Nb1,Nb2);
+
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+numpowerfuns=length(powerfuns);
+
+skewvalsD=zeros(1,its);
+skewvalsR=zeros(1,its);
+skewvalsDelta=zeros(1,its);
+skewvalsH=zeros(1,its);
+skewvalsPi=zeros(1,its);
+skewvalsprobs=zeros(1,its);
+skewvalsC=zeros(1,its);
+finalthresh=zeros(N,its);
+
+keepprobmat=zeros(its,N*N);
+keeptimemat=zeros(its,N*N);
+keepsigmat=zeros(its,N*N);
+
+pairwiseaccuracy=zeros(1,its);
+pairwisetime=zeros(1,its);
+
+for c=1:its
+
+    opt_its=10;
+
+    fighting_abilities=zeros(1,N);
+    
+    switch dist
+        case 'unif'
+            fighting_abilities=20*rand(1,N);
+        case 'norm'
+            fighting_abilities=normrnd(0,10,1,N);
+        case 'lognorm'
+            fighting_abilities=lognrnd(0,1,1,N);
+    end
+    
+    fighting_abilities=sort(fighting_abilities,'descend');
+
+    dmat=zeros(N); %turn difference in fighting abilities into probabilitise of winning
+    extendeddomvals=[1-domvals(end:-1:2) domvals];
[…]

### 2014-10-22 group_props_comparison.m

@@ -1,0 +1,547 @@
+% function toreturn=group_props_parallelized(cvecs,its,N,dist,Nd,Nt,twomat,domvals,threshvals)
+% 
+% if nargin==3
+%     N=30;
+%     its=25;
+% end
+dist='unif';
+ivals=[1 11 12 39 52 57];
+for i=1:length(ivals)
+    cvecs(i,:)=[X(ivals(i)), Y(ivals(i)), Z(ivals(i))];
+end
+
+Nb1=10;
+Nb2=10;
+l=1;
+weight=.05;
+
+Nc=size(cvecs,1);
+
+keeptrackofinfoD=zeros(its,Nc,Nb1*Nb2);
+keeptrackofinfoR=zeros(its,Nc,Nb1*Nb2);
+keeptrackofinfoDelta=zeros(its,Nc,Nb1*Nb2);
+keeptrackofinfoH=zeros(its,Nc,Nb1*Nb2);
+keeptrackofinfoPi=zeros(its,Nc,Nb1*Nb2);
+keeptrackofinfoC=zeros(its,Nc,Nb1*Nb2);
+keeptrackofinfoprobs=zeros(its,Nc,Nb1*Nb2);
+infomatD=zeros(Nc,Nb1,Nb2);
+infomatR=zeros(Nc,Nb1,Nb2);
+infomatDelta=zeros(Nc,Nb1,Nb2);
+infomatH=zeros(Nc,Nb1,Nb2);
+infomatPi=zeros(Nc,Nb1,Nb2);
+infomatC=zeros(Nc,Nb1,Nb2);
+infomatprobs=zeros(Nc,Nb1,Nb2);
+powervecD=zeros(its,Nc,N);
+powervecR=zeros(its,Nc,N);
+powervecDelta=zeros(its,Nc,N);
+powervecH=zeros(its,Nc,N);
+powervecPi=zeros(its,Nc,N);
+powervecC=zeros(its,Nc,N);
+powervecprobs=zeros(its,Nc,N);
+
+powerfuns={'D','R','Delta','H','Pi','C','probs'};
+numpowerfuns=length(powerfuns);
+
+skewvalsD=zeros(its,Nc);
+skewvalsR=zeros(its,Nc);
+skewvalsDelta=zeros(its,Nc);
+skewvalsH=zeros(its,Nc);
+skewvalsPi=zeros(its,Nc);
+skewvalsprobs=zeros(its,Nc);
+skewvalsC=zeros(its,Nc);
+
[…]

### 2014-10-22 group_props_parallelized.m

@@ -1,4 +1,4 @@
-function toreturn=group_props_parallelized(c1,c2,c3,its,N,dist,Nd,Nt,twomat,domvals,threshvals)
+function toreturn=group_props_parallelized(c1,c2,c3,its,N,distribution,Nd,Nt,twomat,domvals,threshvals,l)
 
 if nargin==3
     N=30;
@@ -7,7 +7,7 @@ end
 
 Nb1=10;
 Nb2=10;
-l=1;
+% l=1;
 weight=.05;
 
 keeptrackofinfoD=zeros(its,Nb1*Nb2);
@@ -25,8 +25,7 @@ infomatPi=zeros(Nb1,Nb2);
 infomatC=zeros(Nb1,Nb2);
 infomatprobs=zeros(Nb1,Nb2);
 
-powerfuns={'D','R','Delta','H','Pi','probs'};
-numpowerfuns=length(powerfuns);
+%powerfuns={'D','R','Delta','H','Pi''C','probs'};
 
 skewvalsD=zeros(1,its);
 skewvalsR=zeros(1,its);
@@ -50,7 +49,7 @@ for c=1:its
 
     fighting_abilities=zeros(1,N);
     
-    switch dist
+    switch distribution
         case 'unif'
             fighting_abilities=20*rand(1,N);
         case 'norm'
@@ -66,6 +65,7 @@ for c=1:its
     v=1:(Nd+Nd-1);
     deps=.05;
 
+    % find differences in abilities --> dominance
     for i=1:N
         for j=[1:(i-1),(i+1):N]
             abilitydiff=fighting_abilities(i)-fighting_abilities(j);
@@ -88,8 +88,8 @@ for c=1:its
     Tvals=zeros(N,opt_its);
     Tvals(:,1)=2*ones(N,1);
     % Tvals(:,1)=randi([1 Nt],N,1);
-    perfvals=zeros(N,opt_its);
-
+     
+    %find optimal thresholds
     for i=1:N
         opp_thresh=Tvals([1:(i-1),(i+1):N],1);
         ds=dmat(i,[1:(i-1),(i+1):N]);
@@ -101,7 +101,7 @@ for c=1:its
                         perfsum=perfsum+perf(1,ds(q)-Nd+1,Tvals(i,1),opp_thresh(q));
                     end
                 end  
-        perfvals(i,1)=perfsum;
+        
     end
 
     count=1;
@@ -121,29 +121,27 @@ for c=1:its
[…]

### 2014-10-22 group_props_time.m

@@ -1,0 +1,582 @@
+function toreturn=group_props_time(c1,c2,c3,its,N,distribution,Nd,Nt,twomat,domvals,threshvals,l,timevec)
+
+if nargin==3
+    N=30;
+    its=25;
+end
+
+Nb1=10;
+Nb2=10;
+% l=1;
+
+Ntime=length(timevec);
+
+keeptrackofinfoD=zeros(its,Nb1*Nb2,Ntime);
+keeptrackofinfoR=zeros(its,Nb1*Nb2,Ntime);
+keeptrackofinfoDelta=zeros(its,Nb1*Nb2,Ntime);
+keeptrackofinfoH=zeros(its,Nb1*Nb2,Ntime);
+keeptrackofinfoPi=zeros(its,Nb1*Nb2,Ntime);
+keeptrackofinfoC=zeros(its,Nb1*Nb2,Ntime);
+
+% infomatD=zeros(Nb1,Nb2,Ntime);
+% infomatR=zeros(Nb1,Nb2,Ntime);
+% infomatDelta=zeros(Nb1,Nb2,Ntime);
+% infomatH=zeros(Nb1,Nb2,Ntime);
+% infomatPi=zeros(Nb1,Nb2,Ntime);
+% infomatC=zeros(Nb1,Nb2,Ntime);
+
+%powerfunlabs={'D','R','Delta','H','Pi','C','probs'};
+
+skewvalsD=zeros(1,its);
+skewvalsR=zeros(1,its);
+skewvalsDelta=zeros(1,its);
+skewvalsH=zeros(1,its);
+skewvalsPi=zeros(1,its);
+skewvalsC=zeros(1,its);
+finalthresh=zeros(N,its);
+
+keepprobmat=zeros(its,N*N);
+keeptimemat=zeros(its,N*N);
+keepsigmat=zeros(its,N*N);
+
+pairwiseaccuracy=zeros(1,its);
+pairwisetime=zeros(1,its);
+
+for c=1:its
+
+    opt_its=10;
+
+    fighting_abilities=zeros(1,N);
+    
+    switch distribution
+        case 'unif'
+            fighting_abilities=20*rand(1,N);
+        case 'norm'
+            fighting_abilities=normrnd(0,10,1,N);
+        case 'lognorm'
+            fighting_abilities=lognrnd(0,1,1,N);
+    end
+    
+    fighting_abilities=sort(fighting_abilities,'descend');
+
[…]

### 2014-10-22 infogrowth.m

@@ -1,0 +1,295 @@
+N=20;
+its=100;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+indices=1:Nt;
+leak=2;
+
+% timevec=.1:.1:5;
+% Ntime=length(timevec);
+%%
+mut_time=zeros(Nc,6,Ntime);
+
+for i=1:Nc
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    tmax=max(max(timemats(:,:,i)));
+    timevec=tmax/50:tmax/50:tmax;
+    t=group_props_time(c1,c2,c3,its,N,'unif',Nd,Nt,twomat2,domvals,threshvals,leak,timevec);
+    mut_time(i,:,:)=t;
+end
+%%
+filename='/Users/eleanorbrush/Desktop/infogrowth_output.mat';
+% save(filename,'mut_time');
+%%
+% diag=1:(N+1):N^2;
+% offdiag=setdiff(1:(N^2),diag);
+layerstoplot=[1 3 5];
+Nlays=length(layerstoplot);
+Ntime=100;
+mut_time=zeros(Nlays,6,Ntime+1);
+for I=1:Nlays
+    i=layers{layerstoplot(I)}(1);
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    reshapedtime=reshape(timemats(:,:,i),N,N);
+    tmin=min(min(reshapedtime(reshapedtime~=0)));
+    tmax=max(max(reshapedtime));
+    step=(tmax-tmin)/(Ntime-1);
+    timevec=(tmin-step):step:tmax;
+    t=group_props_time(c1,c2,c3,its,N,'unif',Nd,Nt,twomat2,domvals,threshvals,leak,timevec);
+    mut_time(I,:,:)=t;
+end
+%% horizontal time plots look like crap
+% figure;
[…]

### 2014-10-22 lognormal_abilities.m

@@ -1,0 +1,58 @@
+
+expowervecs_lognorm=zeros(N,7,Nc);
+skewvals_lognorm=zeros(7,Nc);
+mutvals_lognorm=zeros(7 ,Nc);
+accuracy_lognorm=zeros(1,Nc);
+time_lognorm=zeros(1,Nc);
+avgthresh_lognorm=zeros(N,Nc);
+timemats_lognorm=zeros(N,N,Nc);
+probmats_lognorm=zeros(N,N,Nc);
+
+bottomquart=16:20;
+bottomaccuracy_lognorm=zeros(1,Nc);
+
+topquart=1:5;
+topaccuracy_lognorm=zeros(1,Nc);
+
+for i=1:Nc
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    t=group_props_parallelized(c1,c2,c3,its,N,'lognorm',Nd,Nt,twomat,domvals,threshvals);
+    expowervecs_lognorm(:,:,i)=t{1};
+    mutvals_lognorm(:,i)=t{2};
+    skewvals_lognorm(:,i)=t{3};
+    accuracy_lognorm(i)=t{4};
+    time_lognorm(i)=t{5};
+    avgthresh_lognorm(:,i)=t{7};
+    timemats_lognorm(:,:,i)=t{8};
+    probmats_lognorm(:,:,i)=t{9};
+    totalsigmat=t{10};
+    bottomaccuracy_lognorm(i)=sum(sum(triu(totalsigmat(bottomquart,bottomquart),1)))/sum(sum(totalsigmat(bottomquart,bottomquart)));
+    topaccuracy_lognorm(i)=sum(sum(triu(totalsigmat(topquart,topquart),1)))/sum(sum(totalsigmat(topquart,topquart)));
+end
+
+%%
+scaledskewvals_lognorm=skewvals_lognorm-(repmat(min(skewvals_lognorm,[],2),1,Nc));
+scaledskewvals_lognorm=scaledskewvals_lognorm./repmat(max(scaledskewvals_lognorm,[],2),1,Nc);
+
+scaledmutvals_lognorm=mutvals_lognorm-(repmat(min(mutvals_lognorm,[],2),1,Nc));
+scaledmutvals_lognorm=scaledmutvals_lognorm./repmat(max(scaledmutvals_lognorm,[],2),1,Nc);
+
+[~,order]=sort(mutvals_lognorm(1:6,:));
+bestfun_lognorm=order(6,:);
[…]

### 2014-10-22 matlab_solver.m

@@ -1,0 +1,35 @@
+%% geometry
+maxX=1;
+maxY=1;
+Tx=1;
+Ty=1;
+width = 1; height = 1;
+% define  the square by giving the 4 x-locations followed by the 4
+% y-locations of the corners.
+gdmTrans = [3 4 -Tx maxX maxX -Tx -Ty -Ty maxY maxY];
+g = decsg(gdmTrans', 'S1', ('S1')');
+% Create the triangular mesh on the square with approximately
+% ten elements in each direction.
+hmax = .01; % element size
+[p, e, t] = initmesh(g, 'Hmax', hmax);
+
+%%
+%%parameters
+
+l=1;
+b=1;
+d=.7;
+a=b*(2*d-1);
+c=b^2;
+
+%%
+A = @(p,t,u,time) acoeffunction(p,t,l,a);
+C = @(p,t,u,time) ccoeffunction(p,t,c);
+F = [0;0;0];
+B = @(p,e,u,time) boundaryFile(p,e);
+
+u = assempde(B, p, e, t, C, A, F);
+n = size(p,2); % number of finite element nodes
+uu = reshape(u, n, []);
+
+pdeplot(p, e, t, 'xydata', uu(:,1), 'contour', 'on');

### 2014-10-22 my_colors.R

@@ -1,0 +1,33 @@
+library(colorspace)
+setwd("/Users/eleanorbrush/Dropbox/signaling_network")
+
+n=101;
+
+cols=sequential_hcl(n=n,h=120,l=c(30,100),c.=c(100,0),power=.5)
+cols=col2rgb(cols)
+cols=cols/255
+cols=matrix(cols,nrow=3)
+cols=cols[,n:1]
+write(cols,file='green.txt',ncolumns=3)
+ 
+cols=sequential_hcl(n=n,h=10,l=c(30,100),c.=c(100,0),power=.5)
+cols=col2rgb(cols)
+cols=cols/255
+cols=matrix(cols,nrow=3)
+cols=cols[,n:1]
+write(cols,file='red.txt',ncolumns=3)
+ 
+cols=sequential_hcl(n=n,h=260,l=c(30,100),c.=c(100,0),power=.5)
+cols=col2rgb(cols)
+cols=cols/255
+cols=matrix(cols,nrow=3)
+cols=cols[,n:1]
+write(cols,file='blue.txt',ncolumns=3)
+ 
+cols=sequential_hcl(n=n,l=c(30,100),c.=c(0,0),power=.5)
+cols=col2rgb(cols)
+cols=cols/255
+cols=matrix(cols,nrow=3)
+cols=cols[,n:1]
+write(cols,file='black.txt',ncolumns=3)
+

### 2014-10-22 mycolors.m

@@ -1,0 +1,25 @@
+function colorvec=mycolors(n,col)
+
+green=importdata('green.txt');
+blue=importdata('blue.txt');
+red=importdata('red.txt');
+black=importdata('black.txt');
+
+switch col
+    case 'green'
+        base=green;
+    case 'blue'
+        base=blue;
+    case 'red'
+        base=red;
+    case 'black'
+        base=black;
+end
+
+colorvec=zeros(n,3);
+v=0:(1/(n-1)):1;
+
+for i=1:3
+    colorvec(:,i)=interp1(0:.01:1,base(:,i),v);
+end
+end

### 2014-10-22 myfzero.m

@@ -1,0 +1,674 @@
+function [b,fval,exitflag,output] = myfzero(FunFcnIn,x,options,varargin)
+%FZERO  Single-variable nonlinear zero finding. 
+%   X = FZERO(FUN,X0) tries to find a zero of the function FUN near X0, 
+%   if X0 is a scalar.  It first finds an interval containing X0 where the 
+%   function values of the interval endpoints differ in sign, then searches 
+%   that interval for a zero.  FUN is a function handle.  FUN accepts real 
+%   scalar input X and returns a real scalar function value F, evaluated 
+%   at X. The value X returned by FZERO is near a point where FUN changes 
+%   sign (if FUN is continuous), or NaN if the search fails.  
+%
+%   X = FZERO(FUN,X0), where X0 is a vector of length 2, assumes X0 is a 
+%   finite interval where the sign of FUN(X0(1)) differs from the sign of 
+%   FUN(X0(2)). An error occurs if this is not true.  Calling FZERO with a
+%   finite interval guarantees FZERO will return a value near a point where
+%   FUN changes sign.
+%
+%   X = FZERO(FUN,X0), where X0 is a scalar value, uses X0 as a starting 
+%   guess. FZERO looks for an interval containing a sign change for FUN and 
+%   containing X0.  If no such interval is found, NaN is returned.  
+%   In this case, the search terminates when the search interval 
+%   is expanded until an Inf, NaN, or complex value is found. Note: if
+%   the option FunValCheck is 'on', then an error will occur if an NaN or 
+%   complex value is found.
+%
+%   X = FZERO(FUN,X0,OPTIONS) solves the equation with the default optimization
+%   parameters replaced by values in the structure OPTIONS, an argument
+%   created with the OPTIMSET function.  See OPTIMSET for details.  Used
+%   options are Display, TolX, FunValCheck, OutputFcn, and PlotFcns. 
+%
+%   X = FZERO(PROBLEM) finds the zero of a function defined in PROBLEM. 
+%   PROBLEM is a structure with the function FUN in PROBLEM.objective, 
+%   the start point in PROBLEM.x0, the options structure in PROBLEM.options,
+%   and solver name 'fzero' in PROBLEM.solver. The structure PROBLEM must have 
+%   all the fields.
+%
[…]

### 2014-10-22 numerical_pde.R

@@ -1,0 +1,26 @@
+Ni=4
+Nj=5
+u=array(,(Ni-1)*(Nj-1))
+A=array(,c(Ni-1)*(Nj-1),(Ni-1)*(Nj-1))
+
+fval=0
+f=array(fval,(Ni-1)*(Nj-1))
+
+deltax=.01
+Lx=0
+Ux=10
+xvals=seq(Lx,Ux,by=deltax)
+
+deltay=.01
+L
+
+bl=1
+br=0
+bd=0
+du=1
+b=array(0,(Ni-1)*(Nj-1))
+down=1+(1:(Ni-2))*(Ni-1)
+up=Ni-1+(1:(Ni-2))*(Ni-1)
+left=2:(Ni-2)
+right=(Nj-2)*(Ni-1)+2:(Ni-2)
+

### 2014-10-22 numerical_pde.m

@@ -1,0 +1,281 @@
+%% parameters 
+D1=1;
+Dd=-1;
+D2=1;
+l=1;
+a=.5;
+
+%% setting up vectors I need
+
+deltax=.05;
+Lx=-1;
+Ux=5;
+xvals=Lx:deltax:Ux;
+
+deltay=.05;
+Ly=-.5;
+Uy=5;
+yvals=Ly:deltay:Uy;
+
+Ni=length(yvals);
+Nj=length(xvals);
+
+c=xvals(2:(end))';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+xvec = cc(:)';
+
+c=[1:(Nj-1)]';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+jvals = cc(:)';
+
+c=yvals(2:(end))';
+cc=c(:,ones(Nj-1,1));
+yvec = cc(:)';
+
+c=[1:(Ni-1)]';
+cc=c(:,ones(Nj-1,1));
+ivals = cc(:)';
+
+%% construct adjacency matrix
+A=sparse((Ni-1)*(Nj-1),(Ni-1)*(Nj-1));
+
+Aleft=D1/(deltax^2);
+Adown=D2/(deltay^2);
+Aupright=2*Dd/deltax/deltay;
+
+bl_prob=0;
+br_prob=1;
+bd_prob=1;
+bu_prob=0;
+b_prob=zeros(1,(Ni-1)*(Nj-1));
+
+bl_time=0;
+br_time=0;
+bd_time=0;
+bu_time=0;
+b_time=zeros(1,(Ni-1)*(Nj-1));
+
+down=1+(1:(Nj-3))*(Ni-1);
+up=Ni-1+(1:(Nj-3))*(Ni-1);
+left=2:(Ni-2);
+right=(Nj-2)*(Ni-1)+(2:(Ni-2));
+SW=1;
+SE=(Ni-1)*(Nj-2)+1;
+NW=Ni-1;
+NE=(Ni-1)*(Nj-1);
+interior=1:((Ni-1)*(Nj-1));
+interior([down up left right SW SE NW NE])=[];
+
+for k=interior
+    x=xvec(k);
+    y=yvec(k);
+    fx=-l*x+a;
+    gy=-l*y-a;
[…]

### 2014-10-22 nummat2logprobmat.m

@@ -1,0 +1,14 @@
+function logprobmat = nummat2logprobmat(nummat)
+[Nb1,Nb2]=size(nummat);
+tot=sum(sum(nummat));
+logprobmat=zeros(Nb1,Nb2);
+for i=1:Nb1
+    px=sum(nummat(i,:))/(tot);
+    for j=1:Nb2
+        py=sum(nummat(:,j))/(tot);
+        pxy=nummat(i,j)/(tot);
+        if pxy~=0
+            logprobmat(i,j)=pxy*log(pxy/px/py);
+        end
+    end
+end

### 2014-10-22 old_code/basins_of_attraction.m

@@ -1,0 +1,73 @@
+global f eps e t b
+f=0.1;
+eps=2;
+e=.5;
+b=-2*eps*e*(eps+f/e)/(f/e-eps);
+t=(2*eps*e*(eps+f/e)+b*(f/e-eps))/(2*eps*e-b);
+
+xvals = -t-eps-1:.1:f/e+rate(-f/e)/e+3;
+yvals = -t-eps-1:.1:-f/e+rate(f/e)/e+3;
+
+stepsx = min(xvals):2:max(xvals);
+stepsy = min(yvals):2:max(yvals);
+
+
+xgradient=zeros([length(yvals),length(xvals),2]);
+ygradient=zeros([length(yvals),length(xvals),2]);
+zgradient=zeros([length(yvals),length(xvals),2]);
+for i=1:length(xvals)
+    for j=1:length(yvals)
+        xgradient(j,i,:) = -e*xvals(i)+f+rate(yvals(j));
+        ygradient(j,i,:) = -e*yvals(j)-f+rate(xvals(i));
+    end
+end
+
+[X,Y,Z]=meshgrid(xvals,yvals,1:2);
+[sx,sy,sz]=meshgrid(stepsx,stepsy,1);
+figure
+h=streamslice(X,Y,Z,xgradient,ygradient,zgradient,[],[],[1]);
+set(h,'Color','black')
+set(gca,'box','off')
+axis tight
+hold on
+
+plot (f/e,-f/e+rate(f/e)/e,'ro')
+plot (f/e+rate(-f/e)/e,-f/e,'ro')
+xhat = b(eps-t)/(2*eps*e+b)-2*eps*f/(b-2*eps*e);
+yhat = b(eps-t)/(2*eps*e+b)+2*eps*f/(b-2*eps*e);
+plot (xhat,yhat,'ro')
+
+plotx = -t-eps-1:.01:timefunx(-t-eps);
+xhat = timefunx(-t-eps);
+yhat = -t-eps;
+ploty = (yhat-(b-f)/e)/(xhat-(b+f)/e)*(plotx-(b+f)/e)+(b-f)/e;
+plot(plotx,ploty,'r')
+
+ploty = -t-eps:.1:-t-eps+4*eps*f/(b-2*eps*e);
+plotx = timefunx(ploty);
+plot(plotx,ploty,'r')
+
+plotx = -t-eps:.1:-t+eps-4*eps*f/(b-2*eps*e);
[…]

### 2014-10-22 old_code/bifurcation.R

@@ -1,0 +1,114 @@
+e=0.5
+eps=0.5
+
+bifurc<-function(b,f){ #if T<bifurc(b) and b>2*eps*e a bifurcation occurs
+	if(b>2*eps*error){
+	a=(2*eps*error+b)*f/error+2*eps^2*error-eps*b
+	a=a/(2*eps*error-b)
+	} else{ a=-1}
+	return(max(a,-1))
+}
+
+
+both<-function(b,f){ #if T<both(b) and b<2*eps*e both signal at non zero rates
+	if(b<2*eps*error){
+		a=(2*eps*error)*(eps-f/error)+b*(-f/error-eps)
+		a=a/(2*eps*error-b)
+	} else{a=-1}
+	return(max(a,-1))
+}
+
+lowerline<-function(b,f){
+	a=f/error-eps
+	return(a)
+}
+
+
+upperline<-function(b,f){
+	a=eps+f/error
+	return(a)
+}
+
+error=0.5
+eps=.1
+lb=200
+lf=200
+bvals=seq(0.1,10,length.out=lb)
+fvals=seq(0,2,length.out=lf)
+
+bifurcmat<-array(,c(lb,lf))
+bothmat<-array(,c(lb,lf))
+lowermat<-array(,c(lb,lf))
+uppermat<-array(,c(lb,lf))
+
+for(i in 1:lb){
+	for(j in 1:lf){
+		bifurcmat[i,j]=bifurc(bvals[i],fvals[j])
+		bothmat[i,j]=both(bvals[i],fvals[j])
+		lowermat[i,j]=lowerline(bvals[i],fvals[j])
+		uppermat[i,j]=upperline(bvals[i],fvals[j])
+	}
+}
+
+
+
+error=0.5
+eps=.1
+lb=500
+fmax=.3
+bvals=seq(0.01,1,length.out=lb)
+
+quartz()
+pdf(file='bifurc_diagram.pdf')
+
+layout(matrix(1:2,nrow=1))
+f=0.01
+plot(bvals,apply(matrix(bvals,nrow=1),2,bifurc,f=f),type='l',ylim=c(0,eps+eps*error+.25),xlab="Feedback level, b",ylab="Threshold, T",cex.lab=1.5,mgp=c(2.2,1,0))
+# ,main=paste("f = ",round(f,2),", e = ",e,", eps = ",eps)
+title(main=paste("f =",f),cex.main=2,line=1)
[…]

### 2014-10-22 old_code/bistable_plot.R

@@ -1,0 +1,82 @@
+f=0.05
+eps=2
+e=.5
+t=1
+b=5
+
+rate<-function(x){
+	if(x<=(-t-eps)){
+		a=b
+	} 
+	if(x>=(-t+eps)){
+		a=0
+	} 
+	if(x>(-t-eps)&&x<(-t+eps)){
+	a=b*(-1/(2*eps)*(x+t)+1/2)
+	}
+	return(a)
+}
+
+
+timefun<-function(y){
+	c=1/e*(f-b/(2*eps)*(b-f)/e-b*t/(2*eps)+b/2)
+	ty=1/e*log((y-(b-f)/e)/(-t-eps+4*eps*f/(b-2*eps*e)-(b-f)/e))
+	x=(-t-eps-c)*(y-(b-f)/e)/(-t-eps+4*eps*f/(b-2*eps*e)-(b-f)/e)+c+b/(2*eps)*(y-(b-f)/e)*ty
+	return(x)
+}
+
+timefun2<-function(x){
+	c=1/e*(-f-b/(2*eps)*f/e-b*t/(2*eps)+b/2)
+	tx=1/e*log((x-f/e)/(-t+eps-4*eps*f/(b-2*eps*e)-f/e))
+	y=(-t+eps-c)*(x-f/e)/(-t+eps-4*eps*f/(b-2*eps*e)-f/e)+c+b/(2*eps)*(x-f/e)*tx
+	return(y)
+}
+
+xrange=c(-t-eps-1,f/e+rate(-f/e)/e+1)
+yrange=c(-t-eps-1,-f/e+rate(f/e)/e+1)
+
+plot(xrange,yrange,col='white',xlab="",ylab="")
+
+xvals=matrix(seq(min(xrange),max(xrange),length.out=500),nrow=1)
+yvals=matrix(seq(min(yrange),max(yrange),length.out=500),nrow=1)
+
+xisocline=f/e+apply(yvals,2,rate)/e
+yisocline=-f/e+apply(xvals,2,rate)/e
+# lines(xisocline,yvals)
+# lines(xvals,yisocline)
+
+xvals=matrix(seq(-t-eps,-t+eps-4*eps*f/(b-2*eps*e),length.out=500),nrow=1)
+midline=xvals+4*eps*f/(b-2*eps*e)
+lines(xvals,midline)
+
+yvals=matrix(seq(-t-eps,-t-eps+4*eps*f/(b-2*eps*e),length.out=500),nrow=1)
+line=apply(yvals,2,timefun)
[…]

### 2014-10-22 old_code/checking_gillespie_algorithm.R

@@ -1,0 +1,185 @@
+re1<-1 #rate of error in animal 1
+re2<-re1 #rate of error in animal 2
+me1<-0 #mean error in animal 1
+me2<-me1 #mean error in animal 2
+se1<-1 #variance of error in animal 1
+se2<-se1 #variance of error in animal 2
+
+rf<-1 #rate of fighting
+mf<-0.1 #mean fight outcome
+sf<-0.0 #variance of fight outcome
+
+bs1=0 #boost to animal 2 when animal 1 signals
+bs2=0 #boost to animal 1 when animal 2 signals
+
+rs<-1 #maximum signaling rate
+
+T<-2.5  #threshold
+
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	return(rs*(-erf((x-(-T)))+1)/2)
+	}
+
+#for drift diffusion model (size of error does not depend on state)	
+error_rate<-function(T,me1,se1,mf,sf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=1/(1+exp(2*z*a)) #from psych.pdf
+#	x=1-(1-exp(-2*(me1+mf)*(se1^2+sf^2)))/(1-exp(-4*(me1+mf)*(se1^2+sf^2))) #from stoch calc class
+	return(x)
+	}
+	
+dec_time<-function(T,me1,se1,mf,sf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=z*tanh(z*a)
+	return(x)
+	}
+	
+
+timesteps=100 #how many reactions to have
+runs=500 #how many separate runs to do
+bins=seq(0,timesteps,by=0.5) #bin time so we can compare evolution in time between runs
+
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2){ #Gillespie algorithm
+
+
+hitting=NULL
+decision=NULL
+
+where1=list() #each element of the list will be an array of dominance values for animal 1 at a point in time
+for(i in 1:length(bins)){
+	where1[[i]]=array()
+	}
+
+
+for(k in 1:runs){
+#time=array(,timesteps+1) #keep track of timepoints
+#time[1]=0 #initialize
+#X=array(,c(timesteps+1))
+#X[1]=0
[…]

### 2014-10-22 old_code/dist_of_exit_points.R

@@ -1,0 +1,14 @@
+f=.5
+e=.5
+T=1
+eps=.1
+
+c=1
+
+sd=c^2/(2*e)
+mu=f/e
+h=2
+
+Sx=(f-e*(-T+eps))*dnorm(-T+eps,mean=mu,sd=sd)*(pnorm(h,mean=-mu,sd=sd)-pnorm(-T+eps,mean=-mu,sd=sd))+(-f-e*h)*dnorm(h,mean=-mu,sd=sd)*(pnorm(h,mean=mu,sd=sd)-pnorm(h,mean=mu,sd=sd))
+
+Sy=(f-e*h)*dnorm(h,mean=mu,sd=sd)*(pnorm(h,mean=-mu,sd=sd)-pnorm(-T+eps,mean=-mu,sd=sd))+(-f-e*(-T-eps))*dnorm(-T-eps,mean=-mu,sd=sd)*(dnorm(h,mean=mu,sd=sd)-dnorm(-T+eps,mean=mu,sd=sd))

### 2014-10-22 old_code/fixed_points.R

@@ -1,0 +1,155 @@
+library(fields)
+library(gplots)
+
+rate<-function(x){
+	y=(1-tanh(x))/2
+	return(y)
+}
+
+rate<-function(x){
+	y=0
+	if(x<=(-eps)){y=1}
+	if(x>(-eps)&&x<eps){y=-x/(2*eps)+.5}
+	return(y)
+}
+
+fixedy<-function(x,b,t){
+	r=-f/error+b/error*rate(x+t)
+	return(r)
+}
+
+findy<-function(y,b,t){
+	r=-f/error+b/error*rate(f/error+b/error*rate(y+t)+t)-y
+	return(r)
+}
+
+fixedx<-function(y,b,t){
+	r=f/error+b/error*rate(y+t)
+	return(r)
+}
+
+yvals<-function(b,t){
+	yvec=matrix(seq(-f/error+b/error,-f/error,length.out=1000),nrow=1)
+	seek=apply(yvec,2,findy,b=b,t=t)
+	seek=round(seek,5)
+	zeros=array(0,3)
+	if(sum(abs(diff(seek<=0)))<=1){
+	# if(length(setdiff(unique(sign(seek)),0))==1){
+		zeros[1]=uniroot(findy,lower=-f/error,upper=-f/error+b/error,b=b,t=t)$root
+		zeros[2]=zeros[1]
+		zeros[3]=zeros[1]
+	} else{
+	Y1=min(which(seek>0))
+	Y2=min(which(seek[-(1:Y1)]<0))+Y1
+	zeros[1]=uniroot(findy,upper=-f/error+b/error,lower=yvec[Y1],b=b,t=t)$root
+	zeros[2]=uniroot(findy,upper=yvec[Y1],lower=yvec[Y2],b=b,t=t)$root
+	zeros[3]=uniroot(findy,upper=yvec[Y2],lower=-f/error,b=b,t=t)$root
+	}
+	return(zeros)
+}
+
+# type<-function(x,y,b,t){
+	# k=0
+	# if(x[1]!=x[3]){k=7
+	# } else{
+	# x=x[1]
+	# y=y[1]
+	# if(fixedy(x,b,t)>(-f/error)&& fixedy(x,b,t)<(-f/error)+b/error&& fixedx(y,b,t)==f/error){k=1}	#1 kind of, 2 not
+	# if(fixedy(x,b,t)==(-f/error)&&fixedx(y,b,t)==f/error){k=2} #1 not, 2 not
[…]

### 2014-10-22 old_code/linear_learning.R

@@ -1,0 +1,55 @@
+binsig<-function(name){ #puts 1 where there nonzero entries of name and 0s elsewhere
+	n<-dim(name)[1]
+	x<-array(0,c(n,n))
+for(i in 1:n){for(j in 1:n){if(name[i,j]!=0){x[i,j]=1}}}
+return(x)
+}
+
+laplace<-function(i,mat){
+	M=mat
+	diag(M)=1
+	w=c(which(M[i,]!=0))
+	M=M[w,w]
+	B=binsig(M)
+	new=array(0,c(length(w)-1,length(w)-1))
+	for(j in 1:length(w)-1){
+		new[j,j]=-sum(B[j,])-B[j,length(w)]
+		new[j,-j]=B[j,-c(j,length(w))]-B[j,length(w)]
+		}
+	diag(new)=diag(new)+1
+	return(new)
+	}
+
+mat=matrix(sample(0:10,100,replace=TRUE),nrow=8)	
+N=50
+mat=matrix(0,nrow=N,nc=N)
+mat[sample(1:N^2,floor(N^2/2))]=sample(1:N^2,floor(N^2/2),replace=TRUE)
+mat[lower.tri(mat)]=t(mat)[lower.tri(mat)]
+
+mat=prox
+
+mat=get.adjacency(erdos.renyi.game(100,p=.4,type="gnp"))
+
+N=dim(mat)[2]
+learning<-array(,N)
+for(i in 1:N){
+	M=laplace(i,mat)
+	values<-Re(eigen(M)$values)
+	learning[i]=max(values)
+	}
+	
+L=laplace(1,prox)
+eigen(L)$values
+x=rep(0,31)
+y=runif(31,0,0.001)
+for(i in 1:10){
+	#print(sum((L%*%y)^2))
+	print(y[1])
+	y=L%*%y
+	}
+	
+lap<-function(i,mat){
+	B=binsig(mat)
+	diag(B)=-RR(B)
+	return(B)
+	}

### 2014-10-22 old_code/resulting_distribution.R

@@ -1,0 +1,96 @@
+rate<-function(x){
+	if(x<=(-T-eps)){
+		a=b
+	} 
+	if(x>=(-T+eps)){
+		a=0
+	} 
+	if(x>(-T-eps)&&x<(-T+eps)){
+	a=b*(-1/(2*eps)*(x+T)+1/2)
+	}
+	return(a)
+}
+
+# p<-function(n,N,T,sd=NULL){
+	# if(!is.null(sd)){sd=sd
+		# } else(sd=1/(2*sqrt(3)))
+	# L=sd*sqrt(3)
+	# K=0:n
+	# J=0:(N-1-n)
+	# mat=array(,c(n+1,N-n))
+	# for(k in 1:(n+1)){
+		# for(j in 1:(N-n)){
+			# mat[k,j]=choose(n,K[k])*choose(N-1-n,J[j])*L^(K[k]+J[j])*(-1)^(N-1-n-J[j])*1/(N-K[k]-J[j])*((L-e*(T-eps))^(N-K[k]-J[j])-(-L)^(N-K[k]-J[j]))
+		# }
+	# }
+	# s=sum(mat)
+	# s=1/(2*L)^N*choose(N-1,n)*s
+	# if(n==0){
+		# s=s+e*(T-eps)/(2*L)
+	# }
+	# return(s)
+# }
+
+p<-function(n,N,T,sd=NULL){ #given N animals, threshold T, sd of uniform distribution, return probability of receiving signals from n animals
+	if(!is.null(sd)){sd=sd
+		} else(sd=1/(2*sqrt(3)))
+	L=sd*sqrt(3)
+	c=1-e*(T-eps)/(2*L)
+	s=pbeta(c,n+1,N-n)
+	s=s/N
+	if(n==0){
+		s=s+e*(T-eps)/(2*L)
+	}
+	return(s)
+}
+
+step_size=0.01
+int<-function(v){
+	l=length(v)
+	step_size/2*(sum(2*v)-v[1]-v[l])
+}
+
+N=50
+L=1
+eps=.5
+e=.1
+T=eps+.1
+b=5
+f=.1
+
+
+probs<-function(N,T,sd=NULL){ #return vector of probabilities of receiving signals from 0 to N-1 animals
+	if(!is.null(sd)){sd=sd
+		} else(sd=1/(2*sqrt(3)))
+	L=sd*sqrt(3)
+	poss=matrix(0:(N-1),nrow=1)
[…]

### 2014-10-22 old_code/signal_model_03_20.R

@@ -1,0 +1,131 @@
+lambda=-.01
+alpha=0
+a1=1
+a2=1
+sigma2=2
+T=-5
+
+steps=100000
+runs=1
+x1=array(0,c(runs,steps))
+x2=array(0,c(runs,steps))
+s1=array(0,c(runs,steps))
+s2=array(0,c(runs,steps))
+
+for(j in 1:runs){
+for(i in 3:steps){
+ 	if(x2[j,i-2]>T&&x2[j,i-1]<=T){s2[j,i]=1} else s2[j,i]=0
+ 	if(x1[j,i-2]>T&&x1[j,i-1]<=T){s1[j,i]=1} else s1[j,i]=0
+ 	fight=rnorm(1,a1-a2,sqrt(sigma2))
+	x1[j,i]=(1+lambda)*x1[j,i-1]+fight+alpha*s2[j,i]
+	x2[j,i]=(1+lambda)*x2[j,i-1]-fight+alpha*s1[j,i]
+	}
+}
+received1=apply(s2,1,sum)
+received2=apply(s1,1,sum)
+
+layout(matrix(1:1,nrow=1))
+
+#plot(x1[1,],type="l",ylim=range(cbind(x1,x2)))
+#for(j in 2:runs){lines(x1[j,])}
+#lines(x2[1,],col="red")
+#for(j in 2:runs){lines(x2[j,],col="red")}
+
+#plot(s1[1,])
+#points(s2[1,],col="red")
+#hist(received1)
+#hist(received2,col="red")
+#hist(received1-received2)
+
+both_exchanged=intersect(which(received1!=0),which(received2!=0))
+
+lags=NULL
+
+for(j in both_exchanged){
+	
+	exchanged=array(0,steps)
+	exchanged[which(s1[j,]==1)]=1
+	exchanged[which(s2[j,]==1)]=-1
+	
+	k=min(which(exchanged!=0))
+	direction=exchanged[k]
+	walked=0
+	
+	while(k<steps){
+		k=k+1
+		if(exchanged[k]==0){walked=walked+1}
+		if(exchanged[k]==direction){walked=walked+1}
+		if(exchanged[k]==-1*direction){
+			direction=-1*direction
+			lags=c(lags,walked)
+			walked=0			
+			}
+		}
+	
+	}
+	
+hist(lags)
+print(mean(lags))
+
+
+##adrian's model
+#
+#delta=.01
+#psi=1
+#alpha=50
[…]

### 2014-10-22 old_code/signal_model_07_05.R

@@ -1,0 +1,232 @@
+re1<-1 #rate of error in animal 1
+re2<-re1 #rate of error in animal 2
+me1<-0 #mean error in animal 1
+me2<-me1 #mean error in animal 2
+se1<-1 #variance of error in animal 1
+se2<-se1 #variance of error in animal 2
+
+rf<-1 #rate of fighting
+mf<-0.1 #mean fight outcome
+sf<-0.0 #variance of fight outcome
+
+bs1=0 #boost to animal 2 when animal 1 signals
+bs2=0 #boost to animal 1 when animal 2 signals
+
+rs<-1 #maximum signaling rate
+
+T<-2.5  #threshold
+
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	return(rs*(-erf((x-(-T)))+1)/2)
+	}
+
+#for drift diffusion model (size of error does not depend on state)	
+error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=1/(1+exp(2*z*a)) #from psych.pdf
+#	x=1-(1-exp(-2*(me1+mf)*(se1^2+sf^2)))/(1-exp(-4*(me1+mf)*(se1^2+sf^2))) #from stoch calc class
+	return(x)
+	}
+	
+dec_time<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=z*tanh(z*a)
+	return(x)
+	}
+	
+#for ornstein-uhlenbeck model (size of error DOES depend on state)
+#error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+#	a=(me1+mf)^2/(se1^2+sf^2)
+	z=T/(me1+mf)
+#	}
+
+timesteps=100 #how many reactions to have
+runs=500 #how many separate runs to do
+bins=seq(0,timesteps,by=0.5) #bin time so we can compare evolution in time between runs
+
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2){ #Gillespie algorithm
+
+
+#fight_rates=array(,runs) #time spent fighting divided by total time each run
+#error1_rates=array(,runs) #rates of errors for each run
+#error2_rates=array(,runs)
[…]

### 2014-10-22 old_code/signal_model_07_05_forces.R

@@ -1,0 +1,189 @@
+re1<-1 #rate of error in animal 1
+re2<-re1 #rate of error in animal 2
+me1<-0 #mean error in animal 1
+me2<-me1 #mean error in animal 2
+se1<-1 #variance of error in animal 1
+se2<-se1 #variance of error in animal 2
+
+rf<-1 #rate of fighting
+mf<-0.1 #mean fight outcome
+sf<-0.0 #variance of fight outcome
+
+bs1=0 #boost to animal 2 when animal 1 signals
+bs2=0 #boost to animal 1 when animal 2 signals
+
+rs<-1 #maximum signaling rate
+
+T<-2.5  #threshold
+
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	return(rs*(-erf((x-(-T)))+1)/2)
+	}
+
+
+timesteps=100 #how many reactions to have
+runs=500 #how many separate runs to do
+bins=seq(0,timesteps,by=0.5) #bin time so we can compare evolution in time between runs
+
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2,to_plot){ #Gillespie algorithm
+
+
+fight_rates=array(,runs) #time spent fighting divided by total time each run
+error1_rates=array(,runs) #rates of errors for each run
+error2_rates=array(,runs)
+signal1_rates=array(,runs)
+signal2_rates=array(,runs)
+dom1=array(,runs) #sum of intervals in which animal 1 is dominant
+dom2=array(,runs)
+hitting1=NULL
+hitting2=NULL
+decision1=NULL
+decision2=NULL
+
+where1=list() #each element of the list will be an array of dominance values for animal 1 at a point in time
+where2=list()
+for(i in 1:length(bins)){
+	where1[[i]]=array()
+	where2[[i]]=array()
+	}
+
+for(k in 1:runs){
+time=array(,timesteps+1) #keep track of timepoints
+time[1]=0 #initialize
+X=array(,c(2,timesteps+1))
+X[,1]=c(0,0)
+rxns=array(,timesteps) #keep track of which reactions occur
+
+for(i in 1:timesteps+1){
+	X[,i]=X[,i-1]
[…]

### 2014-10-22 old_code/signal_model_07_06.R

@@ -1,0 +1,262 @@
+re1<-1 #rate of error in animal 1
+re2<-re1 #rate of error in animal 2
+me1<-0 #mean error in animal 1
+me2<-me1 #mean error in animal 2
+se1<-1 #variance of error in animal 1
+se2<-se1 #variance of error in animal 2
+
+rf<-1 #rate of fighting
+mf<-0.1 #mean fight outcome
+sf<-0.0 #variance of fight outcome
+
+bs1=0 #boost to animal 2 when animal 1 signals
+bs2=0 #boost to animal 1 when animal 2 signals
+
+rs<-1 #maximum signaling rate
+
+T<-2.5  #threshold
+
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	return(rs*(-erf((x-(-T)))+1)/2)
+	}
+
+#for drift diffusion model (size of error does not depend on state)	
+error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=1/(1+exp(2*z*a)) #from psych.pdf
+#	x=1-(1-exp(-2*(me1+mf)*(se1^2+sf^2)))/(1-exp(-4*(me1+mf)*(se1^2+sf^2))) #from stoch calc class
+	return(x)
+	}
+	
+dec_time<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=z*tanh(z*a)
+	return(x)
+	}
+	
+#for ornstein-uhlenbeck model (size of error DOES depend on state)
+#error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+#	a=(me1+mf)^2/(se1^2+sf^2)
+	z=T/(me1+mf)
+#	}
+
+timesteps=100 #how many reactions to have
+runs=500 #how many separate runs to do
+bins=seq(0,timesteps,by=0.5) #bin time so we can compare evolution in time between runs
+
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2){ #Gillespie algorithm
+
+
+#fight_rates=array(,runs) #time spent fighting divided by total time each run
+#error1_rates=array(,runs) #rates of errors for each run
[…]

### 2014-10-22 old_code/signal_model_07_10.R

@@ -1,0 +1,198 @@
+re1<-1 #rate of error in animal 1
+re2<-re1 #rate of error in animal 2
+me1<-0 #mean error in animal 1
+me2<-me1 #mean error in animal 2
+se1<-1 #variance of error in animal 1
+se2<-se1 #variance of error in animal 2
+
+rf<-1 #rate of fighting
+mf<-0.1 #mean fight outcome
+sf<-0.0 #variance of fight outcome
+
+bs1=0 #boost to animal 2 when animal 1 signals
+bs2=0 #boost to animal 1 when animal 2 signals
+
+rs<-1 #maximum signaling rate
+
+T<-2.5  #threshold
+
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	return(rs*(-erf((x-(-T)))+1)/2)
+	}
+
+#for drift diffusion model (size of error does not depend on state)	
+error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=1/(1+exp(2*z*a)) #from psych.pdf
+#	x=1-(1-exp(-2*(me1+mf)*(se1^2+sf^2)))/(1-exp(-4*(me1+mf)*(se1^2+sf^2))) #from stoch calc class
+	return(x)
+	}
+	
+dec_time<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=z*tanh(z*a)
+	return(x)
+	}
+
+timesteps=100 #how many reactions to have
+runs=500 #how many separate runs to do
+deltat=0.01
+bins=seq(0,timesteps*deltat,by=deltat) #bin time so we can compare evolution in time between runs
+
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2){ #Gillespie algorithm
+
+
+hitting=NULL
+decision=NULL
+switches=NULL
+
+where1=list() #each element of the list will be an array of dominance values for animal 1 at a point in time
+where2=list()
+for(i in 1:length(bins)){
+	where1[[i]]=array()
+	where2[[i]]=array()
+	}
+
+for(k in 1:runs){
[…]

### 2014-10-22 old_code/signal_model_07_13.R

@@ -1,0 +1,257 @@
+re1<-1 #rate of error in animal 1
+re2<-re1 #rate of error in animal 2
+me1<-0 #mean error in animal 1
+me2<-me1 #mean error in animal 2
+se1<-1 #variance of error in animal 1
+se2<-se1 #variance of error in animal 2
+
+rf<-1 #rate of fighting
+mf<-0.1 #mean fight outcome
+sf<-0.0 #variance of fight outcome
+
+bs1=0 #boost to animal 2 when animal 1 signals
+bs2=0 #boost to animal 1 when animal 2 signals
+
+rs<-1 #maximum signaling rate
+
+T<-2.5  #threshold
+
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	return(rs*(-erf((x-(-T)))+1)/2)
+	}
+
+#for drift diffusion model (size of error does not depend on state)	
+error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=1/(1+exp(2*z*a)) #from psych.pdf
+#	x=1-(1-exp(-2*(me1+mf)*(se1^2+sf^2)))/(1-exp(-4*(me1+mf)*(se1^2+sf^2))) #from stoch calc class
+	return(x)
+	}
+	
+dec_time<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=z*tanh(z*a)
+	return(x)
+	}
+
+timesteps=100 #how many reactions to have
+runs=500 #how many separate runs to do
+deltat=0.01
+bins=seq(0,timesteps*deltat,by=deltat) #bin time so we can compare evolution in time between runs
+
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2){ #Gillespie algorithm
+
+
+hitting=NULL
+decision=NULL
+switches=NULL
+
+where1=list() #each element of the list will be an array of dominance values for animal 1 at a point in time
+where2=list()
+for(i in 1:length(bins)){
+	where1[[i]]=array()
+	where2[[i]]=array()
+	}
+
+for(k in 1:runs){
[…]

### 2014-10-22 old_code/signal_model_07_19.R

@@ -1,0 +1,320 @@
+#needed to run:
+erf<-function(x){
+	return(2*pnorm(x*sqrt(2))-1)
+	}
+
+f<-function(x){
+	rs=1
+	return(rs*(-erf(10*(x-(-T)))+1)/2)
+	}
+
+#for drift diffusion model (size of error does not depend on state)	
+error_rate<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=1/(1+exp(2*z*a)) #from psych.pdf
+#	x=1-(1-exp(-2*(me1+mf)*(se1^2+sf^2)))/(1-exp(-4*(me1+mf)*(se1^2+sf^2))) #from stoch calc class
+	return(x)
+	}
+	
+dec_time<-function(re1,re2,me1,me2,se1,se2,rf,mf){
+	z=T/(me1+mf)
+	a=(me1+mf)^2/(se1^2+sf^2)
+	x=z*tanh(z*a)
+	return(x)
+	}
+
+
+################
+################
+################
+B<-function(re1,re2,me1,me2,se1,se2,rf,mf,sf,bs1,bs2){ 
+
+rs=1 # max signaling rate
+
+hitting=NULL
+decision=NULL
+switches=NULL
+
+where1=list() #each element of the list will be an array of dominance values for animal 1 at a point in time
+where2=list()
+for(i in 1:length(bins)){
+	where1[[i]]=array()
+	where2[[i]]=array()
+	}
+
+for(k in 1:runs){
+time=array(,timesteps+1) #keep track of timepoints
+time[1]=0 #initialize
+X=array(,c(2,timesteps+1))
+X[,1]=c(0,0)
+
+for(i in 1:timesteps+1){
+	z=rnorm(5,0,1)
+	X[1,i]=X[1,i-1]+(me1*X[1,i-1]+mf+bs2*f(X[2,i-1]))*deltat+sf*sqrt(deltat)*z[1]+se1*sqrt(deltat)*z[2]#+bs2*sqrt(f(X[2,i-1]))*z[3]
[…]

### 2014-10-22 old_code/signaling_dynamics.R

@@ -1,0 +1,110 @@
+a=.1
+b=10
+c=1
+
+
+
+delta=0
+x=seq(a-b+delta,a+b-delta,length.out=100)
+y=seq(-a-b+delta,-a+b-delta,length.out=100)
+
+fixedy<-function(y,a,b,c){
+	r=-a+b-b*tanh(c*(a+b-b*tanh(c*(y+1))+1))-y
+	return(r)
+}
+
+fixedx<-function(y,a,b,c){
+	r=a+b-b*tanh(c*(y-1))
+	return(r)
+}
+
+# yvals<-function(a,b,c){
+	# d=2*b/10
+	# zeros=array(0,3)
+	# zeros[1]=uniroot(fixedy,lower=-a-b-1,upper=-a-b+d,a=a,b=b,c=c)$root
+	# if(sign(fixedy(-a-b+d,a,b,c))==sign(fixedy(-a+b-d,a,b,c))){
+		# zeros[2]=zeros[1]
+		# zeros[3]=zeros[1]
+	# } else{
+	# zeros[2]=uniroot(fixedy,lower=-a-b+d,upper=-a+b-d,a=a,b=b,c=c)$root
+	# zeros[3]=uniroot(fixedy,lower=-a+b-d,upper=-a+b+1,a=a,b=b,c=c)$root
+	# }
+	# return(zeros)
+# }
+
+yvals<-function(a,b,c){
+	yvec=matrix(seq(-a-1,2*b,length.out=100),nrow=1)
+	seek=apply(yvec,2,fixedy,a=a,b=b,c=c)
+	d=.1
+	zeros=array(0,3)
+	if(sum(abs(diff(seek>0)))==1){
+		zeros[1]=uniroot(fixedy,lower=-a-1,upper=2*b,a=a,b=b,c=c)$root
+		zeros[2]=zeros[1]
+		zeros[3]=zeros[1]
+	} else{
+	Y1=min(which(seek<0))
+	Y2=min(which(seek[-(1:Y1)]>0))+Y1
+	zeros[1]=uniroot(fixedy,lower=-a-1,upper=yvec[Y1],a=a,b=b,c=c)$root
+	zeros[2]=uniroot(fixedy,lower=yvec[Y1],upper=yvec[Y2],a=a,b=b,c=c)$root
+	zeros[3]=uniroot(fixedy,lower=yvec[Y2],upper=2*b,a=a,b=b,c=c)$root
+	}
[…]

### 2014-10-22 old_code/timefunx.m

@@ -1,0 +1,6 @@
+function x=timefunx(y)
+global f b eps t e 
+c = 1/e*(f-b/(2*eps)*(b-f)/e-b*t/(2*eps)+b/2);
+ty = 1/e*log((y-(b-f)/e)/(-t-eps+4*eps*f/(b-2*eps*e)-(b-f)/e));
+x = (-t-eps-c)*(y-(b-f)/e)/(-t-eps+4*eps*f/(b-2*eps*e)-(b-f)/e)+c+b/(2*eps)*(y-(b-f)/e).*ty;
+end

### 2014-10-22 old_code/timefuny.m

@@ -1,0 +1,6 @@
+function y=timefuny(x)
+global f b eps t e 
+c=1/e*(-f-b/(2*eps)*f/e-b*t/(2*eps)+b/2);
+tx=1/e*log((x-f/e)/(-t+eps-4*eps*f/(b-2*eps*e)-f/e));
+y = (-t+eps-c)*(x-f/e)/(-t+eps-4*eps*f/(b-2*eps*e)-f/e)+c+b/(2*eps)*(x-f/e).*tx;
+end

### 2014-10-22 old_code/wireframe_bifurcation.m

@@ -1,0 +1,25 @@
+global e eps;
+eps = .5;
+e = 1;
+bvals=linspace(0,4*eps*e,30);
+tvals=linspace(0,3*eps,30);
+threshold = meshgrid(bvals,tvals);
+for i=1:length(bvals)
+    for j=1:length(tvals)
+        threshold(j,i)=e*(tvals(j)+eps);
+    end
+end
+figure
+surf(tvals,bvals,transpose(threshold))
+hold on
+
+for i=1:length(bvals)
+    for j=1:length(tvals)
+        if tvals(j)>eps
+            threshold(j,i)=e*(tvals(j)-eps);
+        else threshold(j,i)=abs(e*(2*eps*e-bvals(i))/(2*eps*e+bvals(i))*(tvals(j)-eps));
+        end
+    end
+end
+
+surf(tvals,bvals,transpose(threshold))

### 2014-10-22 optimization_for_group_distribution.m

@@ -1,7 +1,7 @@
 
 %%
-its=10;
-N=30;
+opt_its=10;
+N=10;
 
 fighting_abilities=20*rand(1,N);
 % fighting_abilities=normrnd(0,10,1,N);
@@ -23,7 +23,7 @@ for i=1:N
 end
 
 c1=0;
-c2=1;
+c2=.1;
 c3=1;
 
 %twomat=zeros(Nl,Nd,Nt,Nt,2);
@@ -39,10 +39,10 @@ perf(2,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(twomat(:,1,:,:,1));
 %%
 l=2;
 
-Tvals=zeros(N,its);
-Tvals(:,1)=ones(N,1);
+Tvals=zeros(N,opt_its);
+Tvals(:,1)=2*ones(N,1);
 % Tvals(:,1)=randi([1 Nt],N,1);
-perfvals=zeros(N,its);
+perfvals=zeros(N,opt_its);
 
 for i=1:N
     opp_thresh=Tvals([1:(i-1),(i+1):N],1);
@@ -59,7 +59,7 @@ for i=1:N
 end
 
 count=1;
-while count<=its
+while count<=opt_its
     for i=1:N
         opp_thresh=Tvals([1:(i-1),(i+1):N],count);
         ds=dmat(i,[1:(i-1),(i+1):N]);
@@ -81,7 +81,7 @@ while count<=its
     end
     if sum(Tvals(:,count+1)==Tvals(:,count))==N
         maxit=count+2;
-        count=its+1;
+        count=opt_its+1;
     end
     count=count+1;
 end
@@ -119,7 +119,8 @@ imagesc(timemat)
 colorbar
 
 subplot(2,2,3)
-hist(Tvals(:,end))
+% hist(Tvals(:,end))
+hist(sum(probmat,2))
 
 subplot(2,2,4)
 plot(fighting_abilities,sum(probmat,2))

### 2014-10-22 optimization_group.m

@@ -1,0 +1,65 @@
+N=50;
+
+% varval=.1;
+% fighting_abilities=randn(1,N)*sqrt(varval);
+
+fighting_abilities=rand(1,N);
+fighting_abilities=sort(fighting_abilities,'descend');
+
+Lvals=[-.5:.1:0];
+NL=length(Lvals);
+group_learning=zeros(3,NL);
+
+for l=1:NL
+    Lx=Lvals(l);
+    Ly=Lvals(l);
+%% define function from dominance to probability of signaling
+
+deps=.05;
+dvals=0:deps:1;
+Nd=length(dvals);
+fun_d=zeros(3,Nd);
+
+for j=1:Nd
+    d=dvals(j);
+    [x,y,z]=solve_pde(Lx,Ly,d,5,5,.05,.05,1,1);
+    fun_d(:,j)=[x; y; z];
+end
+
+%% generate 'true' dominance matrix 
+
+dvec=zeros(1,N*(N-1)/2);
+probs=dvec;
+times=dvec;
+
+k=0;
+for i=1:N
+    for j=(i+1):N
+        k=k+1;
+        diff=fighting_abilities(i)-fighting_abilities(j);
+        d=exp(diff)/(exp(diff)+1);
+        d=floor(round(d/deps))*deps;
+        dvec(k)=d;
+        probs(k)=fun_d(1,abs(d-dvals)<0.001);
+        times(k)=fun_d(2,abs(d-dvals)<0.001);
+    end
+end
+
+%% simulate signaling matrices
+
+its=1000;
+metrics=zeros(2,its);
+
+for r=1:its
+    flips=rand(1,N*(N-1)/2);
+    right=zeros(1,N*(N-1)/2);
+    right(flips<=probs)=1;
+    metrics(1,r)=sum(right)/length(right);
+end
+
+%%
+mean_frac_right=mean(metrics(1,:));
+mean_time=mean(times);
+
+group_learning(:,l)=[mean_frac_right,mean_time, mean_frac_right/mean_time];
+end

### 2014-10-22 optimization_group_v2.m

@@ -1,0 +1,84 @@
+N=10;
+
+% varval=.1;
+% fighting_abilities=randn(1,N)*sqrt(varval);
+
+fighting_abilities=rand(1,N);
+fighting_abilities=sort(fighting_abilities,'descend');
+
+Lvals=[-.5:.1:0];
+NL=length(Lvals);
+group_learning=zeros(3,NL);
+
+    Lx=-3;
+    Ly=Lx;
+%% define function from dominance to probability of signaling
+
+deps=.05;
+dvals=0:deps:1;
+Nd=length(dvals);
+fun_d=zeros(3,Nd);
+
+for j=1:Nd
+    d=dvals(j);
+    [x,y,z]=solve_pde(Lx,Ly,d,5,5,.05,.05,1,1);
+    fun_d(:,j)=[x; y; z];
+end
+
+%% generate 'true' dominance matrix 
+
+dvec=zeros(1,N*(N-1)/2);
+probs=dvec;
+times=dvec;
+
+k=0;
+for i=1:N
+    for j=(i+1):N
+        k=k+1;
+        diff=fighting_abilities(i)-fighting_abilities(j);
+        d=exp(diff)/(exp(diff)+1);
+        d=floor(round(d/deps))*deps;
+        dvec(k)=d;
+        probs(k)=fun_d(1,abs(d-dvals)<0.001);
+        times(k)=fun_d(2,abs(d-dvals)<0.001);
+    end
+end
+
+%% simulate signaling matrices
+
+its=1000;
+
+obs_ranks=int16.empty(N+1,0);
+metrics=zeros(1,its);
+
+for r=1:its
+    flips=rand(1,N*(N-1)/2);
+    sigmat=zeros(N,N);
+    k=1;
+    for j=1:N
+        for i=1:(j-1)
+            if flips(k)<=probs(k)
+                sigmat(i,j)=1;
+            else sigmat(j,i)=1;
+            end
+            k=k+1;
+        end
+    end
+    v=tiedrank(sum(sigmat,2));
+    l=size(obs_ranks,2);
+    replicate=0;
+    while i<=l
+        if isequal(v,obs_ranks(:,i))
+            obs_ranks(N+1,i)=obs_ranks(N+1,i);
+            replicate=1;
+            i=l+1;
+        else i=i+1;
+        end
+    end
+    if replicate==0
+        obs_ranks=[obs_ranks, [v;1]];
+    end
[…]

### 2014-10-22 output_of_2d_process.m

@@ -23,10 +23,11 @@ figure
 set(gcf,'Color','w')
 v=get(gcf,'Position');
 ratio=v(4)/v(3);
-w=6.83*2;
+w=6.83;
 h=w/2;
 set(gcf,'Units','inches');
 set(gcf,'Position',[v(1) v(2) w h]);
+marg=[.12 .09];
 
 textlabels={'a' 'b'};
 
@@ -37,7 +38,7 @@ currred=mycolors(Nd,'red');
 currblack=mycolors(Nd,'black');
 
 for k=1:2
-    subplot_tight(1,4,(1:2)+2*(k-1),[.1 .07])
+    subplot_tight(1,4,(1:2)+2*(k-1),marg)
     set(gca,'FontSize',labfontsz)
     hold on
     switch k
@@ -55,6 +56,9 @@ for k=1:2
     set(gca,'xlim',[0 5],'ylim',[.5 1]);
     ylabel('Probability of correct output','FontSize',textfontsz)
     xlabel('Expected time to output','FontSize',textfontsz)
+    xlab=get(gca,'xlabel');
+    v=get(xlab,'Position');
+    set(xlab,'Position',[v(1) .46 v(3)]);
     text(-.4, 1.03,textlabels(k),'HorizontalAlignment','center','Color','k','Clipping','off','FontSize',textfontsz)
 end
 
@@ -62,7 +66,7 @@ leg=legend([p([3:2:Nd])],['d=' num2str(domvals(3))],['d=' num2str(domvals(5))],[
 legend boxoff
 
 % 
-set(leg,'Position',[.9 .75 .1 .1])
+set(leg,'Position',[.8 .75 .1 .1])
 %     
 set(gcf,'PaperSize',[w h]);
 set(gcf,'PaperPosition',[0 0 w h]);
@@ -81,8 +85,8 @@ twoflexthresh=zeros(Nr,Nt,2);
 
 for k=1:Nr
     r=ratios(k);
-    twoflexthresh(k,:,1)=interp2(threshvals,threshvals,reshape(twomat(i,j,:,:,1),Nt,[]),threshvals,r*threshvals);
-    twoflexthresh(k,:,2)=interp2(threshvals,threshvals,reshape(twomat(i,j,:,:,2),Nt,[]),threshvals,r*threshvals);
[…]

### 2014-10-22 pairwise_dynamics.R

@@ -1,0 +1,72 @@
+setwd("/Users/eleanorbrush/Dropbox/signaling_network")
+source("gillespie.R")
+library(fields)
+library(lattice)
+
+e=5
+bs=2
+bf=1
+rf=1
+dominance=.5
+leak=.2
+ramp=.5
+maxtime=100
+
+threshvals=seq(0,20,by=.5)
+Lt=length(threshvals)
+domvals=seq(.5,1,by=0.1)
+Ld=length(domvals)
+
+exp.waits=array(,c(Lt,Ld))
+prob.weaker.sigs=array(,c(Lt,Ld))
+trans.probs=array(,c(Lt,Ld,4,4))
+trans.times=array(,c(Lt,Ld,4,4))
+for(i in 1:Lt){
+	for(j in 1:Ld){
+		thresh=threshvals[i]
+		dominance=domvals[j]
+		simulation=sims(e,maxtime,ramp,thresh,thresh,rf,dominance,leak,bs,bf)
+		s=stats(simulation)
+		exp.waits[i,j]=s$exp.wait
+		prob.weaker.sigs[i,j]=s$prob.weaker.sigs
+		trans.probs[i,j,,]=s$trans.probs
+		trans.times[i,j,,]=s$trans.times
+		
+	}
+}
+exp.waits[is.nan(exp.waits)]=NA
+prob.weaker.sigs[is.nan(prob.weaker.sigs)]=NA
+
+L=list(threshvals=threshvals,domvals=domvals,exp.waits,prob.weaker.sigs=prob.weaker.sigs,trans.probs=trans.probs,trans.times=trans.times)
+save(file="/Users/eleanorbrush/Desktop/signetsims.rdata",L)
+
+domvals=c(1-domvals[length(domvals):2],domvals)
+
+waits=L$exp.waits
+waits[is.na(waits)]=maxtime
+waits=cbind(waits[,dim(waits)[2]:2],waits)
+rownames(waits)=paste("thresh",threshvals)
+colnames(waits)=paste("dom",domvals)
+
+focalreceives=L$prob.weaker.sigs
+focalreceives[is.na(focalreceives)]=0.5
+focalreceives=cbind(1-focalreceives[,dim(focalreceives)[2]:2],focalreceives)
+rownames(right)=paste("thresh",threshvals)
+colnames(right)=paste("dom",domvals)
+
+layout(matrix(1:2,nrow=1))
+image.plot(threshvals,domvals,waits,xlab="Threshold",ylab="Probability that focal animal wins a fight")
[…]

### 2014-10-22 paradox_of_mutual_info.m

@@ -1,31 +1,22 @@
-its=2;
+function [seqoptthresh, dmatssaved, pairwiseaccuracy] = paradox_of_mutual_info(cvals,domvals,twomat)
+
+its=100;
 N=20;
 dist='unif';
+Nd=size(twomat,2);
+Nt=size(twomat,3);
 extendeddomvals=[1-domvals(end:-1:2) domvals];
 v=1:(Nd+Nd-1);
 deps=.05;
 l=1;
 opt_its=50;
 
-% c11=.4;c21ç=.6;c31=0;
-% c12=.9;c22=.1;c32=0;
-% c11=0;c21=.4;c31=.6;
-% c12=.9;c22=.1;c32=0;
-% c11=0;c21=.55;c31=.45;
-% c12=.85;c22=.2;c32=.05;
-
-c11=0;c21=.3;c31=.7;
-c12=.7;c22=.3;c32=0;
-c13=0;c23=.3;c33=.7;
-
-cvals=[c11 c21 c31; c12 c22 c32];
-% cvals=[c12,c22,c32];
-% cvals=[c11 c21 c31];
 totalc=size(cvals,1);
 
 seqoptthresh=zeros(its,totalc,N);
 dmatssaved=zeros(its,N,N);
 perfvalssaved=zeros(its,totalc,N);
+pairwiseaccuracy=zeros(its,totalc,N,N);
 
 for c=1:its
 
@@ -51,453 +42,98 @@ for c=1:its
         c2=cvals(ic,2);
         c3=cvals(ic,3);
         
+
+        perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+        perf(1,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(1-twomat(l,2:Nd,:,:,1));
+        perf(2,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(twomat(l,2:Nd,:,:,1));
+
+        perf(1,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(1-twomat(l,1,:,:,1));
+        perf(2,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(twomat(l,1,:,:,1));
+
         Tvals=zeros(N,opt_its);
[…]

### 2014-10-22 parallel_test.m

@@ -1,0 +1,11 @@
+function t=parallel_test(n,parallel)
+t=zeros(1,n);
+if strcmp(parallel,'yes')
+    parfor i=1:n
+        t(i)=i^2;
+    end
+else 
+    for i=1:n
+        t(i)=i^2;
+    end
+end

### 2014-10-22 parallel_test_fun.m

@@ -1,0 +1,363 @@
+Nb1=10;
+Nb2=10;
+l=1;
+keeptrackofinfoD=zeros(its,Nb1*Nb2);
+keeptrackofinfoR=zeros(its,Nb1*Nb2);
+keeptrackofinfoDelta=zeros(its,Nb1*Nb2);
+keeptrackofinfoH=zeros(its,Nb1*Nb2);
+keeptrackofinfoPi=zeros(its,Nb1*Nb2);
+infomatD=zeros(Nb1,Nb2);
+infomatR=zeros(Nb1,Nb2);
+infomatDelta=zeros(Nb1,Nb2);
+infomatH=zeros(Nb1,Nb2);
+infomatPi=zeros(Nb1,Nb2);
+
+powerfuns={'D','R','Delta','H','Pi'};
+numpowerfuns=length(powerfuns);
+
+skewvalsD=zeros(1,its);
+skewvalsR=zeros(1,its);
+skewvalsDelta=zeros(1,its);
+skewvalsH=zeros(1,its);
+skewvalsPi=zeros(1,its);
+finalthresh=zeros(N,its);
+
+keepprobmat=zeros(its,N*N);
+keeptimemat=zeros(its,N*N);
+keepsigmat=zeros(its,N*N);
+
+tic
+for c=1:its
+
+    opt_its=10;
+
+    fighting_abilities=zeros(1,N);
+    
+    switch dist
+        case 'unif'
+            fighting_abilities=20*rand(1,N);
+        case 'norm'
+            fighting_abilities=normrnd(0,10,1,N);
+        case 'lognorm'
+            fighting_abilities=lognrnd(0,1,1,N);
+    end
+    
+    fighting_abilities=sort(fighting_abilities,'descend');
+
+    dmat=zeros(N); %turn difference in fighting abilities into probabilitise of winning
+    extendeddomvals=[1-domvals(end:-1:2) domvals];
+    v=1:(Nd+Nd-1);
+    deps=.05;
+
+    for i=1:N
+        for j=[1:(i-1),(i+1):N]
+            diff=fighting_abilities(i)-fighting_abilities(j);
+            d=exp(diff)/(exp(diff)+1);
+            d=floor(round(d/deps))*deps;
+            dmat(i,j)=v(abs(d-extendeddomvals)<=deps/2);
+        end
+    end
+
+    %size(twomat)=[Nl,Nd,Nt,Nt,2];
+    
[…]

### 2014-10-22 pdesolver.m

@@ -1,0 +1,63 @@
+function [u]=pdesolver(a,b,c,z,deltax,bl,br)
+%PDE written as 0=(a(y+b)df/dy+c*f/dy^2
+
+
+%% setting up vectors I need
+
+xvals=-z:deltax:z;
+
+Nj=length(xvals);
+
+jvals = reshape(1:(Nj-2),1,Nj-2);
+
+%% construct adjacency matrix
+A=sparse(Nj-2,Nj-2);
+bvec=zeros(1,Nj-2);
+
+left=1;
+right=Nj-2;
+interior=2:(Nj-3);
+
+for k=interior
+    x=xvals(k+1);
+    fx=a*x+b;
+    A(k,k)=-fx/deltax-2*c/(deltax^2);
+    A(k,k+1)=fx/deltax+c/(deltax^2);
+    A(k,k-1)=c/(deltax^2);
+end
+
+for k=left
+    x=xvals(k+1);
+    fx=a*x+b;
+    A(k,k)=-fx/deltax-2*c/(deltax^2);
+    A(k,k+1)=fx/deltax+c/(deltax^2);
+    bvec(k)=c/(deltax^2)*bl;
+end
+
+for k=right
+    x=xvals(k+1);
+    fx=a*x+b;
+    A(k,k)=-fx/deltax-2*c/(deltax^2);
+    A(k,k-1)=c/(deltax^2);
+    bvec(k)=(fx/deltax+c/(deltax^2))*br;
+end
+
+%% solve for u
+%solve A*u+b=f
+
+fval=0;
+f=fval*ones(1,(Nj-2));
+
+rightvec=reshape(f-bvec,(Nj-2),1);
+u=A\rightvec;
+u=[bl; u; br];
+
+check=zeros(1,Nj-2);
+for j=2:(Nj-1)
+    check(j-1)=abs((a*xvals(j)+b)*(u(j+1)-u(j))/deltax+c*(u(j+1)+u(j-1)-2*u(j))/deltax^2-fval);
+end
+
+
+end
+
+

### 2014-10-22 powerfuns.m

@@ -1,0 +1,20 @@
+function [powervecD,powervecR,powervecDelta,powervecH,powervecPi,powervecC] = powerfuns(binsigmat,sigmat)
+weight=.05;
+N=size(sigmat,2);
+todivide=sum(sigmat,2);
+todivide(todivide==0)=1;
+entmat=sigmat./repmat(todivide,1,N);
+logmat=entmat;
+logmat(entmat~=0)=-log(entmat(entmat~=0));
+entmat=entmat.*logmat;
+
+%         powervec=sum(probmat,2);
+powervecD=sum(binsigmat,2);
+powervecR=sum(sigmat,2);
+powervecDelta=(powervecD).*(powervecR);
+powervecH=sum(entmat,2);
+powervecH=real(powervecH);
+powervecPi=(powervecR).*(powervecH);
+powervecC=eigcentrality(sigmat,weight);
+% powervecprobs=sum(probmat,2);
+% power=[powervecD,powervecR,powervecDelta,powervecH,powervecPi,powervecC];

### 2014-10-22 solve_pde.m

@@ -1,0 +1,285 @@
+function [prob time perf]=solve_pde(Lx,Ly,d,Ux,Uy,deltax,deltay,l,b)
+if nargin==2 
+    d=.5;
+    Ux=5;
+    Uy=5;
+    deltax=.01;
+    deltay=.05;
+    l=1;
+    b=1;
+end
+
+if nargin==3
+    Ux=5;
+    Uy=5;
+    deltax=.01;
+    deltay=.05;
+    l=1;
+    b=1;
+end
+
+D1=b^2;
+Dd=-b^2;
+D2=b^2;
+a=b*(2*d-1);
+
+%% setting up vectors I need
+
+xvals=Lx:deltax:Ux;
+
+yvals=Ly:deltay:Uy;
+
+Ni=length(yvals);
+Nj=length(xvals);
+
+c=xvals(2:(end))';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+xvec = cc(:)';
+
+c=[1:(Nj-1)]';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+jvals = cc(:)';
+
+c=yvals(2:(end))';
+cc=c(:,ones(Nj-1,1));
+yvec = cc(:)';
+
+c=[1:(Ni-1)]';
+cc=c(:,ones(Nj-1,1));
+ivals = cc(:)';
+
+%% construct adjacency matrix
+A=sparse((Ni-1)*(Nj-1),(Ni-1)*(Nj-1));
+
+Aleft=D1/(deltax^2);
+Adown=D2/(deltay^2);
+Aupright=2*Dd/deltax/deltay;
+
+bl_prob=0;
+br_prob=1;
+bd_prob=1;
+bu_prob=0;
+b_prob=zeros(1,(Ni-1)*(Nj-1));
+
+bl_time=0;
+br_time=0;
+bd_time=0;
+bu_time=0;
+b_time=zeros(1,(Ni-1)*(Nj-1));
+
+down=1+(1:(Nj-3))*(Ni-1);
+up=Ni-1+(1:(Nj-3))*(Ni-1);
+left=2:(Ni-2);
+right=(Nj-2)*(Ni-1)+(2:(Ni-2));
+SW=1;
+SE=(Ni-1)*(Nj-2)+1;
+NW=Ni-1;
+NE=(Ni-1)*(Nj-1);
+interior=1:((Ni-1)*(Nj-1));
[…]

### 2014-10-22 solve_pde_1d.m

@@ -1,0 +1,102 @@
+function [prob, time, perf]=solve_pde_1d(Lx,d,Ux,deltax,l,b)
+%dY=(-lY+2*b*(2d-1))dt+2*b*sqrt(d)dW1t-2*b*sqrt(1-d)dW2t
+%simplified PDE written as 0=(-ly+2*a)df/dy+4D1d^2f/dy^2
+if nargin==1 
+    d=.5;
+    Ux=5;
+    deltax=.01;
+    l=1;
+    b=1;
+end
+
+if nargin==2
+    Ux=-Lx;
+    deltax=.05;
+    l=1;
+    b=1;
+end
+%%
+D1=1/2*b^2;
+a=2*b*(2*d-1);
+
+%% setting up vectors I need
+
+xvals=Lx:deltax:Ux;
+
+Nj=length(xvals);
+
+jvals = reshape(1:(Nj-2),1,Nj-2);
+
+%% construct adjacency matrix
+A=sparse(Nj-2,Nj-2);
+
+bl_prob=0;
+br_prob=1;
+b_prob=zeros(1,Nj-2);
+
+bl_time=0;
+br_time=0;
+b_time=zeros(1,(Nj-2));
+
+left=1;
+right=Nj-2;
+interior=2:(Nj-3);
+
+for k=interior
+    x=xvals(k+1);
+    fx=-l*x+a;
+    A(k,k)=-fx/deltax-8*D1/(deltax^2);
+    A(k,k+1)=fx/deltax+4*D1/(deltax^2);
+    A(k,k-1)=4*D1/(deltax^2);
+end
+
+for k=left
+    x=xvals(k+1);
+    fx=-l*x+a;
+    A(k,k)=-fx/deltax-8*D1/(deltax^2);
+    A(k,k+1)=fx/deltax+4*D1/(deltax^2);
+    b_prob(k)=4*D1/(deltax^2)*bl_prob;
+    b_time(k)=4*D1/(deltax^2)*bl_time;
+end
+
+for k=right
+    x=xvals(k+1);
+    fx=-l*x+a;
+    A(k,k)=-fx/deltax-8*D1/(deltax^2);
+    A(k,k-1)=4*D1/(deltax^2);
+    b_prob(k)=(fx/deltax+4*D1/(deltax^2))*br_prob;
+    b_time(k)=(fx/deltax+4*D1/(deltax^2))*br_time;
+end
+
+%% solve for u
+%solve A*u+b=f
+
+fval_prob=0;
[…]

### 2014-10-22 solve_pde_1d_prob.m

@@ -1,0 +1,100 @@
+function prob =solve_pde_1d_prob(t,d,l,b,deltax)
+%dY=(-lY+2*b*(2d-1))dt+2*b*sqrt(d)dW1t-2*b*sqrt(1-d)dW2t
+%simplified PDE written as 0=(-ly+2*a)df/dy+4D1d^2f/dy^2
+Lx=-t;
+Ux=t;
+
+if nargin==1 
+    d=.5;
+    l=1;
+    b=1;
+end
+
+if nargin==2
+    l=1;
+    b=1;
+end
+%%
+D1=1/2*b^2;
+a=2*b*(2*d-1);
+
+%% setting up vectors I need
+
+xvals=Lx:deltax:Ux;
+
+Nj=length(xvals);
+
+jvals = reshape(1:(Nj-2),1,Nj-2);
+
+%% construct adjacency matrix
+A=sparse(Nj-2,Nj-2);
+
+bl_prob=0;
+br_prob=1;
+b_prob=zeros(1,Nj-2);
+
+bl_time=0;
+br_time=0;
+b_time=zeros(1,(Nj-2));
+
+left=1;
+right=Nj-2;
+interior=2:(Nj-3);
+
+for k=interior
+    x=xvals(k+1);
+    fx=-l*x+a;
+    A(k,k)=-fx/deltax-8*D1/(deltax^2);
+    A(k,k+1)=fx/deltax+4*D1/(deltax^2);
+    A(k,k-1)=4*D1/(deltax^2);
+end
+
+for k=left
+    x=xvals(k+1);
+    fx=-l*x+a;
+    A(k,k)=-fx/deltax-8*D1/(deltax^2);
+    A(k,k+1)=fx/deltax+4*D1/(deltax^2);
+    b_prob(k)=4*D1/(deltax^2)*bl_prob;
+    b_time(k)=4*D1/(deltax^2)*bl_time;
+end
+
+for k=right
+    x=xvals(k+1);
+    fx=-l*x+a;
+    A(k,k)=-fx/deltax-8*D1/(deltax^2);
+    A(k,k-1)=4*D1/(deltax^2);
+    b_prob(k)=(fx/deltax+4*D1/(deltax^2))*br_prob;
+    b_time(k)=(fx/deltax+4*D1/(deltax^2))*br_time;
+end
+
+%% solve for u
+%solve A*u+b=f
+
+fval_prob=0;
+f_prob=fval_prob*ones(1,(Nj-2));
+
[…]

### 2014-10-22 solve_pde_2d.m

@@ -1,0 +1,296 @@
+function [prob time perf]=solve_pde_2d(Lx,Ly,d,Ux,Uy,deltax,deltay,l,b)
+% dX1=(-lX1+b*(2d-1))dt+b*sqrt(d)dW1t-b*sqrt(1-d)dW2t, similar for dX2
+% simplified PDE written as
+% 0=(-lX+a)df/dx+(-ly+a)df/dy+D1d^2f/dx^2-2Ddd^2f/dxdy+D2d^2f/dy^2
+if nargin==2 
+    d=.5;
+    Ux=5;
+    Uy=5;
+    deltax=.01;
+    deltay=.05;
+    l=1;
+    b=1;
+end
+
+if nargin==3
+    Ux=5;
+    Uy=5;
+    deltax=.05;
+    deltay=.05;
+    l=1;
+    b=1;
+end
+
+D1=1/2*b^2;
+Dd=-1/2*b^2;
+D2=1/2*b^2;
+a=b*(2*d-1);
+
+%% setting up vectors I need
+
+xvals=Lx:deltax:Ux;
+
+yvals=Ly:deltay:Uy;
+
+Ni=length(yvals);
+Nj=length(xvals);
+
+c=xvals(2:(end))';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+xvec = cc(:)';
+
+c=[1:(Nj-1)]';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+jvals = cc(:)';
+
+c=yvals(2:(end))';
+cc=c(:,ones(Nj-1,1));
+yvec = cc(:)';
+
+c=[1:(Ni-1)]';
+cc=c(:,ones(Nj-1,1));
+ivals = cc(:)';
+
+%% construct adjacency matrix
+A=sparse((Ni-1)*(Nj-1),(Ni-1)*(Nj-1));
+
+Aleft=D1/(deltax^2);
+Adown=D2/(deltay^2);
+Aupright=2*Dd/deltax/deltay;
+
+bl_prob=0;
+br_prob=1;
+bd_prob=1;
+bu_prob=0;
+b_prob=zeros(1,(Ni-1)*(Nj-1));
+
+bl_time=0;
+br_time=0;
+bd_time=0;
+bu_time=0;
+b_time=zeros(1,(Ni-1)*(Nj-1));
+
[…]

### 2014-10-22 test.m

@@ -1,0 +1,129 @@
+
+%%
+axislw=1.25;
+lw=3;
+labfontsz=15 ;
+textfontsz=15;
+alpha=.7;
+ticklength=[.02,.0250];
+markersz=5;
+markercol='white';
+
+
+%%
+
+c1=0;
+c3=1;
+
+c2vals=0:.05:1;
+Nc2=length(c2vals);
+
+nasheq_ind2=cell(Nc2,Nl,Nd);
+giveup2=zeros(Nc2,Nl,Nd);
+
+objective_best2=cell(Nc2,Nl,Nd);
+
+for i=1:Nc2
+    c2=c2vals(i);
+    perf2=zeros(2,Nl,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+    perf2(1,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2))+c3*(1-twomat(:,2:Nd,:,:,1));
+    perf2(2,:,2:Nd,:,:)=c1*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2))+c3*(twomat(:,2:Nd,:,:,1));
+
+    perf2(1,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(1-twomat(:,1,:,:,1));
+    perf2(2,:,1,:,:)=c2*(twomat(:,1,:,:,2))+c3*(twomat(:,1,:,:,1));
+    
+    objective_perf2=zeros(Nl,Nd,Nt,Nt);
+    
+    objective_perf2(:,2:Nd,:,:)=max([c1 c3])*(1-twomat(:,2:Nd,:,:,1))+c2*(twomat(:,2:Nd,:,:,2));
+    
+    objective_perf2(:,1,:,:)=c2*(twomat(:,1,:,:,2));
+    
+%     for j=1:Nl
+    for j=1:2
+        for k=1:Nd
+            T1_best=zeros(1,Nt); 
+            T2_best=zeros(1,Nt);
+            for u=1:Nt
+            [~,m]=min(perf2(1,j,k,:,u));
+            T1_best(u)=m;
+            [~,m]=min(perf2(2,j,k,u,:));
+            T2_best(u)=m;
+            end
+            pp1=interp1(indices,reshape(T1_best,1,[]),'linear','pp');
[…]

### 2014-10-22 triangle_heatmap_multiple_funs.m

@@ -8,6 +8,7 @@ ticklength=[.02,.0250];
 markersz=5;
 markercol='white';
 fontname='Times New Roman';
+cornerlabels={'ER','SP','DT'};
 
 %%
 its=1000;
@@ -32,199 +33,93 @@ transformed=transform([X',Z',Y']);
 
 %%
 l=length(X);
-skewvals=zeros(6,l);
-mutvals=zeros(6 ,l);
+expowervecs=zeros(N,7,l);
+skewvals=zeros(7,l);
+mutvals=zeros(7 ,l);
 accuracy=zeros(1,l);
+time=zeros(1,l);
 
 for i=1:l
     c1=X(i);
     c2=Y(i);
     c3=Z(i);
     t=group_props_parallelized(c1,c2,c3,its,N,'unif',Nd,Nt,twomat,domvals,threshvals);
-    mutvals(:,i)=t{1};
-    skewvals(:,i)=t{2};
-    accuracy(i)=t{3};
+    expowervecs(:,:,i)=t{1};
+    mutvals(:,i)=t{2};
+    skewvals(:,i)=t{3};
+    accuracy(i)=t{4};
+    time(i)=t{5};
 end
+
+%%
+filename='triangle_heatmap_output.mat';
+save(filename,'skewvals','mutvals','accuracy','time');
 %%
 scaledskewvals=skewvals-(repmat(min(skewvals')',1,l));
 scaledskewvals=scaledskewvals./repmat(max(scaledskewvals')',1,l);
 
-scaledmutvals=mutvals-(repmat(min(mutvals')',1,l));
-scaledmutvals=scaledmutvals./repmat(max(scaledmutvals')',1,l);
+% scaledmutvals=mutvals-(repmat(min(mutvals')',1,l));
+% scaledmutvals=scaledmutvals./repmat(max(scaledmutvals')',1,l);
+scaledmutvals=mutvals-min(min(mutvals));
+scaledmutvals=scaledmutvals/max(max(scaledmutvals));
+
+[~,order]=sort(mutvals(1:6,:));
+bestfun=order(6,:);
+scaledbestfun=bestfun-min(min(bestfun));
+scaledbestfun=scaledbestfun/max(max(scaledbestfun));
+
+[~,order]=sort(skewvals(1:6,:));
+mostskewed=order(6,:);
+scaledmostskewed=mostskewed-min(min(mostskewed));
[…]

### 2014-10-22 triimage_discrete.m

@@ -1,0 +1,57 @@
+function triimage_discrete(scaledvals,vals,coords,cornerlabels,labels,fourcols)
+n=length(unique(vals));
+indices=1:n;
+
+red=evalin('base','red');
+yellow=evalin('base','yellow');
+green=evalin('base','green');
+blue=evalin('base','blue');
+textfontsz=evalin('base','textfontsz');
+labfontsz=evalin('base','labfontsz');
+fontname=evalin('base','fontname');
+powerfuns=evalin('base','powerfuns');
+powerfuns{strcmp(powerfuns,'Delta')}='\Delta';
+layers=evalin('base','layers');
+
+% c=colormap(jet(n));
+c=fourcols;
+
+l=size(coords,1);
+cstep=(coords(1,1)-coords(2,1))/2;
+
+xoffset=.25;
+yoffset=.2;
+
+% v=get(axes,'position');
+% set(axes,'position',[0.02 v(2:end)],'off');
+hold on
+axis off
+axis equal
+
+
+for i=1:l
+    p=coords(i,:);
+    valnow=scaledvals(i);
+    valcol=c(valnow==unique(scaledvals),:);
+    fill([p(1),p(1)+cstep,p(1)+cstep,p(1),p(1)-cstep,p(1)-cstep],[p(2)-2/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)+2/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep],valcol,'EdgeColor','none');
+end
+xl=get(gca,'xlim');
+set(gca,'xlim',[-.5 xl(2)]) 
+
+text(0-xoffset,-yoffset,cellstr(cornerlabels{1}),'Color','k','HorizontalAlignment','left','FontSize',textfontsz,'FontName',fontname)
+text(2+xoffset,-yoffset,cellstr(cornerlabels{2}),'Color','k','HorizontalAlignment','right','FontSize',textfontsz,'FontName',fontname)
+text(1,sqrt(3)+yoffset,cellstr(cornerlabels{3}),'HorizontalAlignment','center','Color','k','Clipping','off','FontSize',textfontsz,'FontName',fontname)
+
+% [x1,y1]=data2norm(coords(layers{10}(1),1),coords(layers{10}(1),2));
[…]

### 2014-10-22 triimage_discrete2.m

@@ -1,0 +1,65 @@
+function triimage_discrete2(scaledvals,vals,coords,cornerlabels,labels,fourcols,highlight,tohighlight,lw,pointlabels)
+n=length(unique(vals));
+indices=1:n;
+
+red=evalin('base','red');
+yellow=evalin('base','yellow');
+green=evalin('base','green');
+blue=evalin('base','blue');
+textfontsz=evalin('base','textfontsz');
+labfontsz=evalin('base','labfontsz');
+fontname=evalin('base','fontname');
+powerfuns=evalin('base','powerfuns');
+powerfuns{strcmp(powerfuns,'Delta')}='\Delta';
+layers=evalin('base','layers');
+
+% c=colormap(jet(n));
+c=fourcols;
+
+l=size(coords,1);
+cstep=(coords(1,1)-coords(2,1))/2;
+
+xoffset=.25;
+yoffset=.2;
+
+% v=get(axes,'position');
+% set(axes,'position',[0.02 v(2:end)],'off');
+hold on
+axis off
+axis equal
+
+
+for i=1:l
+    p=coords(i,:);
+    valnow=scaledvals(i);
+    valcol=c(valnow==unique(scaledvals),:);
+    fill([p(1),p(1)+cstep,p(1)+cstep,p(1),p(1)-cstep,p(1)-cstep],[p(2)-2/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)+2/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep],valcol,'EdgeColor','none');
+end
+xl=get(gca,'xlim');
+set(gca,'xlim',[-.5 xl(2)]) 
+
+text(0-xoffset,-yoffset,cellstr(cornerlabels{1}),'Color','k','HorizontalAlignment','left','FontSize',textfontsz,'FontName',fontname)
+text(2+xoffset,-yoffset,cellstr(cornerlabels{2}),'Color','k','HorizontalAlignment','right','FontSize',textfontsz,'FontName',fontname)
+text(1,sqrt(3)+yoffset,cellstr(cornerlabels{3}),'HorizontalAlignment','center','Color','k','Clipping','off','FontSize',textfontsz,'FontName',fontname)
+
+if strcmp(highlight,'on') 
+    for i=1:length(tohighlight)
[…]

### 2014-10-22 vec2indices.m

@@ -1,0 +1,9 @@
+function powerindices=vec2indices(powervec,Nb2)
+%     powerdiff=ceil(max(powervec))-floor(min(powervec))+.1;
+    %     powerbins=(floor(min(powervec))-.05):(powerdiff/Nb2):(ceil(max(powervec))+.05);
+q=quantile(powervec,0:1/Nb2:1);
+diffvec=diff(q);
+eps=min(diffvec(diffvec>=1e-5))/10;
+q(1)=q(1)-eps;q(end)=q(end)+eps;
+powerbins=q;
+[~,powerindices]=histc(powervec,powerbins);

### 2014-10-22 write_and_save_data.m

@@ -17,12 +17,18 @@ deltax1=0.01;
 threshvals1=.1:.1:8;
 Nt1=length(threshvals1);
 
-%%
+indices1=1:Nt1;
+
+
+% filename=strcat('variables','.mat');
+% save(filename,'Nl','Nd','Nt','lvals','domvals','threshvals');
+
+
 twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time
 
-for k=1:Nl
+for k=2
     l=lvals(k);
-    for u=1:Nd
+    for u=4:Nd
         d=domvals(u);
         for i=1:Nt
             T1=threshvals(i);
@@ -35,21 +41,24 @@ for k=1:Nl
     end
 end
 
-%%
-onemat=zeros(Nl,Nd,Nt1,2);
+% 
+% onemat=zeros(Nl,Nd,Nt1,Nt1,2);
+% 
+% for k=1:Nl
+%     l=lvals(k);
+%     for u=1:Nd
+%         d=domvals(u);
+%         for i=1:Nt1
+%             t1=threshvals1(i);
+%             for j=1:Nt1
+%                 t2=threshvals1(j);
+%                 [x,y,~]=solve_pde_1d(-t1,d,t2,deltax1,l,b);
+%                 onemat(k,u,i,j,:)=[x y];
+%             end
+%         end
+%     end
+% end
 
-for k=1:Nl
-    l=lvals(k);
-    for i=1:Nd
-        d=domvals(i);
-        for j=1:Nt1
-            t=threshvals1(j);
-            [x,y,~]=solve_pde_1d(-t,d,t,deltax1,l,b);
-            onemat(k,i,j,:)=[x y];
-        end
-    end
-end
 
-%%
-filename=strcat('twomat_long_run','.mat');
+filename=strcat('/Users/eleanorbrush/Desktop/twomat_long_run','.mat');
 save(filename,'twomat','lvals','domvals','threshvals','threshvals');

### 2014-10-22 write_and_save_data_parallel.m

@@ -1,0 +1,47 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+lvals=0:.1:1;
+Nl=length(lvals);
+
+domvals=.5:.05:1;
+Nd=length(domvals);
+
+threshvals=.5:.1:2;
+Nt=length(threshvals);
+
+b=1;
+deltax=0.025;
+
+% twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time
+twomat_par=zeros((Nd-3)*Nt*Nt,2);
+
+k=2;
+l=lvals(k);
+
+parfor loopvar=1:(Nd-3)*Nt*Nt
+% parfor loopvar=1:3
+% for loopvar=1
+    [u,i,j]=ind2sub([(Nd-3),Nt,Nt],loopvar);
+    u=u+3;
+    
+    d=domvals(u);
+    
+    T1=threshvals(i);
+
+    T2=threshvals(j);
+    [x,y,~]=solve_pde_2d(-T1,-T2,d,5,5,deltax,deltax,l,b);
+%     storevar=sub2ind([Nd,Nt,Nt],u,i,j);
+    twomat_par(loopvar,:)=[x y];
+       
+end
+
+twomat_par=reshape(twomat_par,Nd-3,Nt,Nt,2);
+
+save('/home/brush/signaling_network/twomat_par.mat','twomat_par')
+
+delete(dellapool);
+
+exit ;

### 2015-05-06 BrushKrakauerFlack03_31_15.tex
writing ms and supplemental and making pretty figures

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath,amsfonts}
\usepackage{custom2}
\usepackage{graphicx} 
\usepackage{epstopdf} 
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{natbib}

\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
\newtheorem{ass}{Assumption}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\title{Conflict and Strategy in Collective Computation}
\author{Eleanor Brush, David Krakauer, Jessica Flack}

\begin{document}
\maketitle

\section*{Abstract}
Function in biological systems emerges from the behavior of tens (e.g. animal social groups) to millions of components (e.g. neural systems), which use imperfect information and have only partially-aligned interests. Components make behavioral decisions in pair-wise interactions. The combined decisions produce system output. The roles of noise and error in collective computation are poorly understood. The leaky-integrator model (LIM) has been used to study neural pair-wise decision-making. Here we modify LIM to study pair-wise decision-making under conflict and error in animal social systems. We extend LIM to study the collective dynamics generating a group level output. Our model system is a macaque society. The decision at the pair-wise level is whether to emit a subordination signal. The output of the collective computation is the distribution of social power (DSP). We study how properties (weighting time, accuracy) of the decision to signal are influenced by signaling conflicts due to variation in strength of the stimulus input (here, asymmetry in fighting ability), and how these decision-level properties impact output properties. We find conflicts in the decision process strongly affect individuals for whom [the average strength of stimulus input] is large by incentivizing accurate signaling. When pairwise decisions are accurate local properties of the decision network are informative about individual state (here, fighting ability). Global properties become informative as accuracy declines, indicating that collective computation becomes nontrivial under error. DSP shape can be controlled by changing the weighting time to decision. The successful use of the LIM in the neural and social contexts suggests general principles of collective computation due to decision-making tradeoffs common across substrates and scales.

\section*{Keywords }
collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power

\section*{Introduction}

In many biological systems, functional patterns emerge from the interactions of components making decisions under uncertainty. This includes the decision-making behavior of an animal emerging from the firing decisions of billions of neurons, social structures like the distribution of power arising from the status-signaling decisions in chimpanzee and some macaque social groups, the collective motion of fish schools emerging from the velocity and alignment decisions of individual fish, and the toxins produced by quorum sensing bacteria once the bacteria have aggregated at sufficient density (Table \ref{examples}). In each of these examples the components make decisions based on accumulating information extracted from noisy environments. Their joint behavior produces an aggregate-level pattern with fitness consequences for the components and the collective. This two part process, which has now been described in a number of systems, constitutes a collective computation.

An open question is how components should choose their decision-making strategies when there are conflicts of interest between components, perceptual errors, and noisy inputs, in order to collectively compute outputs with positive payoffs. This requires understanding (1) how the strength of the  conflicts of interest and the importance given to properties of a decision, including waiting time and decision accuracy, determine what strategies should be used, (2) how these decision-making strategies impact the collective computation, and (3) how the decisions of the components combine to produce the computation.

A variety of models, including the sequential probability ratio test (SPRT) and the leaky-integrator model (LIM), have been developed to study how components choose among alternatives in a noisy environment, for example, the firing decisions of neurons during a motion dot coherence task. Both the SPRT and the LIM track the amount of evidence supporting alternative choices. The LIM has the advantage over the SPRT of allowing for memory loss but application of the LIM to explain, for example, neural firing, has been largely phenomenological. Here we develop a LIM by deriving stochastic differential equations that mechanistically specify how information is accumulated by components and used in decision-making. We then extent the LIM in two ways: we introduce a game-theoretic component and we study the network of interactions generated by our stochastic model to explore how decisions combine to produce the output and how the output is influenced by the components' strategies. We use published empirical and computational results from work on collective computation in a well-studied model system to justify the form of our equations. The model system is a captive macaque society (\emph{Macaca nemestrina}, n=48, see Appendix Sec. \ref{empirical}) that is characterized by 
%the \emph{minimum degree of relevant complexity}: 
social learning at the individual level, social structures that arise from nonlinear processes and feedback to influence individual behavior, frequent non-kin interactions and multiplayer conflict interactions 
(Appendix, Sec.\ref{XX})\cite {Flack:2007ir, Flack:2006jh, Flack:2006vi, Flack:2005ih, Flack:2005dg, Thierry:2004tj}.

\section*{Model} 
\subsection*{Model System }
\ In our model system the decision is whether one individual in a pair should emit a silent-bared teeth display signaling agreement to the subordinate role in the relationship. The decision depends on the perceived magnitude of agonistic asymmetry by the weaker animal. If the asymmetry is perceived large, the cost of subordination is smaller than the cost of continued aggression. Encoded in the directed network of subordination signaling decisions is the output of the collective computation--the distribution of social power (DSP). Social power is operationalized as the degree of agreement in the group that an individual is capable of successfully using force during fights. The DSP is known in this system to effect interaction and conflict management cost with, for example, heavy-tailed distributions making otherwise costly conflict management strategies like policing accessible at least to individuals who sit in the tail. The DSP is obtained by quantifying consensus in the network about a node's capacity to use force successfully. Consensus can be measured using several algorithms that return consensus scores for each individual. These algorithms can be viewed as alternative hypotheses for how power is computed in the system (Sec.\ref{computation}. We study how properties (weighting time, accuracy) of the decision to signal are influenced by signaling conflicts due to variation in strength of the stimulus input (here, asymmetry in fighting ability), and how these decision-level properties impact the shape of the DSP.

\subsection*{Stochastic differential equations }    
\ In chemical systems, stochastic differential equations are used to describe the dynamics of the concentrations of various solutes, and these Langevin equations can be derived from a mechanistic description of chemical reactions \cite{Gillespie:1992vn,Gillespie:2000fk}. In the decision-making literature the SDEs used to model noisy decision processes are usually formulated without any mechanistic justification. By following the mathematical derivation of the SDEs in chemical systems given by Gillespie \cite{Gillespie:2000fk},  we can derive the SDEs used in the leaky integrator model.

In a decision between two alternatives, there are two decision variables, $X_1$ and $X_2$, each indicating the evidence accumulated for one of the options. In the absence of input, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$. If there is input, each decision variable will incorporate the new evidence by responding positively to one type of input and negatively to the other.  In the case of our model system, the input is the accumulating history of fights won and lost.

$X_1$ increases by an amount $b$ when it sees evidence for option $1$ and decrements by $-b$ when it sees evidence for option $2$, and conversely for $X_2$.  In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.  In the social case, one animal's estimate of its own fighting ability increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite.  [ELEANOR- SHOULD DO WINS AND LOSSES NOT DOTS] For clarity, in the following we will refer to the inputs as left or right dots, but they could also be referred to as wins or losses]. To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
\end{align*}

Our first assumption is that the rates at which the two types of inputs occur is constant over time.  We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$, giving 
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+bN_\text{L}-bN_\text{R}
\\ X_1(t+\tau)&=(1-\ell\tau)X_2(t)-bN_\text{L}+bN_\text{R}.
\end{align*}
If events happen at a rate $r$ and the event is a left dot with probability $c$ and a right dot with probability $1-c$, then the expectation of $N_L$ and $N_R$ in a period of length $\tau$ are, respectively, $\tau r c$ and $\tau r(1-c)$. In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.  A value of $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.

If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with normal random variables with mean and variance equal to the mean of the Poisson random variables.  Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation. Let $Z_\text{L}$ and $Z_\text{R}$, be independent standard Normal random variables, i.e. with mean $0$ and standard deviation $1$, giving
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)-b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg)
\\X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)+b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg).
\end{align*}
Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{L}t-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{L}t+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\end{array}
\end{equation}
w here $dW_{\text{L}}$ and $dW_{\text{R}}$ are independent Brownian motions representing, respectively, the left and right inputs.  The assumptions about the timescales on which inputs occur are reasonable in the social system, and the successful application of this type of model to neural populations implies they are not unreasonable in that system.  In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the neural and social systems.

\subsection*{Reaching a decision }
\ There are two ways to use the decision variables $X_1$ and $X_2$ to make a decision.  If the difference in the variables can be observed, it indicates the relative strengths of the evidence for each option. If $Y=X_1-X_2$, the system should decides on $X_1$ when $Y$ is large and positive and on $X_2$ when $Y$ is large and negative.  Specifically, there are two thresholds, $T_1$ and $T_2$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_1$ and $T_2$, and if $X_2<-T_2$ the decision is for $X_1$ and if the $X_1<-T_1$ the decision is for $X_2$.

For neural systems, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}. The reduced model is easier to analyze but it assumes a neural mechanism for calculating the difference in activity rates and there is no empirical basis for this. In social systems like our model system, the one dimensional simplification implies a third party evaluating the difference in the opinions of the two animals rather than allowing for the animals to decide themselves based on their accumulating history--a decision, i.e. the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below its threshold. Our social system (and probably the neural system as well) is therefore inherently two-dimensional and so our model is formulated this way.  In Table \ref{differences}, we compare properties of the decision process in both systems.

\subsection*{Utility of decision }
\ The thresholds affect how long the decision takes and the probabilities of a given decision.  A ``good'' decision is one that reaches the correct output quickly, i.e.\ the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.  However, it is impossible to minimize both simultaneously as waiting longer and accumulating more evidence makes the decision more accurate.  We also allow for individuals to have preferences for the decisions independent of correctness.  In the social case, for example, the weaker individual may agree to signal because it has perceived a large asymmetry but its preference may be to be the dominant individual and not to signal. In the neural case, different neural populations may receive different rewards if the decision reaches different outcomes, regardless of which of the alternatives is correct. Allowing for decision preference introducing a mechanism for maximizing the probability that the preferred outcome is reached (probability of preference being chosen, $\text{PP}_i$) and captures the extent to which the weaker animal perceives the subordination contract to be less costly than continued fighting.  ELEANOR I THINK DECISION PREFERENCE SOUNDS BETTER THAN PERSONAL PREFERENCE. WE COULD CHANGE PP TO DP.

To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{PP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  In the neural system, $w_2$ depends on whatever costs the brain or the whole animal incurs by waiting for a decision. Given that $\text{PP}_1=1-\text{PP}_2$, it is impossible for them both to maximize $\text{PP}_i$, so decision preference weight $w_3$ indicates the strength of the conflict of interests between individuals and how stubborn the individuals are about their preferences. In the social system, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.

We show in the Appendix (Sec.\ref{X}) each of the three decision properties, ER, DT, and DP satisfies a partial differential equation that depends on the SDEs and the parameters of the model. There are analytical solutions to these equations when the system of SDEs is reduced to one dimension. However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.

\subsection*{Nash equilibrium thresholds }
\ In a group of individuals, each with a value $a_i$, the difficulty with which the pair $i,j$ makes a decision increases with the difference in value, i.e. $c_{ij}$ is an increasing function of $|a_i-a_j|$. In the social system, the value $a_i$ reflects an animal's fighting ability. In the neural system, the value $a_i$ is high if the property (e.g.\ left movement or red color) a neural population responds to is abundant in the visual stimulus. We assume each individual has the same decision threshold for all the decisions processes with each of its peers and, given those thresholds, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$ and a total utility given by the average of these, $\langle U_{ij}\rangle _j$. For each set of values $\{a_1,\dots,a_N\}$, we find the Nash equilibrium thresholds such that no individual has an incentive to choose another threshold to improve its utility.  Since the Nash thresholds depend on the values $\{a_1,\dots,a_N\}$,  we repeatedly draw a set of values from a uniform distribution and find the Nash thresholds for each set. Then we find the average Nash threshold as a function of an animal's rank in the set of fighting abilities.

\subsection*{Collective computation }
\label{computation}
\ The decision to signal and the decision to fire can be viewed as opinions. The decider neuron fires or not in response to some input by an upstream neuron. This decision can be viewed as the opinion of the decider node about the `value' of the upstream neuron. In the subordination signal case, the opinion of the sender is that the receiver is capable of using force. Encoded in these directed decision networks is the collective opinion or degree of consensus in the group about the value of a given node. In the social case--the degree of consensus about the ability to use force successfully gives a node's social power \cite{Brush:2013fk, Flack:2006uq}.

There are a number of algorithms that can be used to compute how much agreement there is in the decision network about the value of a component. These algorithms compute a consensus score for each node in the network. In the case of our model system, the distribution of these scores gives the distribution of power (Sec.refX) and the integration over the opinions via the algorithm is the collective computation.

Here we ask which consensus algorithm best captures the collective computation in our model system. We use the algorithms to compute a DSP from the decision network generated by our stochastic model. We ask which model generated DSP is most informative about the underlying distribution of fighting abilities assumed in the model--this is how we assess which algorithm best describes how information in the network is integrated to produce the collective computation (Appendix, Sec.\ref{X}).

The simplest algorithm is the unweighted in-degree of a node, i.e. the number of individuals deciding in favor of each individual.  The second algorithm we consider is the weighted in-degree of a node, which is a finer measure than the unweighted in-degree since it takes into consideration the strength of each decision made in favor of the node.  The third is the entropy of the distribution of the number of decisions made in favor of each individual, which gives a coarse measurement of how uniformly all other individuals in the group behave with respect to a focal individual.  The fourth is the eigenvector centrality of the decision network, which measures how central each node is in the global structure of the network.  For more details see \cite{Brush:2013fk}. Using many random sets of fighting abilities, we find the mutual information between the consensus scores from the network at each time $t$ and the true values, for each consensus formalism. We also find the average skewness of the set of consensus scores.

\ For each set of values $\{a_1,\dots,a_N\}$ and Nash thresholds $\{T_1,\dots,T_N\}$, we find the probability that each pair will reach either outcome and the expected time it will take.  We use these probabilities to generate a set of decisions between all pairs.  We use the convention that a decision is sent from $i$ to $j$ if $j$'s preference prevailed.  To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be made, at each point in time $t$, we define the weight of the edge from $i$ to $j$ to be
$$
\begin{array}{lllll}
0 & \text{ if the decision was made in favor of } i\text{ or } DT_{ij}>t 
\\t-DT_{ij} & \text{ if the decision was made in favor of } j\text{ and } DT_{ij}>t 
\end{array}
$$
Once $t$ is greater than the maximum decision time between all pairs of individuals, all edges of the network will have formed.

\section*{Results }
\ We confirm that decision-making is accurate under the two dimensional model (Appendix, Sec.\ref{onevstwoD}).

\subsection*{The decision to signal }
\subsubsection*{Conflicts of interest}
\ We start with a pair of individuals, i.e. a group of size two, to build intuition for how the optimization weights affect the Nash equilibrium thresholds.  If accuracy of the decision is important ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible. When only personal preference matters and there is a strong conflict of interest between the individuals($w_3=1)$, both will set their threshold high since there is no incentive for the individuals to stop accumulating evidence. As the importance of decision time increases (i.e. waiting costs increase), both individuals will lower their thresholds in order to reach the decision more quickly, whatever that decision may be (Figure \ref{nasheq}). The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and decreases as either decision time or personal preference become important (Figure \ref{nasheq}).

ELEANOR: I FEEL MOST OF THE NEXT TWO PARAGRAPHS SHOULD GO IN THE APPENDIX. CAN YOU REDUCE TO THREE-FOUR SENTENCES FOR THE PAPER? In a group with more than two individuals, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time. Prioritizing accuracy tends to lead strong fighters to set high thresholds and weak fighters to set low thresholds. Prioritizing decision preference tends to lead all individuals to set high thresholds (Figure \ref{groupeq}). As one might expect, weak individuals set different thresholds depending on whether accuracy is important or decision preference is important, with low thresholds set when accuracy is prioritized. Strong fighters set high thresholds in either case. This is because choosing a higher threshold increases waiting time in decisions with respect to all individuals, decreases the accuracy of its decisions with respect to stronger animals, and only slightly increases the accuracy of the decisions with respect to other weak individuals. In other words, weak individuals sacrifice accuracy in their decisions with respect to each other in order to decrease overall decision times. This leads to the counterintuitive result that when decision preference is important and by definition accuracy less important, average decision accuracy actually increases for weak pairs.

This in turn leads to a very slight increase in the average accuracy of all pairs in the group (Figure \ref{groupeq}). In summary in a group with more than two members, increasing the importance of decision time decreases the average accuracy, but increasing the strength of conflicts of interest leads to a (slight) increase in accuracy. While the improvement is small, it is very different from the decrease in accuracy caused by increasing the strength of conflicts of interest in a pair.

\subsubsection*{Waiting time }
\ When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with different abilities also take a fairly long time.  When waiting costs are high, all pairs reach decisions quickly.  At intermediate costs, most pairs can reach a decision quite quickly, except those who have similar and high abilities (Figure \ref{groupeq}).  Pairs with similar and high abilities always take as long or longer to make a decision than any other pairs do.

\subsection*{Collective computation of DSP }
\subsubsection*{Algorithms }
E: LETS MAKE A LITTLE MORE CONCISE, MOVE SOME DETAIL TO APPENDIX
\ When waiting costs are low, the ``finer'' consensus algorithms--- weighted in-degree and eigenvector centrality---outperform the ``coarser'' algorithms--in-degree and entropy---because in these circumstances, the edges of the decision network tend to be accurate and the finer measures make use of more information in the network. When there are no waiting costs and decision preference is important, all individuals have high thresholds and weighted in-degree outperforms eigenvector centrality (Figure \ref{bestmetric}).

When the pairwise decisions are less accurate, global structure of the decision network provides information about values that the individual relationships do not, so eigenvector centrality is the the most informative about the individuals' true values (Figure \ref{bestmetric}).  When accuracy is quite low, weighted in-degree and eigenvector centrality correctly rank the very strongest individuals but cannot distinguish among the weakest individuals. On the other hand, in-degree assigns the very strongest individuals the same scores but correctly resolves the differences between the weakest individuals, making it slightly more informative over all. When accuracy is very low, none of the measures are very informative but entropy correctly gives the weakest individuals low scores and the very strongest individuals high scores, giving it an edge over the other measures.

The information content of the different formalisms show different patterns as the network is forming (Figure \ref{bestmetric}). The information content of all formalisms show an initial rise as edges start to form in the decision network. When there are waiting costs and hence some inaccurate edges in the network there is a decrease in the information content of the three local measures---weighted in-degree, in-degree, and entropy---but not in eigenvector centrality. As the network becomes more fully developed, the information content of the three local measures recovers, in some cases even surpassing the information content of eigenvector centrality. Therefore, regardless of which measure is most informative on a fully developed network, as long as there are non-zero waiting costs, eigenvector centrality has the advantage of never losing information content as edges are added to the network and consistently performing well on networks that are not fully formed. We also find that for each consensus algorithm mutual information between the DSP and the underlying distribution of fighting ability increases with average pairwise accuracy. Hence the information content of the consensus scores is maintained when decision preference is prioritized over error rate.

\subsubsection*{Aggregate-level output properties as a function of pair-wise decision properties. }
MORE CONCISE?
\ Properties of the DSP like skewness effect the accessibility of conflict management strategies like policing. Hence an important question is how these properties can be tuned or controlled. For each measure of consensus in the decision network we find the average skewness of the distribution of consensus scores of a group using Nash thresholds, as they depend on optimization weight. We find that, for all measures, skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and preference (Figure \ref{skewness}). (Results for number of signalers are shown in Figure \ref{skewness} and others are similar).  When waiting costs are low and accuracy is high, the decisions accurately represent the individuals' true  abilities, so the distribution of number of signalers accurately represents the distribution of abilities, which is not very right-skewed.  When waiting costs are high and accuracy is low, decisions are so noisy that all individuals receive a similar number of decisions, giving a distribution that is quite uniform and not right-skewed.  At intermediate waiting costs, the decisions between individuals with low to middle abilities are very noisy, so that the individuals in the bottom and middle of the group receive a similar number of signals, but all individuals make accurate decisions with the top few individuals, which gives them high scores and results in a right-skewed distribution.

\section*{Discussion}
ELEANOR: We have ~9000 characters with spaces left for the discussion. Why don't you take the first stab at condensing and making a bit more general what you wrote in the first draft. Try not to repeat the results. Some of the text in your draft can go in the appendix discussion.

\section*{Acknowledgements }
This research was supported by a grant to the Santa Fe Institute from the John Templeton Foundation for the study of complexity and by ARO contract contract W911NF-13-1-0340. EB acknowledges support from NIH training grant 5T32HG003284. The authors thank Bryan Daniels, Chris Ellison, Philip Poon, and Eddie Lee for helpful discussion.

\pagebreak
\nocite{*}
\bibliographystyle{plain}
\bibliography{signaling_model}

\appendix
\section*{Supplementary Information}
\renewcommand{\thesubsection}{\Alph{subsection}.}

\section*{Model system}
\label{empirical}
We may delete this section as I covered most of it in the main paper. I will see how it all looks next iteration.

\section*{Model}
\label{modeldetails}
Additional details on the model to go here if necessary. Math issues? Relation to other models?

\section*{Results}
\subsection*{Decision to signal}
\subsubsection*{Decision making is as accurate without a third party arbiter. }
\label{onevstwoD}
\ First we compare the decision process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one. To make the systems directly comparable, we assume symmetric thresholds, i.e. $T_1=T_2$.  We find that two-dimensional process is as accurate as the one-dimensional process--for a given expected time to decision, the two-dimensional process reaches the correct decision with the same probability (Supplemental Figure \ref{dimensionality}). We use the two-dimensional system in the rest of our analyses as it is  
%Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.

\subsubsection*{Conflicts of interests}
Move some of paper results section here. Expand on nash issues here?

\subsubsection*{Decision time}
Move or expand on some of these results here.

\subsection*{Computation of output}
\subsubsection*{Algorithms }
Move some of paper results section here

\subsubsection*{Testing the algorithms and notion of `correctness' }
JESS HAS IDEAS FOR HOW TO WRITE THIS SECTION AND CAN FILL IN NEXT INTERATION

\subsubsection*{Properties of output }
Move some paper results here.

\section*{Additional discussion}
Some of the text in your first draft can go here, divided into subsections.
\label{additionaldiscussion}

\end{document}

### 2015-05-06 BrushKrakauerFlack_PNASformatted_05_01_15.tex
writing ms and supplemental and making pretty figures

%% PNAStwoS.tex
%% Sample file to use for PNAS articles prepared in LaTeX
%% For two column PNAS articles
%% Version1: Apr 15, 2008
%% Version2: Oct 04, 2013

%% BASIC CLASS FILE
\documentclass{pnastwo}

%% ADDITIONAL OPTIONAL STYLE FILES Font specification

%\usepackage{pnastwoF}

%% OPTIONAL MACRO DEFINITIONS
\def\s{\sigma}
%%%%%%%%%%%%
%% For PNAS Only:
\url{}
\copyrightyear{}
\issuedate{Issue Date}
\volume{Volume}
\issuenumber{Issue Number}
%\setcounter{page}{} %Set page number here if desired
%%%%%%%%%%%%

\begin{document}

\title{Conflict and Strategy in Collective Computation}

\author{Eleanor Brush\affil{1}{Princeton University, Princeton, NJ, USA}\affil{2}{Wisconsin Institute for Discovery, University of Wisconsin, Madison, USA},
David Krakauer \affil{2}{}\affil{3}{Santa Fe Institute, Santa Fe, USA},
\and
Jessica Flack\affil{2}{}\affil{3}{}}

\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

%%%Newly updated.
%%% If significance statement need, then can use the below command otherwise just delete it.
\significancetext{EB did blah DK did blah JC did blah}

\maketitle

\begin{article}
\begin{abstract}
{Function in biological systems emerges from the behavior of tens (e.g. animal social groups) to millions of components (e.g. neural systems), which use imperfect information and have only partially-aligned interests. Components make behavioral decisions in pair-wise interactions. The combined decisions produce system output. The roles of noise and error in collective computation are poorly understood. The leaky-integrator model (LIM) has been used to study neural pair-wise decision-making. Here we modify LIM to study pair-wise decision-making under conflict and error in animal social systems. We extend LIM to study the collective dynamics generating a group level output. Our model system is a macaque society. The decision at the pair-wise level is whether to emit a subordination signal. The output of the collective computation is the distribution of social power (DSP). We study how properties (waiting time, accuracy) of the decision to signal are influenced by signaling conflicts due to variation in strength of the stimulus input (here, asymmetry in fighting ability), and how these decision-level properties impact output properties. We find conflicts in the decision process strongly affect individuals for whom [the average strength of stimulus input] is large by incentivizing accurate signaling. When pairwise decisions are accurate local properties of the decision network are informative about individual state (here, fighting ability). Global properties become informative as accuracy declines, indicating that collective computation becomes nontrivial under error. DSP shape can be controlled by changing the waiting time to decision. The successful use of the LIM in the neural and social contexts suggests general principles of collective computation due to decision-making tradeoffs common across substrates and scales.
}
\end{abstract}

\keywords{collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power}

\dropcap{I}n many biological systems, functional patterns emerge from the interactions of components making decisions under uncertainty. This includes the decision-making behavior of an animal emerging from the firing decisions of billions of neurons, social structures like the distribution of power arising from the status-signaling decisions in chimpanzee and some macaque social groups, the collective motion of fish schools emerging from the velocity and alignment decisions of individual fish, and the toxins produced by quorum sensing bacteria once the bacteria have aggregated at sufficient density (Table \ref{examples}). In each of these examples the components make decisions based on accumulating information extracted from noisy environments. Their joint behavior produces an aggregate-level pattern with fitness consequences for the components and the collective. This two part process, which has now been described in a number of systems, constitutes a collective computation.

An open question is how components should choose their decision-making strategies when there are conflicts of interest between components, perceptual errors, and noisy inputs, in order to collectively compute outputs with positive payoffs. This requires understanding (1) how the strength of the  conflicts of interest and the importance given to properties of a decision, including waiting time and decision accuracy, determine what strategies should be used, (2) how these decision-making strategies impact the collective computation, and (3) how the decisions of the components combine to produce the computation.

A variety of models, including the sequential probability ratio test (SPRT) and the leaky-integrator model (LIM), have been developed to study how components choose among alternatives in a noisy environment, for example, the firing decisions of neurons during a motion dot coherence task. Both the SPRT and the LIM track the amount of evidence supporting alternative choices. The LIM has the advantage over the SPRT of allowing for memory loss but application of the LIM to explain, for example, neural firing, has been largely phenomenological. Here we develop a LIM by deriving stochastic differential equations that mechanistically specify how information is accumulated by components and used in decision-making. We then extend the LIM in two ways: we introduce a game-theoretic component and we study a network of pairwise decisions generated by our stochastic model to explore how pairwise interactions combine to produce the output and how the output is influenced by the components' strategies. We use published empirical and computational results from work on collective computation in a well-studied model system to justify the form of our equations. The model system is a captive macaque society (\emph{Macaca nemestrina}, n=48, see Appendix Sec. \ref{empirical}) that is characterized by 
%the \emph{minimum degree of relevant complexity}: 
social learning at the individual level, social structures that arise from nonlinear processes and feedback to influence individual behavior, frequent non-kin interactions and multiplayer conflict interactions 
(Appendix, Sec.\ref{XX})\cite {Flack:2007ir, Flack:2006jh, Flack:2006vi, Flack:2005ih, Flack:2005dg, Thierry:2004tj}.

\section{Model}
\subsection{Model System}
In our model system, a component is an individual monkey and the decision it makes is whether or not to emit a silent-bared teeth display signaling agreement to another monkey to the subordinate role in their relationship. The decision to signal depends on the perceived magnitude of the asymmetry in their fighting abilities. If the asymmetry is perceived to be large, the cost of subordination is smaller than the cost of continued aggression. Encoded in the directed network of subordination signaling decisions is the output of the collective computation: the distribution of social power (DSP). Social power is operationalized as the degree of agreement in the group that an individual is capable of successfully using force during fights. The DSP is known in this system to affect interaction and conflict management cost with, for example, heavy-tailed distributions making otherwise costly conflict management strategies like policing accessible at least to individuals who sit in the tail. The DSP is obtained by quantifying consensus in the network about a node's capacity to use force successfully. Consensus can be measured using several algorithms that return consensus scores for each individual. These algorithms can be viewed as alternative hypotheses for how power is computed in the system (Sec.\ref{computation}). We study how the strength of the stimulus (asymmetry in fighting ability) and the importance of different properties of the decision to signal (waiting time, accuracy) influence what decision-making strategies the animals should use, and how these decision-level properties impact the shape of the DSP.

\subsection{Stochastic differential equations}    
We first derive a stochastic model describing decisions made about or between pairs of components. In chemical systems, stochastic differential equations are used to describe the dynamics of the concentrations of various solutes, and these Langevin equations can be derived from a mechanistic description of chemical reactions \cite{Gillespie:1992vn,Gillespie:2000fk}. In the decision-making literature the SDEs used to model noisy decision processes are usually formulated without any mechanistic justification. By following the mathematical derivation of the SDEs in chemical systems given by Gillespie \cite{Gillespie:2000fk},  we can derive the SDEs used in the leaky integrator model. In our model system, each component is a monkey accumulating evidence about its ability relative to another monkey by keeping track of the fights it has won and lost. We could use the same derivation to reach a set of SDEs describing the firing rates of neural populations responding to left and right motion in the visual field, the system to which the LIM has been applied most frequently. For more details about this system, see Appendix Sec. XX.

In our model system, for each pair of monkeys, each has a decision variable, $X_1$ and $X_2$, indicating the evidence that it has accumulated about its ability relative to the other monkey. In the absence of new information, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ each decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$.

If there is input from the environment, each decision variable incorporates the new evidence.  In our model system,  $X_1$ increases by an amount $b$ when monkey $A$ wins a fight against monkey $B$ and decreases by $b$ when it loses, and conversely for $X_2$.  To calculate the variables at time $t+\tau$, we  count how many times each type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of times $A$  wins}[t,t+\tau)\\&-b\times\# \text{ of times $B$ wins }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of times $A$ wins }[t,t+\tau)\\&+b\times\# \text{ of times $B$ wins }[t,t+\tau). 
\end{align*}

Our first assumption is that the rates at the inputs (fights) occur is constant over time.  We can thus describe the number of each type of event with a Poisson random variable, $N_\text{A}$ and $N_\text{B}$, giving 
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+bN_\text{A}-bN_\text{B}
\\ X_1(t+\tau)&=(1-\ell\tau)X_2(t)-bN_\text{A}+bN_\text{B}.
\end{align*}
If events happen at a rate $r$ and the event is a win for $A$ with probability $c$ and a win for $B$ with probability $1-c$, then the expectation of $N_A$ and $N_B$ in a period of length $\tau$ are, respectively, $\tau r c$ and $\tau r(1-c)$. In our model system, $c$ is related to the strength of the asymmetry in the monkeys' abilities. In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.  A value of $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.

If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with normal random variables with mean and variance equal to the mean of the Poisson random variables.  Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation. Let $Z_\text{A}$ and $Z_\text{B}$, be independent standard Normal random variables, i.e. with mean $0$ and standard deviation $1$, giving
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{A}}\bigg)\\&-b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{B}}\bigg)
\\X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{A}}\bigg)\\&+b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{B}}\bigg).
\end{align*}
Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
\begin{equation*}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{A}t\\&-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{B}t
\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{A}t\\&+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{B}t,
\end{array}
\end{equation*}
where $dW_{\text{A}}$ and $dW_{\text{B}}$ are independent Brownian motions representing, respectively, the wins and losses for monkey $A$.  The assumptions about the timescales on which inputs occur are reasonable in the social system, and the successful application of this type of model to neural populations implies they are not unreasonable in that system.  In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the social and neural systems.

\subsection{Reaching a decision}
There are two ways to use the decision variables $X_1$ and $X_2$ to make a decision.  If the difference in the variables can be observed, it indicates the relative strengths of the evidence for each option. If $Y=X_1-X_2$, the system should decides on $X_1$ when $Y$ is large and positive and on $X_2$ when $Y$ is large and negative.  Specifically, there are two thresholds, $T_1$ and $T_2$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_1$ and $T_2$, and if $X_2<-T_2$ the decision is for $X_1$ and if the $X_1<-T_1$ the decision is for $X_2$.

For neural systems, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}. The reduced model is easier to analyze, but it assumes a neural mechanism for calculating the difference in activity rates and there is no empirical basis for this. In social systems like our model system, the one dimensional simplification implies a third party evaluating the difference in the evidence each animal has accumulated rather than allowing for the animals to decide themselves based on their accumulating history.  A decision, i.e. the emission of a signal from one of the two animals, is only reached when one of the two animals' decision variable goes below its threshold. Our social system (and probably the neural system as well) is therefore inherently two-dimensional and so our model is formulated this way.  In Table \ref{differences}, we compare properties of the decision process in both systems.

\subsection{Utility of decision}
The thresholds affect how long the decision takes and the probabilities of either outcome.  A ``good'' decision is one that reaches the correct output quickly, i.e.\ the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.  However, it is impossible to minimize both simultaneously as waiting longer and accumulating more evidence makes the decision more accurate.  In the social system, an accurate decision occurs when a weaker animal agrees to be subordinate to a stronger one. We also allow for individuals to have preferences for the decisions independent of correctness.  In the social case, the weaker individual may agree to signal because it has perceived a large asymmetry, but its preference may be to be the dominant individual and not to signal. In the neural case, different neural populations may receive different rewards if the decision reaches different outcomes, regardless of which of the alternatives is correct. Each component involved in the decision prefers a different outcome and wants to maximize the probability that the preferred outcome is reached (probability of decision preference being chosen, $\text{DP}_i$). 
To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{DP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  In the neural system, $w_2$ depends on whatever costs the brain or the whole animal incurs by waiting for a decision. Given that $\text{DP}_1=1-\text{DP}_2$, it is impossible for them both to maximize $\text{DP}_i$, so the decision preference weight $w_3$ indicates the strength of the conflict of interests between components and how stubborn they are about their preferences. In the social system, $w_3$ captures the benefit from being the dominant animal in a pair and the extent to which the weaker animal perceives the subordination contract to be more costly than continued fighting.

We show in the Appendix (Sec.\ref{X}) that each of the three decision properties, ER, DT, and DP satisfies a partial differential equation that depends on the SDEs and the parameters of the model. There are analytical solutions to these equations when the system of SDEs is reduced to one dimension. However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.

\subsection{Nash equilibrium thresholds}
In a group of individuals, each with a value $a_i$, the difficulty with which the pair $i,j$ makes a decision increases with the difference in value, i.e. $c_{ij}$ is an increasing function of $|a_i-a_j|$. In the social system, the value $a_i$ reflects an animal's fighting ability. In the neural system, the value $a_i$ is high if the property (e.g.\ left movement or red color) a neural population responds to is abundant in the visual stimulus. We assume each individual has the same decision threshold for all the decisions processes with each of its peers and, given those thresholds, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$ and a total utility given by the average of these, $\langle U_{ij}\rangle _j$. For each set of values $\{a_1,\dots,a_N\}$, we find the Nash equilibrium thresholds such that no individual has an incentive to choose another threshold to improve its utility.  Since the Nash thresholds depend on the values $\{a_1,\dots,a_N\}$,  we repeatedly draw a set of values from a uniform distribution and find the Nash thresholds for each set. Then, for each possible rank $i=1,\dots,N$, we find the average Nash threshold for an individual with the $i^{\text{th}}$ highest ability across these draws.

\subsection{Collective computation}
\label{computation}
In both the neural system and the social system, the decision can also be thought of as opinions. In the neural system, a decider neuron fires or not in response to some input by upstream neurons, which can be viewed the opinion of the decider node that the relative `value' of one upstream neurons is higher than another. In our model system, the opinion of the sender is that the receiver is capable of using force. In each case, the opinions  in favor of one of each pair of components form a directed network and the collective opinion or degree of consensus in the group about the value of a given component is encoded in this network. In the social case, the degree of consensus about the ability to use force successfully gives a monkey's social power and the distribution of these consensus scores across the group is the distribution of power (Sec. refX) \cite{Brush:2013fk, Flack:2006uq}.

The integration of the opinions about or between each pair of components into a set of consensus scores is the collective computation. There are a number of algorithms that can be used to compute how much consensus there is in a network about the value of each component \cite{Brush:2013fk, Flack:2006uq}.  Here we ask which consensus algorithm is most informative about the components' values and would therefore be the best with which to perform the collective computation. To do this, we use the algorithms to compute a DSP from networks generated by our stochastic model of decision-making and ask which model-generated DSP is most informative about the underlying fighting abilities in the model.

The simplest algorithm is the unweighted in-degree of a node, i.e. the number of opinions in favor of each individual.  The second algorithm we consider is the weighted in-degree of a node, which is a finer measure than the unweighted in-degree since it takes into consideration the strength of each opinion in favor of the node.  The third is the entropy of the distribution of the number of opinions in favor of each individual, which gives a coarse measurement of how uniformly all other individuals in the group behave with respect to a focal individual.  The fourth is the eigenvector centrality of the network, which measures how central each node is in the global structure of the network.  For more details see \cite{Brush:2013fk}. Using many random sets of values $\{a_1,\dots,a_n\}$ to generate thresholds $\{T_1,\dots,T_N\}$ and then a directed network, we find the mutual information between the consensus scores from the networks and the values, for each consensus algorithm. We also find the average skewness of the set of consensus scores given by each algorithm.

To create a directed network, for each set of values $\{a_1,\dots,a_N\}$ and Nash thresholds $\{T_1,\dots,T_N\}$, we find the probability that each pair will reach either outcome and the expected time it will take to make a decision.  We draw decisions between all pairs according to these probabilities, using the convention that a decision is sent from $i$ to $j$ if $j$'s preference prevailed.  To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be made, at each point in time $t$, we define the weight of the edge from $i$ to $j$ to be
$$
\begin{array}{ll}
0 & \text{if the decision is in favor of } i\text{ or } DT_{ij}>t 
\\t-DT_{ij} & \text{if the decision is in favor of } j\text{ and } DT_{ij}<t 
\end{array}
$$
Once $t$ is greater than the maximum decision time between all pairs of individuals, all edges of the network will have formed.

\section{Results}
 We confirm that the two-dimensional process is as accurate as the one-dimensional process (Appendix, Sec. \ref{onevstwoD}, Supplemental Figure \ref{dimensionality}). Because they perform equally well and because the two-dimensional system is more appropriate for the social system (and perhaps the neural system as well), we use the two-dimensional system in the rest of our analyses.

\subsection{The decision to signal}
\subsubsection{Conflicts of interest}
We start with a pair of individuals, i.e. a group of size two, to build intuition for how the optimization weights affect the Nash equilibrium thresholds.  If the accuracy of the decision is important ($w_1=1$), the Nash strategies are for the weaker individual to set his threshold as low as possible and the stronger animal to set his threshold as high as possible. This will lead to the weaker animal signaling, which is the correct outcome. When only decision preference matters and there is a strong conflict of interest between the individuals($w_3=1)$, the Nash strategies are for both to set their threshold high since there is no incentive for the individuals to stop accumulating evidence. As the importance of decision time increases (i.e. waiting costs increase), the Nash thresholds of both individuals decrease in order to reach the decision more quickly, whatever that decision may be (Figure \ref{nasheq}). The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and decreases as either decision time or decision preference becomes more important (Figure \ref{nasheq}).

In a group with more than two individuals, the Nash thresholds respond in a similar way to the optimization weights: when accuracy is most important, the Nash strategies are for strong animals to have higher thresholds and weak animals to have lower thresholds, whereas, when decision preference is most important, the Nash strategies are for all animals to have high thresholds. The change brought about by introducing more components is that, as long as there are non-zero waiting costs, the accuracy of some decisions between pairs in a group using Nash thresholds increases as decision preference becomes more important (Figure \ref{groupeq}).  In particular, when the group's Nash thresholds increases as decision preference becomes more important, low-valued individuals make more accurate decisions that take only slightly longer.

If a single individual raises its threshold, it increases its decision times with \emph{all} individuals, increases the error rate of its decisions with the individuals \emph{above} it, and decreases the error rate of its decisions with the individuals \emph{below} it. When only error rate and decision time matter, the net effect is to increase $\langle U_{ij}\rangle _j$, and the individual has no incentive to benefit the individuals below it by incurring this cost. Increasing the weight given to decision preference provides this incentive. When high- to intermediate-valued individuals raise their thresholds in response to an increase in the weight given to decision preference, they both increase the accuracy of their decisions with low-valued individuals and provide incentive for those individuals to raise their thresholds, resulting in an improvement in the accuracy of decisions made my low-valued components.

\subsubsection{Waiting time }
When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with different abilities also take a fairly long time.  When waiting costs are high, all pairs reach decisions quickly.  At intermediate costs, most pairs can reach a decision quite quickly, except those who have similar and high abilities (Figure \ref{groupeq}).  Pairs with similar and high abilities always take as long or longer to make a decision than any other pairs do.

\subsection{Collective computation of DSP}
\subsubsection{Algorithms}
For each consensus algorithm, the mutual information between the DSP and the underlying distribution of fighting ability increases with average pairwise accuracy. Hence the information content of the consensus scores produced by every algorithm is maintained when decision preference is prioritized over error rate, for any group with more than two components.

We first consider the information content of the algorithms applied to fully developed networks. When waiting costs are low, the ``finer'' consensus algorithms--- weighted in-degree and eigenvector centrality---outperform the ``coarser'' algorithms--in-degree and entropy---because, in these circumstances, the edges of the decision network tend to be accurate and the finer measures make use of more information in the network. In these relatively high-accuracy circumstances, low-valued individuals are particularly prone to make inaccurate decisions about each other when only decision preference matters. This  throws off eigenvector centrality and makes weighted in-degree outperform eigenvector centrality (Figure \ref{bestmetric}), but otherwise eigenvector centrality is the the most informative about the individuals' true values (Figure \ref{bestmetric}).  When accuracy is quite low, in-degree is the most informative by a slim margin. When accuracy is very low, none of the measures are very informative but entropy is marginally better than the rest.

When we look at how the information content of the consensus scores change as the network is developing, we find that the information content of the scores produced by eigenvector centrality always increase as more edges are formed. The information content of the other algorithms may surpass eigenvector centrality when the network becomes fully formed, but eigenvector centrality has the advantage of never losing information content and consistently performing well on networks that are not fully formed.

\subsubsection{Aggregate-level output properties as a function of pair-wise decision properties.}
Skewness, and other properties of the DSP, affect the accessibility of conflict management strategies like policing since the few components in the tail of a highly right-skewed distribution can play a different role in the group than the rest of the components. Hence an important question is how these properties can be tuned or controlled. For each measure of consensus in the decision network, we find the average skewness of the distribution of consensus scores of a group using Nash thresholds, as a function of the optimization weights. We find that, for all measures, skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and preference (Figure \ref{skewness}). (Results for unweighted in-degree are shown in Figure \ref{skewness} and others are similar).  When waiting costs are low and accuracy is high, the decisions accurately represent the individuals' true  abilities, so the distribution of number of signalers accurately represents the distribution of abilities, which is not very right-skewed.  When waiting costs are high and accuracy is low, decisions are so noisy that all individuals receive a similar number of decisions, giving a distribution that is quite uniform and not right-skewed.  At intermediate waiting costs, the decisions between individuals with low to middle abilities are very noisy, so that the individuals in the bottom and middle of the group receive a similar number of signals, but all individuals make accurate decisions with the top few individuals, which gives them high scores and results in a right-skewed distribution.

\section{Discussion}
In order for the difference between the decisions variables to be used to determine the outcome of a decision, a third party arbiter needs to be present to evaluate the value of each decision variable and find their difference.  In the brain, it might be the case that there is a third neural population whose activity depends on the two populations gathering external evidence. In a social group, there is no third party gathering information about the opinions of two monkeys in order to dictate when the decision establishing their subordination-dominance relationship should be made.  The fact that the existence of this third party arbiter does not affect the model output highlights the similarity between the two systems and it means that the results of our model without the arbiter apply in either case.

Usually, only two neural populations are included in models of neural decision making (although see XXX).  Only by considering a group with more than two components do we find the surprising result that the accuracy of a collective computation can be improved by strengthening the conflicts interests between components.  Conflicts of interest in which different components of a system prefer different outcomes are common in biological systems. Even if two individuals engaged in a decision agree that they would like to make an accurate decision,  it is rarely the case, except in experiment situations, that they would know which of the two possible decisions is the ``right'' one, since if they were to, they would not to engage in the decision process to begin with. While in some situations, conflicts of interest between members of a group can sow discord and lead to gridlock, when they incentivize individuals to make slower more accurate decisions, they improve the group's ability to perform collective computations. This suggests that if the accuracy of the collective computation leads to selection pressure at the group-level, we might expect conflicts of interest between components in a biological system to increase over evolutionary time. It also suggests that, if we were to try to design a system with which to perform collective computation, it might be advantageous to introduce conflicts of interest between its components.

As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities. In previous work, we found that both eigenvector centrality and weighted in-degree (and variants thereof) produced scores that were informative about animals' functional behavior in a group of pigtailed macaques.  This may suggest that decision time is less important than the outcome of the decision. As network scientists, if we are presented with an interaction network, we would like to know how to measure the consensus present in the network about each node.  If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes or how complete the network is, our results give a rule of thumb for the measure of consensus we might expect the components of a system are using or that would be most informative about the components' values in the system.

In our model system, we observe a right-skewed distribution of power, which allows the very few individuals in the right tail to police the aggressive interactions between other members of the group and thereby reduce conflict \cite{Brush:2013fk,Flack:2006uq,Flack:2006fk}. If skewness in the SDP results from skewness in the distribution of fighting abilities, the animals would not be able to construct a social structure with this advantage. On the other hand, if the skewness results from changing the fighting costs in the time until a subordination signal is exchanged, as we find in our model, the animals might be able to tune these costs to construct a beneficial SDP. In a spatially explicitly simulation model of a macaque social group, the ``aggressiveness'' of the animals was found to affect the shape of the dominance hierarchy \cite{Hemelrijk:2011fk}.  While the mechanisms of that model are quite different, ``aggressiveness'' is similar to disregarding fighting costs and prioritizing preference. Our results are therefore consistent with their finding that changing the underlying distribution of fighting abilities is not required to change the distribution of dominance or power.

While our finding that the consensus measures the monkeys use in our model system are the best for much of parameter space suggests a coarse agreement between our model and the data, there are discrepancies between the waiting times our model predicts and those observed in the data.  In our data, we observe that pairs of animals with high abilities reach agreements about their relative dominance but that pairs of animals with low abilities sometimes do not reach such agreements (Flack, unpublished data).  In our model, decisions between pairs who have similar and high abilities always  take as long or longer than decisions between any other pairs.  This discrepancy indicates that our simple model is missing a level of social complexity that could explain this empirical pattern.

There are (at least) two modifications to the model that we expect would produce results closer to the empirical observations.  First, in the model, the rates at which animals fight each other do not depend on the abilities of the animals.  However, it is likely that animals with lower abilities are less mobile and encounter their peers at a lower rate than animals with higher abilities who are more free to range throughout the group.  In the model, this heterogeneity in encounter rate would make the decisions between animals with low encounter rates take longer.  Second, in the model, all animals value preference, error rate, and waiting costs equally.  However, it is likely that the benefits from receiving a subordination signal from a partner and the costs of fighting depend on both animals' abilities.  If animals with low ability  do not value an agreement with other animals with similar abilities, there would be less incentive for them to reach an agreement.  These types of heterogeneity would make our model more biologically realistic when applied to the social system, but without them it is general enough to be applied to a number of systems in which collective computations are being performed.

We assumed that the individuals in a group use Nash equilibrium thresholds, but in order to do so, they would first have to determine what those are.  By assuming that they use these optimal strategies, we are able to analyze the consequences of what it means to be ``optimal," but it neglects the interesting question of how those optimal strategies might be reached. Another way to make our model more realistic would be to include a way for the individuals to learn what thresholds to use. Finding the appropriate timescale separation between the dynamics of how strategies are learned and the decision dynamics and the consequences of this additional level of learning are left for future work.

We apply a noisy learning model to two different biological systems, one neural and one social, and our model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two individuals are equal.  The similarities between these three frameworks---neural decision making, social decision making, and game theory---suggest that there are common principles of decision making through conflict that are applicable to a number of systems. Conflict and strategy are not usually ascribed to the neural model, and the war of attrition is not usually thought of as a process through which the pair of competitors reaches a decision about an external truth.  Our work is a first step toward unifying these different frameworks and better understanding the importance of conflict and strategy in collective computation.
%\appendix

\appendix[Model System] \label{empirical}
We may delete this section as I covered most of it in the main paper. I will see how it all looks next iteration.

\appendix[Model]
%\label{modeldetails}
Additional details on the model to go here if necessary. Math issues? Relation to other models?  Application to neural case.  How to reach Nash equilibrium?  We have no mechanism.  Equilibrium approach. Derivation of PDEs.

\appendix[Results]

\subsection{Decision to signal}

\subsubsection{Decision making is as accurate without a third party arbiter. }
\label{onevstwoD}
\ First we compare the decision process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one. To make the systems directly comparable, we assume symmetric thresholds, i.e. $T_1=T_2$.  We find that two-dimensional process is as accurate as the one-dimensional process--for a given expected time to decision, the two-dimensional process reaches the correct decision with the same probability (Supplemental Figure \ref{dimensionality}).

\subsubsection{Conflicts of interests}

\subsubsection{Decision time}
Move or expand on some of these results here.

\subsection{Collective computation}
\subsubsection{Algorithms}
When accuracy is quite low, weighted in-degree and eigenvector centrality correctly rank the very strongest individuals but cannot distinguish among the weakest individuals. On the other hand, in-degree assigns the very strongest individuals the same scores but correctly resolves the differences between the weakest individuals, making it slightly more informative over all. When accuracy is very low, none of the measures are very informative but entropy correctly gives the weakest individuals low scores and the very strongest individuals high scores, giving it an edge over the other measures.

The information content of the different formalisms show different patterns as the network is forming (Figure \ref{bestmetric}). The information content of all formalisms show an initial rise as edges start to form in the decision network. When there are waiting costs and hence some inaccurate edges in the network there is a decrease in the information content of the three local measures---weighted in-degree, in-degree, and entropy---but not in eigenvector centrality. As the network becomes more fully developed, the information content of the three local measures recovers, in some cases even surpassing the information content of eigenvector centrality. Therefore, regardless of which measure is most informative on a fully developed network, as long as there are non-zero waiting costs, eigenvector centrality has the advantage of never losing information content as edges are added to the network and consistently performing well on networks that are not fully formed.

\subsubsection{Testing the algorithms and notion of `correctness' }
JESS HAS IDEAS FOR HOW TO WRITE THIS SECTION AND CAN FILL IN NEXT INTERATION

\subsubsection{Properties of output}
Move some paper results here.

\appendix{Additional discussion}
Some of the text in your first draft can go here, divided into subsections.
\label{additionaldiscussion}

\begin{acknowledgments}
This research was supported by a grant to the Santa Fe Institute from the John Templeton Foundation for the study of complexity and by ARO contract contract W911NF-13-1-0340. EB acknowledges support from NIH training grant 5T32HG003284. The authors thank Bryan Daniels, Chris Ellison, Philip Poon, and Eddie Lee for helpful discussion. 
\end{acknowledgments}

%\begin{thebibliography}{10}
%\bibitem{BN}
%M.~Belkin and P.~Niyogi, {\em Using manifold structure for partially
%  labelled classification}, Advances in NIPS, 15 (2003).
%  
%\end{thebibliography}
%\nocite{*}
\bibliographystyle{plain}
\bibliography{signaling_model}

\end{article}

\begin{figure*}[ht]
\includegraphics[width=6.83in]{Figure1.pdf}
\caption{\label{nasheq} Nash thresholds depend on the optimization weights. The accuracy of a pair using Nash equilibrium thresholds decreases as the weight given to either decision time or decision preference increases.  In (a)-(d) the lines show the Nash equilibrium thresholds for a pair of animals as a function of the difficulty of the decision ($c$). The optimization weights for each panel are indicated in the simplex with the corresponding letter. In (e) the color indicates the accuracy of a decision made by a pair making a difficult decision ($c=.55$) using Nash equilibrium thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.  In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$). Parameters: $w_1=1$, $w_2=0$, $w_3=0$ (a), $w_1=0$, $w_2=0$, $w_3=1$ (b), $w_1=0$, $w_2=0.2$, $w_3=0.8$ (c), $w_1=0$, $w_2=0$, $w_3=1$ (d), $b=1$, $r=1$, $\ell=0.1$ in all panels. }
\end{figure*}

\begin{figure*}[ht]
\includegraphics[width=6.83in]{Figure2.pdf}
\caption{\label{groupeq}   The accuracy of a group using Nash thresholds decreases as the weight given to decision time increases, but is not affected or increases as the weight given to decision preference increases.  In (a) the lines show the Nash thresholds as a function of position in the group for different optimization weights,  where $1$ is the strongest individual and $20$ is the weakest. The optimization weights for each panel are indicated in the simplex with the corresponding letter. The color in the simplex indicates (b) the average accuracy of the whole group and (c) the average accuracy of the decisions between individuals in the bottom quartile of a group using Nash thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.  In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$).   In (d)-(g) the color indicates the expected time it takes each pair of components to reach a decision in a group using Nash thresholds. Parameters: $w_1=1$, $w_2=0$, $w_3=0$ (i), $w_1=0$, $w_2=0$, $w_3=1$ (ii), $w_1=0.9$, $w_2=0.1$, $w_3=0$ (iii), $w_1=0$, $w_2=0.1$, $w_3=0.9$ (iv), $N=20$, $b=1$, $r=1$, $\ell=0.1$ in all panels. }

\end{figure*}

\begin{figure}[ht]
\includegraphics[width=3.4in]{Figure3.pdf}
\caption{\label{bestmetric} The best measure of consensus in the decision network depends on the average error rate and the types of errors being made.  In (a) the color indicates the most informative consensus measure to apply to a network constructed by a group using Nash thresholds, as a function of the optimization weights. In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$). In (b)-(d) we show how the mutual information of each consensus formalism changes over time as the network forms. The optimization weights for each panel are indicated in the simplex with the corresponding letter. Parameters: $w_1=0$, $w_2=0$, $w_3=1$ (c), $w_1=0$, $w_2=0.2$, $w_3=0.8$ (d), $w_1=0$, $w_2=0.4$, $w_3=0.6$ (e),  $N=20$, $b=1$, $r=1$, $\ell=0.1$ in all panels.}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=3.4in]{Figure4}
%\includegraphics[width=\textwidth]{skewness_histograms.pdf}
\caption{\label{skewness} The average skewness of the distribution of unweighted in-degree is maximized at intermediate waiting costs. In the upper panel, the color indicates the average skewness of the distribution of a group using Nash thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$. In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$). Parameters: $N=20$, $b=1$, $r=1$, $\ell=0.1$.}
\end{figure}

\begin{table}[h]
\caption{\label{examples}{\bf  Examples of collective computation.} A collective computation can occur when a group of components, each of which gathers information and makes a decision in a noisy environment, produces an output that depends in a complex non-linear way on the individuals' decisions. This phenomenon is found in many biological systems. }
\begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}llllll@{}}
System & Information & Component &   Component decision & Collective computation \\
\cline{1-5} 
brain & moving dots& neural population & to stop firing & decision of the brain
\\  social group & fights won or lost & monkey & to emit a subordination signal & power structure
\\  quorum sensing & signaling molecule & bacteria & to produce signaling molecule & production of toxin
\\ flock movement & environment & bird & where to fly and how quickly & flock movement
\\
\hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{\label{variables}{ Variables in the model and their interpretations in neural and social systems.} }
\begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}llllll@{}}
Variable & Definition & Neural &   Social \\
\cline{1-4} 
$a$  & value &  degree to which the property  & fighting ability
\\ & & a neural population responds to 
\\ & & is present in the environment&
\\ $b$ & change due to new evidence
\\$c$ & strength of input & coherence of dots & prob. of stronger animal winning
\\ $\ell$ & leak rate
\\ $r$ & interaction rate 
\\ $T_1,T_2$ & decision thresholds
\\ $w_1$ & error rate weight & reward from being ``right" & reward from being ``right"
\\ $w_2$ & decision time weight & penalty for taking a long time & costs of fighting
\\ $w_3$ & prob. of preference weight &  & benefit from receiving signal
\\$X_1,X_2$ & decision variables &  firing rates of neural populations & evidence accumulated about relative dominance
\\
\hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{\label{differences}{\bf  Comparison of model in different application.} In its original application to neural systems, the leaky integrator model was reduced to one dimension and the only optimality criteria considered were accuracy and decision time. We study a two-dimensional version of the model and consider the possibility of decision preferences as an additional optimality criterion. This model can be applied to a collection of neural populations in the brain and to a social group of monkeys. Differences between the original neural application and the extended neural and social applications are highlighted in red.}
\begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}llllll@{}}
  &Original neural & Extended neural & Social \\
\cline{1-3} \cline{4-4} 
decision & difference hits a threshold  & ? & \fcolorbox{red}{white}{one var. hits a threshold}
\\dimensionality & $1$  & $?$ & $2$  
\\ group size & $2$ &\fcolorbox{red}{white}{$N$}& \fcolorbox{red}{white}{$N$}
\\ optimality criterion &  reward from being ``right" & reward from being ``right"& reward from being ``right"
\\ & decision time &decision time & decision time
\\ & &\fcolorbox{red}{white}{reward from  receiving signal} & \fcolorbox{red}{white}{reward from  receiving signal}
\\optimization & no &  \fcolorbox{red}{white}{yes} & \fcolorbox{red}{white}{yes} 
\\ depends on other
\\  individual's threshold
\\ \hline
\end{tabular}
\end{table}

\end{document}

### 2015-05-06 Brush_collectivecomputation_ms.tex
writing ms and supplemental and making pretty figures

--- 2014-10-22 ms_draft_10_10_14.tex
@@
 \begin{document}
-\begin{center}
-\Large
 
-\end{center}
+\title{Conflict and Strategy in Collective Computation}
+\author{Eleanor Redstart Brush, David C. Krakauer, Jessica C. Flack}
+\maketitle
 
-\vspace{0pt}
-
-\begin{center}
-{\bf \LARGE{Conflict and Strategy in Collective Computation}}
-\end{center}
-
-\section{Abstract}
-Groups of individuals making decisions in noisy environments perform collective computations in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys.
-In each of these systems, pairs of individuals make decisions about their relative values.
-We use a leaky integrator model, a set of stochastic differential equations describing how decision variables change over time, to model the decision process.
+\section*{Abstract}
+Collective computations occur in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys.
+In each of these systems, a group-level output emerges from the behavior of individuals making decisions in noisy environments.
+In particular, in the brain and in primate social groups, pairs of individuals must make decisions about their relative quality or ability.
+We use a leaky integrator model, a set of stochastic differential equations describing the dynamics of two decision variables, to model this decision process.
 The decisions made between each pair form a directed network over the whole group.
 The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values.
-We find that differences in the model when it is applied to neural and social systems do not affect the accuracy of decisions.
-We find that conflicts of interest in the decision process affect individuals with low values the most and that conflicts of interest can incentivize these individuals to make more accurate decisions.
-We find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative.
+First, we show that the same model can be used to describe both neural and social systems.
+Next, we find that conflicts of interest in the decision process most strongly affect individuals at the bottom of the group and that conflicts of interest can incentivize these individuals to make more accurate decisions.
+Then we find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative.
 Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.
 
-\section{Keywords }
-collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power, social structure
+\section*{Keywords }
+collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power
 
-\section{Introduction} 
+\section*{Introduction} 
 Many biological systems perform collective computation.
-When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether the dots are moving left or right, even if the difference in the sizes of the groups moving either direction is not very large.
-A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules only when the density is sufficiently high.
+When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether more dots are moving left or right, even when the numerical superiority of one direction is very small.
+A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules once the density is sufficiently high.
 Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion.
-In social groups of primates, the distribution of the degree of consensus about each monkeyÕs ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave.
-In each of these examples, the individual members of the group make decisions based on noisy environmental information and a function of or algorithm acting on those individual decisions determines the groupÕs output.
-This two-step process allows the group to produce a successful output or behavior, despite computational limitations of its members and the conflicts of interest between them.
+In social groups of primates, the distribution of the degree of consensus about each monkey's ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave.
+In each of these examples, the individual members of the group make decisions based on noisy environmental information and the collective computation emerges from those individual decisions.
+Through this two-step process the group produces a successful output or behavior, despite the computational limitations of its members and the conflicts of interest between them.
 While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them.
 
+%Both individuals involved in the decision value making a correct and fast decision.
+Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed.
+Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher ability.
+This introduces a conflict of interest.
+Each individual will have a strategy to negotiate accuracy, speed, and their personal preferences.
+
+In this paper, we study a group of individuals, each of which is trying to make a decision about its quality or ability relative to its peers.
 The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment in the sense that for a given amount of time the SPRT is the most accurate algorithm and for a given error rate the SPRT reaches a decision most quickly.
-The algorithm works by keeping track of the amount of evidence supporting one alternative and deciding in favor of that alternative once that amount is sufficiently high or deciding against it once that amount is sufficiently low.
-While this algorithm assumes the decider has perfect memory, in reality this is rarely the case and a simple way to model memory loss is to assume that the amount of evidence decays back toward 0 at teach point in time.
+The algorithm works by keeping track of the amount of evidence supporting each alternative and deciding in favor of one once the evidence for it is sufficiently high.
+This algorithm assumes the individuals making the decision have perfect memory, but in reality this is rarely the case.
+A simple way to model memory loss is to assume that the amount of evidence decays back toward $0$ at teach point in time.
 This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time.
@@
 This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which he has gathered more support.
-The leaky integrator model, while a powerful for modeling noisy learning dynamics, has been applied rather phenomenologically, and we seek here to justify its use with a mechanistic derivation.
 
-In this paper, we will study a group of individuals, each pair of which is trying to make a decision about their relative values.
-Each individual will have a decision variable indicating how much evidence it has accumulated and a threshold.
-If the decision variable reaches the threshold, the individual will signal that a decision has been made.
-Both individuals involved in the decision value making a correct and fast decision.
-Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed.
-Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher value.
-This introduces a conflict of interest.
-The thresholds are the strategies they use to negotiate accuracy, speed, and their personal preferences.
-
-The leaky integrator model has been applied to neural decision-making, and it successfully explains and predict performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
-In this application, each of two neural populations responds to evidence for either left or right motion.
+The leaky integrator model has been applied to neural decision-making, and it successfully predicts performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
+In this system, each of two neural populations responds to evidence for either left or right motion.
 The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high.
@@
 The bias in the SDEs is in favor of the direction in which the coherent dots are moving.
-However, this process could be used by neural populations deciding between any two possibilities, for example between up and down motion or between two colors, and a group of neural populations could be used to distinguish between or rank a whole set of possibilities.
+The model could also describe neural populations deciding between other possibilities, for example between up and down motion or between two colors.
+In order to distinguish between or rank a whole set of possibilities, the brain could rely on a group of neural populations, each responding to one of the options.
 
 While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been.
-In this application, each of two animals has a belief about its dominance with respect to the other which increases or decreases as the animal wins or loses fights against its opponent.
-The SDEs describe how these beliefs change over time.
-The source of evidence is the fight history between the two animals.
-Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise.
-The bias in the SDEs is in favor of the animal that is more likely to win.
+In this application, each of two animals has an opinion about its dominance with respect to the other which goes up or down as the animal wins or loses fights against its opponent.
+Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise that is captured by the SDEs.
+The bias in the model is in favor of the animal that is more likely to win.
 Once an animals' opinion goes below a threshold, it decides to agree to be subordinate to the other.
 A correct decision is one in which the weaker individual makes this agreement.
-In pigtailed macaques ({\it Macaca nemestrina}) and other species of macaques, there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.
+In pigtailed macaques ({\it Macaca nemestrina}), there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.
 This signal marks the end of the decision process.
 
-The leaky integrator model can thus be used to describe a decision being made between each pair of individuals in a group.
-These decisions form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects.
-In many systems, the degree of consensus in a group about each individuals' state is informative about the individuals'  true states and affects the individuals' functional behavior \cite{Brush:2013fk}.
-In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functional behaviors \cite{Brush:2013fk,Flack:2006uq}.
-The output of one of these consensus measures being applied to a decision network would be a set of consensus scores, a collective computation which would be informative about the individuals' true states.
+In both systems, the decisions between pairs of individuals form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects.
+In many systems, the degree of consensus in a group about each individual's value is informative about the individuals'  true values and affects the individuals' functional behaviors \cite{Brush:2013fk}.
+In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functions in the group \cite{Brush:2013fk,Flack:2006uq}.
+The consensus scores that result from applying one of these consensus measures to the decision network is the collective computation that we are interested in.
+There are three factors affecting this collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.
 
-In the social case, this consensus is referred to as social power.
-In macaques, power is a useful variable since it reflects the animals' true abilities and the animals themselves do use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.
-The distribution of power scores in an empirical group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions in the social group than the rest of its members \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.
-Further, the shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms.
+In the social system, this consensus is referred to as social power.
+In macaques, power is a useful variable since it reflects the animals' true fighting abilities and the animals themselves use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.
+The distribution of power scores in a group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions  than the rest of the animals \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.
+The shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms.
 This kind of consensus may be important in neural systems, as well.
@@
 
-There are thus three factors affecting the collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.
-In this paper, first we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}.
-We then use this model to answer the following questions.
+In this paper, we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}.
+While the model has been successful used to describe neural learning dynamics, its application has been rather phenomenological and this mechanistic derivation justifies its application in any system where the assumptions are not unreasonable.
+We then use the leaky integrator model to answer the following questions.
 Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process?
@@
 
-\section{Model} 
-\subsection{Stochastic differential equations }    
+\section*{Model} 
+\subsection*{Stochastic differential equations }    
 %% make debt to Gillespie clearer
-
-There are two possible types of environmental evidence: in the neural system, a dot can move either left or right and in the social system, a fight can be won by one animal or another.
-For clarity in the following we will refer to these left or right dots, but the biological interpretation of the two types of inputs depends on the system.
-There are two decision variables, $X_1$ and $X_2$.
+\ In a decision between two alternatives, there are two decision variables, $X_1$ and $X_2$, each indicating the evidence accumulated for one of the options.
 In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.
 (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$.
-
-Each decision variable responds positively to one type of input and negatively to the other.
+If there is input, each decision variable will incorporate the new evidence by responding positively to one type of input and negatively to the other.
+Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.
 In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.
 In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite.
-Neither decision variable has perfect memory.
-Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.
+For clarity, in the following we will refer to the inputs as left or right dots, but they can be interpreted as wins or losses in the social system.
 To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
@@
 
-\subsection{Reaching a decision }
-Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.
-The sign and size of the two variables indicate how much evidence there is for each possible decision outcome.
-When a lot of evidence for the first possibility has been observed, $X_1$ will be large and positive, and similarly for the second possibility.
+\subsection*{Reaching a decision }\ There are two ways to use the decision variables $X_1$ and $X_2$ to actually make a decision.
 If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility.
-If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.
-Specifically, there are two thresholds, $T_2$ and $T_1$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.
+If $Y=X_1-X_2$, the system should decides on $X_1$ when $Y$ is large and positive on $X_2$ when $Y$ is large and negative.
+Specifically, there are two thresholds, $T_1$ and $T_2$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.
 However, it may be the case that the difference in the variables cannot be observed.
 In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.
-Again, there are two thresholds, $T_2$ and $T_1$, and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.
+Again, there are two thresholds, $T_1$ and $T_2$, and if $X_2<-T_2$ the decision is for $X_1$ and if the $X_1<-T_1$ the decision is for $X_2$.
 
@@
 
-\subsection{Utility of the decision }
-The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.
-A good decision is one that reaches the correct output quickly, i.e.
-the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.
+\subsection*{Utility of the decision }
+\ The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.
+A good decision is one that reaches the correct output quickly, i.e.\ the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.
 However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.
 It may also be the case that each individual prefers one output over the other, regardless of which is ``correct''.
-In the social case, an animal would prefer that its partner agreed to be subordinate, whether or not it is truly stronger.
-In the neural case, different neural populations may receive different rewards if the decision reaches different outputs, regardless of which of the alternatives is ``true."  In these cases, a variable wants to maximize the probability that his preference is reached (probability of preference being chosen, $\text{PP}_i$).
+In the social case, each animal would prefer that its partner agree to be subordinate, whether or not it is truly stronger.
+In the neural case, different neural populations may receive different rewards if the decision reaches different outcomes, regardless of which of the alternatives is ``correct."  Therefore, an individual may want to maximize the probability that its preferred outcome is reached (probability of preference being chosen, $\text{PP}_i$).
 
@@
 In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.
-Similarly, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.
-In both cases, $w_3$ represents how stubborn the individuals are about their preferences.
-If $w_3\gneq 0$ and the individuals prefer different outcomes, neither can simultaneously be satisfied since $\text{PP}_1=1-\text{PP}_2$.
+In the neural system, $w_2$ depends on whatever costs the brain or the whole animal incurs by waiting for a decision.
+The personal preference weight $w_3$ indicates the strength of the conflict of interests between individuals since it is impossible for them both to maximize $\text{PP}_i$, given that $\text{PP}_1=1-\text{PP}_2$.
+In the social system, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.
+In both systems, $w_3$ represents how stubborn the individuals are about their preferences.
 
-As show in the Appendix, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model.
+Each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model.
 There are analytical solutions to these equations when the system of SDEs is reduced to one dimension.
@@
 
-\subsection{Optimal thresholds }
-An individual-- either a neural population or a monkey-- can be engaged in either a single decision or in many decisions at once.
-In the social system, for example, an animal wants to decide whether it is dominant or subordinate to every other member of its social group.
-For any pair of individuals, each has a utility, $U_1$ and $U_2$, and consequently each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.
-Similarly, in a group of individuals, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$.
-His total utility is given by the average of these, $\langle U_{ij}\rangle _j$.
-These lead to an optimized set of thresholds such that no individual has an incentive to choose another threshold.
+\subsection*{Nash equilibrium thresholds }
+\ In a group of individuals, each with a value $a_i$, the difficulty with which the pair $i,j$ makes a decision increases with the difference in value, i.e.
+$c_{ij}$ is an increasing function of $|a_i-a_j|$.
+In the social system, the value $a_i$ reflects an animal's fighting ability.
+In the neural system, the value $a_i$ is high if the property (e.g.\ left movement or red color) a neural population responds to is abundant in the visual stimulus.
+We assume that each individual has the same decision threshold for all the decisions processes with each of its peers and, given those thresholds, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$ and a total utility given by the average of these, $\langle U_{ij}\rangle _j$.
+For each set of values $\{a_1,\dots,a_N\}$, we find the Nash equilibrium thresholds such that no individual has an incentive to choose another threshold to improve its utility.
+Since the Nash thresholds depend on the values $\{a_1,\dots,a_N\}$,  we repeatedly draw a set of values from a uniform distribution and find the Nash thresholds for each set.
+Then we find the average Nash threshold as a function of an animal's rank in the set of fighting abilities.
 
-Since an individual is trying to use the same algorithm to learn about every one else, he is constrained to have only one decision threshold for each of these decision processes.
-In order to consistently assign the difficulties with which all pairs in a group make decisions, we assign each individual a value $a_i$ and the difficulty with which the pair $i,j$ makes a decision increases with the difference in abilities, i.e.
-$c_{ij}$ is an increasing function of $|a_i-a_j|$.
-In the social system, these values can be interpreted as the fighting abilities of the animals.The optimal decision threshold in this group context depends on the distribution of difficulties ($c_{ij}$) of the decision an individual will have to make and the thresholds of the other individuals.
-
- We use the following procedure to find optimal thresholds as a function of position in the group.
-We repeatedly draw a set of fighting abilities from a uniform distribution; we find the optimal thresholds for each set, and then find the average optimal threshold as a function of an animal's order in the set of fighting abilities.
-To find the optimal thresholds for a given set of fighting abilities, we turn the difference in fighting abilities between each pair into the difficulty of the decision they have to make ($c$).
-For an initial set of thresholds, we find $U_{ij}$ for each animal $i$ in each of its decisions and then average $\langle U_{ij}\rangle_j$ across all the animals with which it is engaged in a decision process.
-We then allow each animal to choose a threshold that optimizes $\langle U_{ij}\rangle_j$ with respect to the rest of the group's thresholds and we iterate these choices until no animal chooses a new threshold.
-
-\subsection{Collective computation }
-We are interested in how the consensus about the value of each individual emerges out of the decision process.
-(In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq}.)  Given an interaction network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}.
-These formalisms assign a consensus score to each node in the network (interpreted as a power score in the social system).
-In the following we consider four  consensus formalisms.
-The first and simplest is simply the number of nodes sending decisions to each animal (number of signalers).
-The second is the number of decisions received by each node, which is a finer measure than the first.
-The third is the entropy of the distribution of the number of decisions sent to each individual, which measures gives a coarse measurement of how uniformly all other animals in the group behave with respect to a focal individual.
-The fourth is the eigenvector centrality of the signaling network, which measures how central each node is in the global structure of the network.
-For more details see \cite{Brush:2013fk}.
-
-To use these measures, we need to construct an interaction network between individuals in the group.
-For each random set of fighting abilities, we let the animals optimize their thresholds, which give the probability that each pair will reach either decision and the expected time it will take.
+\subsection*{Collective computation }
+\ For each set of values $\{a_1,\dots,a_N\}$ and Nash thresholds $\{T_1,\dots,T_N\}$, we find the probability that each pair will reach either outcome and the expected time it will take.
 We use these probabilities to generate a set of decisions between all pairs.
-We use the convention that a decision is sent from $i$ to $j$ if in their decision $j$'s preference prevailed.
-(In the social system, $i$ sends a subordination signal to $j$ to communicate his agreement to be subordinate.) To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be sent, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.
+We use the convention that a decision is sent from $i$ to $j$ if $j$'s preference prevailed.
+To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be made, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.
 At each point in time $t$, the weight of the edge from $i$ to $j$ is $0$ if the expected decision time is greater than $t$ (i.e.
 a decision has not been made) and is given by the number of signals sent from $i$ to $j$ otherwise.
-Using many random sets of fighting abilities, for each consensus formalism and each point in time $t$, we can find the mutual information between the consensus scores from the network at $t$ and the true values.
-We can also find the average skewness of the set of consensus scores.
 
-\section{Results}
-\subsection{Decision making is as accurate without a third party arbiter. }
-First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
+Given a decision network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}.
+These formalisms assign a consensus score to each node in the network.
+In the following we consider four  consensus formalisms.
+The first and simplest is simply the unweighted in-degree of a node, i.e.
+the number of individuals deciding in favor of each individual.
+The second is the weighted in-degree of a node, i.e.
+the number of decisions made in favor of each node, which is a finer measure than the first.
+The third is the entropy of the distribution of the number of decisions made in favor of each individual, which gives a coarse measurement of how uniformly all other individuals in the group behave with respect to a focal individual.
+The fourth is the eigenvector centrality of the decision network, which measures how central each node is in the global structure of the network.
+For more details see \cite{Brush:2013fk}.
+Using many random sets of fighting abilities, we find the mutual information between the consensus scores from the network at each time $t$ and the true values, for each consensus formalism.
+We also find the average skewness of the set of consensus scores.
+
+\section*{Results}
+\subsection*{Decision making is as accurate without a third party arbiter. }
+\ First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
 To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e.
@@
 
-\subsection{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
-The Nash equilibrium thresholds for a pair of individuals depend on whether the individuals are minimizing error rate, maximizing probability of their preference, or minimizing decision time.
+\subsection*{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
+\ We start with a pair of individuals, i.e.
+a group of size two, to build intuition for how the optimization weights affect the Nash equilibrium thresholds.
 If accuracy of the decision is the only thing that matters ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible since if the decision is correct the weaker animal gives up first.
@@
 
-In a group, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.
+In a group with more than two individuals, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.
 As in a pair, prioritizing accuracy tends to lead animal's at the top of the group to set high thresholds and those at the bottom to set low thresholds.
@@
 
-\subsection{Global network structure provides the most information when accuracy is high but imperfect and when the network is developing. }
-First, we consider the mutual information between the consensus formalisms applied to a fully formed decision network and the true values.
+\subsection*{Global network structure provides the most information when accuracy is high but imperfect and when the network is developing. }
+\ First, we consider the mutual information between the consensus formalisms applied to a fully formed decision network and the true values.
 It is interesting to note that for each of the four types of formalisms we consider, there are regions of the simplex given by the optimization weights that make that the most informative formalism.
@@
 
-\subsection{The skewness of the power distribution is maximized when decisions are moderately accurate. }
-For each measure of consensus in the decision network, we find the average skewness of the distribution of consensus scores of a group using optimal thresholds, as they depend on optimization weight.
+\subsection*{The skewness of the power distribution is maximized when decisions are moderately accurate. }
+\ For each measure of consensus in the decision network, we find the average skewness of the distribution of consensus scores of a group using Nash thresholds, as they depend on optimization weight.
 We find that for all measures skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and preference (Figure \ref{skewness}).
@@
 
-\subsection{Decisions in a group take longest between individuals with similarly high abilities. }
-Difficult decisions tend to (but do not always) take longer than easier decisions.
+\subsection*{Decisions in a group take longest between individuals with similarly high abilities. }
+\ Difficult decisions tend to (but do not always) take longer than easier decisions.
 In a social system, the difficult of a decision increases as a pair becomes more similar.
@@
 
-\section{Discussion}
+\section*{Discussion}
 
@@
 In the brain, it might be the case that there is a third neural population whose activity depends on the two populations gathering external evidence, but in a social group, there is not a third party informing how a pair of animals should establish their subordination-dominance relationship.
-The fact, then, that the model does not depend on the existence of this third party arbiter means that it can be equally well applied whether or not it is present.
+The fact, then, that the existence of this third party arbiter does not affect the model output means that the results of a model without the arbiter should apply in either case and it highlights the similarity between the two systems.
 
-The decisions between pairs of individuals give rise to a network of decisions between all members of a group.
-A group of individuals will have some degree of consensus about the status of each individual.
-This consensus is the result of a collective computation that depends on individual behaviors.
-There are many different ways of measuring this consensus and each represents a different way of performing the collective computation \cite{Brush:2013fk,Flack:2006uq}.
-When the accuracy of pairwise decisions is relatively high but imperfect, then a global metric like eigenvector centrality is the most informative about the true states of the nodes.
-When pairwise accuracy is quite low, then a coarse measure like the number of signalers or the entropy of decisions received outperforms a finer measure like the number of decisions received.
-
-As network scientists, if we are presented with an interaction network, we would like to know how to measure the consensus present in the network about each node.
-If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes, this result gives a rule of thumb for how to pick a measure of consensus.
-However, since eigenvector centrality always performs optimally or close to optimally on a fully formed network and it performs optimally on a network that is still forming, eigenvector centrality would would generally be a good choice.
-
-As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities.
-In previous work, we found that both eigenvector centrality and measures of the number of decisions received produced scores that were informative about animals' functional behavior in a group of pigtailed macaques.
-This may be informative about the optimization weights the animals in that group are using.
-If so, it would suggest that decision time is less important than a combination of error rate and the probability of being dominant.
-If animals in different species do prioritize fighting costs and the benefits from being dominant (i.e.
-decision time and probability of preference)  differently, we might expect to find that those animals use different properties of their social networks to estimate the power structure of the group.
-
-Individuals in biological systems are likely to be selfish and stubborn, both in order to maximize their own welfare and because prioritizing accuracy requires knowing which the ``right'' answer is.
+Individuals in biological systems are likely to be selfish and stubborn, both in order to maximize their own welfare and because prioritizing accuracy requires knowing which the ``right'' answer is beforehand.
 The accuracy of the collective computation depends on how the individuals make the tradeoff between personal preference and accuracy, regardless of which measure is used to make the computation.
@@
 In a pair, this holds true.
-However, when there is a group of individuals all making this tradeoff, valuing preferences over accuracy improves the accuracy with which individuals in the bottom of the group resolve their differences, increasing the average accuracy of the whole group.
-This overall accuracy translates into a more accurate collective computation.
+However, when there is a group of individuals all making this tradeoff, valuing preferences over accuracy improves the accuracy with which individuals in the bottom of the group resolve their differences.
+This slightly increases the average accuracy of the whole group, which translates into a slightly more informative collective computation.
+While the increase in accuracy is not huge, it is encouraging that, as long as there are non-zero waiting costs, conflicts of interest do not undermine the group's ability to perform a collective computation.
 
-Usually only two neural populations are included in models of neural decision making (although see XXX).
+Usually, only two neural populations are included in models of neural decision making.
 Further, it is usually assumed that neural populations in the brain value accuracy, since the whole organism receives a benefit from making a correct decision in a short time.
@@
 Our result also has interesting implications for designing networks whose properties we would like to use to make computations.
+
+There are many different ways of measuring the consensus about individuals' values from a decision network.
+Each network metric represents a different way of performing a collective computation \cite{Brush:2013fk,Flack:2006uq}.
+When the accuracy of pairwise decisions is relatively high but imperfect, then a global metric like eigenvector centrality is the most informative about the true states of the nodes.
+When pairwise accuracy is quite low, a coarse measure like the number of signalers or the entropy of decisions received outperforms a finer measure like the number of decisions received.
+Eigenvector centrality tends to outperform all other measures on a network that is not fully formed.
+
+As network scientists, if we are presented with an interaction network, we would like to know how to measure the consensus present in the network about each node.
+If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes or how complete the network is, this result gives a rule of thumb for how to pick a measure of consensus.
+Since eigenvector centrality always performs optimally or close to optimally on a fully formed network and it performs optimally on a network that is still forming, eigenvector centrality seems to be generally a good algorithm to use.
+
+As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities.
+In previous work, we found that both eigenvector centrality and measures of the number of decisions received produced scores that were informative about animals' functional behavior in a group of pigtailed macaques.
+This may be tell us something about the optimization weights the animals in that group are using.
+If so, it would suggest that decision time is less important than a combination of error rate and the probability of being dominant.
+It also suggests that if animals in different species prioritize fighting costs and the benefits from being dominant differently, we might expect to find that those animals use different properties of their social networks to estimate the power structure of the group.
 
@@
 
-We consider two systems, a neural and a social, and our model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two individuals are equal.
-The similarities between these three frameworks-- neural decision making, social decision making, and game theory-- suggest that there are common principles of decision making through conflict that are applicable to a number of systems.
+We apply a noisy learning model to two different biological systems, one neural and one social, and our model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two individuals are equal.
+The similarities between these three frameworks---neural decision making, social decision making, and game theory---suggest that there are common principles of decision making through conflict that are applicable to a number of systems.
 Conflict and strategy are not usually ascribed to the neural model, and the war of attrition is not usually thought of as a process through which the pair of competitors reaches a decision about an external truth.
@@
 \includegraphics[width=\textwidth]{groupeq_thresholds.pdf}
-\caption{\label{groupeq}   The accuracy of a group using optimal thresholds decreases as the weight given to decision time increases.
-In the upper panels, the color indicates (a) the average accuracy of the whole group and (b) the average accuracy of the decisions between individuals in the bottom quartile using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
+\caption{\label{groupeq}   The accuracy of a group using Nash thresholds decreases as the weight given to decision time increases.
+In the upper panels, the color indicates (a) the average accuracy of the whole group and (b) the average accuracy of the decisions between individuals in the bottom quartile using Nash thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
@@
 Both bottom-quartile and total accuracy decrease as $w_2$ increases, and the increase in bottom-quartile accuracy as $w_3$ increases drives a slight increase in total accuracy.
-In (c)-(f) we show how the group-context optimal thresholds (left panels) and decision time for each pair (right panels) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest.
+In (c)-(f) we show how the group-context Nash thresholds (left panels) and decision time for each pair (right panels) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest.
 The optimization weights for each panel are indicated in the simplex with the corresponding letter.
@@
 
-\section{Figures}
+\section*{Acknowledgements }
+This publication was made possible through the support of a subaward from the Santa Fe Institute under a grant from the John Templeton Foundation for the study of complexity.
+The opinions expressed in this publication are those of the author(s) and do not necessarily reflect the views of the John Templeton Foundation.
+EB acknowledges support from the training grant NIH 5T32HG003284.
+
+\section*{Figures}
 \begin{figure}[ht]
-%\begin{minipage}{.66\linewidth}
-%\includegraphics[width=\linewidth]{mostinformative.pdf}
-%\end{minipage}
-%\begin{minipage}{.44\linewidth}
-%\includegraphics[width=.66\linewidth]{infogrowth.pdf}
-%\end{minipage}
 \includegraphics[width=\textwidth]{info.pdf}
@@
 \caption{\label{skewness} The average skewness of the distribution of the number of signalers is maximized at intermediate waiting costs.
-In the upper panel, the color indicates the average skewness of the distribution of a group using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
+In the upper panel, the color indicates the average skewness of the distribution of a group using Nash thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
@@
 \centering
-\caption{\label{variables}{\bf  Collective computations.} }
+\caption{\label{variables}{\bf  Variables and interpretations in neural and social systems.} }
 \ra{1.3}
 \begin{tabular}{@{}lllll@{}}
-System & Environmental information & Individual &   Individual decision & Collective computation \\
-\cmidrule{1-5} 
-brain & dots moving left or right & neural population & to stop firing & brain-level decision for left or right
-\\  social group & fights won or lost & monkey & to emit a subordination signal & power structure
-\\  quorum sensing & signaling molecule & bacteria & to produce signaling molecule & toxicity or phosphorescence
-\\ flock movement & environment & bird & where to fly and how quickly & flock movement and navigation
+Variable & Definition & Neural &   Social \\
+\cmidrule{1-4} 
+ $b$ & change caused by new evidence
+\\$c$ & strength of input & coherence of dots & prob. of stronger animal winning
+\\ $\ell$ & leak rate
+\\ $r$ & interaction rate 
+\\ $w_1$ & error rate weight & reward from being ``right" & reward from being ``right"
+\\ $w_2$ & decision time weight & penalty for taking a long time & costs of fighting
+\\ $w_3$ & prob. of preference weight & ? & benefit from receiving signal
+\\ $T_1,T_2$ & decision thresholds
+\\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
 \end{tabular}
@@
 \centering
-\caption{\label{variables}{\bf  Variables and interpretations in neural and social systems.} }
-\ra{1.3}
-\begin{tabular}{@{}lllll@{}}
-Variable & Interpretation & Neural &   Social \\
-\cmidrule{1-4} 
-& input & dots moving left or right & fights won or lost
-\\$c$ & strength of input & coherence of dots & prob. of stronger animal winning
-\\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
-\\ $w_1$ & error rate weight & reward from being ``right" & reward from being ``right"
-\\ $w_2$ & decision time weight & penalty for taking a long time & costs of fighting
-\\ $w_3$ & prob. of preference weight & ? & benefit from receiving signal
-\\ & decision & subject indicates left or right & one animal emits subordination signal
-\\ & correct decision & subject chooses correct direction & weaker animal signals
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{\label{differences}{\bf  Comparison of model in its original application to neural systems, in an extended application to neural systems, and its application to social systems.} Differences are highlighted in red.}
+\caption{\label{differences}{\bf  Comparison of model in its original application to neural systems and its application to social systems.} Differences are highlighted in red.}
 \ra{1.3}
@@
 dimensionality & $1$  && $2$
+\\ group size & $2$ && \fcolorbox{red}{white}{$N$}
 \\decision & difference hits a threshold  && \fcolorbox{red}{white}{one var. hits a threshold}
@@
 \pagebreak
-\section{Appendix}
-
-\subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}
-
-\pagebreak
 \nocite{*}

### 2015-05-06 ms_draft_10_10_14.tex
writing ms and supplemental and making pretty figures

@@
 \usepackage{latexsym}
-\usepackage{amssymb,amsmath}
+\usepackage{amssymb,amsmath,amsfonts}
 \usepackage{custom2}
-\usepackage{graphicx} % for figures
-\usepackage{epstopdf} % so can use EPS or PDF figures
+\usepackage{graphicx} 
+\usepackage{epstopdf} 
 \usepackage{caption}
@@
 \usepackage{url}
-\usepackage{amssymb,amsfonts}
 \usepackage[all,arc]{xy}
@@
 \newcommand{\Tr}{\text{Tr}}
-%\newtheorem{claim}{Claim}
 \newtheorem{ass}{Assumption}
@@
 
+\title{Conflict and Strategy in Collective Computation}
+\author{Eleanor Brush, David Krakauer, Jessica Flack}
+
 \begin{document}
-\begin{center}
-\Large
+\maketitle
 
-\end{center}
-
-\vspace{0pt}
-
-\begin{center}
-{\bf \LARGE{Conflict and Strategy in Collective Computation}}
-\end{center}
-
-\section{Abstract}
-Groups of individuals making decisions in noisy environments perform collective computations in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys.
-In each of these systems, pairs of individuals make decisions about their relative values.
-We use a leaky integrator model, a set of stochastic differential equations describing how decision variables change over time, to model the decision process.
+\section*{Abstract}
+Collective computations occur in many biological systems, ranging in size from a brain made of billions of neurons to a social group made up of tens of monkeys.
+In each of these systems, a group-level output emerges from the behavior of individuals making decisions in noisy environments.
+In particular, in the brain and in primate social groups, pairs of individuals must make decisions about their relative relevance or ability.
+We use a leaky integrator model, a set of stochastic differential equations describing the dynamics of two decision variables, to model this decision process.
 The decisions made between each pair form a directed network over the whole group.
 The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values.
-We find that differences in the model when it is applied to neural and social systems do not affect the accuracy of decisions.
-We find that conflicts of interest in the decision process affect individuals with low values the most and that conflicts of interest can incentivize these individuals to make more accurate decisions.
-We find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative.
+First, we show that the same model can be used to describe both neural and social systems.
+Next, we find that conflicts of interest in the decision process most strongly affect individuals at the bottom of the group and that conflicts of interest can incentivize these individuals to make more accurate decisions.
+Then, we find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but global properties are more informative when decisions are less accurate or when the decision network is not fully formed.
 Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.
 
-\section{Keywords }
-collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power, social structure
+\section*{Keywords }
+collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power
 
-\section{Introduction} 
+\section*{Introduction} 
 Many biological systems perform collective computation.
-When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether the dots are moving left or right, even if the difference in the sizes of the groups moving either direction is not very large.
-A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules only when the density is sufficiently high.
+A group of quorum sensing bacteria can sense its density and start to produce toxic, phosphorescent, or colorful molecules once the density is sufficiently high.
 Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion.
-In social groups of primates, the distribution of the degree of consensus about each monkeyÕs ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave.
-In each of these examples, the individual members of the group make decisions based on noisy environmental information and a function of or algorithm acting on those individual decisions determines the groupÕs output.
-This two-step process allows the group to produce a successful output or behavior, despite computational limitations of its members and the conflicts of interest between them.
+When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether more dots are moving left or right, even when the numerical superiority of one direction is very small.
+In social groups of primates, the distribution of the degree of consensus about each monkey's ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave.
+In each of these examples, the individual members of the group make decisions based on noisy environmental information and the collective computation emerges from those individual decisions.
+Through this two-step process the group produces a successful output or behavior, despite the computational limitations of its members and the conflicts of interest between them.
 While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them.
+(In Table \ref{examples} we identify the components of the collective computation--- the environmental information, the individual, the decision that each individual must make, and the resulting collective computation---in each of these systems.)
 
- The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment in the sense that for a given amount of time the SPRT is the most accurate algorithm and for a given error rate the SPRT reaches a decision most quickly.
-The algorithm works by keeping track of the amount of evidence supporting one alternative and deciding in favor of that alternative once that amount is sufficiently high or deciding against it once that amount is sufficiently low.
-While this algorithm assumes the decider has perfect memory, in reality this is rarely the case and a simple way to model memory loss is to assume that the amount of evidence decays back toward 0 at teach point in time.
+%Both individuals involved in the decision value making a correct and fast decision.
+Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed.
+Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher ability.
+This introduces a conflict of interest.
+Each individual will have a strategy to negotiate accuracy, speed, and their personal preferences.
+
+To understand these collective computations, we must understand how the individuals make decisions.
+The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment, in the sense that, for a given amount of time, the SPRT is the most accurate algorithm and, for a given error rate, the SPRT reaches a decision most quickly.
+The algorithm works by keeping track of the amount of evidence supporting each alternative and deciding in favor of one once the evidence for it is sufficiently high.
+This algorithm assumes the individuals making the decision have perfect memory, but in reality this is rarely the case.
+A simple way to model memory loss is to assume that the amount of evidence decays back toward zero at each point in time.
 This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time.
 The random walk described by the SDEs is biased by the true state of the environment.
-This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which he has gathered more support.
-The leaky integrator model, while a powerful for modeling noisy learning dynamics, has been applied rather phenomenologically, and we seek here to justify its use with a mechanistic derivation.
+This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which it has gathered more support.
 
-In this paper, we will study a group of individuals, each pair of which is trying to make a decision about their relative values.
-Each individual will have a decision variable indicating how much evidence it has accumulated and a threshold.
-If the decision variable reaches the threshold, the individual will signal that a decision has been made.
-Both individuals involved in the decision value making a correct and fast decision.
-Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed.
-Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher value.
-This introduces a conflict of interest.
-The thresholds are the strategies they use to negotiate accuracy, speed, and their personal preferences.
-
-The leaky integrator model has been applied to neural decision-making, and it successfully explains and predict performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
-In this application, each of two neural populations responds to evidence for either left or right motion.
+The leaky integrator model has been applied to neural decision-making, and it successfully predicts performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}.
+In this system, each of two neural populations responds to evidence for either left or right motion.
 The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high.
-Usually, the visual stimulus driving this decision process is a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.
+The visual stimulus driving this decision process is usually a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.
 The bias in the SDEs is in favor of the direction in which the coherent dots are moving.
-However, this process could be used by neural populations deciding between any two possibilities, for example between up and down motion or between two colors, and a group of neural populations could be used to distinguish between or rank a whole set of possibilities.
+The model could also describe neural populations deciding between other possibilities, for example between up and down motion or between two colors.
+In order to distinguish between or rank a whole set of possibilities, the brain could rely on a group of neural populations, each responding to one of the options.
 
 While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been.
-In this application, each of two animals has a belief about its dominance with respect to the other which increases or decreases as the animal wins or loses fights against its opponent.
-The SDEs describe how these beliefs change over time.
-The source of evidence is the fight history between the two animals.
-Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise.
-The bias in the SDEs is in favor of the animal that is more likely to win.
-Once an animals' opinion goes below a threshold, it decides to agree to be subordinate to the other.
+In this application, each of two animals has an opinion about its dominance with respect to the other, which goes up or down as the animal wins or loses fights against its opponent.
+Even if one animal is stronger, the outcome of a fight is affected by the identity and presence of each animal's allies, their motivation levels,  external environmental variables, and other stochastic factors, which introduce noise that is captured by the SDEs.
+The bias in the model is in favor of the animal that is more likely to win.
+Once an animal's opinion goes below a threshold, it decides to agree to be subordinate to the other.
 A correct decision is one in which the weaker individual makes this agreement.
-In pigtailed macaques ({\it Macaca nemestrina}) and other species of macaques, there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.
+In pigtailed macaques ({\it Macaca nemestrina}), there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.
 This signal marks the end of the decision process.
 
-The leaky integrator model can thus be used to describe a decision being made between each pair of individuals in a group.
-These decisions form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects.
-In many systems, the degree of consensus in a group about each individuals' state is informative about the individuals'  true states and affects the individuals' functional behavior \cite{Brush:2013fk}.
-In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functional behaviors \cite{Brush:2013fk,Flack:2006uq}.
-The output of one of these consensus measures being applied to a decision network would be a set of consensus scores, a collective computation which would be informative about the individuals' true states.
+In both systems, the decisions between pairs of individuals form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects.
+In many systems, the degree of consensus in a group about each individual's value is informative about the individuals'  true values and affects the individuals' functional behaviors \cite{Brush:2013fk}.
+In previous work, we developed a number of network metrics to measure the degree of consensus in a group about each individual and we identified a set of algorithms that produced scores that were predictive of individuals' functions in a group in multiple empirical systems \cite{Brush:2013fk,Flack:2006uq}.
+The consensus scores that result from applying one of these consensus measures to the decision network is the collective computation that we are interested in.
+There are three factors affecting this collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.
 
-In the social case, this consensus is referred to as social power.
-In macaques, power is a useful variable since it reflects the animals' true abilities and the animals themselves do use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.
-The distribution of power scores in an empirical group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions in the social group than the rest of its members \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.
-Further, the shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms.
-This kind of consensus may be important in neural systems, as well.
+In the social system, this consensus is referred to as social power.
+In macaques, power is a useful variable since it reflects the animals' true fighting abilities and the animals themselves use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.
+The distribution of power scores in a group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions  than the rest of the animals \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.
+The shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or by other mechanisms.
+Consensus may be important in neural systems, as well.
 A number of neural populations may respond to or encode different environmental stimuli or variables.
@@
 
-There are thus three factors affecting the collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.
-In this paper, first we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}.
-We then use this model to answer the following questions.
+In this paper, we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}.
+While the model has been successfully used to describe neural learning dynamics, its application has been rather phenomenological and this mechanistic derivation justifies its application in any system where the assumptions are reasonable.
+We then use the leaky integrator model to answer the following questions.
 Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process?
-What threshold strategies should individuals with conflicting interests use to make decisions in a group context?
-How do the individualsÕ threshold strategies affect the collective computation?
+What strategies should individuals with conflicting interests use to make decisions in a group context?
+How do the individuals' strategies affect the collective computation?
 What function of the decision network produces the most informative collective computation?
@@
 
-\section{Model} 
-\subsection{Stochastic differential equations }    
-%% make debt to Gillespie clearer
+\section*{Model} 
+\subsection*{Stochastic differential equations }    
+\ In chemical systems, stochastic differential equations are used to describe the dynamics of the concentrations of various solutes, and these Langevin equations can be derived from a mechanistic description of chemical reactions \cite{Gillespie:1992vn,Gillespie:2000fk}.
+However, the SDEs used to model noisy decision processes are usually applied without any mechanistic justification.
+By following the mathematical derivation of the SDEs in chemical systems given by Gillespie \cite{Gillespie:2000fk},  we can derive the SDEs used in the leaky integrator model.
 
-There are two possible types of environmental evidence: in the neural system, a dot can move either left or right and in the social system, a fight can be won by one animal or another.
-For clarity in the following we will refer to these left or right dots, but the biological interpretation of the two types of inputs depends on the system.
-There are two decision variables, $X_1$ and $X_2$.
+In a decision between two alternatives, there are two decision variables, $X_1$ and $X_2$, each indicating the evidence accumulated for one of the options.
 In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.
 (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$.
-
-Each decision variable responds positively to one type of input and negatively to the other.
+If there is input, each decision variable will incorporate the new evidence by responding positively to one type of input and negatively to the other.
+Specifically, $X_1$ increases by an amount $b$ when it sees evidence for option $1$ and decrements by $-b$ when it sees evidence for option $2$, and conversely for $X_2$.
 In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.
 In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite.
-Neither decision variable has perfect memory.
-Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.
+For clarity, in the following we will refer to the inputs as left or right dots, but they could also be referred to as wins or losses.
 To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
@@
 Our first assumption is that the rates at which the two types of inputs occur is constant over time.
-We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$.
-We rewrite our basic equations to give
+We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$, giving
 \begin{align*}
@@
 In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.
-A $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.
+A value of $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.
 
-If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with Normal random variables with mean and variance equal to the mean of the Poisson random variables.
+If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with normal random variables with mean and variance equal to the mean of the Poisson random variables.
 Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation.
 Let $Z_\text{L}$ and $Z_\text{R}$, be independent standard Normal random variables, i.e.
-with mean $0$ and standard deviation $1$.
-Then our equations for $X_i(t+\tau)$ become
+with mean $0$ and standard deviation $1$, giving
 \begin{align*}
@@
 where $dW_{\text{L}}$ and $dW_{\text{R}}$ are independent Brownian motions representing, respectively, the left and right inputs.
-The assumptions about the timescales on which inputs occur are particularly reasonable in the social system, although the successful application of this type of model to neural populations implies they are not unreasonable in that system.
+The assumptions about the timescales on which inputs occur are reasonable in the social system, and the successful application of this type of model to neural populations implies they are not unreasonable in that system.
 In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the neural and social systems.
 
-\subsection{Reaching a decision }
-Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.
-The sign and size of the two variables indicate how much evidence there is for each possible decision outcome.
-When a lot of evidence for the first possibility has been observed, $X_1$ will be large and positive, and similarly for the second possibility.
-If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility.
-If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.
-Specifically, there are two thresholds, $T_2$ and $T_1$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.
+\subsection*{Reaching a decision }\ There are two ways to use the decision variables $X_1$ and $X_2$ to make a decision.
+If the difference in the variables can be observed, it indicates the relative strengths of the evidence for each option.
+If $Y=X_1-X_2$, the system should decides on $X_1$ when $Y$ is large and positive and on $X_2$ when $Y$ is large and negative.
+Specifically, there are two thresholds, $T_1$ and $T_2$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.
 However, it may be the case that the difference in the variables cannot be observed.
 In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.
-Again, there are two thresholds, $T_2$ and $T_1$, and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.
+Again, there are two thresholds, $T_1$ and $T_2$, and if $X_2<-T_2$ the decision is for $X_1$ and if the $X_1<-T_1$ the decision is for $X_2$.
 
-In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels to make the system easier to analyze \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.
-However, in the social system we are considering, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.
-A ``decision," i.e.
-the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below a certain threshold.
-The social system is therefore inherently two-dimensional, whereas the neural system may be adequately described by one dimension.
+In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.
+The reduced model is easier to analyze, but there is no empirical evidence for a mechanism in the brain that measures the difference in activity rates and the full two-dimensional model should apply equally well to neural decision making, a possibility we consider here.
+In the social system, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.
+A decision, i.e.
+the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below its threshold.
+The social system is therefore inherently two-dimensional.
 In Table \ref{differences}, we compare how the decision is made and how the decision process is optimized in the two systems.
 
-\subsection{Utility of the decision }
-The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.
-A good decision is one that reaches the correct output quickly, i.e.
-the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.
+\subsection*{Utility of the decision }
+\ The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.
+A good decision is one that reaches the correct output quickly, i.e.\ the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.
 However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.
-It may also be the case that each individual prefers one output over the other, regardless of which is ``correct''.
-In the social case, an animal would prefer that its partner agreed to be subordinate, whether or not it is truly stronger.
-In the neural case, different neural populations may receive different rewards if the decision reaches different outputs, regardless of which of the alternatives is ``true."  In these cases, a variable wants to maximize the probability that his preference is reached (probability of preference being chosen, $\text{PP}_i$).
+It may also be the case that each individual prefers one output over the other, regardless of which is correct.
+In the social case, each animal would prefer that its partner agree to be subordinate, whether or not it is truly stronger.
+In the neural case, different neural populations may receive different rewards if the decision reaches different outcomes, regardless of which of the alternatives is correct.
+Therefore, an individual may want to maximize the probability that its preferred outcome is reached (probability of preference being chosen, $\text{PP}_i$).
 
@@
 In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.
-Similarly, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.
-In both cases, $w_3$ represents how stubborn the individuals are about their preferences.
-If $w_3\gneq 0$ and the individuals prefer different outcomes, neither can simultaneously be satisfied since $\text{PP}_1=1-\text{PP}_2$.
+In the neural system, $w_2$ depends on whatever costs the brain or the whole animal incurs by waiting for a decision.
+Given that $\text{PP}_1=1-\text{PP}_2$, it is impossible for them both to maximize $\text{PP}_i$, so the personal preference weight $w_3$ indicates the strength of the conflict of interests between individuals and how stubborn the individuals are about their preferences.
+In the social system, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.
 
-As show in the Appendix, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model.
+As show in the Supplementary Information, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model.
 There are analytical solutions to these equations when the system of SDEs is reduced to one dimension.
@@
 
-\subsection{Optimal thresholds }
-An individual-- either a neural population or a monkey-- can be engaged in either a single decision or in many decisions at once.
-In the social system, for example, an animal wants to decide whether it is dominant or subordinate to every other member of its social group.
-For any pair of individuals, each has a utility, $U_1$ and $U_2$, and consequently each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.
-Similarly, in a group of individuals, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$.
-His total utility is given by the average of these, $\langle U_{ij}\rangle _j$.
-These lead to an optimized set of thresholds such that no individual has an incentive to choose another threshold.
+\subsection*{Nash equilibrium thresholds }
+\ In a group of individuals, each with a value $a_i$, the difficulty with which the pair $i,j$ makes a decision increases with the difference in value, i.e.
+$c_{ij}$ is an increasing function of $|a_i-a_j|$.
+In the social system, the value $a_i$ reflects an animal's fighting ability.
+In the neural system, the value $a_i$ is high if the property (e.g.\ left movement or red color) a neural population responds to is abundant in the visual stimulus.
+We assume that each individual has the same decision threshold for all the decisions processes with each of its peers and, given those thresholds, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$ and a total utility given by the average of these, $\langle U_{ij}\rangle _j$.
+For each set of values $\{a_1,\dots,a_N\}$, we find the Nash equilibrium thresholds such that no individual has an incentive to choose another threshold to improve its utility.
+Since the Nash thresholds depend on the values $\{a_1,\dots,a_N\}$,  we repeatedly draw a set of values from a uniform distribution and find the Nash thresholds for each set.
+Then we find the average Nash threshold as a function of an animal's rank in the set of fighting abilities.
 
-Since an individual is trying to use the same algorithm to learn about every one else, he is constrained to have only one decision threshold for each of these decision processes.
-In order to consistently assign the difficulties with which all pairs in a group make decisions, we assign each individual a value $a_i$ and the difficulty with which the pair $i,j$ makes a decision increases with the difference in abilities, i.e.
-$c_{ij}$ is an increasing function of $|a_i-a_j|$.
-In the social system, these values can be interpreted as the fighting abilities of the animals.The optimal decision threshold in this group context depends on the distribution of difficulties ($c_{ij}$) of the decision an individual will have to make and the thresholds of the other individuals.
+\subsection*{Collective computation }
+\ For each set of values $\{a_1,\dots,a_N\}$ and Nash thresholds $\{T_1,\dots,T_N\}$, we find the probability that each pair will reach either outcome and the expected time it will take.
+We use these probabilities to generate a set of decisions between all pairs.
+We use the convention that a decision is sent from $i$ to $j$ if $j$'s preference prevailed.
+To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be made, at each point in time $t$, we define the weight of the edge from $i$ to $j$ to be
+$$
+\begin{array}{lllll}
+0 & \text{ if the decision was made in favor of } i\text{ or } DT_{ij}>t 
+\\t-DT_{ij} & \text{ if the decision was made in favor of } j\text{ and } DT_{ij}>t 
+\end{array}
+$$
+Once $t$ is greater than the maximum decision time between all pairs of individuals, all edges of the network will have formed.
 
- We use the following procedure to find optimal thresholds as a function of position in the group.
-We repeatedly draw a set of fighting abilities from a uniform distribution; we find the optimal thresholds for each set, and then find the average optimal threshold as a function of an animal's order in the set of fighting abilities.
-To find the optimal thresholds for a given set of fighting abilities, we turn the difference in fighting abilities between each pair into the difficulty of the decision they have to make ($c$).
-For an initial set of thresholds, we find $U_{ij}$ for each animal $i$ in each of its decisions and then average $\langle U_{ij}\rangle_j$ across all the animals with which it is engaged in a decision process.
-We then allow each animal to choose a threshold that optimizes $\langle U_{ij}\rangle_j$ with respect to the rest of the group's thresholds and we iterate these choices until no animal chooses a new threshold.
+Given a decision network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}.
+These formalisms assign a consensus score to each node in the network.
+In the following we consider four  consensus formalisms.
+The first and simplest is simply the unweighted in-degree of a node, i.e.
+the number of individuals deciding in favor of each individual.
+The second is the weighted in-degree of a node, which is a finer measure than the unweighted in-degree since it takes into consideration the strength of each decision made in favor of the node.
+The third is the entropy of the distribution of the number of decisions made in favor of each individual, which gives a coarse measurement of how uniformly all other individuals in the group behave with respect to a focal individual.
+The fourth is the eigenvector centrality of the decision network, which measures how central each node is in the global structure of the network.
+For more details see \cite{Brush:2013fk}.
+Using many random sets of fighting abilities, we find the mutual information between the consensus scores from the network at each time $t$ and the true values, for each consensus formalism.
+We also find the average skewness of the set of consensus scores.
 
-\subsection{Collective computation }
-We are interested in how the consensus about the value of each individual emerges out of the decision process.
-(In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq}.)  Given an interaction network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}.
-These formalisms assign a consensus score to each node in the network (interpreted as a power score in the social system).
-In the following we consider four  consensus formalisms.
-The first and simplest is simply the number of nodes sending decisions to each animal (number of signalers).
-The second is the number of decisions received by each node, which is a finer measure than the first.
-The third is the entropy of the distribution of the number of decisions sent to each individual, which measures gives a coarse measurement of how uniformly all other animals in the group behave with respect to a focal individual.
-The fourth is the eigenvector centrality of the signaling network, which measures how central each node is in the global structure of the network.
-For more details see \cite{Brush:2013fk}.
-
-To use these measures, we need to construct an interaction network between individuals in the group.
-For each random set of fighting abilities, we let the animals optimize their thresholds, which give the probability that each pair will reach either decision and the expected time it will take.
-We use these probabilities to generate a set of decisions between all pairs.
-We use the convention that a decision is sent from $i$ to $j$ if in their decision $j$'s preference prevailed.
-(In the social system, $i$ sends a subordination signal to $j$ to communicate his agreement to be subordinate.) To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be sent, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.
-At each point in time $t$, the weight of the edge from $i$ to $j$ is $0$ if the expected decision time is greater than $t$ (i.e.
-a decision has not been made) and is given by the number of signals sent from $i$ to $j$ otherwise.
-Using many random sets of fighting abilities, for each consensus formalism and each point in time $t$, we can find the mutual information between the consensus scores from the network at $t$ and the true values.
-We can also find the average skewness of the set of consensus scores.
-
-\section{Results}
-\subsection{Decision making is as accurate without a third party arbiter. }
-First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
+\section*{Results}
+\subsection*{Decision making is as accurate without a third party arbiter. }
+\ First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.
 To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e.
 $T_1=T_2$.
-We find that the two-dimensional decision is as accurate as the one-dimensional decision.
-That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability.
+We find that the two-dimensional process is as accurate as the one-dimensional process.
+That is, for a given expected time to reach a decision, the two-dimensional process will reach the correct decision with the same probability (Supplemental Figure \ref{dimensionality}).
 Because they perform equally well and because the two-dimensional system is more appropriate for the social system (and perhaps the neural system as well), we use the two-dimensional system in the rest of our analyses.
@@
 
-\subsection{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
-The Nash equilibrium thresholds for a pair of individuals depend on whether the individuals are minimizing error rate, maximizing probability of their preference, or minimizing decision time.
-If accuracy of the decision is the only thing that matters ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible since if the decision is correct the weaker animal gives up first.
-When personal preference matters ($w_3>0)$ and there is a conflict of interest between the individuals, both will set their threshold high since there is no incentive for the individuals to stop accumulating evidence.
+\subsection*{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
+\ We start with a pair of individuals, i.e.
+a group of size two, to build intuition for how the optimization weights affect the Nash equilibrium thresholds.
+If accuracy of the decision is the only thing that matters ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible, since the decision is correct when the weaker animal gives up first.
+When only personal preference matters and there is a strong conflict of interest between the individuals($w_3=1)$, both will set their threshold high since there is no incentive for the individuals to stop accumulating evidence.
 As the importance of decision time increases (i.e.
 waiting costs increase), both individuals will lower their thresholds in order to reach the decision more quickly, whatever that decision may be (Figure \ref{nasheq}).
-Notice that decision time and personal preference lead both animals to have the same thresholds whereas the individuals have different thresholds when accuracy is important.
 The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and decreases as either decision time or personal preference become important (Figure \ref{nasheq}).
 
-In a group, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.
-As in a pair, prioritizing accuracy tends to lead animal's at the top of the group to set high thresholds and those at the bottom to set low thresholds.
-Prioritizing personal preferences, on the other hand, tends to lead all individuals to set high thresholds (Figure \ref{groupeq}).
+In a group with more than two individuals, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.
+As in a pair, prioritizing accuracy tends to lead individuals with high value to set high thresholds and those with low values to set low thresholds and prioritizing personal preferences tends to lead all individuals to set high thresholds (Figure \ref{groupeq}).
 The difference in an individual's threshold between when accuracy is important and when personal preferences are important is largest for the individuals at the bottom of the group.
-Individuals towards the top will set high thresholds in either case.
+Individuals at the top of the group will set high thresholds in either case.
 Individuals toward the bottom, however, will set lower thresholds when they prioritize accuracy then when they prioritize personal preferences.
-This is because choosing a higher threshold would increase waiting time in decisions with respect to all individuals and decrease the accuracy of its decisions with respect to higher animals, and it would only slightly increase the accuracy of its decisions with respect to other low individuals.
+This is because choosing a higher threshold increases waiting time in decisions with respect to all individuals, decreases the accuracy of its decisions with respect to higher animals, and only slightly increases the accuracy of its decisions with respect to other low individuals.
 In other words, individuals at the bottom of the group sacrifice some accuracy in their decisions with respect to each other in order to decrease their overall decision times.
+This leads to the counterintuitive result that making personal preferences more important (i.e.
+having a larger conflict of interest) and making accuracy less important actually increases the average accuracy of the decisions made between individuals at the bottom of the group.
 
-This leads to the slightly counterintuitive result that making personal preferences less important (i.e.
-having a larger conflict of interest) and making accuracy more important actually decreases the average accuracy of the decisions made between individuals at the bottom of the group.
 This in turn leads to a very slight increase in the average accuracy of all pairs in the group (Figure \ref{groupeq}).
+Therefore, in a group with more than two members, increasing the importance of decision time decreases the average accuracy, but increasing the strength of conflicts of interest leads to a (slight) increase in accuracy.
+While the improvement is small, it is very different from the decrease in accuracy caused by increasing the strength of conflicts of interest in a pair.
 For every measure of consensus, the mutual information between the consensus scores and the true values increases with average pairwise accuracy.
-Thus, the information content of the consensus scores is also slightly higher when the individuals prioritize personal preference rather than error rate.
+Thus, the information content of the consensus scores is also maintained when the individuals prioritize personal preference rather than error rate.
+We show in Supplemental Figure \ref{groupsize} that the size of the group does not affect the accuracy with which the group, the top quartile of the group, or the bottom quartile of the group makes decisions.
 
-\subsection{Global network structure provides the most information when accuracy is high but imperfect and when the network is developing. }
-First, we consider the mutual information between the consensus formalisms applied to a fully formed decision network and the true values.
-It is interesting to note that for each of the four types of formalisms we consider, there are regions of the simplex given by the optimization weights that make that the most informative formalism.
-When pairwise decisions are on average quite accurate, the number of decisions received is the most informative about the underlying values (Figure \ref{bestmetric}).
-This makes sense since when each edge in the decision network accurately reflects the difference in value between individuals, the number of edges received is informative about an individual's position in the group.
-When the pairwise decisions are less accurate, global structure of the decision network provides information about values that the individual relationships do not, so eigenvector centrality is the measure that is the most informative about values.
+\subsection*{Eigenvector centrality provides the most information when accuracy is high but imperfect and when the network is developing. }
+\ First, we consider the mutual information between the consensus formalisms applied to a fully formed decision network and the true values.
+For each of the four measures we consider, there are combinations of  the optimization weights that make it the most informative formalism (Figure \ref{bestmetric}).
+When waiting costs are low, the finer measures of consensus--- weighted in-degree and eigenvector centrality---outperform the coarser measures---in-degree and entropy---because in these circumstances, the edges of the decision network tend to be accurate and the finer measures make use of more information from the network and make finer distinctions.
+When there are no waiting costs and personal preferences are important, all individuals have high thresholds,  weighted in-degree outperforms eigenvector centrality (Figure \ref{bestmetric}).
+%%WHY?!?!?
+%This makes sense since when each edge in the decision network accurately reflects the difference in value between individuals, weighted in-degree received is informative about an individual's position in the group.
+In particular, weighted in-degree out performs eigenvector centrality by correctly resolving small differences between the highest value individuals.
+When the pairwise decisions are less accurate, global structure of the decision network provides information about values that the individual relationships do not, so eigenvector centrality is the the most informative about the individuals' true values (Figure \ref{bestmetric}).
+When accuracy is quite low, weighted in-degree and eigenvector centrality correctly rank the very highest individuals, but cannot distinguish between the lower half of the group.
+On the other hand, in-degree assigns the same scores to the few highest individuals, but correctly resolves the differences between the bottom half, making it slightly more informative over all.
+When accuracy is very low, none of the measures are very informative, but entropy correctly gives the very lowest individuals low scores and the very highest individuals high scores, giving it an edge over the other measures.
 
-The differences in mutual information given by each formalism are very small, but the clear pattern that the best formalism depends on waiting costs needs explanation.
-The number of signalers consistently overestimates individuals' value.
-When accuracy is fairly high, this bias makes it uninformative about value.
-However, when accuracy is quite low, this consistent bias actually make the number of signalers more informative than measures whose errors are less predictable.
-When accuracy is lower still, none of the measures are very informative, but entropy correctly gives the lowest individuals low scores, giving it an edge over the other measures.
+The information content of the different formalisms show different patterns as the network is forming (Figure \ref{bestmetric}).
+The information content of all formalisms show an initial rise as edges start to form in the decision network.
+When there are waiting costs and hence some inaccurate edges in the network, as these edges develop, there is a decrease in the information content of the three local measures---weighted in-degree, in-degree, and entropy---but not in eigenvector centrality.
+As the network becomes more fully developed, the information content of the three local measures recovers, in some cases even surpassing the information content of eigenvector centrality.
+Therefore, regardless of which measure is most informative on a fully developed network, as long as there are non-zero waiting costs, eigenvector centrality has the advantage of never losing information content as edges are added to the network and consistently performing well on networks that are not fully formed.
 
-When the number of signals or number of signalers out-perform the other formalisms, the differences in performance are very small.
-When eigenvector centrality is the best, on the other hand, it clearly outperforms the others (Figure \ref{bestmetric}).
-Eigenvector centrality is consistently the most informative formalism on a network that is not fully formed (Figure \ref{bestmetric}).
+\subsection*{The skewness of the power distribution is maximized when decisions are moderately accurate. }
+\ For each measure of consensus in the decision network, we find the average skewness of the distribution of consensus scores of a group using Nash thresholds, as they depend on optimization weight.
+We find that, for all measures, skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and preference (Figure \ref{skewness}).
+(Results for number of signalers are shown in Figure \ref{skewness} and others are similar).
+When waiting costs are low and accuracy is high, the decisions accurately represent the individuals' true  abilities, so the distribution of number of signalers accurately represents the distribution of abilities, which is not very right-skewed.
+When waiting costs are high and accuracy is low, decisions are so noisy that all individuals receive a similar number of decisions, giving a distribution that is quite uniform and not right-skewed.
+At intermediate waiting costs, the decisions between individuals with low to middle abilities are very noisy, so that the individuals in the bottom and middle of the group receive a similar number of signals, but all individuals make accurate decisions with the top few individuals, which gives them high scores and results in a right-skewed distribution.
 
-\subsection{The skewness of the power distribution is maximized when decisions are moderately accurate. }
-For each measure of consensus in the decision network, we find the average skewness of the distribution of consensus scores of a group using optimal thresholds, as they depend on optimization weight.
-We find that for all measures skewness is maximized at intermediate waiting costs and does not depend strongly on the tradeoff between error rate and preference (Figure \ref{skewness}).
-(Results for number of signalers are shown in Figure \ref{skewness} and others are similar).
-When waiting costs are low and accuracy is high, the decisions accurately represent the individuals' true  abilities so the distribution of number of signalers accurately represents the distribution of abilities, which is not very right-skewed.
-When waiting costs are high and accuracy is low, all individuals receive decisions from a similar number of others, giving a distribution that is quite uniform and not right-skewed.
-At intermediate waiting costs, the decisions sent to individuals with low to middle abilities are very noisy so that the individuals in the bottom and middle of the group receive signals from similar numbers of individuals.
-However, all individuals make accurately decisions with the top few individuals, giving them high scores, resulting in a right-skewed distribution.
-
-\subsection{Decisions in a group take longest between individuals with similarly high abilities. }
-Difficult decisions tend to (but do not always) take longer than easier decisions.
-In a social system, the difficult of a decision increases as a pair becomes more similar.
-In a group of decision makers, the absolute value of two individuals, as well as the difference in their values, affects how quickly they can make a decision.
-When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with different abilities also take a fairly long time.
+\subsection*{Decisions in a group take longest between individuals with similarly high abilities. }
+\ When waiting costs are low, the pairs that take the longest to reach a decision are those with similar fighting abilities, but pairs with different abilities also take a fairly long time.
 When waiting costs are high, all pairs reach decisions quickly.
@@
 
-\section{Discussion}
+\section*{Discussion}
 
-We derive a pair of stochastic differential equations that describe how two individuals accumulate evidence from a noisy environment.
+From a mechanistic description of a decision-making process, we derive a set of stochastic differential equations that describe how two individuals accumulate evidence from a noisy environment.
 The model can be equally well applied to the firing rates of neural populations observing a visual stimulus and the beliefs of two monkeys about their relative dominance based on their fighting history.
-We first compare the simplified version of the model that is frequently used to describe the neural system and the full version we need to describe the social system.
-We then consider a population of individuals using this model to make decisions to investigate the best way of making a collective computation, how to improve the accuracy of this computation and how the shape of the power structure in the social system depends on the accuracy of decisions.
-Finally, we find a model output that is comparable to empirical observations.
-
 Either the absolute value of the decision variables or the difference in their values can determine the outcome of the decision.
@@
 In order for the difference in variables to be used, a third party arbiter needs to be present to evaluate the value of each decision variable and find their difference.
-In the brain, it might be the case that there is a third neural population whose activity depends on the two populations gathering external evidence, but in a social group, there is not a third party informing how a pair of animals should establish their subordination-dominance relationship.
-The fact, then, that the model does not depend on the existence of this third party arbiter means that it can be equally well applied whether or not it is present.
+In the brain, it might be the case that there is a third neural population whose activity depends on the two populations gathering external evidence.
+In a social group, there is no third party gathering information about the opinions of two monkeys in order to dictate when the decision establishing their subordination-dominance relationship should be made.
+The fact that the existence of this third party arbiter does not affect the model output highlights the similarity between the two systems and it means that the results of a model without the arbiter apply in either case.
 
-The decisions between pairs of individuals give rise to a network of decisions between all members of a group.
-A group of individuals will have some degree of consensus about the status of each individual.
-This consensus is the result of a collective computation that depends on individual behaviors.
-There are many different ways of measuring this consensus and each represents a different way of performing the collective computation \cite{Brush:2013fk,Flack:2006uq}.
-When the accuracy of pairwise decisions is relatively high but imperfect, then a global metric like eigenvector centrality is the most informative about the true states of the nodes.
-When pairwise accuracy is quite low, then a coarse measure like the number of signalers or the entropy of decisions received outperforms a finer measure like the number of decisions received.
+Individuals in biological systems are likely to be selfish and stubborn, both in order to maximize their own welfare and because prioritizing accuracy requires knowing which the right answer is beforehand.
+One might expect that individuals valuing their preference at the expense of the accuracy would tend to make decisions that were less accurate.
+In a pair, this holds true.
+However, when there is a group of individuals all making this tradeoff, valuing preferences over accuracy improves the accuracy with which individuals in the bottom of the group resolve their differences.
+This slightly increases the average accuracy of the whole group, which translates into a slightly more informative collective computation.
+While the increase in accuracy is not very big, it is encouraging that, as long as there are non-zero waiting costs, conflicts of interest do not undermine the group's ability to perform a collective computation.
+In every biological system, there are constraints on the amount of time an individual can take to make a decision so the condition of non-zero waiting costs is likely to be satisfied.
 
+Usually, only two neural populations are included in models of neural decision making (although see XXX).
+Further, it is usually assumed that neural populations in the brain value accuracy, since the whole organism receives a benefit from making a correct decision in a short time.
+But it is possible that different neural populations receive different rewards from making different decisions, regardless of which decision is the correct one.
+Only by considering a group with more than two members do we find the surprising result that the accuracy of a collective computation can be improved by introducing conflicting interests.
+This might suggest that in situations where a group of neural populations are recruited to make a decision, we could expect to find that they have different preferences, rather than all valuing accuracy.
+Our result also has  implications for designing networks with which to make computations.
+
+There are many different ways of measuring the consensus about individuals' values from a decision network \cite{Brush:2013fk,Flack:2006uq}.
+Each network metric represents a different way of performing a collective computation.
+We find that when the accuracy of pairwise decisions high, finer measures like weighted in-degree and eigenvector centrality are the most informative about the true states of the nodes, but  when pairwise accuracy is quite low, coarser measures like in-degree and entropy perform best.
+We also find that eigenvector centrality performs well on network that are not fully formed.
 As network scientists, if we are presented with an interaction network, we would like to know how to measure the consensus present in the network about each node.
-If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes, this result gives a rule of thumb for how to pick a measure of consensus.
-However, since eigenvector centrality always performs optimally or close to optimally on a fully formed network and it performs optimally on a network that is still forming, eigenvector centrality would would generally be a good choice.
+If we have a sense of the accuracy with which the edges reflect true differences in the states of the nodes or how complete the network is, these results give a rule of thumb for how to pick a measure of consensus.
+Since eigenvector centrality always performs optimally or close to optimally on a fully formed network and it performs optimally on a network that is still forming, eigenvector centrality seems to be generally a good algorithm to use.
 
 As members of a social group about which they would like to have information, the monkeys may use these measures (or approximations of them) to estimate their own and their peers' abilities.
-In previous work, we found that both eigenvector centrality and measures of the number of decisions received produced scores that were informative about animals' functional behavior in a group of pigtailed macaques.
-This may be informative about the optimization weights the animals in that group are using.
+In previous work, we found that both eigenvector centrality and weighted in-degree (and variants thereof)  produced scores that were informative about animals' functional behavior in a group of pigtailed macaques.
+This may be tell us something about the optimization weights the animals in that group are using.
 If so, it would suggest that decision time is less important than a combination of error rate and the probability of being dominant.
-If animals in different species do prioritize fighting costs and the benefits from being dominant (i.e.
-decision time and probability of preference)  differently, we might expect to find that those animals use different properties of their social networks to estimate the power structure of the group.
-
-Individuals in biological systems are likely to be selfish and stubborn, both in order to maximize their own welfare and because prioritizing accuracy requires knowing which the ``right'' answer is.
-The accuracy of the collective computation depends on how the individuals make the tradeoff between personal preference and accuracy, regardless of which measure is used to make the computation.
-We expected to find that individuals valuing their preference at the expense of the correct output would tend to make decisions that were less accurate.
-In a pair, this holds true.
-However, when there is a group of individuals all making this tradeoff, valuing preferences over accuracy improves the accuracy with which individuals in the bottom of the group resolve their differences, increasing the average accuracy of the whole group.
-This overall accuracy translates into a more accurate collective computation.
-
-Usually only two neural populations are included in models of neural decision making (although see XXX).
-Further, it is usually assumed that neural populations in the brain value accuracy, since the whole organism receives a benefit from making a correct decision in a short time.
-But it is possible that different neural populations receive different rewards from making different decisions, regardless of which decision is the ``correct'' one.
-Only by considering a group of decision makers do we find the surprising result that the accuracy of a collective computation can be improved by introducing conflicting interests.
-This might suggest that in situations where a group of neural populations are recruited to make a decision, we could expect to find that they have different preferences, rather than all valuing accuracy.
-Our result also has interesting implications for designing networks whose properties we would like to use to make computations.
+It also suggests that if animals in different species prioritize fighting costs and the benefits from being dominant differently, we might expect to find that those animals use different properties of their social networks to estimate the power structure of the group.
 
@@
 In a group, each pair of individuals experiences the difference in their abilities, but their absolute positions within the group also matter.
-In data from a captive population of macaques, we observe that pairs of animals with high abilities reach agreements about their relative dominance but that pairs of animals with low abilities sometimes do not reach such agreements (JC Flack, unpublished data).
+In data from a captive population of macaques, we observe that pairs of animals with high abilities reach agreements about their relative dominance but that pairs of animals with low abilities sometimes do not reach such agreements (Flack, unpublished data).
 In our model, decisions between pairs who have similar and high abilities always  take as long or longer than decisions between any other pairs.
@@
 
-We consider two systems, a neural and a social, and our model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two individuals are equal.
-The similarities between these three frameworks-- neural decision making, social decision making, and game theory-- suggest that there are common principles of decision making through conflict that are applicable to a number of systems.
+We assumed that the individuals in a group use Nash equilibrium thresholds, but in order to do so, they would first have to determine what those are.
+By assuming that they use these optimal strategies, we are able to analyze the consequences of what it means to be ``optimal," but it neglects the interesting question of how those optimal strategies might be reached.
+Another way to make our model more realistic would be to include a way for the individuals to learn what thresholds to use.
+Finding the appropriate timescale separation between the dynamics of how strategies are learned and the decision dynamics and the consequences of this additional level of learning are left for future work.
+
+We apply a noisy learning model to two different biological systems, one neural and one social, and our model reduces to previously studied ``war of attrition'' games in the game theory literature when the abilities of the two individuals are equal.
+The similarities between these three frameworks---neural decision making, social decision making, and game theory---suggest that there are common principles of decision making through conflict that are applicable to a number of systems.
 Conflict and strategy are not usually ascribed to the neural model, and the war of attrition is not usually thought of as a process through which the pair of competitors reaches a decision about an external truth.
@@
 %%other applications?
+
+\section*{Acknowledgements }
+This publication was made possible through the support of a subaward from the Santa Fe Institute under a grant from the John Templeton Foundation for the study of complexity.
+The opinions expressed in this publication are those of the author(s) and do not necessarily reflect the views of the John Templeton Foundation.
+EB acknowledges support from the training grant NIH 5T32HG003284.
+
+\section*{Figures}
 
@@
 The optimization weights for each panel are indicated in the simplex with the corresponding letter.
+Parameters: $w_1=1$, $w_2=0$, $w_3=0$ (b), $w_1=0.8$, $w_2=0.2$, $w_3=0$ (c), $w_1=0$, $w_2=1$, $w_3=0$ (d), $w_1=0$, $w_2=0$, $w_3=1$ (e), $w_1=0$, $w_2=0.2$, $w_3=0.8$ (f), $b=1$, $r=1$, $\ell=0.1$.
 }
@@
 Both bottom-quartile and total accuracy decrease as $w_2$ increases, and the increase in bottom-quartile accuracy as $w_3$ increases drives a slight increase in total accuracy.
-In (c)-(f) we show how the group-context optimal thresholds (left panels) and decision time for each pair (right panels) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest.
+In (c)-(f) we show how the group-context Nash thresholds (left panels) and decision time for each pair (right panels) depends on the true order of value where $1$ is the strongest individual and $20$ is the weakest.
 The optimization weights for each panel are indicated in the simplex with the corresponding letter.
 For low waiting costs (c and e), it takes longest for pairs with similar values to reach a decision.
-For slightly higher waiting costs (d and f), it takes longer for pairs with similar and high values to reach a decision.}
+For slightly higher waiting costs (d and f), it takes longer for pairs with similar and high values to reach a decision.
+Parameters: $w_1=1$, $w_2=0$, $w_3=0$ (c), $w_1=0.9$, $w_2=0.1$, $w_3=0$ (d), $w_1=0$, $w_2=0$, $w_3=1$ (e), $w_1=0$, $w_2=0.1$, $w_3=0.9$ (f), $N=20$, $b=1$, $r=1$, $\ell=0.1$.
+}
 
@@
 
-\section{Figures}
 \begin{figure}[ht]
-%\begin{minipage}{.66\linewidth}
-%\includegraphics[width=\linewidth]{mostinformative.pdf}
-%\end{minipage}
-%\begin{minipage}{.44\linewidth}
-%\includegraphics[width=.66\linewidth]{infogrowth.pdf}
-%\end{minipage}
 \includegraphics[width=\textwidth]{info.pdf}
@@
 In (b)-(d) we show how the mutual information of each consensus formalisms increases over time.
-The optimization weights for each panels are indicated in the simplex with the corresponding letter.}
+The optimization weights for each panels are indicated in the simplex with the corresponding letter.
+Parameters: $w_1=0$, $w_2=0$, $w_3=1$ (c), $w_1=0$, $w_2=0.2$, $w_3=0.8$ (d), $w_1=0$, $w_2=0.4$, $w_3=0.6$ (e),  $N=20$, $b=1$, $r=1$, $\ell=0.1$.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{skewness_histograms.pdf}
-\caption{\label{skewness} The average skewness of the distribution of the number of signalers is maximized at intermediate waiting costs.
-In the upper panel, the color indicates the average skewness of the distribution of a group using optimal thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
+\caption{\label{skewness} The average skewness of the distribution of in-degree is maximized at intermediate waiting costs.
+In the upper panel, the color indicates the average skewness of the distribution of a group using Nash thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$.
 In the lower left corner of the simplex, only error rate matters ($w_1=1$).
@@
 In the lower right corner, only preference matters ($w_3=1$).
-In (b)-(d) we show a histogram for a representative distribution of the number of signalers for a group using the optimization weights as indicated in the simplex with the corresponding letter.}
+In (b)-(d) we show a histogram for a representative in-degree distribution for a group using the optimization weights as indicated in the simplex with the corresponding letter.
+Parameters: $w_1=0$, $w_2=0$, $w_3=1$ (c), $w_1=0$, $w_2=0.4$, $w_3=0.6$ (d), $w_1=0$, $w_2=0.6$, $w_3=0.4$ (e),  $N=20$, $b=1$, $r=1$, $\ell=0.1$.}
 \end{figure}
@@
 \centering
-\caption{\label{variables}{\bf  Collective computations.} }
+\caption{\label{examples}{\bf  Examples of collective computation.} A collective computation can occur when a group of individuals, each of which gathers information and makes a decision in a noisy environment, produces an output that depends in a complex non-linear way on the individuals' decisions.
+This phenomenon is found in many biological systems.
+}
 \ra{1.3}
 \begin{tabular}{@{}lllll@{}}
-System & Environmental information & Individual &   Individual decision & Collective computation \\
+System & Information & Individual &   Individual decision & Collective computation \\
 \cmidrule{1-5} 
-brain & dots moving left or right & neural population & to stop firing & brain-level decision for left or right
+brain & moving dots& neural population & to stop firing & decision of the brain
 \\  social group & fights won or lost & monkey & to emit a subordination signal & power structure
-\\  quorum sensing & signaling molecule & bacteria & to produce signaling molecule & toxicity or phosphorescence
-\\ flock movement & environment & bird & where to fly and how quickly & flock movement and navigation
+\\  quorum sensing & signaling molecule & bacteria & to produce signaling molecule & production of toxin
+\\ flock movement & environment & bird & where to fly and how quickly & flock movement
 \end{tabular}
@@
 \centering
-\caption{\label{variables}{\bf  Variables and interpretations in neural and social systems.} }
+\caption{\label{variables}{\bf  Variables in the model and their interpretations in neural and social systems.} }
 \ra{1.3}
 \begin{tabular}{@{}lllll@{}}
-Variable & Interpretation & Neural &   Social \\
+Variable & Definition & Neural &   Social \\
 \cmidrule{1-4} 
-& input & dots moving left or right & fights won or lost
+$a$  & value &  degree to which the property  & fighting ability
+\\ & & a neural population responds to 
+\\ & & is present in the environment&
+\\ $b$ & change due to new evidence
 \\$c$ & strength of input & coherence of dots & prob. of stronger animal winning
-\\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
+\\ $\ell$ & leak rate
+\\ $r$ & interaction rate 
 \\ $w_1$ & error rate weight & reward from being ``right" & reward from being ``right"
@@
 \\ $w_3$ & prob. of preference weight & ? & benefit from receiving signal
-\\ & decision & subject indicates left or right & one animal emits subordination signal
-\\ & correct decision & subject chooses correct direction & weaker animal signals
+\\ $T_1,T_2$ & decision thresholds
+\\$X_1,X_2$ & decision variables &  firing rates of neural populations & opinions about relative dominance
 \end{tabular}
@@
 
+%\begin{table}[ht]
+%\centering
+%\caption{\label{differences}{\bf  Comparison of model in its original application to neural systems and its application to social systems.} Differences are highlighted in red.}
+%\ra{1.3}
+%\begin{tabular}{@{}lllll@{}}
+%&Original neural & & Social \\
+%\cmidrule{2-3} \cmidrule{4-4} 
+%dimensionality & $1$  && $2$
+%\\ group size & $2$ && \fcolorbox{red}{white}{$N$}
+%\\decision & difference hits a threshold  && \fcolorbox{red}{white}{one var. hits a threshold}
+%\\ optimality criterion &  reward from being ``right" && reward from being ``right"
+%\\ & decision time && decision time
+%\\ & && \fcolorbox{red}{white}{reward from  receiving signal}
+%\\optimization depends on & input strength && input strength
+%\\ & noise && noise
+%\\ & leak rate && leak rate
+%\\ & && \fcolorbox{red}{white}{other animal's threshold}
+%\end{tabular}
+%\end{table}
+
 \begin{table}[ht]
 \centering
-\caption{\label{differences}{\bf  Comparison of model in its original application to neural systems, in an extended application to neural systems, and its application to social systems.} Differences are highlighted in red.}
+\caption{\label{differences}{\bf  Comparison of model in different application.} In its original application to neural systems, the leaky integrator model was reduced to one dimension and the only optimality criteria considered were accuracy and decision time.
+We study a two-dimensional version of the model and consider the possibility of personal preferences as an additional optimality criterion.
+This model can be applied to a collection of neural populations in the brain and to a social group of monkeys.
+Differences between the original neural application and the extended neural and social applications are highlighted in red.}
 \ra{1.3}
 \begin{tabular}{@{}lllll@{}}
-&Original neural & & Social \\
-\cmidrule{2-3} \cmidrule{4-4} 
-dimensionality & $1$  && $2$
-\\decision & difference hits a threshold  && \fcolorbox{red}{white}{one var. hits a threshold}
-\\ optimality criterion &  reward from being ``right" && reward from being ``right"
-\\ & decision time && decision time
-\\ & && \fcolorbox{red}{white}{reward from  receiving signal}
-\\optimization depends on & input strength && input strength
-\\ & noise && noise
-\\ & leak rate && leak rate
-\\ & && \fcolorbox{red}{white}{other animal's threshold}
+&Original neural & Extended neural & Social \\
+\cmidrule{1-3} \cmidrule{4-4} 
+decision & difference hits a threshold  & ? & \fcolorbox{red}{white}{one var. hits a threshold}
+\\dimensionality & $1$  & $?$ & $2$  
+\\ group size & $2$ &\fcolorbox{red}{white}{$N$}& \fcolorbox{red}{white}{$N$}
+\\ optimality criterion &  reward from being ``right" & reward from being ``right"& reward from being ``right"
+\\ & decision time &decision time & decision time
+\\ & &\fcolorbox{red}{white}{reward from  receiving signal} & \fcolorbox{red}{white}{reward from  receiving signal}
+\\optimization & no &  \fcolorbox{red}{white}{yes} & \fcolorbox{red}{white}{yes} 
+\\ depends on other
+\\  individual's threshold
+\\ & 
 \end{tabular}
@@
 
-\pagebreak
-\section{Appendix}
+\appendix
+\section*{Supplementary Information}
+\renewcommand{\thesubsection}{\Alph{subsection}.}
 
 \subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}
+
+\subsection{Analysis of homogeneous abilities }
+
+\subsection{Supplemental Figures }
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}  
+\begin{figure}[ht]
+\includegraphics[width=3.415in]{dimensionality_comparison.pdf}
+
+\caption{\label{dimensionality} The accuracy of decision-making is the same for the reduced one-dimensional model and the full two-dimensional model.
+In each version of the model, changing the decision thresholds changes the probability of a correct decision and the expected time to a decision.
+Here we show the probability of a correct output as a function of the expected time to a decision, as these thresholds are varied.
+The red lines correspond to the reduced model and the dashed blue lines correspond to the full model.
+Decisions of different difficulty ($c$) are represented with lines of different intensity.
+Parameters: $N=20$, $b=1$, $r=1$, $\ell=0.1$.}
+\end{figure}
+
+\begin{figure}[ht]
+\includegraphics[width=6.83in]{group_size.pdf}
+\caption{\label{groupsize} In each panel, group size is on the horizontal axis and average accuracy is on the vertical axis.
+In (a) we show the average accuracy of the whole group, in (b) we show the average accuracy of the top quartile, and in (c) we show the average accuracy of the bottom quartile.
+Each line corresponds to a different value of $w_2$.
+Parameters: $N=20$, $b=1$, $r=1$, $\ell=0.1$, $w_1=0$, $w_3=1-w_2$.}
+\end{figure}
 

### 2015-05-06 suppinfo_12_01_14.tex
writing ms and supplemental and making pretty figures

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath,amsfonts}
\usepackage{custom2}
\usepackage{graphicx} 
\usepackage{epstopdf} 
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{natbib}
\usepackage{tabularx}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}

\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
\newcommand{\mb}{\mathbf}
\newtheorem{ass}{Assumption}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\title{Conflict and Strategy in Collective Computation}
\author{Eleanor Brush, David Krakauer, Jessica Flack}

\begin{document}
\maketitle
\tableofcontents

%\appendix
\section*{Supplementary Information}
\renewcommand{\thesubsection}{\Alph{subsection}.}
\renewcommand{\thesubsubsection}{\thesubsection \arabic{subsubsection}}
%\renewcommand{\thetheorem}{\thesubsection .\arabic{theorem}}
\renewcommand{\thetheorem}{\arabic{theorem}}

\subsection{Derivation of PDEs for waiting time and accuracy \label{pdes_deriv}}
\subsubsection{SDEs }
%(Section 4.3.6 Stochastic Differential Equations, Multivariable Systems in Gardiner) 
In the following we will assume we have a set of stochastic differential equation written as
\begin{equation}
d{\mathbf x} ={\mb A}({\mb x},t)dt+{\mb B}({\mb x},t)d{\mb W}(t), \label{sde}
\end{equation}
where ${\mb x}\in\R^N$, $A:\R^N\times \R\to\R^N$, ${\mb B}:\R^N\times\R\to\R^N\times\R^N$, and $d{\mb W}(t)$ is an $N$ variable Weiner process. The forward equation describing the probability density $p({\mb x},t|{\mb y},s)$ is 
%% i'm trying to be consistent about my noise terms: I'll try to always use B*B^T even though Gardiner seems to switch back and forth between B and B*^T
\begin{equation}
\frac{\partial p}{\partial t} =-\sum_{i=1}^N\frac{\partial \big(A_i({\mb x},t)p\big)}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N\frac{\partial^2\big([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{ij}p\big)}{\partial x_i\partial x_j} \label{forward}
\end{equation}
and the backward equation is 
\begin{equation}
-\frac{\partial p}{\partial s}=\sum_{i=1}^NA_i({\mb y},s)\frac{\partial p}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},s){\mb B}^T({\mb y},s)]_{ij}\frac{\partial^2 p}{\partial y_i\partial y_j}. \label{backward}
\end{equation}

\subsubsection{Expected first exit time }
%(section 5.5 the Fokker-Planck Equation, First Passage Times for Homogeneous Processes in Gardiner) 
Let $G({\mathbf y},t)$ be the probability that the particle is still in the region $R$ at time $t$ given that it started at ${\mathbf y}$ so that $$G({\mathbf y},t)=\int_Rp({\mathbf x},t|{\mathbf y},0)d{\mathbf x}.$$

\begin{claim} If the system is time homogeneous,
\begin{equation}
\frac{\partial G}{\partial t}=\sum_{i=1}^NA_i({\mb y})\frac{\partial G}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y}){\mb B}^T({\mb y})]_{ij}\frac{\partial^2 G}{\partial y_i\partial y_j}. \label{Geq}
\end{equation}
%%Gardiner has a 2 instead of 1/2 in Section 6.6 but that must be a typo
\end{claim}
\begin{pf}
If the system is time homogeneous, then $p({\mb x},t|{\mb y},0)=p({\mb x},0|{\mb y},-t)$ so that we can rewrite $G$ as 
\begin{align*}
G({\mb y},t)&=\int_Rp({\mb x},0|{\mb y},-t)d{\mb x},
\\ \text{ which gives that } \frac{\partial G}{\partial t}\bigg|_{t=t}&=\int_R\frac{\partial p({\mb x},0|{\mb y},-t)}{\partial t}\bigg|_{t=t}d{\mb x}
\\&=\int_R -\frac{\partial p({\mb x},0|{\mb y},s)}{\partial s}\bigg|_{s=t} d{\mb x}
\\&=\int_R\bigg(\sum_{i=1}^NA_i({\mb y})\frac{\partial p}{\partial y_{i}}\bigg|_{s=t}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y}){\mb B}^T({\mb y})]_{ij}\frac{\partial^2 p}{\partial y_{i}\partial y_{j}}\bigg|_{s=t}\bigg) d{\mb x} \text{ using (\ref{backward})}
\\&=\sum_{i=1}^NA_i({\mb y})\bigg(\int_R\frac{\partial p}{\partial y_{i}}\bigg|_{s=t}d{\mb x}\bigg)+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y}){\mb B}^T({\mb y})]_{ij}\bigg(\int_R\frac{\partial^2 p}{\partial y_{i}\partial y_{j}}\bigg|_{s=t}\bigg) d{\mb x}
\\&=\sum_{i=1}^NA_i({\mb y})\frac{\partial G}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y}){\mb B}^T({\mb y})]_{ij}\frac{\partial^2 G}{\partial y_i\partial y_j}
\end{align*}
\end{pf}

We can now use Equation (\ref{Geq}) to find an equation that the expected hitting time satisfies. If $T$ is the time at which the particle hits the boundary $S=\partial R$, then $P(T\geq t|{\mb y})=G({\mathbf y},t)$.  Therefore
\begin{align*}
E_{{\mathbf y}}(T) &=\int_0^\infty t P(T\in[t,t+dt)|{\mathbf y})
\\&= \int_0^\infty t \bigg(P(T\geq t|{\mathbf y})-P(T\geq t+dt|{\mathbf y})\bigg) 
\\&=\int_0^\infty t \bigg(G({\mathbf y},t)-G({\mathbf y},t+dt)\bigg)
\\&=\int_0^\infty t \frac{G({\mathbf y},t)-G({\mathbf y},t+dt)}{dt}dt
\\&=-\int_0^\infty t\frac{\partial G}{\partial t}dt
\\&=-\bigg(\big[tG({\mathbf y},t)\big]_0^{\infty}-\int_0^\infty G({\mathbf y},t)dt\bigg)  \text{ by integration by parts }
\\&=\int_0^\infty G({\mathbf y},t)dt \text{ if $\lim_{t\to \infty}G({\mathbf y},t)=0$ quickly enough}.
\end{align*}

If we let $f({\mathbf y})=E_{\mathbf y}(T)$, then 
\begin{align*}
\sum_{i=1}^NA_i({\mb y},t)\frac{\partial f}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},t){\mb B}^T({\mb y},t)]_{ij}\frac{\partial^2 f}{\partial y_i\partial y_j}&=\int_0^\infty\bigg(\sum_{i=1}^NA_i({\mb y},t)\frac{\partial G}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},t){\mb B}^T({\mb y},t)]_{ij}\frac{\partial^2 G}{\partial y_i\partial y_j}\bigg)dt
\\&=\int_0^\infty\frac{\partial G}{\partial t}dt \text{ using (\ref{Geq})}
\\&=G({\mb y},\infty)-G({\mb y},0)
\\&=-1
\end{align*}
Finally, this gives us a PDE for the expected hitting time of the boundary of the region $R$:
\begin{equation}
\sum_{i=1}^NA_i({\mb y})\frac{\partial f}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y}){\mb B}^T({\mb y})]_{ij}\frac{\partial^2 f}{\partial y_i\partial y_j}=-1 \label{DT}
\end{equation}
with boundary conditions $f({\mb x})=0 \text{ for } {\mb x}\in\partial R$.

\subsubsection{Distribution of exit points }
%(section 6.6.2 The Fokker-Planck Equation in Several Dimensions, Distribution of Exit Points and section 5.5.4 Probability of Exit Through a Particular End of the Interval in Gardiner )
Let ${\mb J}({\mb x},t|{\mb y},s):(\R^N\times\R)\times(\R^N\times\R)\to \R^N$ be such that 
$$J_k({\mb x},t|{\mb y},s)=A_k({\mb x},t)p({\mb x},t|{\mb y},s)-\frac{1}{2}\sum_{l=1}^N\frac{\partial \big([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{kl}p({\mb x},t|{\mb y},s)\big)}{\partial x_j}.$$ 
If $p$ satisfies the forward equation (\ref{forward}), then 
\begin{equation}
\frac{\partial p}{\partial t} =-\sum_{i=1}^N\frac{\partial \big(A_i({\mb x},t)p\big)}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N\frac{\partial^2\big([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{ij}p\big)}{\partial x_i\partial x_j}=-\sum_{i=1}^N\frac{\partial J_i({\mb x},t|{\mb y},s)}{\partial x_i}. \label{jequals}
\end{equation}

\begin{claim}
Each $J_k$  satisfies the backward equation.
\end{claim}
\begin{pf}
{\tiny
\begin{align*}
\sum_{i=1}^NA_i({\mb y},s)\frac{\partial J_k({\mb x},t|{\mb y},s)}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},s){\mb B}^T({\mb y},s)]_{ij}\frac{\partial^2 J_k({\mb x},t|{\mb y},s)}{\partial y_i\partial y_j}
&=\sum_{i=1}^NA_i({\mb y},s)\bigg[A_k({\mb x},t)\frac{\partial p}{\partial y_i}-\frac{1}{2}\sum_{l=1}^N\frac{\partial \bigg([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{kl}\frac{\partial p({\mb x},t|{\mb y},s)}{\partial y_i}\bigg)}{\partial x_l}\bigg]
\\&+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},s){\mb B}^T({\mb y},s)]_{ij}\bigg[A_k({\mb x},t)\frac{\partial^2 p}{\partial y_i\partial y_j}-\frac{1}{2}\sum_{l=1}^N\frac{\partial \bigg([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{kl}\frac{\partial^2 p({\mb x},t|{\mb y},s)}{\partial y_iy_j}\bigg)}{\partial x_l}\bigg]
\\&=A_k({\mb x},t)\bigg[\sum_{i=1}^NA_i({\mb y},s)\frac{\partial p}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},s){\mb B}^T({\mb y},s)]_{ij}\frac{\partial^2 p}{\partial y_i\partial y_j}\bigg]
\\&-\frac{1}{2}\sum_{l=1}^N\frac{\partial }{\partial x_l}\bigg([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{kl}\bigg[\sum_{i=1}^NA_i({\mb y},s)\frac{\partial p}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},s){\mb B}^T({\mb y},s)]_{ij}\frac{\partial^2 p}{\partial y_i\partial y_j}\bigg)\bigg]
\\&=A_k({\mb x},t)\bigg(-\frac{\partial p}{\partial s}\bigg)-\frac{1}{2}\sum_{l=1}^N\frac{\partial }{\partial x_l}\bigg(-[{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{kl}\frac{\partial p}{\partial s}\bigg) \text{ using (\ref{backward})}
\\&=-\frac{\partial }{\partial s}\bigg(A_k({\mb x},t)p({\mb x},t|{\mb y},s)-\frac{1}{2}\sum_{l=1}^N\frac{\partial \big([{\mb B}({\mb x},t){\mb B}^T({\mb x},t)]_{kl}p({\mb x},t|{\mb y},s)\big)}{\partial x_l}\bigg)
\\&=-\frac{\partial J_k}{\partial s}.
\end{align*}
}
\end{pf}

It follows that any linear combination $q({\mb x},t|{\mb y},s)=\sum_{k=1}^Nc_kJ_k({\mb x},t|{\mb y},s)$ also satisfies the backward equation:
\begin{equation}
-\frac{\partial q}{\partial s}=\sum_{i=1}^NA_i({\mb y},s)\frac{\partial q}{\partial y_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb y},s){\mb B}^T({\mb y},s)]_{ij}\frac{\partial ^2 q}{\partial y_i\partial y_j}. \label{jbackeq}
\end{equation}

\begin{claim} \label{probcurrent}
If $S_{12}$ represents the boundary between two regions $R_2$ and $R_2$, then the net flow of probability from $R_2$ to $R_1$ can be written as $\int_{S_{12}} dS {\mb n}\cdot {\mb J}({\mb x},t) $ where ${\mb n}$ is a normal vector from $R_2$ to $R_1$.  
\end{claim}

\begin{pf}
Let $S_1$ be the part of the boundary of $R_1$ not shared with $R_2$ and $S_2$ be the part of the boundary of $R_2$ not shared with $R_1$ so that  $\partial R_1=S_1\cup S_{12}$, $R_2=S_2\cup S_{12}$, and $S_1\cap S_{12}=S_2\cap S_{12}=\emptyset$. The probability of crossing from $R_2$ to $R_1$ across $S_{12}$ in the interval of time $[t,t+dt)$ is 
\begin{align*}
\int_{R_1}d{\mb x}\int_{R_2}d{\mb y} p({\mb x},t+dt|{\mb y},t).
\end{align*}

The probability of crossing from $R_1$ to $R_2$ across $S_{12}$  in the interval of time $[t,t+dt)$ can be found analogously so the net flow $F$ from $R_2$ to $R_1$ at time $t$ is 
\begin{align*}
F=\lim_{dt \to 0}\frac{1}{dt}\int_{R_1}d{\mb x}\int_{R_2}d{\mb y} \big(p({\mb x},t+dt|{\mb y},t)-p({\mb y},t+dt|{\mb x},t)\big).
\end{align*}
Since $p({\mb x},t|{\mb y},t)=p({\mb y},t|{\mb x},t)=0$ for ${\mb y}\in R_2$ and ${\mb x}\in R_1$,
\begin{align*}
F&=\lim_{dt \to 0}\frac{1}{dt}\int_{R_1}d{\mb x}\int_{R_2}d{\mb y} \big(p({\mb x},t+dt|{\mb y},t)-p({\mb x},t|{\mb y},t)-p({\mb y},t+dt|{\mb x},t)+p({\mb y},t|{\mb x},t)\big)
\\&=\int_{R_1}d{\mb x}\int_{R_2}d{\mb y} \bigg(\frac{\partial p({\mb x},t'|{\mb y},t)}{\partial t'}\bigg|_{t'=t}-\frac{\partial p({\mb y},t'|{\mb x},t)}{\partial t'}\bigg|_{t'=t}\bigg)
\\&=\int_{R_1}d{\mb x}\int_{R_2}d{\mb y}\bigg(-\sum_{i=1}^N\frac{\partial J_i({\mb x},t|{\mb y},t)}{\partial x_i}+\sum_{i=1}^N\frac{\partial J_i({\mb y},t|{\mb x},t)}{\partial y_i}\bigg) \text{using (\ref{jequals})}
\\&=\int_{R_1}d{\mb x}\int_{R_2}d{\mb y}\sum_{i=1}^N\frac{\partial J_i({\mb y},t|{\mb x},t)}{\partial y_i}-\int_{R_2}d{\mb y}\int_{R_1}d{\mb x}\sum_{i=1}^N\frac{\partial J_i({\mb x},t|{\mb y},t)}{\partial x_i}
\\&=\int_{R_1}d{\mb x}\bigg[\int_{S_2}dS {\mb n_2}\cdot {\mb J}({\mb y},t|{\mb x},t)+\int_{S_{12}}dS {\mb n_2}\cdot {\mb J}({\mb y},t|{\mb x},t)\bigg]-\int_{R_2}d{\mb y}\bigg[\int_{S_1}dS{\mb n}_1\cdot{\mb J}({\mb x},t|{\mb y},t)+\int_{S_{12}}dS{\mb n}_1\cdot{\mb J}({\mb x},t|{\mb y},t)\bigg]
\\ &\text{ where ${\mb n}_2$ is a normal pointing out of $R_2$ and similarly for ${\mb n}_1$, by the divergence theorem}
\end{align*}
For ${\mb y}\in S_2$ and ${\mb x}\in R_1$, $J({\mb y},t|{\mb x},t)=0$. Similarly, for ${\mb x}\in S_1$ and ${\mb y}\in R_2$, ${\mb J}({\mb x},t|{\mb y},t)=0$.  Therefore, the integrals over $S_2$ and $S_1$ disappear and $F$ simplifies to
\begin{align*}
F&=\int_{R_1}d{\mb x}\bigg[\int_{S_{12}}dS {\mb n_2}\cdot {\mb J}({\mb y},t|{\mb x},t)\bigg]-\int_{R_2}d{\mb y}\bigg[\int_{S_{12}}dS{\mb n}_1\cdot{\mb J}({\mb x},t|{\mb y},t)\bigg]
\\&=\int_{S_{12}}dS\bigg[\int_{R_1}d{\mb y}\ {\mb n}\cdot \mb{J}({\mb x},t|{\mb y},t)+\int_{R_2}d{\mb y}\ {\mb n}\cdot {\mb J}({\mb x},t|{\mb y},t)\bigg]
\\ & \text{ where ${\mb n}$ is a normal pointing from $R_2$ to $R_1$}
\\&=\int_{S_{12}}dS\bigg[\int_{R_1\cup R_2}d{\mb y}\ {\mb n}\cdot \mb{J}({\mb x},t|{\mb y},t)\bigg]
\\&=\int_{S_{12}}dS\ {\mb n}\cdot {\mb J}({\mb x},t) \text{ since ${\mb x}\in S_{12}\subset R_1\cup R_2$}
\end{align*}
\end{pf}

Now consider a region $R$ with boundary $S$.  Let $g({\mb a},{\mb x},t)|d{\mb S}({\mb a})|$ be the probability of exiting through an element $d{\mb S}({\mb a})$ of the boundary $S$ after time $t$.

\begin{claim}
If the system is time homogeneous, 
\begin{equation}
\frac{\partial g}{\partial t}=\sum_{i=1}^NA_i({\mb x})\frac{\partial g}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial ^2 g}{\partial x_i\partial x_j}. \label{geq}
\end{equation}
\end{claim}

\begin{pf}
By the definition of $g$,
{\tiny
%% Gardiner has a negative sign in the following but I don't know why and it seems to work out just fine without it?
\begin{align*}
g({\mb a},{\mb x},t)|d{\mb S}({\mb a})|&=\int_t^\infty dt' {\mb J}({\mb a},t'|{\mb x},0)\cdot d{\mb S}(\mb{a}) \text{using Claim \ref{probcurrent}}
\\&=\int_t^\infty dt' {\mb J}({\mb a},0|{\mb x},-t')\cdot d{\mb S}(\mb{a})
\\&=\int_{-\infty}^{-t}dt' {\mb J}({\mb a},0|{\mb x},t')\cdot d{\mb S}(\mb{a}) \text{ using integration by substitution}
\\ \text{ so that }  \sum_{i=1}^NA_i({\mb x})\frac{\partial g}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial ^2 g}{\partial x_i\partial x_j}
&=\int_{-\infty}^{-t} dt'\bigg(\sum_{i=1}^NA_i({\mb x})\frac{\partial ({\mb J}({\mb a},0|{\mb x},t')\cdot d{\mb S}({\mb a}))}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial ^2 ({\mb J}({\mb a},0|{\mb x},t')\cdot d{\mb S}({\mb a}))}{\partial x_i\partial x_j}\bigg)
\\&=\int_{-\infty}^{-t}dt' \bigg(-\frac{\partial ({\mb J}({\mb a},0|{\mb x},t')\cdot d{\mb S}({\mb a}))}{\partial t'}\bigg) \text{ using (\ref{jbackeq})}
\\&=\big(-{\mb J}({\mb a},0|{\mb x},t')\cdot d{\mb S}({\mb a})\big)\bigg|_{-\infty}^{-t}
\\&=\big(-{\mb J}({\mb a},-t'|{\mb x},0)\cdot d{\mb S}({\mb a})\big)\bigg|_{-\infty}^{-t}
\\&=-{\mb J}({\mb a},t|{\mb x},0)\cdot d{\mb S}({\mb a})\
\\&=\frac{\partial g}{\partial t}
\end{align*}
}
\end{pf}

We are interested in finding the total probability of exiting through $d{\mb S}({\mb a})$, $h({\mb a},{\mb x})=g({\mb a},{\mb x},0)$.  If we let $t\to 0$ in (\ref{geq}) and notice that ${\mb J}({\mb a},0|{\mb x},0)=\delta(|{\mb a}-{\mb x}|)$, we find that 
\begin{equation*}
\sum_{i=1}^NA_i({\mb x})\frac{\partial h}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial ^2 h}{\partial x_i\partial x_j}=0
\end{equation*}
with boundary conditions $h({\mb a},{\mb x})=\delta(|{\mb a}-{\mb x}|)$ for ${\mb x}$ in $S$. To find the total probability of exiting through a section $S'\subset S$, $H(S',{\mb x})$ we can simply integrate the solutions $h({\mb a},{\mb x})$ for ${\mb a}\in S'$ to find 
\begin{equation}
\sum_{i=1}^NA_i({\mb x})\frac{\partial H}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial ^2 H}{\partial x_i\partial x_j}=0 \label{ER}
\end{equation}
with boundary conditions $H(S',{\mb x})=\delta({\mb x}\in S')$ for ${\mb x}\in S$.

We will introduce the operator ${\mb L}$ such that 
\begin{equation}
{\mb L}(f)=\sum_{i=1}^NA_i({\mb x})\frac{\partial f}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial ^2 f}{\partial x_i\partial x_j}. \label{operator}
\end{equation}
Then the expected time to hit the boundary $S$ satisfies ${\mb L}f=-1$ with boundary conditions $f({\mb x})=0$ for ${\mb x}\in S$ and the probability of hitting a subset $S'\subset S$ satisfies ${\mb L}H=0$ with boundary conditions $H({\mb x})=\delta({\mb x}\in S')$ for ${\mb x}\in S$.

%We can also consider the expected exit time, given that the particle exits through $d{\mb S}({\mb a})$, $T({\mb a},{\mb x})$:
%
%\begin{align*}
%T({\mb a},{\mb x})&=\int_0^\infty t P(\text{ exit in }[t,t+dt)|\text{ exit through }d{\mb S}({\mb a}))
%\\&=\int_0^\infty t\frac{P(\text{ exit through }d{\mb S}({\mb a}) \text{ in }[t,t+dt))}{P(\text{ exit through } d{\mb S}({\mb a}))}
%\\&=\frac{1}{{P(\text{ exit through } d{\mb S}({\mb a}))}}\int_0^\infty t P(\text{ exit through }d{\mb S}({\mb a}) \text{ in }[t,t+dt))
%\\&=\frac{1}{h({\mb a},{\mb x})}\int_0^\infty t\bigg(g({\mb a},{\mb x},t)-g({\mb a},{\mb x},t+dt)\bigg)
%\\&=\frac{1}{h({\mb a},{\mb x})}\int_0^\infty- t\frac{\partial g}{\partial t}dt
%\\&=-\frac{1}{h({\mb a},{\mb x})}\bigg([tg({\mb a},{\mb x},t)]_0^\infty-\int_0^\infty g({\mb a},{\mb x},t)dt\bigg) \text{ by integration by parts}
%\\&=\frac{1}{h({\mb a},{\mb x})}\int_0^\infty g({\mb a},{\mb x},t)dt \text{ if }\lim_{t\to\infty} g({\mb a},{\mb x},t)=0 \text{ quickly enough}
%\\\text{so that }T({\mb a},{\mb x})h({\mb a},{\mb x})&=\int_0^\infty g({\mb a},{\mb x},t)dt
%\end{align*}
%If we integrate (\ref{geq}) from $0$ to $\infty$ we find that 
%\begin{align*}
%-h({\mb a},{\mb x})=-g({\mb a},{\mb x},0)=\sum_{i=1}^NA_i({\mb x})\frac{\partial (Th)}{\partial x_i}+\frac{1}{2}\sum_{i,j=1}^N[{\mb B}({\mb x}){\mb B}^T({\mb x})]_{ij}\frac{\partial^2(Th)}{\partial x_i\partial x_j}.
%\end{align*}

\subsubsection{Our model }
The stochastic differential equations in our model are
\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{L}t-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{L}t+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t.
\end{array}
\end{equation}

This gives 
\begin{equation}
{\mb A}({\mb x},t)=\left(\begin{array}{ll} 
-\ell x_1+br(2c-1)
\\ -\ell x_2-br(2c-1)
\end{array}\right), \text{ and }
\end{equation}

\begin{equation}
{\mb B}({\mb x},t)=\left(\begin{array}{ll} 
b\sqrt{rc} & -b\sqrt{r(1-c)}
\\ -b\sqrt{rc} & b\sqrt{r(1-c)}
\end{array}\right), \text{ so that }
\end{equation}

\begin{equation}
{\mb B}({\mb x},t){\mb B}^T({\mb x},t)
=\left(\begin{array}{ll}
b^2r & b^2r
\\ b^2r & b^2r
\end{array}\right)
\end{equation}

Thus, the operator ${\mb L}$ is given by

\begin{align}
%\begin{array}{ll}
{\mb L}(f)&=(-\ell x_1+br(2c-1))\frac{\partial f}{\partial x_1}+(-\ell x_2-br(2c-1))\frac{\partial f}{\partial x_2}+\frac{1}{2}b^2r\frac{\partial^2 f}{\partial x_1^2}+\frac{1}{2}b^2r\frac{\partial^2 f}{\partial x_1\partial x_2}+\frac{1}{2}b^2r\frac{\partial^2 f}{\partial x_2\partial x_1}+\frac{1}{2}b^2r\frac{\partial^2 f}{\partial x_2^2} \notag
\\&=(-\ell x_1+br(2c-1))\frac{\partial f}{\partial x_1}+(-\ell x_2-br(2c-1))\frac{\partial f}{\partial x_2}+b^2r\bigg(\frac{1}{2}\frac{\partial^2 f}{\partial x_1^2}+\frac{\partial^2 f}{\partial x_1\partial x_2}+\frac{1}{2}\frac{\partial^2 f}{\partial x_2^2} \bigg)
%\end{array}
\end{align}

The region we are interested in is the one in which both individuals' opinions $(x_1,x_2)$ are greater than or equal to their respective thresholds, i.e. $R=\{(x_1,x_2): x_1\geq -T_1 \text{ and } x_2\geq -T_2\}$ so that $S=\partial R=\{(x_1,-T_2): x_1>-T_1\}\cup\{(-T_1,x_2):x_2>-T_2\}$.  The subset of the boundary we are interested in is the one that indicates a correct decision has been made, i.e. $x_2=-T_2$, so that $S'=\{(x_1,-T_2):x_1>-T_1\}$. The expected decision time is the expected hitting time of $S$ and the probability of a correct decision is the probability that the exit is through $S'$.

In order to  numerically solve the PDEs for decision time and error rate, we artificially close the region $R$ so that we can impose boundary conditions on a finite subset of the plane. In order to this, we pick two upper bounds $B_1$ and $B_2$ and restrict $R$ to $$R_2=\{(x_1,x_2): x_1\in [-T_1,B_1] \text{ and } x_2\in[ -T_2,B_2]\},$$ with boundary $S_2$. The subset of the boundary we are interested in becomes 
$$S'_2=\{(x_1,-T_2):x_1\in [-T_1,B_1]\}\cup\{(B_1,x_2):x_2\in [-T_2,B_2]\}$$
The boundary condition $f({\mb x})=0$ for ${\mb x}\in S$ becomes $f({\mb x})=0$ for ${\mb x}\in S_2$ and the condition $H({\mb x})=\delta({\mb x}\in S')$ for ${\mb x}\in S$ becomes $H({\mb x})=\delta({\mb x}\in S'_2)$ for ${\mb x}\in S_2$.

\subsection{Initial conditions }

\subsection{Analysis of homogeneous abilities }

\subsection{Explanation for most informative measures }

Which measure is the most informative about the individuals' true values depends on how accurately pairwise decisions tend to be and what kind of errors are most frequent. When decisions are quite accurate ($w_2<.4$), the finer measures of consensus---weighted in-degree and eigenvector centrality---can use all of the information in the decision network to make fine distinctions between individuals, leading to consensus scores that are more informative than those from the coarser measures---in-degree and entropy.  Whether weighted in-degree or eigenvector centrality is more informative depends on what type of error is common. When there are no waiting costs ($w_2=0$) and the individuals value their personal preferences over accuracy ($w_1<.8$), low valued individuals can sometimes wait out higher valued individuals leading to an incorrect  decision being made in favor of a low individual at the expense of a high individual. Since eigenvector centrality takes the identity of the source of decisions into account, it will incorrectly assign the low individual a high score, whereas weighted in-degree will ignore the source and assign the low individual an appropriately low score.  Therefore, in these circumstances, weighted in-degree is more informative. When accuracy matters ($w_1\geq .8$, $w_2=0$) or when there are low non-zero waiting costs ($0<w_2\leq .3$) low value individual are less likely to wait out high value individuals, but there are a moderate amount of mistakes between pairs of individuals in the middle and bottom of the group. Consider two middle value individuals, with $A$ having slightly higher value than $B$. Another individual might make a mistake with respect to $B$ so that weighted in-degree gives $B$ a slightly higher score than $A$. But $A$ and $B$ will tend to decide correctly so that, despite errors, $B$ will send more decisions than $A$. Since eigenvector centrality is affected by the decisions sent as well as received, it will correctly give $A$ a higher score. Therefore, in these circumstances, eigenvector centrality is the most informative. As waiting costs increase ($.4 \leq w_2 \leq .5$) and decisions become even less accurate, even the high value individuals start to make mistakes. Now consider two high value individuals, with $A$ having slight higher value than $B$. If $A$ errs and decides strongly in favor of a lower value individual, eigenvector centrality will give $A$ a lower score because of those sent decisions and weighted in-degree will give $A$ a lower score because of the decisions it has prevented itself from receiving. However, $A$ is likely to make at most one more mistake than $B$ so that the number of individuals deciding in favor of either will be the same. Thus the coarseness of unweighted in-degree becomes an advantage by allowing it to group high value individuals together when finer measures would make incorrect fine distinctions. When decisions are even less accurate ($.5<w_2<.7$), weighted in-degree, unweighted in-degree, and eigenvector centrality can all get confused by incorrect decisions between any pair of individuals. What becomes especially difficult is distinguishing between low value individuals, one of whom might incorrectly receive decisions from higher individuals. If a low value individual receives correct decisions, they will be of similar strengths from even lower individuals. However, if it receives decisions from a higher individual, even though its score from any of the other measures will be increased, entropy decreases its score because of the difference in the strengths of signals coming from different types of individuals. In these circumstances, entropy does well at measuring the lack of consensus of decisions with respect of low individuals. When decisions are very inaccurate ($w_2>.6$), weighted in-degree becomes the most reliable.

\subsection{Supplemental Figures }

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}[ht]
\includegraphics[width=6.83in]{cartoon_cropped.pdf}
\caption{\label{cartoon} Schematic of the development of the decision network. 1. First, each individual's value is drawn from a uniform distribution. 2. The difficulty of each pairwise decision increases with the similarity of values, so that $c_{ij}$ increases as $|a_i-a_j|$ increases. 3. Given these difficulties, each individual has a Nash threshold strategy $T_i$. 4. Given the set of Nash thresholds, a pair will reach each decision with some probability and in an expected amount of time. 5. We draw the decisions according to the probabilities of the two possible outcomes and the strength of the decision decreases with increasing decision time. 6. For each decision network, we compute a set of consensus measures to give each individual a consensus score. We repeat this process for many random draws of values to find average Nash thresholds and the mutual information between the consensus scores and the values. }
\end{figure}

\begin{figure}[ht]
\includegraphics[width=\textwidth]{dimensionality_comparison.pdf}
\caption{\label{dimensionality} The accuracy of decision-making is the same for the reduced one-dimensional model and the full two-dimensional model. In each version of the model, changing the decision thresholds changes the probability of a correct decision and the expected time to a decision. Here we show the probability of a correct output as a function of the expected time to a decision, as these thresholds are varied. The red lines correspond to the reduced model and the dashed blue lines correspond to the full model. Decisions of different difficulty ($c$) are represented with lines of different intensity. Parameters: $N=20$, $b=1$, $r=1$, $\ell=0.1$.}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=6.83in]{group_size.pdf}
\caption{\label{groupsize} Group size does not affect the accuracy with each pair of individuals make decisions. In each panel, group size is on the horizontal axis and average accuracy is on the vertical axis. In (a) we show the average accuracy of the whole group, in (b) we show the average accuracy of the top quartile, and in (c) we show the average accuracy of the bottom quartile. Each line corresponds to a different value of $w_2$. Parameters: $N=20$, $b=1$, $r=1$, $\ell=0.1$, $w_1=0$, $w_3=1-w_2$.}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=3.4in]{mutinfo_vs_accuracy.pdf}
\caption{ }
\end{figure}

\begin{figure}[ht]
\includegraphics[width=6.83in]{skewness_histograms.pdf}
\caption{ }
\end{figure}

\begin{figure}[ht]
\includegraphics[width=6.83in]{multi_skewness.pdf}
\caption{\label{supp_skewness} The skewness of the consensus scores is maximized at intermediate waiting costs for every consensus measure. In each panel, the color indicates the average skewness of the consensus scores of a group using Nash thresholds, as a function of the optimization weights, $w_1$, $w_2$, $w_3$. In the lower left corner of the simplex, only error rate matters ($w_1=1$).  In the upper corner, only decision time matters ($w_2=1$).  In the lower right corner, only preference matters ($w_3=1$). In (a) consensus is given by weighted in-degree, in (b) by entropy, and in (c) by eigenvector centrality. Parameters: $N=20$, $b=1$, $r=1$, $\ell=0.1$.} 
\end{figure}

%\begin{figure}[ht]
%\includegraphics[width=3.4in]{skewness_max.pdf}
%\caption{ }
%\end{figure}

\pagebreak
\nocite{*}
\bibliographystyle{plain}
\bibliography{signaling_model}

\end{document}

### 2015-05-06 compare_nasheq.m

@@ -1,22 +1,0 @@
-comp=zeros(Nc,Nd,2,2);
-
-for i=1:Nc
-    for j=1:Nd
-        n=nasheq_ind{i,j};
-        if isempty(n)
-            n=[0; 0];
-        end
-        n2=nasheq_ind2{i,j};
-        if isempty(n2)
-            n2=[0;0];
-        end
-        comp(i,j,1,:)=n;
-        comp(i,j,2,:)=n2;
-    end
-end
-%%
-j=2;
-plot(col(comp(:,j,1,2)),'-o')
-hold on
-plot(col(comp(:,j,2,2)),'r')
-hold off

### 2015-05-06 dimensionality_comparison.m

@@ -1,22 +1,21 @@
-green=importdata('green.txt');
-blue=importdata('blue.txt');
-red=importdata('red.txt');
-black=importdata('black.txt');
-%%
+greens=importdata('green.txt');
+blues=importdata('blue.txt');
+reds=importdata('red.txt');
+blacks=importdata('black.txt');
+load twomat.mat
+load onemat.mat
+load variables.mat
+
 axislw=1.25;
 lw=3;
 labfontsz=15 ;
 textfontsz=15;
+fontname='Times';
 alpha=.7;
 ticklength=[.02,.0250];
 markersz=5;
 markercol='white';
-%%
-k=2;
-i=3;
 
-f2=interp1(reshape(diag(reshape(twomat(k,i,:,:,2),Nt,[])),1,[]),reshape(diag(reshape(twomat(k,i,:,:,1),Nt,[])),1,[]),'linear','pp');
-f1=interp1(reshape(diag(reshape(onemat(k,i,:,:,2),Nt1,[])),1,[]),reshape(diag(reshape(onemat(k,i,:,:,1),Nt1,[])),1,[]),'linear','pp');
 %%
 
 figure
@@ -28,13 +27,16 @@ h=w/2;
 set(gcf,'Units','inches');
 set(gcf,'Position',[v(1) v(2) w h]);
 
-textlabels={'a' 'b'};
+textlabels={'(a)' '(b)'};
 
 p=zeros(1,2*Nd);
 
 currblue=mycolors(Nd,'blue');
 currred=mycolors(Nd,'red');
 currblack=mycolors(Nd,'black');
+% currblue=cbrewer('seq','Blues',Nd);
+% currred=cbrewer('seq','Reds',Nd);
+% currblack=cbrewer('seq','Grays',Nd);
 
 for k=1:2
     subplot_tight(1,4,(1:2)+2*(k-1),[.1 .07])
@@ -43,37 +45,45 @@ for k=1:2
     switch k
         case 2
             for i=2:Nd
-            p(i)=plot(reshape(diag(reshape(twomat(k,i,:,:,2),Nt,[])),1,[]),reshape(diag(reshape(twomat(k,i,:,:,1),Nt,[])),1,[]),'Color',currblue(i,:),'LineWidth',lw);
[…]

### 2015-05-06 effect_of_group_size.m

@@ -1,0 +1,68 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+load variables.mat
+load twomat.mat
+
+sizes=[500];
+Ns=length(sizes);
+
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+endvals=zeros(1,Nc);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    endvals(i+1)=layers{i+1}(end);
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+Nlays=Nc;
+Nc=length(X);
+
+leak=2;
+
+accuracy=zeros(Ns,Nlays,2);
+bottomaccuracy=zeros(Ns,Nlays,2);
+topaccuracy=zeros(Ns,Nlays,2);
+
+for s=1:Ns
+    N=sizes(s);
+    q=N/4;
+    topquart=1:q;
+    bottomquart=fliplr(N+1-(1:q));
+    for l=1:Nlays
+        for k=1:2
+        if k==1
+            i=layers{l}(1);
+        else i=layers{l}(end);
+        end
+        c1=X(i);
+        c2=Y(i);
+        c3=Z(i);
+        t=group_props_parallelized(c1,c2,c3,its,N,'unif',Nd,Nt,twomat,domvals,threshvals,leak);
+        accuracy(s,l,k)=t{4};
+        totalsigmat=t{11};
+        bottomaccuracy(s,l,k)=sum(sum(triu(totalsigmat(bottomquart,bottomquart),1)))/sum(sum(totalsigmat(bottomquart,bottomquart)));
+        topaccuracy(s,l,k)=sum(sum(triu(totalsigmat(topquart,topquart),1)))/sum(sum(totalsigmat(topquart,topquart)));
+        end
+    end
[…]

### 2015-05-06 effect_of_group_size_check.m

@@ -1,0 +1,30 @@
+N=20;
+q=N/4;
+topquart=1:q;
+bottomquart=fliplr(N+1-(1:q));
+
+layerstocheck=[1:3];
+Nlaystocheck=length(layerstocheck);
+acc_tocheck=zeros(Nlaystocheck,1);
+bottomacc_tocheck=zeros(Nlaystocheck,1);
+topacc_tocheck=zeros(Nlaystocheck,1);
+
+for l=1:Nlaystocheck
+    i=layers{l}(1);
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    t=group_props_parallelized(c1,c2,c3,its,N,'unif',Nd,Nt,twomat,domvals,threshvals,leak);
+    acc_tocheck(l)=t{4};
+    totalsigmat=t{11};
+    bottomacc_tocheck(l)=sum(sum(triu(totalsigmat(bottomquart,bottomquart),1)))/sum(sum(totalsigmat(bottomquart,bottomquart)));
+    topacc_tocheck(l)=sum(sum(triu(totalsigmat(topquart,topquart),1)))/sum(sum(totalsigmat(topquart,topquart)));
+end
+
+'Finished'
+
+%%
+bottomquart=(N+1-(1:q));
+a=sum(sum(triu(totalsigmat(bottomquart,bottomquart),1)));
+b=sum(sum(totalsigmat(bottomquart,bottomquart)));
+[a b a/b]

### 2015-05-06 effect_of_group_size_par.m

@@ -1,0 +1,80 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+load variables.mat
+load twomat.mat
+
+sizes=[100 500 1000];
+Ns=length(sizes);
+
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+endvals=zeros(1,Nc);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    endvals(i+1)=layers{i+1}(end);
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+Nlays=Nc;
+Nc=length(X);
+
+leak=2;
+
+accuracy=zeros(Ns,Nlays*2);
+bottomaccuracy=zeros(Ns,Nlays*2);
+topaccuracy=zeros(Ns,Nlays*2);
+
+% N=20;
+% c1=0;c2=0;c3=1;
+% t=group_props_parallelized(c1,c2,c3,its,N,'unif',Nd,Nt,twomat,domvals,threshvals,leak);
+
+for s=1:Ns
+    N=sizes(s);
+    q=N/4;
+    topquart=1:q;
+    bottomquart=fliplr(N+1-(1:q));
+    parfor ind=1:Nlays*2
+        [l,k]=ind2sub([Nlays 2],ind);
+%     for l=1:Nlays
+%         for k=1:2
+        if k==1
+            i=layers{l}(1);
+        else i=layers{l}(end);
+        end
+        c1=X(i);
+        c2=Y(i);
+        c3=Z(i);
+        t=group_props_parallelized(c1,c2,c3,its,N,'unif',Nd,Nt,twomat,domvals,threshvals,leak);
+        accuracy(s,ind)=t{4};
+        totalsigmat=t{11};
+        bottomaccuracy(s,ind)=sum(sum(triu(totalsigmat(bottomquart,bottomquart),1)))/sum(sum(totalsigmat(bottomquart,bottomquart)));
[…]

### 2015-05-06 example_nash_optimization.m

@@ -1,0 +1,269 @@
+function [optthresh, dmatssaved, pairwiseaccuracy, pairwisetime] = example_nash_optimization(cvals,domvals,twomat)
+
+its=100;
+N=20;
+dist='unif';
+Nd=size(twomat,2);
+Nt=size(twomat,3);
+extendeddomvals=[1-domvals(end:-1:2) domvals];
+v=1:(Nd+Nd-1);
+deps=.05;
+l=1;
+opt_its=20;
+
+totalc=size(cvals,1);
+
+optthresh=zeros(its,totalc+5,N);
+dmatssaved=zeros(its,N,N);
+% perfvalssaved=zeros(its,totalc+1,N);
+pairwiseaccuracy=zeros(its,totalc+5,N,N);
+pairwisetime=zeros(its,totalc+5,N,N);
+
+for c=1:its
+
+    fighting_abilities=20*rand(1,N);
+    fighting_abilities=sort(fighting_abilities,'descend');
+
+    dmat=zeros(N);
+    
+    for i=1:N
+        for j=[1:(i-1),(i+1):N]
+            diff=fighting_abilities(i)-fighting_abilities(j);
+            d=exp(diff)/(exp(diff)+1);
+            d=floor(round(d/deps))*deps;
+            dmat(i,j)=v(abs(d-extendeddomvals)<=deps/2);
+        end
+    end
+
+    dmatssaved(c,:,:)=dmat;
+    
+    for ic=1:totalc
+%     for ic=2:-1:1
+ 
+        c1=cvals(ic,1);
+        c2=cvals(ic,2);
+        c3=cvals(ic,3);
+        
+
+        perf=zeros(2,Nd,Nt,Nt); %individual, dominance, thresholds
+
+        perf(1,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(1-twomat(l,2:Nd,:,:,1));
+        perf(2,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(twomat(l,2:Nd,:,:,1));
+
+        perf(1,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(1-twomat(l,1,:,:,1));
+        perf(2,1,:,:)=c2*(twomat(l,1,:,:,2))+c3*(twomat(l,1,:,:,1));
+
[…]

### 2015-05-06 explain_loss_of_average_accuracy.m

@@ -1,0 +1,61 @@
+
+time=0.05;
+c11=0;c21=time;c31=1-time;
+c12=1-time;c22=time;c32=0;
+
+
+% time=0.0;
+% c11=0;c21=time;c31=1-time;
+% c12=1-time;c22=time;c32=0;
+
+cvals=[c11 c21 c31; c12 c22 c32];
+
+its=100;
+[optthresh, dmatssaved, pairwiseaccuracy, pairwisetime] = example_nash_optimization(cvals,domvals,twomat);
+
+Nc=size(cvals,1);
+ER=zeros(its,Nc+5,N,N);
+DT=zeros(its,Nc+5,N,N);
+DP=zeros(its,Nc+5,N,N);
+
+for i=1:its
+    for j=1:(Nc+5)
+        for k=1:N
+            for l=setdiff(1:N,k)
+                m1=min(k,l);
+                m2=max(k,l);
+            ER(i,j,k,l)=pairwiseaccuracy(i,j,m2,m1);
+            DT(i,j,k,l)=pairwisetime(i,j,k,l);
+            DP(i,j,k,l)=pairwiseaccuracy(i,j,k,l);
+            end
+        end
+    end
+end
+
+ERmean=sum(ER,4)/N;
+DTmean=sum(DT,4)/N;
+DPmean=sum(DP,4)/N;
+
+i=1;
+c1=cvals(i,1);
+c2=cvals(i,2);
+c3=cvals(i,3);
+totalperf1=c1*ERmean+c2*DTmean+c3*(1-DPmean);
+
+i=2;
+c1=cvals(i,1);
+c2=cvals(i,2);
+c3=cvals(i,3);
+totalperf2=c1*ERmean+c2*DTmean+c3*(1-DPmean);
+
+%%
+i=i+1;
+k=16;
+M=max(max(max(ER(i,1,k:end,k:end))),max(max(ER(i,2,k:end,k:end))));
+subplot(1,2,1)
+imagesc(reshape(ER(i,1,k:end,k:end),N-k+1,[]));colorbar
+caxis manual;caxis([0 M])
+subplot(1,2,2)
+imagesc(reshape(ER(i,2,k:end,k:end),N-k+1,[]));colorbar
+caxis manual;caxis([0 M])
+

### 2015-05-06 figures_extra_v2.m

@@ -1,0 +1,396 @@
+%%
+axislw=1.25;
+lw=3;
+trilw=2;
+doubletrilw=1;
+markersz=3;
+textfontsz=12;
+labfontsz=9;
+fontname='Times New Roman';
+yellow=[1 1 0];
+% red=[.9 0 0];
+% green=[0 .7 0];
+% blue=[0 0 .9];
+% greens=importdata('green.txt');
+% blues=importdata('blue.txt');
+% reds=importdata('red.txt');
+% blacks=importdata('black.txt');
+mycolormap=cbrewer('seq','Reds',64);
+mycolormap=mycolormap(1:end,:);
+cornerlabels={'Error rate','Personal preference','Decision time'};
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+
+load /Users/eleanorbrush/Dropbox/signaling_network/twomat.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/variables
+load /Users/eleanorbrush/Dropbox/signaling_network/optanalysis_output.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/infogrowth_output.mat
+load twomat.mat
+load onemat.mat
+load variables.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/groupsize.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/homogenabilities.mat
+
+%% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+endvals=zeros(1,Nc);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    endvals(i+1)=layers{i+1}(end);
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+indices=1:Nt;
+leak=2;
+%% figure: dimensionality comparison 
+load /Users/eleanorbrush/Dropbox/signaling_network/variables
+figure
+set(gcf,'Color','w')
[…]

### 2015-05-06 figures_forlabtea.m

@@ -1,0 +1,450 @@
+%% plotting parameters
+axislw=1.25;
+lw=3;
+textfontsz=12;
+labfontsz=10;
+fontname='Times';
+yellow=[1 1 0];
+red=[.9 0 0];
+green=[0 .7 0];
+blue=[0 0 .9];
+% mycolormap=cbrewer('seq', 'YlOrRd', 100);
+% mycolormap=cbrewer('seq','BuPu',100);
+mycolormap=cbrewer('seq','Reds',64);
+mycolormap=mycolormap(1:end,:);
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+numpowerfuns=length(powerfuns);
+load /Users/eleanorbrush/Dropbox/signaling_network/variables
+%% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+indices=1:Nt;
+leak=2;
+
+%% find how nash equilibria depend on c1,c2,c3
+
+nasheq_ind=cell(Nc,Nd);
+giveup=zeros(Nc,Nd);
+nashaccuracy=zeros(Nc,Nd);
+
+for i=1:Nc
+    c1=X(i);
+    c2=Y(i);
+    c3=Z(i);
+    
+    perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+
+    perf(1,2:Nd,:,:)=c1*(1-twomat2(leak,2:Nd,:,:,1))+c2*(twomat2(leak,2:Nd,:,:,2))+c3*(1-twomat2(leak,2:Nd,:,:,1));
[…]

### 2015-05-06 figures_nonzeroleak.m

@@ -26,6 +26,7 @@ funstoplot=[1 2 4 6];
 colorperm=[1 2 4 3];
 numpowerfuns=length(powerfuns);
 load /Users/eleanorbrush/Dropbox/signaling_network/variables
+load /Users/eleanorbrush/Dropbox/signaling_network/optanalysis_output.mat
 %% coordinate set up
 N=20;
 its=1000;
@@ -207,7 +208,7 @@ set(gcf,'PaperSize',[w h]);
 set(gcf,'PaperPosition',[-w*.55 0 w*2 h]);
 
 filename=strcat('/Users/eleanorbrush/Desktop/','pairwise_accuracy','.pdf');
-print(filename,'-dpdf','-r300');
+% print(filename,'-dpdf','-r300');
 
 
 %% nasheq figure
@@ -276,7 +277,7 @@ set(gcf,'PaperSize',[w h]);
 set(gcf,'PaperPosition',[0 0 w h]);
 
 filename=strcat('/Users/eleanorbrush/Desktop/','nasheq_thresholds','.pdf');
-print(filename,'-dpdf','-r300');
+% print(filename,'-dpdf','-r300');
 
 
 %% group accuracy triangle heatmap figure
@@ -305,7 +306,7 @@ set(gcf,'PaperSize',[w h]);
 set(gcf,'PaperPosition',[0 0 w h]);
 
 filename=strcat('/Users/eleanorbrush/Desktop/','groupaccuracy','.pdf');
-print(filename,'-dpdf','-r300');
+% print(filename,'-dpdf','-r300');
 
 %% group eq thresholds figure
 figure
@@ -383,11 +384,11 @@ for I=1:toplot
         case 1
             ylim=get(gca,'Ylim');
             [x,y]=data2norm(0,ylim(2));
-            annotation('textbox',[x-.04 .95 .05 .04],'String',['(' alphabet(I) ')'],'FitBoxToText','on','FontSize',textfontsz,'FontName',fontname,'EdgeColor','none','VerticalAlignment','middle','HorizontalAlignment','left')
+            annotation('textbox',[x-.04 .95 .05 .04],'String',['(' alphabet(I+2) ')'],'FitBoxToText','on','FontSize',textfontsz,'FontName',fontname,'EdgeColor','none','VerticalAlignment','middle','HorizontalAlignment','left')
         case 0
             ylim=get(gca,'Ylim');
             [x,y]=data2norm(0,ylim(2));
[…]

### 2015-05-06 figures_v4.m

@@ -1,0 +1,626 @@
+%% plotting parameters
+axislw=1.25;
+lw=3;
+trilw=1;
+doubletrilw=1;
+markersz=3;
+textfontsz=8;
+labfontsz=6;
+fontname='Helvetica';
+yellow=[1 1 0];
+red=[.9 0 0];
+green=[0 .7 0];
+blue=[0 0 .9];
+% mycolormap=cbrewer('seq', 'YlOrRd', 100);
+% mycolormap=cbrewer('seq','BuPu',100);
+mycolormap=cbrewer('seq','Reds',64);
+mycolormap=mycolormap(1:end,:);
+% fourcols=[yellow;red;green;blue];
+fourcols=cbrewer('qual','Set1',4);
+fourcols=fourcols([3 1 2 4],:);
+fourcols_light=brighten(fourcols,.75);
+alphabet=['a' 'b' 'c' 'd' 'e' 'f' 'g' 'h'];
+romans={'i', 'ii', 'iii', 'iv'}';
+powerfuns={'D','R','Delta','H','Pi''C','probs'};
+fourpowerfuns={'D','R','H','C'};
+funtypes={'number of signals','entropy','number of signalers','eigenvector centrality'};
+cornerlabels={'Error rate','Decision preference','Decision time'};
+cornerlabels_short={'ER','DP','DT'};
+markertypes={'-o','-d','-v','-s'};
+funstoplot=[1 2 4 6];
+colorperm=[1 2 4 3];
+numpowerfuns=length(powerfuns);
+load /Users/eleanorbrush/Dropbox/signaling_network/twomat.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/variables
+load /Users/eleanorbrush/Dropbox/signaling_network/optanalysis_output.mat
+load /Users/eleanorbrush/Dropbox/signaling_network/infogrowth_output.mat
+
+% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+endvals=zeros(1,Nc);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    endvals(i+1)=layers{i+1}(end);
+    X=[X, vec(1:(Nc-i))];
[…]

### 2015-05-06 group_props_time.m

@@ -182,9 +182,22 @@ for c=1:its
     
     for t=1:(Ntime-1)
         timematnow=zeros(N,N);
+        
         timematnow(timemat<=timevec(t))=1;
+        numsigsnow=zeros(N,N);
+        for i=1:N
+            for j=[1:(i-1),(i+1):N]
+                if timemat(i,j)<=timevec(t)
+                    numsigsnow(i,j)=(timevec(t)-timemat(i,j))/(max(max(timemat))+1-timemat(i,j))*sigmat(i,j); %linear growth of signals
+                end
+            end
+        end
+%         numsigsnow=timematnow;
+        
         binsigmatnow=binsigmat.*timematnow;
-        sigmatnow=sigmat.*timematnow;
+        
+        sigmatnow=binsigmatnow.*numsigsnow;
+        
         [powervecDnow,powervecRnow,powervecDeltanow,powervecHnow,powervecPinow,powervecCnow] = powerfuns(binsigmatnow,sigmatnow);
         powervec=powervecDnow;
     %     powerdiff=ceil(max(powervec))-floor(min(powervec))+.1;

### 2015-05-06 homogen_analysis.m

@@ -1,0 +1,44 @@
+meanvec=mean(reshape(finalthreshmat,20000,66),1);
+varvec=var(reshape(finalthreshmat,20000,66),1);
+cornerlabels={'Error rate','Personal preference','Decision time'};
+
+%% coordinate set up
+N=20;
+its=1000;
+cstep=.1;
+vec=0:cstep:1;
+Nc=length(vec);
+endvals=zeros(1,Nc);
+
+% X=repmat(0:cstep:1,1,11);
+% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
+
+X=[];
+Y=[];
+layers=cell(Nc,1);
+for i=0:(Nc-1)
+    layers{i+1}=length(X)+(1:(Nc-i));
+    endvals(i+1)=layers{i+1}(end);
+    X=[X, vec(1:(Nc-i))];
+    Y=[Y,vec(i+1)*ones(1,Nc-i)];   
+end
+Z=1-X-Y;
+
+transformed=transform([X',Z',Y']);
+Nc=length(X);
+
+indices=1:Nt;
+leak=2;
+
+mainlabel='';
+
+triimage(varvec,transformed,cornerlabels,mainlabel,'off');
+
+%%
+k=layers{2}(5);
+figure
+hold on
+
+for i=1:1000
+    plot(sort(finalthreshmat(:,i,k)))
+end

### 2015-05-06 infogrowth.m

@@ -4,9 +4,6 @@ cstep=.1;
 vec=0:cstep:1;
 Nc=length(vec);
 
-% X=repmat(0:cstep:1,1,11);
-% Y=reshape(repmat(0:cstep:1,11,1),1,[]);
-
 X=[];
 Y=[];
 layers=cell(Nc,1);
@@ -23,9 +20,8 @@ Nc=length(X);
 indices=1:Nt;
 leak=2;
 
-% timevec=.1:.1:5;
-% Ntime=length(timevec);
 %%
+Ntime=100;
 mut_time=zeros(Nc,6,Ntime);
 
 for i=1:Nc
@@ -38,14 +34,9 @@ for i=1:Nc
     mut_time(i,:,:)=t;
 end
 %%
-filename='/Users/eleanorbrush/Desktop/infogrowth_output.mat';
-% save(filename,'mut_time');
-%%
-% diag=1:(N+1):N^2;
-% offdiag=setdiff(1:(N^2),diag);
 layerstoplot=[1 3 5];
 Nlays=length(layerstoplot);
-Ntime=100;
+Ntime=50;
 mut_time=zeros(Nlays,6,Ntime+1);
 for I=1:Nlays
     i=layers{layerstoplot(I)}(1);
@@ -60,6 +51,10 @@ for I=1:Nlays
     t=group_props_time(c1,c2,c3,its,N,'unif',Nd,Nt,twomat2,domvals,threshvals,leak,timevec);
     mut_time(I,:,:)=t;
 end
+%%
+filename='/Users/eleanorbrush/Desktop/infogrowth_output.mat';
+% save(filename,'mut_time','layerstoplot','Nlays','Ntime',);
+
 %% horizontal time plots look like crap
 % figure;
 % set(gcf,'Color','w')
@@ -135,6 +130,7 @@ xpos=.3;
 
 
 layerstoplot=[1 3 7];
+Nlays=length(layerstoplot);
 % linesty={'-.','--','-'};
 linesty={'-','-','-'};
 p=zeros(1,7);
@@ -155,7 +151,7 @@ for now=Nlays:-1:1
     fourcolsnow=hsv2rgb(fourcolsnow);
     for i=1:4
     I=funstoplot(i);
-    p(i+3)=plot((log(timevec)),[0; (col(mut_time(now,I,2:end)))],'-','Color',fourcolsnow(colorperm(i),:),'LineStyle',linesty{now},'LineWidth',lw);
[…]

### 2015-05-06 paradox_of_mutual_info.m

@@ -43,7 +43,7 @@ for c=1:its
         c3=cvals(ic,3);
         
 
-        perf=zeros(2,Nd,Nt,Nt); %individual, leak, dominance, thresholds
+        perf=zeros(2,Nd,Nt,Nt); %individual, dominance, thresholds
 
         perf(1,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(1-twomat(l,2:Nd,:,:,1));
         perf(2,2:Nd,:,:)=c1*(1-twomat(l,2:Nd,:,:,1))+c2*(twomat(l,2:Nd,:,:,2))+c3*(twomat(l,2:Nd,:,:,1));

### 2015-05-06 solve_pde_2d_init.m

@@ -1,0 +1,296 @@
+function [prob time perf]=solve_pde_2d_init(Lx,Ly,d,Ux,Uy,deltax,deltay,l,b,x0,y0)
+% dX1=(-lX1+b*(2d-1))dt+b*sqrt(d)dW1t-b*sqrt(1-d)dW2t, similar for dX2
+% simplified PDE written as
+% 0=(-lX+a)df/dx+(-ly+a)df/dy+D1d^2f/dx^2-2Ddd^2f/dxdy+D2d^2f/dy^2
+if nargin==2 
+    d=.5;
+    Ux=5;
+    Uy=5;
+    deltax=.01;
+    deltay=.05;
+    l=1;
+    b=1;
+end
+
+if nargin==3
+    Ux=5;
+    Uy=5;
+    deltax=.05;
+    deltay=.05;
+    l=1;
+    b=1;
+end
+
+D1=1/2*b^2;
+Dd=-1/2*b^2;
+D2=1/2*b^2;
+a=b*(2*d-1);
+
+%% setting up vectors I need
+
+xvals=Lx:deltax:Ux;
+
+yvals=Ly:deltay:Uy;
+
+Ni=length(yvals);
+Nj=length(xvals);
+
+c=xvals(2:(end))';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+xvec = cc(:)';
+
+c=[1:(Nj-1)]';
+cc=c(:,ones(Ni-1,1));
+cc=transpose(cc);
+jvals = cc(:)';
+
+c=yvals(2:(end))';
+cc=c(:,ones(Nj-1,1));
+yvec = cc(:)';
+
+c=[1:(Ni-1)]';
+cc=c(:,ones(Nj-1,1));
+ivals = cc(:)';
+
+%% construct adjacency matrix
+A=sparse((Ni-1)*(Nj-1),(Ni-1)*(Nj-1));
+
+Aleft=D1/(deltax^2);
+Adown=D2/(deltay^2);
+Aupright=2*Dd/deltax/deltay;
+
+bl_prob=0;
+br_prob=1;
+bd_prob=1;
+bu_prob=0;
+b_prob=zeros(1,(Ni-1)*(Nj-1));
+
+bl_time=0;
+br_time=0;
+bd_time=0;
+bu_time=0;
+b_time=zeros(1,(Ni-1)*(Nj-1));
+
[…]

### 2015-05-06 triimage_returnhandles.m

@@ -1,0 +1,61 @@
+function [mybar,yl]=triimage_returnhandles(vals,coords,cornerlabels,mainlabel,highlight,tohighlight,lw,pointlabels,color,m,M,baron,xoffset,yoffset)
+
+if nargin<=9
+    m=min(vals);
+    M=max(vals);
+end
+
+textfontsz=evalin('base','textfontsz');
+labfontsz=evalin('base','labfontsz');
+fontname=evalin('base','fontname');
+
+c=colormap;
+spaced=linspace(0,1,size(c,1));
+l=size(coords,1);
+cstep=(coords(1,1)-coords(2,1))/2;
+
+% xoffset=.25;
+% yoffset=.2;
+
+hold on
+axis off
+axis equal
+
+for i=1:l
+    p=coords(i,:);
+    valnow=vals(i);
+    valnow=interp1([m M],[0 1],valnow);
+    ind=sum(spaced<=valnow);
+    if ind==size(c,1)
+        valcol=c(end,:);
+    else
+        valcol=c(ind,:)+(valnow-spaced(ind))/(spaced(ind+1)-spaced(ind))*(c(ind+1,:)-c(ind,:));
+    end
+    fill([p(1),p(1)+cstep,p(1)+cstep,p(1),p(1)-cstep,p(1)-cstep],[p(2)-2/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)+2/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep],valcol,'EdgeColor','none');
+end
+
+if strcmp(highlight,'on') && ~strcmp(color,'none')
+    for i=1:length(tohighlight)
+        p=coords(tohighlight(i),:);
+        plot([p(1),p(1)+cstep,p(1)+cstep,p(1),p(1)-cstep,p(1)-cstep,p(1)],[p(2)-2/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)+2/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep,p(2)-2/sqrt(3)*cstep],'Color',color(i,:),'LineWidth',lw);
[…]

### 2015-05-06 write_and_save_data.m

@@ -20,8 +20,8 @@ Nt1=length(threshvals1);
 indices1=1:Nt1;
 
 
-% filename=strcat('variables','.mat');
-% save(filename,'Nl','Nd','Nt','lvals','domvals','threshvals');
+filename=strcat('variables','.mat');
+save(filename,'Nl','Nd','Nt','Nt1','lvals','domvals','threshvals','threshvals1','deltax2','deltax1');
 
 
 twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time

### 2015-05-06 writeandsave_init.m

@@ -1,0 +1,53 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+lvals=0:.1:1;
+Nl=length(lvals);
+
+domvals=.5:.05:1;
+Nd=length(domvals);
+
+threshvals=[.1 .5:.1:2 3:5];
+Nt=length(threshvals);
+
+b=1;
+deltax=0.025;
+
+x0=.1;
+y0=1;
+
+% twomat=zeros(Nl,Nd,Nt,Nt,2); %leak, dominance, thresholds, prob, time
+twomat_par=zeros(Nd*Nt*Nt,2);
+onemat_par=zeros(Nd*Nt*Nt,2);
+
+k=2;
+l=lvals(k);
+
+parfor loopvar=1:Nd*Nt*Nt
+% parfor loopvar=1:3
+% for loopvar=1
+    [u,i,j]=ind2sub([Nd,Nt,Nt],loopvar);
+    
+    d=domvals(u);
+    
+    T1=threshvals(i);
+
+    T2=threshvals(j);
+    [x,y,~]=solve_pde_2d_init(-T1,-T2,d,5,5,deltax,deltax,l,b,x0,y0);
+%     storevar=sub2ind([Nd,Nt,Nt],u,i,j);
+    twomat_par(loopvar,:)=[x y];
+    
+    [x,y,~]=solve_pde_1d_init(-T1,d,T2,deltax,l,b,x0-y0);
+    onemat_par(loopvar,:)=[x y];
+end
+
+twomat_par=reshape(twomat_par,Nd,Nt,Nt,2);
+onemat_par=reshape(onemat_par,Nd,Nt,Nt,2);
+
+save('/home/brush/signaling_network/solve_pde_init.mat','twomat_par','onemat_par','threshvals','Nt','x0','y0')
+
+delete(dellapool);
+
+exit ;
